### 2019-08-16 Paper/article.tex
some improvements for sobol codes

\documentclass{article}
\usepackage[utf8]{inputenc}
\title{Referee report}
\author{}
%\date
\begin{document}
\maketitle
This paper presents a numerical procedure to estimate the sampling error of the Multigrid Multilevel Monte Carlo estimator (MG-MLMC) proposed by Kumar et al. (2017). The key idea behind the MG-MLMC estimator is to use coarse grid solutions from the full multigrid (FMG) solver as samples for the MLMC estimator. The computation of the sampling error for the MG-MLMC estimator is not straightforward as estimators appearing in the MLMC telescopic sum are no longer uncorrelated. The paper employs multiple Quasi-Monte Carlo estimators to numerically approximate this sampling variance. Numerical results are presented using Darcy flow problem with different permeability parameters.

The paper is very well-written and well-structured, presenting the ideas in a very clear way. In my opinion, the sampling error estimator proposed for the Multigrid Multilevel (Quasi-) Monte Carlo estimator is a very nice contribution. However, I have some suggestions and comments which should be addressed in my opinion:

\begin{enumerate}
    \item On page 7, line 248, the assumption (A2) should be $V_\ell \leq h^{\beta}_{\ell}$?
    \item Are constants $c_\alpha,c_\beta,c_\gamma,c_\lambda$ dependent on the rates $\alpha,\beta,\gamma,\lambda}$ respectively ?  If not then consider changing the notation to avoid any confusion.

    \item On page 10, instead of explaining the two grid correction scheme, consider describing the FMG algorithm and specifying the inter-grid operators as well as the FMG-interpolation operator that is used to transfer solution field to the next finer level.

    \item On page 10, you say typical values for number of relaxation steps are $\mu_1=2$ and $\mu_2=1$. This is not correct. They are problem dependent parameters and can be decided using Local Fourier analysis, see ref 30 of your paper.

    \item There is no explanation on how the coarse grid operators for FMG $\mathbf{A}_{\ell-1},\mathbf{A}_{\ell-2},$ etc. are  computed. Is it by direct-discretization or using Galerkin appraoach, $A_{\ell-1} =R_{\ell}^{\ell-1}A_\ell I^{\ell}_{\ell-1}$?

    \item Related to the previous comment, you did not mention how do you upscale the random field from the finest grid level to all the coarser grids. This should be done in a way such that the statistical properties of the random field does not change. If you employ direct discretization then using the KL-expansion can avoid this problem. In case of Galerkin approach this is not so trivial. Similar problem will arise if you use an algebraic multigrid solver.

    \item How do you ensure that solutions on the coarser FMG levels converge to respective discretization accuracies? In page 11, you introduce the number of multilevel correction cycles $\mu_0$ but it is not clearly explained whether this parameter is a constant or varies depending on the grid level? Usually, the number of cycles needed for convergence depends on the grid as well as the Mat\'ern parameter as shown in this article:

    Kumar, P., Rodrigo, C., Gaspar, F. J., & Oosterlee, C. W. (2018). On cell-centered multigrid methods and Local Fourier Analysis for PDEs with random coefficients. arXiv preprint arXiv:1803.08864.

    Alternately, the number of multilevel correction cycles can be determined by monitoring residuals on different FMG levels.

    \item On page 14, based on what criteria did you decide the number of random shifts? For smaller tolerances $\epsilon$, $R=20$ may not be enough to approximate $V[\overline{\mathcal{Q}}^{QMC}_{L,R,reuse}]$.

    \item On page 22, line 532, the percentage of recycled samples should be $\frac{N_{\ell+1}}{N_\ell}\cdot 100$ ?

\end{enumerate}
$$
k(x,y) =
\left\{
\begin{array}{l}
10^4, \quad {\mbox {\rm if} } \quad (x,y) \in \left(
\frac{2}{128}\frac{5}{128}\right)\times\left(\frac{2}{128},\frac{3}{128}\right)\cup\left(
\frac{2}{128},\frac{3}{128}\right)\times\left( \frac{2}{128}\frac{5}{128}\right), \\
1, \quad {\mbox {\rm otherwise}}.
\end{array}
\right.
$$
\end{document}

### 2019-08-16 sensitivity_analysis/cases/aero_module/aero_module.m

@@ -10,8 +10,6 @@ randVecChord(P{31}) = X(length(P{30})+1:end);
 
 twist = computeTwist(1, randVecTwist, P{28}, 1); % computeTwist routine uses the specifications of NM80 turbine by default
 chord = computeChord(1, randVecChord, P{29}, 1); % computeChord routine uses the specifications of NM80 turbine by default
-
-
 %% Write the random data into the input file of AERO module
 writeAeroModuleInput(P{1},P{2},P{10}, P{3}, chord, ...
                      P{5},  twist, P{7},  P{8},  P{9},... 

### 2019-09-02 Paper/article.tex
incorporated YAW angle in sensitivity analysis

@@
-\documentclass{article}
+\documentclass[review]{elsarticle}
+\usepackage{lineno,hyperref}
+\modulolinenumbers[5]
+\journal{ }
+\usepackage{microtype}
+\usepackage{algorithm}
+\usepackage[noend]{algpseudocode}
+\usepackage[a4paper, total={6.2in, 9in}]{geometry}
+\usepackage[a4paper]{geometry}
+\renewcommand{\baselinestretch}{1.2} 
+\usepackage{color}
+\usepackage[x11names]{xcolor}
 \usepackage[utf8]{inputenc}
-\title{Referee report}
-\author{}
-%\date
+\usepackage{amsmath}
+\usepackage{tabularx,ragged2e,booktabs,caption}
+\renewcommand{\figurename}{{Figure}}
+\usepackage[labelfont=bf]{caption}
+\usepackage[font=footnotesize]{caption}
+\usepackage{amssymb,latexsym}
+\usepackage{graphicx}
+\usepackage{booktabs}
+\usepackage{amsfonts}
+\usepackage{tikz}
+\usepackage{array}
+\usepackage{mathtools}
+\usepackage{relsize}
+\usepackage[T1]{fontenc}
+\usepackage{float}
+\usepackage{cases}
+\usepackage{amsthm}
+\usepackage{subcaption}
+\usepackage{multirow}
+\usepackage[toc,page]{appendix}
+\newcolumntype{Y}{>{\centering\arraybackslash}X}
+\newtheorem{theorem}{Theorem}
+\newtheorem{remark}{Remark}[section]
+\numberwithin{equation}{section}
+\usepackage{tabularx,ragged2e,booktabs,caption}
+\newcommand{\footremember}[2]{%
+    \footnote{#2}
+    \newcounter{#1}
+    \setcounter{#1}{\value{footnote}}%
+}
+\newcommand{\footrecall}[1]{%
+    \footnotemark[\value{#1}]%
+}
+\usepackage[toc]{appendix}
+\DeclareMathAlphabet{\mathpzc}{OT1}{pzc}{m}{it}
+\newcommand{\lnorm}{\left|\left|}
+\newcommand{\rnorm}{\right|\right|}
+\newcommand{\Ld}{L^2(\mathcal{D})}
+\newcommand{\Lomtd}{L^2(\Omega;\mathcal{T};\mathcal{D})}
+%\newcommand{\Lomd}{L^2(\Omega;L^2(\mathcal{D}))}
+\newcommand{\Lomd}{L^2(\Omega,\mathcal{D})}
+\newcommand{\myreferences}{../Bibliography/references}
+\bibliographystyle{elsarticle-num}
+\usepackage{color}
+\usepackage[utf8]{inputenc}
+\usepackage{amsmath}
+\usepackage{tabularx,ragged2e,booktabs,caption}
+\renewcommand{\figurename}{{Fig.}}
+\usepackage[labelfont=bf]{caption}
+\usepackage[font=footnotesize]{caption}
+\usepackage{amssymb,latexsym}
+\usepackage{graphicx}
+\usepackage{booktabs}
+\usepackage{amsfonts}
+\usepackage{tikz}
+\usepackage{float}
+\usepackage{array}
+\usepackage{mathtools}
+\usepackage{relsize}
+\usepackage[T1]{fontenc}
+\usepackage{float}
+\usepackage{cases}
+\usepackage{amsthm}
+\usepackage{subcaption}
+\usepackage[toc,page]{appendix}
+\newcolumntype{Y}{>{\centering\arraybackslash}X}
+\numberwithin{equation}{section}
+\usepackage{tabularx,ragged2e,booktabs,caption}
+\graphicspath{{./images/}}
+\usepackage[toc]{appendix}
+\DeclareMathAlphabet{\mathpzc}{OT1}{pzc}{m}{it}
+\bibliographystyle{elsarticle-num}
+\usepackage{stix2}
+\usepackage{pgfplots}
+\usepackage{appendix}
+\usepackage[labelfont=bf]{caption}
+\pgfplotsset{compat=newest}
+\pagestyle{myheadings}
+%%%%%%%%%%%%%%%%%%%%%%%
+\title{A comprehensive global sensitivity analysis study of wind turbine aeroelastic models}
 \begin{document}
-\maketitle
-This paper presents a numerical procedure to estimate the sampling error of the Multigrid Multilevel Monte Carlo estimator (MG-MLMC) proposed by Kumar et al.
-(2017).
-The key idea behind the MG-MLMC estimator is to use coarse grid solutions from the full multigrid (FMG) solver as samples for the MLMC estimator.
-The computation of the sampling error for the MG-MLMC estimator is not straightforward as estimators appearing in the MLMC telescopic sum are no longer uncorrelated.
-The paper employs multiple Quasi-Monte Carlo estimators to numerically approximate this sampling variance.
-Numerical results are presented using Darcy flow problem with different permeability parameters.
-
-The paper is very well-written and well-structured, presenting the ideas in a very clear way.
-In my opinion, the sampling error estimator proposed for the Multigrid Multilevel (Quasi-) Monte Carlo estimator is a very nice contribution.
-However, I have some suggestions and comments which should be addressed in my opinion:
-
-\begin{enumerate}
-    \item On page 7, line 248, the assumption (A2) should be $V_\ell \leq h^{\beta}_{\ell}$?
-    \item Are constants $c_\alpha,c_\beta,c_\gamma,c_\lambda$ dependent on the rates $\alpha,\beta,\gamma,\lambda}$ respectively ?  If not then consider changing the notation to avoid any confusion.
-
-    \item On page 10, instead of explaining the two grid correction scheme, consider describing the FMG algorithm and specifying the inter-grid operators as well as the FMG-interpolation operator that is used to transfer solution field to the next finer level.
-
-    \item On page 10, you say typical values for number of relaxation steps are $\mu_1=2$ and $\mu_2=1$.
-This is not correct.
-They are problem dependent parameters and can be decided using Local Fourier analysis, see ref 30 of your paper.
-
-    \item There is no explanation on how the coarse grid operators for FMG $\mathbf{A}_{\ell-1},\mathbf{A}_{\ell-2},$ etc.
-are  computed.
-Is it by direct-discretization or using Galerkin appraoach, $A_{\ell-1} =R_{\ell}^{\ell-1}A_\ell I^{\ell}_{\ell-1}$?
-
-    \item Related to the previous comment, you did not mention how do you upscale the random field from the finest grid level to all the coarser grids.
-This should be done in a way such that the statistical properties of the random field does not change.
-If you employ direct discretization then using the KL-expansion can avoid this problem.
-In case of Galerkin approach this is not so trivial.
-Similar problem will arise if you use an algebraic multigrid solver.
-
-    \item How do you ensure that solutions on the coarser FMG levels converge to respective discretization accuracies?
-In page 11, you introduce the number of multilevel correction cycles $\mu_0$ but it is not clearly explained whether this parameter is a constant or varies depending on the grid level?
-Usually, the number of cycles needed for convergence depends on the grid as well as the Mat\'ern parameter as shown in this article:
-
-    Kumar, P., Rodrigo, C., Gaspar, F. J., & Oosterlee, C. W. (2018). On cell-centered multigrid methods and Local Fourier Analysis for PDEs with random coefficients. arXiv preprint arXiv:1803.08864.
-
-    Alternately, the number of multilevel correction cycles can be determined by monitoring residuals on different FMG levels.
-
-    \item On page 14, based on what criteria did you decide the number of random shifts?
-For smaller tolerances $\epsilon$, $R=20$ may not be enough to approximate $V[\overline{\mathcal{Q}}^{QMC}_{L,R,reuse}]$.
-
-    \item On page 22, line 532, the percentage of recycled samples should be $\frac{N_{\ell+1}}{N_\ell}\cdot 100$ ?
-
-\end{enumerate}
-$$
-k(x,y) =
-\left\{
-\begin{array}{l}
-10^4, \quad {\mbox {\rm if} } \quad (x,y) \in \left(
-\frac{2}{128}\frac{5}{128}\right)\times\left(\frac{2}{128},\frac{3}{128}\right)\cup\left(
-\frac{2}{128},\frac{3}{128}\right)\times\left( \frac{2}{128}\frac{5}{128}\right), \\
-1, \quad {\mbox {\rm otherwise}}.
-\end{array}
-\right.
-$$
+\begin{frontmatter}
+%\address[cwi]{CWI - Centrum Wiskunde \& Informatica, Amsterdam, The Netherlands}
+%\address[aero]{Faculty of Aerospace Engineering, Delft University of Technology, Delft, The Netherlands.}
+%\author[cwi,aero]{Prashant~Kumar}
+\author[]{Prashant~Kumar}
+\ead{pkumar@cwi.nl}
+\begin{abstract}
+\end{abstract}
+\begin{keyword}
+Wind energy, global sensitivity analysis, UQ, aeroelastic models 
+\end{keyword}
+\end{frontmatter}
+\linenumbers
+\section{Introduction}
+\section{BEM model}
+\section{Global sensitivity analysis}
+\subsection{PCE Ordinary Least Square (PCE\_OLS)}
+\subsection{PCE Least Angle Regression (PCE\_LAR)}
+\section{Parametrization of uncertain inputs in the BEM model}
+\subsection{Geometric uncertainty}
+\subsubsection{NURBS based parametrization}
+\subsection{Model uncertainty}
+\subsection{Wind velocity}
+\section{Description of test case}
+\section{Numerical experiment}
+\section{Interpretation of global sensitivity analysis result}
+\section{Conclusions}
+\section*{Acknowledgements}
 \end{document}

### 2019-09-02 sensitivity_analysis/cases/aero_module/aero_module.m

@@ -6,12 +6,12 @@ workingDir = pwd;
 randVecTwist = zeros(1,P{26}); 
 randVecTwist(P{30}) = X(1:length(P{30}));
 randVecChord = zeros(1,P{27});
-randVecChord(P{31}) = X(length(P{30})+1:end);
+randVecChord(P{31}) = X(length(P{30})+1:end-1);
 twist = computeTwist(1, randVecTwist, P{28}, 0); % computeTwist routine uses the specifications of NM80 turbine by default
 chord = computeChord(1, randVecChord, P{29}, 0); % computeChord routine uses the specifications of NM80 turbine by default
 
 %% Get one random samples for YAWANGLE
-YAWANGLE = P{20} + P{32}.*randn(1,1);
+YAWANGLE = P{20} + P{32}.*X(end);
 %% Write the random data into the input file of AERO module
 writeAeroModuleInput(P{1},  P{2},  P{10}, P{3},  chord, ...
                      P{5},  twist, P{7},  P{8},  P{9}, ... 

### 2019-09-02 sensitivity_analysis/cases/aero_module/getParameterAeroModule.m

@@ -8,7 +8,7 @@ P ={};
 [AEROMODEL,TURBINETYPE,zB, ref_chord, t_by_c,ref_twist, C14, xB, yB, vectorLength, ...
     BLADELENGTH, BLADEROOT, HUBHEIGHT, TILTANGLE, PITCHANGLE, XNAC2HUB, ...
     RPM, TEND, TIMESTEP, YAWANGLE, NROFBEMELEMENTS, ZNAC2HUB, NRCP_TWIST, ...
-    NRCP_CHORD, PERTURBATION_TWIST, PERTURBATION_CHORD, INDEX_TWIST,INDEX_CHORD, YAW_MEAN, YAW_VARIANCE]  = turbuineDataNM80();
+    NRCP_CHORD, PERTURBATION_TWIST, PERTURBATION_CHORD, INDEX_TWIST,INDEX_CHORD, YAW_VARIANCE]  = turbuineDataNM80();
 
 %% Other parameters related to AERO module software
 % Top and the end of AeroPower.dat file

### 2019-09-04 Paper/article.tex
updated article

@@
 \usepackage[T1]{fontenc}
-\usepackage{float}
 \usepackage{cases}
@@
 \bibliographystyle{elsarticle-num}
-\usepackage{color}
-\usepackage[utf8]{inputenc}
-\usepackage{amsmath}
-\usepackage{tabularx,ragged2e,booktabs,caption}
-\renewcommand{\figurename}{{Fig.}}
-\usepackage[labelfont=bf]{caption}
-\usepackage[font=footnotesize]{caption}
-\usepackage{amssymb,latexsym}
-\usepackage{graphicx}
-\usepackage{booktabs}
-\usepackage{amsfonts}
-\usepackage{tikz}
-\usepackage{float}
-\usepackage{array}
-\usepackage{mathtools}
-\usepackage{relsize}
-\usepackage[T1]{fontenc}
-\usepackage{float}
-\usepackage{cases}
-\usepackage{amsthm}
-\usepackage{subcaption}
-\usepackage[toc,page]{appendix}
-\newcolumntype{Y}{>{\centering\arraybackslash}X}
 \numberwithin{equation}{section}
@@
 %%%%%%%%%%%%%%%%%%%%%%%
-\title{A comprehensive global sensitivity analysis study of wind turbine aeroelastic models}
+\title{A Global Sensitivity Analysis of Wind Turbine Aeroelastic Models}
 \begin{document}
@@
 \section{Introduction}
-\section{BEM model}
+\section{Aeroelastic models}
+\subsection{BEM models}
+\subsection{Input parameters of aeroelastic models}
 \section{Global sensitivity analysis}
-\subsection{PCE Ordinary Least Square (PCE\_OLS)}
-\subsection{PCE Least Angle Regression (PCE\_LAR)}
+\subsection{Sobol' indices}
+\subsubsection{PCE Ordinary Least Square (PCE\_OLS)}
+\subsubsection{PCE Least Angle Regression (PCE\_LAR)}
 \section{Parametrization of uncertain inputs in the BEM model}
 \subsection{Geometric uncertainty}
-\subsubsection{NURBS based parametrization}
+We use Non-Uniform Rational Basis Splines (NURBS) \cite{nurbs_book} to perturb the geometrical parameters of the turbine blade, such as the reference chord and twist curves.
+The main advantage of using NURBS is that it provides great flexibility to approximate a large variety of curves using a very limited number of control points.
+Further, the set of control points and knots can be directly manipulated to control the smoothness and curvature.
+
+\subsubsection{NURBS based perturbation}
+The value of NURBS curve at each location $x$ is computed using a weighted sum of $N$ basis functions (or B-splines):
+\begin{equation}\label{NURB_curve}
+S(x)  = \sum_{i=0}^{N-1} c_iB_{i,p}(x),
+\end{equation}
+where $S(x)$ is the value of the curve at location $x$, $c_i$ is the weight of control point $i$ and $B_{i,p}(x)$ is the value of B-spline corresponding to  the $i$-th control point at $x$.
+The subscript $p$ denote the polynomial degree of the NURBS curve.
+The total number of control points $N= m+p+1$ where $m$ is the number of knots  and the degree of NURBS curve.
+The above definition can be easily extended to higher dimensions.
+
+The B-splines are recursive in polynomial degree, for example, we can derive quadratic B-spline ($p=2$) using the linear B-spline ($p=1$), cubic ($p=3$) from quadratic B-splines and so on.
+Given $m$ knot locations $t_0,t_1,...,t_{m-1}$, the B-spline of degree 0 is defined as:
+\begin{equation}\label{linearBspline}
+B_{i,0}(x) :=
+\begin{cases}
+1\quad t_i\leq x < t_{i+1},\\
+0\quad\text{elsewhere.}
+\end{cases} 
+\end{equation}
+Higher order B-splines can be then derived using the recurrence relation\cite{deBoor}:
+\begin{equation}\label{NURBS_recurrence}
+B_{i,p}(x):= \frac{x-t_i}{t_{i+p} - t_i}B_{i,p-1}(x) + \frac{t_{i+p+1}  - x}{t_{i+p+1}  - t_{i+1} }B_{i+1,p-1}(x).
+\end{equation}
+These B-splines can be constructed efficiently using the De Boor's algorithm \cite{deBoor}.
+In Fig. \ref{B-splines}, we show the linear, quadratic and cubic spline for $x\in[0,1]$. At the interval boundaries these B-splines go smoothly to zero.
+
+Next, we describe steps to generate perturbed samples of chord from a given reference chord using NURBS based parametrization.
+
 \subsection{Model uncertainty}
@@
 \section{Description of test case}
+\section{Sensitivity analysis workflow}
 \section{Numerical experiment}

### 2019-09-14 Paper/article.tex
added routines for purtubing Thickness and twist curves, and som some results on high dim sensitivity analysis

@@
 \begin{abstract}
+Wind turbines are a complex multi-physics systems. The performance and life-cycle of turbine is influenced by a large number
 \end{abstract}
@@
 
-The B-splines are recursive in polynomial degree, for example, we can derive quadratic B-spline ($p=2$) using the linear B-spline ($p=1$), cubic ($p=3$) from quadratic B-splines and so on.
+The B-splines are recursive in polynomial degree, for example, we can derive quadratic B-spline ($p=2$) using the linear B-spline ($p=1$); cubic ($p=3$) from quadratic B-splines and so on.
 Given $m$ knot locations $t_0,t_1,...,t_{m-1}$, the B-spline of degree 0 is defined as:
@@
 \end{equation}
-Higher order B-splines can be then derived using the recurrence relation\cite{deBoor}:
+Higher order B-splines can then be derived using the recurrence relation \cite{deBoor}:
 \begin{equation}\label{NURBS_recurrence}
-B_{i,p}(x):= \frac{x-t_i}{t_{i+p} - t_i}B_{i,p-1}(x) + \frac{t_{i+p+1}  - x}{t_{i+p+1}  - t_{i+1} }B_{i+1,p-1}(x).
+B_{i,p}(x) := \frac{x-t_i}{t_{i+p} - t_i}B_{i,p-1}(x) + \frac{t_{i+p+1}  -  x}{t_{i+p+1}  -  t_{i+1}}B_{i+1,p-1}(x).
 \end{equation}
 These B-splines can be constructed efficiently using the De Boor's algorithm \cite{deBoor}.
-In Fig. \ref{B-splines}, we show the linear, quadratic and cubic spline for $x\in[0,1]$. At the interval boundaries these B-splines go smoothly to zero.
+In Fig.
+\ref{B-splines}, we show the linear, quadratic and cubic spline for $x\in[0,1]$.
+At the interval boundaries these B-splines go smoothly to zero.
 
-Next, we describe steps to generate perturbed samples of chord from a given reference chord using NURBS based parametrization.
+Next, we describe steps to generate perturbed samples of chord from a given reference chord, $S_{ref}(x)$, using NURBS based parametrization.
+The first step is to approximate the given reference chord using a NURBS curve with a fixed degree $p$.
+For this, we sample $S_{ref}(x)$ at $N$ locations $\{x_i\}_{i=0}^N$ and
+
+ perform inversion to compute $\{c_i\}_{i=0}^N$ by solving the $N\times N$ linear system:
+\begin{equation}\label{nurbs_inversion}
+\mathbf{B}\mathbf{c} = \mathbf{S},
+\end{equation}
+where $\mathbf{B}\in \mathbb{R}^{N\times N}$ with elements $\mathbf{B}_{ij} = B_{}()$
+
+with a \emph{limited} number of control points. The number of control points are decided based on an error threshold criteria between the approximated and the reference chord. 
+\begin{remark}
+How to chose $x$
+\end{remark}
 

### 2019-09-14 Geometry/computeChord.m

@@ -58,7 +58,7 @@ if plotSamples == 1
     hold on
     plot(sampledLocations,c,'marker','o','linewidth',2) % plot control points
     plot(sampledLocations,sampledValues,'marker','x','markersize',8,'linestyle','none','linewidth',2) % plot sampled points
-    plot(interpolationLocations,samplesChord,'linewidth',2,'color','g')
+    plot(interpolationLocations,samplesChord,'linewidth',2,'color','b')
 end
 samplesChord = perturbNURBS(t0,n,interpolationLocations,c, pc,samples,randVec);
 

### 2019-09-14 sensitivity_analysis/cases/aero_module/initialize.m

@@ -17,7 +17,7 @@ addpath('C:\Users\pkumar\Dropbox\WindTrue\windtrue\Geometry\')
 
 % specify a list of options from the following list:
 % methods = {'MC','PCE_Quad','PCE_OLS','PCE_LARS'};
-methods = {'PCE_OLS','PCE_LARS'};
+methods = {'PCE_LARS'};
 % % for Monte Carlo, specify number of times to repeat MC-based methods to obtain 'nice' convergence
 % % graphs
 % MC_repeat = 1;
@@ -27,12 +27,12 @@ methods = {'PCE_OLS','PCE_LARS'};
 % % for PCE-Quad, specify the polynomial degrees to be tested
 % DegreesQuad = 1:4; %[1 2 3 4 5 6];
 
-% % for PCE-OLS:
-NsamplesOLS = [256 512 1024]; % if not specified, the number of samples from Quad is taken
-OLS_repeat = 1; % like MC_repeat
+% % % for PCE-OLS:
+% NsamplesOLS = [16]; % if not specified, the number of samples from Quad is taken
+% OLS_repeat = 1; % like MC_repeat
 % 
 % % % for PCE-LARS:
-NsamplesLARS = [256 512 1024]; % if not specified, the number of samples from Quad is taken
+NsamplesLARS = [256 512 1024 2048]; % if not specified, the number of samples from Quad is taken
 LARS_repeat = 1; % like MC_repeat
 
 %% number of random variables
@@ -46,5 +46,7 @@ for i = 1:ndim-1
     Input.Marginals(i).Bounds = [-0.5 0.5]; 
 end
 
+Input.Marginals(i).Name = 'YAW';
 Input.Marginals(ndim).Type = 'Gaussian'; 
 Input.Marginals(ndim).Parameters = [P{20}, P{35}];
+Input.Marginals(ndim).Bounds = [-10 10]; 

### 2019-09-17 Paper/article.tex
added some text in article

@@
 \begin{abstract}
-Wind turbines are a complex multi-physics systems. The performance and life-cycle of turbine is influenced by a large number
+Wind turbines are a complex multi-physics systems.
+As the performance and the life span of a turbine is highly dependent on  manufacturing, meteorological and operational factors, it is very important to quantify the impact of each of these parameter in the output of interest.
 \end{abstract}
@@
 \section{Global sensitivity analysis}
+The main objective of sensitivity analysis is to quantify the relative significance of individual inputs (or their combinations) and how variations in input parameters (or their combinations) affects the output of interest.
+In engineering, sensitivity analysis can be employed for a number of reasons.
+It can be used to determine the stability and robustness of a computational model with respect to input parameters.
+Sensitivity analysis can be used for simplification of stochastic models by fixing the insensitive parameters.
+It can also be further used to guide data acquisition campaigns and experimental design to refine the data on sensitive parameters.
+
+Sensitivity analysis techniques can be broadly classified as local and global methods, see \cite{ralphSmith}.
+Within local sensitivity analysis, individual parameters are perturbed around their nominal values allowing for the description of output variability in a small neighbourhood of nominal input values.
+% Global sensitivity analysis, on the other hand,
+
 \subsection{Sobol' indices}
@@
 We use Non-Uniform Rational Basis Splines (NURBS) \cite{nurbs_book} to perturb the geometrical parameters of the turbine blade, such as the reference chord and twist curves.
-The main advantage of using NURBS is that it provides great flexibility to approximate a large variety of curves using a very limited number of control points.
+The main advantage of using NURBS is that it provides great flexibility to approximate a large variety of curves with a limited number of control points.
 Further, the set of control points and knots can be directly manipulated to control the smoothness and curvature.
@@
 \begin{equation}\label{NURB_curve}
-S(x)  = \sum_{i=0}^{N-1} c_iB_{i,p}(x),
+S(x)  = \sum_{i=1}^{N} c_iB_{i,p}(x),
 \end{equation}
@@
 The subscript $p$ denote the polynomial degree of the NURBS curve.
-The total number of control points $N= m+p+1$ where $m$ is the number of knots  and the degree of NURBS curve.
+The total number of control points $N = m+p+1$ where $m$ is the number of knots  and the degree of NURBS curve.
 The above definition can be easily extended to higher dimensions.
@@
 The B-splines are recursive in polynomial degree, for example, we can derive quadratic B-spline ($p=2$) using the linear B-spline ($p=1$); cubic ($p=3$) from quadratic B-splines and so on.
-Given $m$ knot locations $t_0,t_1,...,t_{m-1}$, the B-spline of degree 0 is defined as:
+Given $m$ knot locations $t_1, t_2, ..., t_{m}$, the B-spline of degree 0 is defined as:
 \begin{equation}\label{linearBspline}
@@
 \begin{cases}
-1\quad t_i\leq x < t_{i+1},\\
+1\quad t_i\leq x < t_{i+1},\quad i = 1,2,...,m,\\
 0\quad\text{elsewhere.}
@@
 \begin{equation}\label{NURBS_recurrence}
-B_{i,p}(x) := \frac{x-t_i}{t_{i+p} - t_i}B_{i,p-1}(x) + \frac{t_{i+p+1}  -  x}{t_{i+p+1}  -  t_{i+1}}B_{i+1,p-1}(x).
+B_{i,p}(x) := \frac{x-t_i}{t_{i+p} - t_i}B_{i,p-1}(x) + \frac{t_{i+p+1}  -  x}{t_{i+p+1}  -  t_{i+1}}B_{i+1,p-1}(x),\quad p\geq1, i = 1,2,...,m.
 \end{equation} 
+Something about padding...
 These B-splines can be constructed efficiently using the De Boor's algorithm \cite{deBoor}.
 In Fig.
-\ref{B-splines}, we show the linear, quadratic and cubic spline for $x\in[0,1]$.
-At the interval boundaries these B-splines go smoothly to zero.
+\ref{B-splines}, we show linear, quadratic and cubic splines for $x\in[0,1]$.
+At the interval boundaries these B-splines go to zero smoothly.
+The domain of influence of a given control point depends on the respective B-spline (rephrase)??
 
@@
 The first step is to approximate the given reference chord using a NURBS curve with a fixed degree $p$.
-For this, we sample $S_{ref}(x)$ at $N$ locations $\{x_i\}_{i=0}^N$ and
-
- perform inversion to compute $\{c_i\}_{i=0}^N$ by solving the $N\times N$ linear system:
+For this, we need to sample $S_{ref}(x)$ at $N$ locations $\{x_j\}_{j=1}^N$ and compute $B_{i,p}(x_j), \text{ for }i, j = 1,2, ..., N$, such that we can compute the set of control points $\mathbf{c}=\{c_i\}_{i=0}^N$ by solving the following linear system:
 \begin{equation}\label{nurbs_inversion}
@@
 \end{equation}
-where $\mathbf{B}\in \mathbb{R}^{N\times N}$ with elements $\mathbf{B}_{ij} = B_{}()$
+where $\mathbf{S}\in \mathbb{R}^{N}$ is a vector containing sampled values of the reference curve and $\mathbf{B}\in \mathbb{R}^{N\times N}$ is a matrix with $j$-th row consisting of B-splines values at $x_j$ locations, i.e., $B_{i,p}(x_j), i= 1, 2,..., N$.
+Once the control points are obtained, we can derive the approximate reference curve $S_N(x)$ using \eqref{NURB_curve}.
+Note that the accuracy of the approximated curve is dependent on the sample locations $x_j$ as well as the degree $p$ and can be set heuristically \cite{adaptiveNurbs}.
+The number of sampled locations $N$ can be adaptively increased until the following tolerance criteria is met:
+\begin{equation}
+\frac{||S_N - S_{ref}||}{||S_{ref}||} <\varepsilon.
+\end{equation}
+More advanced approaches, monotonicity preserving methods, etc??
+\begin{remark}
+How to chose $x$?
+\end{remark}
+The final step is to introduce uncertainty in the above computed control points by associating a probability distribution.
 
-with a \emph{limited} number of control points. The number of control points are decided based on an error threshold criteria between the approximated and the reference chord. 
-\begin{remark}
-How to chose $x$
-\end{remark}
-
+The steps to obtained perturbed samples from a given reference curve is outlined in Fig. \ref{}.
 \subsection{Model uncertainty}
@@
 \section{Numerical experiment}
-\section{Interpretation of global sensitivity analysis result}
+\subsection{Interpretation of global sensitivity analysis result}
 \section{Conclusions}

### 2019-09-19 Paper/article.tex
added some text in article

@@
 \section{Global sensitivity analysis}
-The main objective of sensitivity analysis is to quantify the relative significance of individual inputs (or their combinations) and how variations in input parameters (or their combinations) affects the output of interest.
-In engineering, sensitivity analysis can be employed for a number of reasons.
-It can be used to determine the stability and robustness of a computational model with respect to input parameters.
-Sensitivity analysis can be used for simplification of stochastic models by fixing the insensitive parameters.
-It can also be further used to guide data acquisition campaigns and experimental design to refine the data on sensitive parameters.
+The main objective of sensitivity analysis is to quantify the relative significance of individual inputs (or in combination) and how variations in input values affects the output of interest.
+In engineering, sensitivity analysis can be employed for a number of reasons: to determine the stability and robustness of a computational model with respect to input parameters, for simplification of stochastic models by fixing the insensitive parameters, and to guide data acquisition campaigns and experimental design to refine the data on sensitive parameters.
 
-Sensitivity analysis techniques can be broadly classified as local and global methods, see \cite{ralphSmith}.
-Within local sensitivity analysis, individual parameters are perturbed around their nominal values allowing for the description of output variability in a small neighbourhood of nominal input values.
-% Global sensitivity analysis, on the other hand,
+Sensitivity analysis techniques can be classified as local and global methods, see \cite{ralphSmith}.
+In local sensitivity analysis, individual parameters are perturbed around their nominal values --- allowing for the description of output variability only in a small neighbourhood of nominal input values.
+Although, local approaches are widely employed due to their ease of implementation and low computational cost, they are unable to quantify global behavior of nonlinearly parametrized models such as aeroelastic models.
+Global sensitivity approaches, on the other hand, consider the entire range of input values to compute output sensitivities.
+Therefore, global sensitivity analysis is more suitable for aeroelastic models considered in this work.
 
-\subsection{Sobol' indices}
+\subsection{Sobol analysis}
+In this work, we employ variance-based Sobol decomposition to perform global sensitivity analysis.
+One of the advantages of this method is that the relative importance of input parameters is quantified on a scale of $[0,1]$ known as \emph{Sobol indices}.
+The main idea of Sobol analysis is to decompose the total variance of the output in term of contributions from individual parameters and their contribution.
+For the ease of exposition, let us consider a nonlinear model:
+\begin{equation}\label{nonlinear_model}
+Y = f(\mathbf{Z}),
+\end{equation}
+where $\mathbf{Z} = [z_1, z_2, ..., z_M]\in \mathcal{D}_{\mathbf{Z}}\in \mathbb{R}^M$ is the input vector.
+For simplicity, we assume input parameters are uniformly distributed i.e.
+$z_i \sim \mathcal{U}(0,1)$ and the support of input set is $\mathcal{D}_\mathbf{Z}  =  [0,1]^M$ where $M$ is the total number input parameters.
+The Sobol decomposition is defined as:
+\begin{equation}\label{sobol_decomp}
+f(z_1, z_2, ..., z_M) = f_0+\sum_{i=1}^M f_i(z_i) + \sum_{1\leq<i<j\leq M} f_{ij}(z_i,z_j) + ... + f_{1, 2, ..., M}(z_1, z_2, ..., z_M),
+\end{equation}
+The above decomposition is only valid for independent input parameters. 
 \subsubsection{PCE Ordinary Least Square (PCE\_OLS)}
@@
 \begin{equation}\label{NURB_curve}
-S(x)  = \sum_{i=1}^{N} c_iB_{i,p}(x),
+S(x)  = \sum_{i=1}^{N} c_i B_{i,p}(x),
 \end{equation}
-where $S(x)$ is the value of the curve at location $x$, $c_i$ is the weight of control point $i$ and $B_{i,p}(x)$ is the value of B-spline corresponding to the $i$-th control point at $x$.
+where $S(x)$ is the value of the curve at location $x, c_i$ is the weight of control point $i$ and $B_{i,p}(x)$ is the value of B-spline corresponding to the $i$-th control point at $x$.
 The subscript $p$ denote the polynomial degree of the NURBS curve.
@@
 
-The B-splines are recursive in polynomial degree, for example, we can derive quadratic B-spline ($p=2$) using the linear B-spline ($p=1$); cubic ($p=3$) from quadratic B-splines and so on.
+The B-splines are recursive in polynomial degree, for example, we can derive quadratic B-splines ($p=2$) using linear B-splines ($p=1$), cubic ($p=3$) from quadratic B-splines and so on.
 Given $m$ knot locations $t_1, t_2, ..., t_{m}$, the B-spline of degree 0 is defined as:
@@
 More advanced approaches, monotonicity preserving methods, etc??
-\begin{remark}
-How to chose $x$?
-\end{remark}
-The final step is to introduce uncertainty in the above computed control points by associating a probability distribution.
-
-The steps to obtained perturbed samples from a given reference curve is outlined in Fig. \ref{}.
+%\begin{remark}
+%How to chose $x$?
+%\end{remark}
+%The final step is to introduce uncertainty in the above computed control points by associating a probability distribution.
+%
+%The steps to obtained perturbed samples from a given reference curve is outlined in Fig. \ref{}.
 \subsection{Model uncertainty}

### 2019-09-25 Paper/article.tex
generalized the sensitivity analysis workflow

@@
 \begin{keyword}
-Wind energy, global sensitivity analysis, UQ, aeroelastic models 
+Global sensitivity analysis, UQ, Sobol indices, aeroelastic models
 \end{keyword}
@@
 \section{Introduction}
+Aeroelastic models such as the Blade Element Momentum (BEM) models \cite{HandBook} play a critical role in the design, development and optimization of modern wind turbines.
+In particular, BEM models are employed to predict turbine response such as the structural loads and power outputs.
+A number of parameters describing meteorological and operational conditions as well as manufacturing specifications are needed as input to simulate BEM models.
+In this work, we seek to quantify the sensitivities of these input parameters for different turbine responses.
 \section{Aeroelastic models}

### 2019-09-25 sensitivity_analysis/AEROmoduleWrapper/AeroPower.m

@@ -1,0 +1,50 @@
+function [Times,Azimuthdeg,PowerWatt,Axial_ForceN] = AeroPower(filename, startRow, endRow)
+%AeroPower imports data from the AeroPower.dat file from the AERO module
+% package
+%   [TIMES,AZIMUTHDEG,POWERWATT,AXIAL_FORCEN] = AeroPower(FILENAME) Reads
+%   data from text file FILENAME for the default selection.
+%
+%   [TIMES,AZIMUTHDEG,POWERWATT,AXIAL_FORCEN] = AeroPower(FILENAME,
+%   STARTROW, ENDROW) Reads data from rows STARTROW through ENDROW of text
+%   file FILENAME.
+%
+% Example:
+%   [Times,Azimuthdeg,PowerWatt,Axial_ForceN] = AeroPower('AeroPower.dat',2, 445);
+
+%% Initialize variables.
+if nargin<=2
+    startRow = 2;
+    endRow = inf;
+end
+
+%% Format for each line of text:
+%   column1: double (%f)
+%	column2: double (%f)
+%   column3: double (%f)
+%	column4: double (%f)
+% For more information, see the TEXTSCAN documentation.
+formatSpec = '%20f%20f%20f%f%[^\n\r]';
+
+%% Open the text file.
+fileID = fopen(filename,'r');
+
+%% Read columns of data according to the format.
+dataArray = textscan(fileID, formatSpec, endRow(1)-startRow(1)+1, 'Delimiter', '', 'WhiteSpace', '', 'TextType', 'string', 'HeaderLines', startRow(1)-1, 'ReturnOnError', false, 'EndOfLine', '\r\n');
+for block=2:length(startRow)
+    frewind(fileID);
+    dataArrayBlock = textscan(fileID, formatSpec, endRow(block)-startRow(block)+1, 'Delimiter', '', 'WhiteSpace', '', 'TextType', 'string', 'HeaderLines', startRow(block)-1, 'ReturnOnError', false, 'EndOfLine', '\r\n');
+    for col=1:length(dataArray)
+        dataArray{col} = [dataArray{col};dataArrayBlock{col}];
+    end
+end
+
+%% Close the text file.
+fclose(fileID);
+
+%% Allocate imported array to column variable names
+Times = dataArray{:, 1};
+Azimuthdeg = dataArray{:, 2};
+PowerWatt = dataArray{:, 3};
+Axial_ForceN = dataArray{:, 4};
+
+

### 2019-09-25 sensitivity_analysis/AEROmoduleWrapper/generateAeroModuleSamples.m

@@ -1,0 +1,116 @@
+%% This routine compute samples of power output and axial force with uncertain chord and twist inputs
+% using the ECN AERO module software package 
+close all
+clearvars
+clc
+workingDir = pwd; % Store the current working directory
+
+%% Add paths for dependent routines located in the directories 'NURBS',
+% 'AEROmoduleWrapper' and 'Geometry'
+addpath('C:\Users\pkumar\Dropbox\WindTrue\windtrue\AEROmoduleWrapper\')
+addpath('C:\Users\pkumar\Dropbox\WindTrue\windtrue\NURBS\')
+addpath('C:\Users\pkumar\Dropbox\WindTrue\windtrue\Geometry\')
+folder ='C:\Users\pkumar\Dropbox\WindTrue\ECNAero2CWI\'; % Directory where the AERO modules executable is stores 
+
+%% Variables of input file extracted from reference test case from DANAERO turbuine NM80
+AEROMODEL = 1;
+TURBINETYPE = 1;
+
+zB = [0 2 4 6 8 10 12 14 16 18 20 22 ...
+      24 26 28 30 32 34 36 37 38 38.4 38.8];
+
+ref_chord = [2.42 2.48 2.65 2.81 2.98 3.14 3.17 2.99 2.79 2.58 2.38 ...
+         2.21 2.06 1.92 1.8 1.68 1.55 1.41 1.18 0.98 0.62 0.48 0.07]; % Replace with random chord vector using NURBS
+
+t_by_c = [0.9999 0.9641 0.8053 0.6508 0.5167 0.403 0.3253 0.284 0.2562 0.2377 0.2225 ...
+          0.2099 0.2003 0.194 0.1903 0.1879 0.186 0.1839 0.1795 0.1739 0.1633 0.157 0.1484]; % Replace with ?? 
+
+ref_twist = [0 5.37 6.69 7.9 9.11 10.19 9.39 7.16 5.45 4.34 3.5 2.86 ... 
+         2.31 1.77 1.28 0.9 0.55 0.23 0.03 0.02 0.93 2.32 6.13]; % Replace with random twist vector using NURBS
+
+C14 = [-25 -25 -25 -25 -25 -25 -25 -25 -25 -25 -25 -25 -25 -25 ...
+       -25 -25 -25 -25 -25 -25 -25 -25 -25];
+
[…]

### 2019-09-25 sensitivity_analysis/Geometry/computeThickness.m

@@ -1,0 +1,74 @@
+function  samplesThickness = computeThickness(samples,randVec, pc, plotSamples,bladeLength,...
+                                      interpolationLocations,referenceThickness,...
+                                      t0,n,sampledLocations,sampledValues)
+
+ % This routine computes the random samples of airfoil Thickness using the purturbed
+% control points of NURBS curve
+
+% Input arguments
+% 'samples' Number of samples of perturbed Chord
+% 'randVec' a samples-by-numOfControlPoints matrix of random numbers
+% 'pc' is vector with each element bw [0,1] representing fraction of perturbation for each control point from their baseline value.
+% 'plotSamples' 0 for no plot, 1 (default) to plot the generated samples, the baseline curve and the control points
+% 'interpolationLocations' Points along the blade length according to ECNAERO module input file 
+% 'referenceThickness' Reference values of Chord*(t/c) obtained from ECNAERO module input file
+% 't0' Knot vector b/w [0,1], the number of resulting basis function is numel(t0)+1
+% 'n' NURBS order: 2 for linear B-splines, 3 for Quadratic, so on. The polynomial degree of B-spline is n-1.
+% 'sampledLocations' Locations where the value of chord is sample. NOTE: numel(sampleLocations) =  numel(t0) + 1
+% 'sampledValues' Vales of chord at sampled location
+% 'pc' is vector with each element bw [0,1] representing fraction of perturbation for control points from their baseline value. The numel(pc) = numel(sampledLocations)  
+
+% Output argument
+% 'samplesThickness' are samples of Thickness generated by perturbing the baseline control points. Each row corresponds to one sample.
+
+if nargin < 5 % Default function argument values corresponding to NM80 turbine
+    bladeLength = 38.8; % To normalize the blade length bw [0,1]   
+    interpolationLocations = [0 2 4 6 8 10 12 14 16 18 20 22 ...
+                          24 26 28 30 32 34 36 37 38 38.4 38.8]; 
+
+    referenceChord = [2.42 2.48 2.65 2.81 2.98 3.14 3.17 2.99 2.79 2.58 2.38 ...
+                  2.21 2.06 1.92 1.8 1.68 1.55 1.41 1.18 0.98 0.62 0.48 0.07];
+    
[…]

### 2019-09-25 sensitivity_analysis/NURBS/getControlPoints.m

@@ -1,0 +1,20 @@
+function c = getControlPoints(x,S,t,n)
+%% This routine computes control points c for B-splines that corresponds to sampled data.
+% Input arguments:
+% 'x' is locations of known values of curve. Should be normalized between [0,1]
+% 'S' known function values at x
+% 't' is the knot vector
+% 'n' is the NURBS order: 2 for linear B-splines, 3 for Quadratic, so on. The polynomial degree of B-spline is n-1.
+
+% Output arguments:
+% 'c' is the values of control points at 'x' locations (at the sampled locations)
+
+t = [t(1)*ones(1,n-1) t t(end)*ones(1,n-1)]; % padded knot vector obtained by padding n-1 elements at front and end. 
+j = 0: numel(t)- n-1; % Index of B-spline from 0 =< j < numel(t)-n
+
+B = zeros(numel(j),numel(j)); % Bspline matrix at 
+
+for i = 1:numel(j)
+    [B(:,i),x] = bspline_basis(j(i),n,t,x);
+end
+c = B\S'; % control points derived from sampled data 

### 2019-09-25 sensitivity_analysis/NURBS/perturbNURBS.m

@@ -1,0 +1,26 @@
+function S = perturbNURBS(t0,n,xu,c, pc,samples, randVec)
+% This routine perturbs (using a uniform distribution) the control points 
+% obtained from the baseline curves and generates perturbed smooth curves
+
+% Input:
+% 't' is the knot vector
+% 'n' is the NURBS order: 2 for linear B-splines, 3 for Quadratic, so on. The polynomial degree of B-spline is n-1.
+% 'c' control point vector
+% 'pc' vector containing perturbation fraction for each control points
+% 'xu' discrete query points
+% 'samples' number of samples of 
+% 'randVec' a samples-by-numel(c) matrix of random numbers
+
+% Output:
+% 'S' a samples-by-numel(xu) matrix where elements in each row corresponds to one sample 
+
+t = [t0(1)*ones(1,n-1) t0 t0(end)*ones(1,n-1)]; % padded knot vector obtained by padding n-1 elements at front and end. 
+j = 0: numel(t)- n - 1; % Index of B-spline from 0 = < j < numel(t)-n
+S = zeros(samples,numel(xu));
+
+for k = 1:samples
+    for i = 1:numel(j)
+        [y,xu] = bspline_basis(j(i),n,t,xu);
+        S(k,:) = S(k,:) + (c(i)+ pc(i).*c(i)*(randVec(k,i)))*y;
+    end
+end

### 2019-09-25 sensitivity_analysis/NURBS/test_perturbNURBS.m

@@ -1,0 +1,37 @@
+% This routine computes perturbed samples obtained from perturbing control
+% points of baseline curve
+
+close all
+clearvars
+clc
+
+x = [1.5 15.42 32.25 52.75 63]/63; % locations of known values of curve (e.g. locations where values of chord or twist are sampled)
+S = [3.28 4.62 3.70 2.55 1.37]; % known function values at x
+t0 = [0 1/3 2/3 1]; % Original knot vectors 
+n = 3; % NURBS order: 2 for linear B-splines, 3 for Quadratic, so on. The polynomial degree of B-spline is n-1.
+
+c = getControlPoints(x,S,t0,n); % control points for NURB curve
+
+t = [t0(1)*ones(1,n-1) t0 t0(end)*ones(1,n-1)]; % Padded knot vector obtained by padding n-1 elements at front and end. 
+j = 0: numel(t)- n-1; % Index of B-spline from 0 =< j < numel(t) - n
+
+plot(x,c,'marker','o','linewidth',2) % plot control points
+hold on
+plot(x,S,'marker','x','markersize',8,'linestyle','none','linewidth',2) % plot sampled points
+xu  = linspace(x(1), x(end), 100); % discrete query points. Interpolation is performed at these points.
+Sxu = zeros(1,numel(xu)); % 'Sxu' is the func tion values of NURB curve interpolated at xu
+
+for i = 1:numel(j)
+    [y,xu] = bspline_basis(j(i),n,t,xu);
+    Sxu = Sxu + c(i)*y;
+end
+
+plot(xu,Sxu,'linewidth',2,'color','r')
+samples=10;
+pc = 0.1*ones(numel(j),1); % Amount of purturbation for each control points, for example, 0.1 corresponds to plus-minus 5% of perturbation.
+randVec = rand(samples,numel(j)) - 0.5; % Uniform distribution between [-0.5,0.5]
+S = perturbNURBS(t0,n,xu,c, pc,samples, randVec);
+plot(xu,S','color','k','linestyle','--','HandleVisibility','off')
+plot(xu,S(1,:),'color','k','linestyle','--')
+legend('control points','sampled data', 'NURB curve','random samples')
+hold off

### 2019-09-25 sensitivity_analysis/UQLabCore_Rel1.0.0/Examples/Inversion/uq_Example_Inversion_03_Hydro.m

@@ -1,0 +1,183 @@
+%% INVERSION: HYDROLOGICAL MODEL CALIBRATION
+%
+% In this example, the HYMOD conceptual watershed model is calibrated
+% using measured discharge records from the Leaf River $(1950\,\mathrm{km^2})$
+% watershed in Mississippi between January 1, 1952 and September 30, 1954.
+% The calibration is carried out with a custom additive discrepancy and 
+% solved using an MCMC sampler.
+%
+% The problem setup of this example is adapted from Example 4 of the DREAM
+% Suite developed by Miroslav Sejna and Jasper Vrugt. For details, see:
+% PC-Progress. (2016). DREAM Suite version 1: User Manual, 
+% <http://www.pc-progress.com/en/onlinehelp/dream1/DREAM_Suite.html URL>
+% (last accessed: 14/02/2019). 
+
+%% 1 - INITIALIZE UQLAB
+%
+% Clear all variables from the workspace, set the random number generator
+% for reproducible results, and initialize the UQLab framework:
+clearvars
+rng(100,'twister')
+uqlab
+
+%%
+% Load the model constants |ModelSetup| and |Data|:
+load('uq_Example_BayesianHydro')
+
+%% 2 - FORWARD MODEL
+%
+% The forward model used in the calibration is the HYMOD watershed model.
+% It models the discharge using five parameters:
+%
+% # $c_{\mathrm{max}}$: maximum interception (mm)
+% # $b_{\mathrm{exp}}$: soil water storage capacity (-)
+% # $\alpha$: maximum percolation rate (-)
+% # $r_s$: recession constant slow-flow reservoir (-)
+% # $r_q$: recession constant fast-flow reservoir (-)
+%
+% The HYMOD model has been implemented in the function
+% |uq_hymod(X,ModelConstants)| supplied with UQLab. The function
+% evaluates the model using the input parameters gathered in the columns
+% of the matrix |X| following the order above and constant parameters 
+% given in the structure |ModelConstants|.
+%
+% Define the forward model as a UQLab MODEL object:
+ModelOpts.Name = 'HYMOD Model';
+ModelOpts.mHandle = @(x) uq_hymod(x,ModelSetup);
+ModelOpts.isVectorized = true;
+
[…]

### 2019-09-25 sensitivity_analysis/UQLabCore_Rel1.0.0/Examples/MetaModelling/Kriging/uq_Example_Kriging_03_TrendTypes.m

@@ -1,0 +1,205 @@
+%% KRIGING METAMODELING: TREND TYPES
+%
+% This example showcases how to perform Kriging for a 
+% simple one-dimensional function using various trend types. 
+
+%% 1 - INITIALIZE UQLAB
+%
+% Clear all variables from the workspace,
+% set the random number generator for reproducible results,
+% and initialize the UQLab framework:
+clearvars
+rng(100,'twister')
+uqlab
+
+%% 2 - COMPUTATIONAL MODEL
+%
+% The computational model is a simple analytical function defined by:
+%
+% $$Y = x \sin(x), \; x \in [0, 15]$$
+% 
+% Specify this model in UQLab using a string with vectorized operation:
+ModelOpts.Name = 'XsinX';
+ModelOpts.mString = 'X.*sin(X)';
+ModelOpts.isVectorized = true;
+
+myModel = uq_createModel(ModelOpts);
+
+%% 3 - PROBABILISTIC INPUT MODEL
+%
+% The probabilistic input model consists of one uniform random variable:
+% 
+% $X \sim \mathcal{U}(0, 15)$
+%
+% Specify the marginal and create a UQLab INPUT object:
+InputOpts.Marginals.Type = 'Uniform';
+InputOpts.Marginals.Parameters = [0 15];
+
+myInput = uq_createInput(InputOpts);
+
+%% 4 - EXPERIMENTAL DESIGN AND MODEL RESPONSES
+%
+% Generate $10$ samples of $X$ using the Sobol' sequence sampling:
+X = uq_getSample(10,'Sobol');
+
+%%
+% Evaluate the corresponding model responses:
+Y = uq_evalModel(X);
+
+%% 5 - KRIGING METAMODELS
+%
+% Three trend types are considered in this example:
+%
+% * Constant (so-called _ordinary Kriging_)
+% * 3rd-degree polynomial
+% * Custom type 
+
+%% 5.1 Ordinary Kriging
+%
+% Select the metamodeling tool and the Kriging module:
+MetaOpts.Type = 'Metamodel';
+MetaOpts.MetaType = 'Kriging';
+
+%%
+% Assign the experimental design and the corresponding model responses:
+MetaOpts.ExpDesign.X = X;
+MetaOpts.ExpDesign.Y = Y;
+
+%%
+% Set the Kriging trend type to ordinary Kriging:
+Metaopts.Trend.Type = 'ordinary';
+
+%%
[…]

### 2019-09-25 sensitivity_analysis/UQLabCore_Rel1.0.0/Examples/MetaModelling/Kriging/uq_Example_Kriging_07_MultipleOutputs.m

@@ -1,0 +1,234 @@
+%% KRIGING METAMODELING: MULTIPLE OUTPUTS
+%
+% This example showcases an application of Kriging to the metamodeling
+% of a simply supported beam model with multiple outputs.
+% The model computes the deflections at several points along the length 
+% of the beam subjected to a uniform random load.
+
+%% 1 - INITIALIZE UQLAB
+%
+% Clear all variables from the workspace, set the random number generator
+% for reproducible results, and initialize the UQLab framework:
+clearvars
+rng(102,'twister')
+uqlab
+
+%% 2 - COMPUTATIONAL MODEL
+%
+% The simply supported beam problem is shown in the following figure:
+uq_figure('Position', [50 50 500 400])
+[I,~] = imread('SimplySupportedBeam.png');
+image(I)
+axis equal
+set(gca, 'visible', 'off');
+
+%% 
+% The (negative) deflection of the beam at any longitudinal coordinate
+% $s$ is given by:
+%
+% $$V(s) = -\frac{p \,s (L^3 - 2\, s^2 L + s^3) }{2E b h^3}$$
+%
+% This computation is carried out by the function
+% |uq_SimplySupportedBeam9Points(X)| supplied with UQLab.
+% The function evaluates the inputs gathered in the $N \times M$
+% matrix |X|, where $N$ and $M$ are the numbers of realizations
+% and inputs, respectively.
+% The inputs are given in the following order:
+%
+% # $b$: beam width $(m)$
+% # $h$: beam height $(m)$
+% # $L$: beam length $(m)$
+% # $E$: Young's modulus $(Pa)$
+% # $p$: uniform load $(N/m)$
+%
+% The function returns the beam deflection $V(s_i)$
+% at nine equally-spaced points along the length
+% $s_i = i \cdot L/10, \; i=1,\ldots,9.$ 
+%
+% Create a MODEL object from the |uq_SimplySupportedBeam9points| function:
+ModelOpts.mFile = 'uq_SimplySupportedBeam9points';
+
+myModel = uq_createModel(ModelOpts);
+
+%% 3 - PROBABILISTIC INPUT MODEL
+%
[…]

### 2019-09-25 sensitivity_analysis/UQLabCore_Rel1.0.0/Examples/MetaModelling/LRA/uq_Example_LRA_02_Ishigami.m

@@ -1,0 +1,125 @@
+%% LRA METAMODELING: ISHIGAMI FUNCTION
+%
+% This example showcases an application of a canonical 
+% low-rank approximation (LRA) to the metamodeling 
+% of the Ishigami function.
+
+%% 1 - INITIALIZE UQLAB
+%
+% Clear all variables from the workspace, set the random number generator
+% for reproducible results, and initialize the UQLab framework:
+clearvars
+rng(150,'twister')
+uqlab
+
+%% 2 - COMPUTATIONAL MODEL
+%
+% The Ishigami function is defined as:
+%
+% $$Y(\mathbf{x}) = \sin(x_1) + 7 \sin^2(x_2) + 0.1 x_3^4 \sin(x_1)$$
+
+%%
+% This computation is carried out by the function
+% |uq_ishigami(X)| supplied with UQLab.
+% The function evaluates the inputs gathered in the $N \times M$ matrix
+% |X|, where $N$ and $M$ are the numbers of realizations and input
+% variables, respectively.
+% 
+% Create a MODEL object from the |uq_ishigami| function:
+ModelOpts.mFile = 'uq_ishigami';
+
+myModel = uq_createModel(ModelOpts);
+
+%% 3 - PROBABILISTIC INPUT MODEL
+%
+% The probabilistic input model consists of three independent uniform 
+% random variables:
+%
+% $$X_i \sim \mathcal{U}(-\pi, \pi), \quad i = 1,2,3$$
+
+%%
+% Specify these marginals in UQLab:
+for ii = 1:3
+    InputOpts.Marginals(ii).Type = 'Uniform';
+    InputOpts.Marginals(ii).Parameters = [-pi pi]; 
+end
+
+%%
+% Create an INPUT object based on the marginals:
+myInput = uq_createInput(InputOpts);
+
+%% 4 - LOW-RANK APPROXIMATION METAMODEL
+%
+% Select the metamodeling tool and the LRA module:
+MetaOpts.Type = 'Metamodel';
+MetaOpts.MetaType = 'LRA';
+
+%%
+% Specify the range for the univariate polynomials degree selection:
+MetaOpts.Degree = 3:20;             
+
+%%  
+% In this example, the rank is adapted and the optimal rank is selected 
+% based on the 3-fold cross-validation.
[…]

### 2019-09-25 sensitivity_analysis/UQLabCore_Rel1.0.0/Examples/MetaModelling/PCE/uq_Example_PCE_03_SimplySupportedBeam.m

@@ -1,0 +1,235 @@
+%% PCE METAMODELING: ESTIMATION OF STATISTICAL MOMENTS
+%
+% This example showcases an application of sparse
+% polynomial chaos expansion (PCE) to the estimation of
+% the statistical moments of a computational model with random inputs.
+% The model is a simply supported beam model that computes
+% deflections at mid-span of a beam subjected to a uniform random load.
+% The convergence behavior of the PCE-based and pure Monte-Carlo-based 
+% estimation of the statistical moments is compared.
+
+%% 1 - INITIALIZE UQLAB
+%
+% Clear all variables from the workspace, set the random number generator
+% for reproducible results, and initialize the UQLab framework:
+clearvars
+rng(100,'twister')
+uqlab
+
+%% 2 - COMPUTATIONAL MODEL
+%
+% The simply supported beam model is shown in the following figure:
+uq_figure('Position', [50 50 500 400])
+[I,~] = imread('SimplySupportedBeam.png');
+image(I)
+axis equal
+set(gca,'visible','off')
+
+%% 
+% The forward model computes the deflection of the beam $V$
+% at mid-span location, which reads:
+%
+% $$ V = \frac{ 5 p L^4 }{32 E b h^3}$$
+%
+% This computation is carried out by the function
+% |uq_SimplySupportedBeam(X)| supplied with UQLab.
+% The input variables of this function are gathered into the $N \times M$
+% matrix |X|, where $N$ and $M$ are the numbers of realizations
+% and input variables, respectively.
+% The input variables are given in the following order:
+%
+% # $b$: beam width $(m)$
+% # $h$: beam height $(m)$
+% # $L$: beam length $(m)$
+% # $E$: Young's modulus $(MPa)$
+% # $p$: uniform load $(kN/m)$
+
+%%
+% Create a MODEL object from the function:
+ModelOpts.mFile = 'uq_SimplySupportedBeam';
+ModelOpts.isVectorized = true;
+
+myModel = uq_createModel(ModelOpts);
+
+%% 3 - PROBABILISTIC INPUT MODEL
+%
+% The simply supported beam model has five inputs,
[…]

### 2019-09-25 sensitivity_analysis/UQLabCore_Rel1.0.0/Examples/MetaModelling/PCE/uq_Example_PCE_04_MultipleOutputs.m

@@ -1,0 +1,240 @@
+%% PCE METAMODELING: MULTIPLE OUTPUTS
+%
+% This example showcases an application of polynomial chaos expansion
+% (PCE) to the metamodeling of a simply supported beam model
+% with multiple outputs.
+% The model computes the deflections at several points along the length 
+% of the beam subjected to a uniform random load.
+
+%% 1 - INITIALIZE UQLAB
+%
+% Clear all variables from the workspace, set the random number generator
+% for reproducible results, and initialize the UQLab framework:
+clearvars
+rng(100,'twister')
+uqlab
+
+%% 2 - COMPUTATIONAL MODEL
+%
+% The simply supported beam problem is shown in the following figure:
+uq_figure('Position', [50 50 500 400])
+[I,~] = imread('SimplySupportedBeam.png');
+image(I)
+axis equal
+set(gca, 'visible', 'off');
+
+%% 
+% The (negative) deflection of the beam at any longitudinal coordinate
+% $s$ is given by:
+%
+% $$V(s) = -\frac{p \,s (L^3 - 2\, s^2 L + s^3) }{2E b h^3}$$
+%
+% This computation is carried out by the function
+% |uq_SimplySupportedBeam9Points(X)| supplied with UQLab.
+% The function evaluates the inputs gathered in the $N \times M$
+% matrix |X|, where $N$ and $M$ are the numbers of realizations
+% and input variables, respectively.
+% The input variables are given in the following order:
+%
+% # $b$: beam width $(m)$
+% # $h$: beam height $(m)$
+% # $L$: beam length $(m)$
+% # $E$: Young's modulus $(Pa)$
+% # $p$: uniform load $(N/m)$
+%
+% The function returns the beam deflection $V(s_i)$
+% at nine equally-spaced points along the length
+% $s_i = i \cdot L/10, \; i=1,\ldots,9.$ 
+%
+% Create a MODEL object from the |uq_SimplySupportedBeam9points| function:
+ModelOpts.mFile = 'uq_SimplySupportedBeam9points';
+
[…]

### 2019-09-25 sensitivity_analysis/UQLabCore_Rel1.0.0/Examples/MetaModelling/PCE/uq_Example_PCE_05_TrussDataSet.m

@@ -1,0 +1,140 @@
+%% PCE METAMODELING: TRUSS DATA SET
+%
+% This example showcases how to perform polynomial chaos expansion (PCE)
+% metamodeling using existing data sets.
+% The data sets come from a finite element model of a truss structure
+% and are retrieved from different MAT-files.
+% The files consist of an experimental design of size $200$
+% and a validation basis of size $10^4$.
+%
+% More information about the truss structure model can be found in the
+% |README.txt| file located in the same folder as the truss data set.
+
+%% 1 - INITIALIZE UQLAB
+%
+% Clear all variables from the workspace 
+% and initialize the UQLab framework:
+clearvars
+uqlab
+
+%% 2 - RETRIEVE DATA SETS
+%
+% The experimental design and the validation basis are stored
+% in two separate files in the following location:
+FILELOCATION = fullfile(uq_rootPath, 'Examples', 'SimpleDataSets',...
+    'Truss_Matlab_FEM');
+
+%%
+% Read the experimental design data set file and store the contents 
+% in matrices:
+load(fullfile(FILELOCATION,'Truss_Experimental_Design.mat'), 'X', 'Y');
+
+%%
+% Read the validation basis data set file and store the contents
+% in matrices:
+load(fullfile(FILELOCATION,'Truss_Validation_Basis.mat'), 'Xval', 'Yval');
+
+%% 3 - INPUT MODEL
+%
+% Because PCE requires a choice of polynomial basis,
+% a probabilistic input model needs to be defined.
+% Specify the marginals of the probabilistic input model:
+
+% Young's modulus of cross-sections
+for i=1:2
+    InputOpts.Marginals(i).Name = sprintf('E%d',i);
+    InputOpts.Marginals(i).Type = 'Lognormal';    
+    InputOpts.Marginals(i).Moments = [2.1e11 2.1e10];  
+end
+
+% Cross-section of horizontal elements
+InputOpts.Marginals(3).Name = 'A1';
+InputOpts.Marginals(3).Type = 'Lognormal';
+InputOpts.Marginals(3).Moments = [2.0e-3 2.0e-4];
+
+% Cross-section of diagonal elements
[…]

### 2019-09-25 sensitivity_analysis/UQLabCore_Rel1.0.0/Examples/MetaModelling/PCE/uq_Example_PCE_06_ArbitraryBasis.m

@@ -1,0 +1,180 @@
+%% PCE METAMODELING: CONSTRUCTION OF ORTHOGONAL BASIS FOR ARBITRARY DISTRIBUTION
+%
+% This example demonstrates the numerical basis construction with 
+% Stieltjes procedure. The Ishigami function with Gaussian distributions 
+% restricted to the $[-\pi, \pi]$ interval is used.
+% The Stieltjes procedure is expected to produce a basis
+% that performs better than the Legendre polynomials,
+% the default basis for all bounded distribution.
+
+%% 1 - INITIALIZE UQLAB
+%
+% Clear all variables from the workspace, set the random number generator
+% for reproducible results, and initialize the UQLab framework:
+clearvars
+rng(100,'twister')
+uqlab
+
+%% 2 - COMPUTATIONAL MODEL
+%
+% The Ishigami function is defined as:
+%
+% $$Y(\mathbf{x}) = \sin(x_1) + 7 \sin^2(x_2) + 0.1 x_3^4 \sin(x_1)$$
+%
+% where $x_i \in [-\pi, \pi], \; i = 1,2,3.$
+
+%%
+% This computation is carried out by the function
+% |uq_ishigami(X)| supplied with UQLab.
+% The function evaluates the inputs given in $N \times M$ matrix |X|,
+% where $N$ and $M$ are the number of realizations and input variables,
+% respectively.
+% 
+% Create a MODEL object from the |uq_ishigami| function:
+ModelOpts.mFile = 'uq_ishigami';      
+myModel = uq_createModel(ModelOpts);  
+
+%% 3 - PROBABILISTIC INPUT MODEL
+%
+% The probabilistic input model consists of three independent uniform 
+% random variables:
+%
+% $$X_i \sim \mathcal{N}(0, \pi), \; X_i \in [-\pi, pi], \; i = 1,2,3$$
+
+%%
+% Specify these marginals:
+for ii = 1:3
+    InputOpts.Marginals(ii).Type = 'Gaussian';
+    InputOpts.Marginals(ii).Parameters = [pi/2 pi];
+    InputOpts.Marginals(ii).Bounds = [-pi pi];
+end
+
+%%
+% Create an INPUT object based on the specified marginals:
+myInput = uq_createInput(InputOpts);
+
+%% 4 - POLYNOMIAL CHAOS EXPANSION METAMODEL
+%
+% Select PCE as the metamodeling tool:
+MetaOpts.Type = 'Metamodel';
+MetaOpts.MetaType = 'PCE';
+
+%% 
[…]

### 2019-09-25 sensitivity_analysis/UQLabCore_Rel1.0.0/Examples/MetaModelling/PCK/uq_Example_PCK_02_Ishigami.m

@@ -1,0 +1,169 @@
+%% PC-KRIGING METAMODELING: MULTIPLE INPUT DIMENSIONS
+%
+% This example illustrates different aspects of polynomial chaos-Kriging
+% (PC-Kriging) metamodel construction
+% using the eight-dimensional Ishigami function 
+% as the full computational model.
+
+%% 1 - INITIALIZE UQLAB
+%
+% Clear all variables from the workspace, set the random number generator
+% for reproducible results, and initialize the UQLab framework:
+clearvars
+rng(1,'twister')
+uqlab
+
+%% 2 - COMPUTATIONAL MODEL
+%
+% The Ishigami function is defined as:
+%
+% $$Y(\mathbf{x}) = \sin(x_1) + 7 \sin^2(x_2) + 0.1 x_3^4 \sin(x_1)$$
+%
+% where $x_i \in [-\pi, \pi], \; i = 1,2,3.$
+
+%%
+% This computation is carried out by the function
+% |uq_ishigami(X)| supplied with UQLab.
+% The function evaluates the inputs gathered in the $N \times M$
+% matrix |X|, where $N$ and $M$ are the numbers of realizations
+% and input variables, respectively.
+% 
+% Create a MODEL object from the |uq_ishigami| function:
+ModelOpts.mFile = 'uq_ishigami';
+
+myModel = uq_createModel(ModelOpts);  
+
+%% 3 - PROBABILISTIC INPUT MODEL
+%
+% The probabilistic input model consists of three independent uniform 
+% random variables:
+%
+% $$X_i \sim \mathcal{U}(-\pi, \pi), \quad i = 1,2,3$$
+
+%%
+% Specify these marginals:
+for ii = 1:3
+    InputOpts.Marginals(ii).Type = 'Uniform';
+    InputOpts.Marginals(ii).Parameters = [-pi pi]; 
+end
+
+%%
+% Create an INPUT object based on the specified marginals:
+myInput = uq_createInput(InputOpts);
+
+%% 4 - EXPERIMENTAL DESIGN
+%
+% To compare different PC-Kriging metamodeling techniques,
+% create a single  experimental design of size $80$
+% using the latin hypercube sampling:
+X = uq_getSample(myInput, 80, 'LHS');
+
+%%
[…]

### 2019-09-25 sensitivity_analysis/UQLabCore_Rel1.0.0/Examples/MetaModelling/PCK/uq_Example_PCK_04_TrussDataSet.m

@@ -1,0 +1,148 @@
+%% PC-KRIGING METAMODELING: TRUSS DATA SET
+%
+% This example showcases how to perform polynomial chaos-Kriging
+% (PC-Kriging) metamodeling using existing data sets.
+% The data sets come from a finite element model of a truss structure
+% and are retrieved from different MAT-files.
+% The files consist of an experimental design of size $200$
+% and a validation basis of size $10^4$.
+
+%% 1 - INITIALIZE UQLAB
+%
+% Clear all variables from the workspace 
+% and initialize the UQLab framework:
+clearvars
+uqlab
+
+%% 2 - RETRIEVE DATA SETS
+%
+% The experimental design and the validation basis are stored
+% in two separate files in the following location:
+FILELOCATION = fullfile(uq_rootPath, 'Examples', 'SimpleDataSets',...
+    'Truss_Matlab_FEM');
+
+%%
+% Read the experimental design data set file and store the contents 
+% in matrices:
+load(fullfile(FILELOCATION,'Truss_Experimental_Design.mat'), 'X', 'Y');
+
+%%
+% Read the validation basis data set file and store the contents
+% in matrices:
+load(fullfile(FILELOCATION,'Truss_Validation_Basis.mat'), 'Xval', 'Yval');
+
+%% 3 - PROBABILISTIC INPUT MODEL
+%
+% Due to the use of polynomial chaos as trend, a probabilistic input model 
+% has to be specified for the PC-Kriging metamodel.
+% Specify the marginals of the probabilistic input model:
+
+% Young's modulus of cross-sections
+for i=1:2
+    InputOpts.Marginals(i).Name = sprintf('E%d',i);
+    InputOpts.Marginals(i).Type = 'Lognormal';    
+    InputOpts.Marginals(i).Moments = [2.1e11 2.1e10];  
+end
+
+% Cross-section of horizontal elements
+InputOpts.Marginals(3).Name = 'A1';
+InputOpts.Marginals(3).Type = 'Lognormal';
+InputOpts.Marginals(3).Moments = [2.0e-3 2.0e-4];
+
+% Cross-section of diagonal elements
+InputOpts.Marginals(4).Name = 'A2';
+InputOpts.Marginals(4).Type = 'Lognormal' ;
[…]

### 2019-09-25 sensitivity_analysis/UQLabCore_Rel1.0.0/Examples/MetaModelling/SVC/uq_Example_SVC_02_VariousMethods.m

@@ -1,0 +1,183 @@
+%% SVC METAMODELING: VARIOUS METHODS
+%
+% This example showcases how to perform Support Vector Machine for
+% Classification (SVC) metamodeling for a simple two-dimensional function,
+% using various hyperparameter estimation and optimization methods.
+
+%% 1 - INITIALIZE UQLAB
+%
+% Clear all variables from the workspace,
+% set the random number generator for reproducible results,
+% and initialize the UQLab framework:
+clearvars
+rng(100,'twister')
+uqlab
+
+%% 2 - COMPUTATIONAL MODEL
+%
+% The computational model is an analytical function defined as:
+%
+% $$Y(\mathbf{x}) = \left\{ \begin{array}{ll}
+%   -1 & \quad Y_{val}(\mathbf{x}) < 0  \\
+%   +1 & \quad Y_{val}(\mathbf{x}) > 0 \\
+% \end{array}\right.$$
+%
+% $Y_{val}(\mathbf{x}) = x_2 - x_1 \sin(x_1) - 1$
+%
+% This model is implemented in the UQLab function |uq_cmsin|.
+
+%%
+% Create a MODEL object from the function file:
+ModelOpts.mFile = 'uq_cmsin';
+
+myModel = uq_createModel(ModelOpts);
+
+%% 3 - PROBABILISTIC INPUT MODEL
+%
+% The probabilistic input model consists of two uniform random variables:
+%
+% $$X_i \in \mathcal{U}(0, 10), \; i = 1,2$$
+
+%%
+% Specify the distributions of the input variables:
+Input.Marginals(1).Type = 'Uniform';
+Input.Marginals(1).Parameters = [0 10];
+Input.Marginals(2).Type = 'Uniform';
+Input.Marginals(2).Parameters = [0 10];
+
+%%
+% Create an INPUT object:
+myInput = uq_createInput(Input);
+
+%% 4 - SVC METAMODELS
+%
+% Select the SVC metamodeling technique:
+MetaOpts.Type = 'Metamodel';
+MetaOpts.MetaType = 'SVC';
+
+%%
+% Specify the sampling strategy and the number of sample points
+% for the experimental design:
+MetaOpts.ExpDesign.Sampling = 'LHS';
+MetaOpts.ExpDesign.NSamples = 100;
+
+%%
+% Assign the previously created INPUT and MODEL objects:
+MetaOpts.Input = myInput;
+MetaOpts.FullModel = myModel;
+%%
[…]

### 2019-09-25 sensitivity_analysis/UQLabCore_Rel1.0.0/Examples/MetaModelling/SVC/uq_Example_SVC_03_BreastCancerData.m

@@ -1,0 +1,174 @@
+%% SVC METAMODELING: BREAST CANCER DIAGNOSIS DATA SET
+%
+% This example showcases how to perform Support Vector Machine
+% for Classification (SVC) metamodeling on the breast 
+% cancer Wisconsin (diagnostic) data set.
+%
+% For more information, see: Street, W.N., W.H. Wolberg, and
+% O.L. Mangasarian. (1993). Nuclear feature extraction for breast tumor 
+% diagnosis. IS&T/SPIE 1993 International Symposium on Electronic Imaging: 
+% Science and Technology (1905), 861-870. 
+
+%% 1 - INITIALIZE UQLAB
+%
+% Clear all variables from the workspace,
+% set the random number generator for reproducible results,
+% and initialize the UQLab framework:
+clearvars
+rng(1,'twister')
+uqlab
+
+%% 2 - RETRIEVE DATA SET
+%
+% The data set is stored in a MAT-file located in:
+FILELOCATION = fullfile(uq_rootPath, 'Examples', 'SimpleDataSets',...
+    'Breast_Cancer');
+
+%%
+% Read the data set file and store the contents in matrices:
+load(fullfile(FILELOCATION,'wdbc.mat'), 'X', 'Y');
+
+%% 3 - TRAINING AND VALIDATION SETS
+%
+% Use $70\%$ of the data set for training
+% and the rest for validation:
+Ntrain = floor(0.7*size(X,1));
+
+%%
+% Get the indices of the training (resp. validation) sample:
+idx_train = randperm(size(X,1), Ntrain);
+Xtrain = X(idx_train,:);
+Ytrain = Y(idx_train,:);
+
+idx_val = setdiff(1:size(X,1),idx_train); 
+Xval = X(idx_val,:);
+Yval = Y(idx_val,:);
+
+%% 4 - SVC METAMODEL
+%
+% Select SVC as the metamodeling tool:
+MetaOpts.Type = 'Metamodel';
+MetaOpts.MetaType = 'SVC';
+
+%% 
+% Use the data set loaded earlier:
+MetaOpts.ExpDesign.X = Xtrain;
+MetaOpts.ExpDesign.Y = Ytrain;
+
+%%
+% Use linear penalization:
+MetaOpts.Penalization = 'linear';
+
+%%
+% Use the Sequential Minimal Optimization (SMO) algorithm to solve the
+% quadratic programming problem:
[…]

### 2019-09-25 sensitivity_analysis/UQLabCore_Rel1.0.0/Examples/MetaModelling/SVR/uq_Example_SVR_04_HousingDataSet.m

@@ -1,0 +1,131 @@
+%% SVR METAMODELING: BOSTON HOUSING DATA SET
+%
+% This example showcases how to use Support Vector Machine for Regression
+% (SVR) metamodeling on an existing data set. 
+% The data set under consideration is related to the housing prices
+% in Boston, a standard machine learning benchmark data set.
+% 
+% For more information, see: Harrison, D. and D. L. Rubinfeld, D.L. (1978).
+% Hedonic prices and the demand for clean air.
+% J. Environ. Economics & Management, vol. 5, pp. 81-102.
+
+%% 1 - INITIALIZE UQLAB
+%
+% Clear all variables from the workspace, set the random number generator
+% for reproducible results, and initialize the UQLab framework:
+clearvars
+rng(1,'twister')
+uqlab
+
+% This example will be using SMO QP optimizer. This feature only appears in Matlab R2015b. 
+% Check its availability. Otherwise, issue a warning and revert to interior point (IP) ;
+QPSolver = 'SMO' ;
+if ~exist('fitrsvm','file')
+    QPSolver = 'IP' ;
+    warning('SMO is not available on this version of Matlab. Setting QP Solver to IP');
+    % IP with a large training set may be slow to converge
+    fprintf('This example will take some time!') ;  
+end
+%% 2 - RETRIEVE DATA SETS
+%
+% The housing data set is stored in a MAT-file in the following location:
+FILELOCATION = fullfile(uq_rootPath, 'Examples', 'SimpleDataSets',...
+    'Boston_Housing');
+
+%%
+% Read the data set and store the contents in matrices:
+load(fullfile(FILELOCATION,'housing.mat'), 'X', 'Y')
+
+%% 3 - TRAINING AND VALIDATION SETS
+%
+% Retrieve the size of the experimental design:
+[N,M] = size(X);
+
+%%
+% Use $85\%$ of the data for training and the rest for validation:
+Ntrain = floor(0.85*N);
+Nval = N - Ntrain;
+
+%% 4 - SVR METAMODELS
+%
+% Three SVR metamodels are created using different
+% training sets of the same size (|Ntrain|) randomly sampled
+% from the whole housing data set.
+% The steps (repeated three times) are as follows:
+%
+% # Randomly split the available data into a traning and a validation sets
[…]

### 2019-09-25 sensitivity_analysis/UQLabCore_Rel1.0.0/Examples/Model/uq_Example_Model_01_ModelDefinition.m

@@ -1,0 +1,145 @@
+%% MODEL MODULE: COMPUTATIONAL MODEL DEFINITION
+%
+% This example showcases the various ways to define a computational
+% model in UQLab using the analytical Ishigami function
+% as the computational model.
+
+%% 1 - INITIALIZE UQLAB
+%
+% Clear all variables from the workspace, set the random number generator
+% for reproducible results, and initialize the UQLab framework:
+clearvars
+rng(100,'twister')
+uqlab
+
+%% 2 - COMPUTATIONAL MODEL
+%
+% The Ishigami function is defined as:
+%
+% $$Y(\mathbf{x}) = \sin(x_1) + 7 \sin^2(x_2) + 0.1 x_3^4 \sin(x_1)$$
+%
+% where $x_i \in [-\pi, \pi], \; i = 1,2,3.$
+
+%%
+% The Model module in UQLab offers three ways
+% to define a computational model:
+%
+% * using an m-file
+% * using a function handle
+% * using a string
+
+%%
+% Each of these will be illustrated next to define the Ishigami function
+% as a computational model in UQLab. 
+
+%% 2.1 Using an m-file
+%
+% The Ishigami function is available in UQLab as |uq_ishigami(X)|.
+% It is located in:
+%
+%   <uq_rootPath>/Examples/SimpleTestFunctions/uq_ishigami.m
+%
+% |uq_rootPath| depends on the user's UQLab local installation.
+%
+% The function evaluates the inputs given in the $N \times M$ matrix |X|,
+% where $N$ and $M$ are the numbers of realizations and inputs,
+% respectively.
+% 
+% Specify the |uq_ishigami| function in the options for the MODEL object:
+Model1Opts.mFile = 'uq_ishigami';
+
+%%
+% Create the MODEL object:
+myModel_mFile = uq_createModel(Model1Opts);
+
+%% 2.2 Using a function handle
+%
+% The already available |uq_ishigami| function can also be defined
+% using function handles:
+Model2Opts.mHandle = @uq_ishigami;
+
+%%
+% Create the MODEL object:
+myModel_mHandle = uq_createModel(Model2Opts);
+
+%% 2.3 Using a string
+%
+% Finally, the Ishigami function can also be specified
+% using a string of MATLAB expression as follows:
[…]

### 2019-09-25 sensitivity_analysis/UQLabCore_Rel1.0.0/Examples/Model/uq_Example_Model_02_ModelParameters.m

@@ -1,0 +1,115 @@
+%% MODEL MODULE: COMPUTATIONAL MODEL PARAMETERS
+%
+% This example showcases how to pass parameters in a computational model.
+
+%% 1 - INITIALIZE UQLAB
+%
+% Clear all variables from the workspace, set the random number generator
+% for reproducible results, and initialize the UQLab framework:
+clearvars
+rng(100,'twister')
+uqlab
+
+%% 2 - COMPUTATIONAL MODEL
+%
+% The Ishigami function is defined as:
+%
+% $$Y(\mathbf{x}) = \sin(x_1) + a \sin^2(x_2) + b x_3^4 \sin(x_1)$$
+%
+% where $x_i \in [-\pi, \pi], \; i = 1,2,3$;
+% and $a$ and $b$ are model parameters.
+%
+% This computation is carried out by the function
+% |uq_ishigami(X,P)| supplied with UQLab.
+% The inputs and model parameters of this function are gathered
+% into the $N \times M$ matrix |X| and into the $2$-dimensional vector |P|,
+% respectively; and where $N$ and $M$ are the numbers of realizations and
+% inputs, respectively.
+
+%%
+% Specify the function in the options for MODEL object:
+ModelOpts.mFile = 'uq_ishigami';
+
+%%
+% First, a typical parameter set is chosen, $[a, b] = [7, 0.1]$:
+ModelOpts.Parameters = [7 0.1];
+
+%%
+% Create a MODEL object:
+myModel = uq_createModel(ModelOpts);
+
+%% 3 - PROBABILISTIC INPUT MODEL
+%
+% The probabilistic input model consists of
+% three independent uniform random variables:
+%
+% $X_i \sim \mathcal{U}(-\pi, \pi), \quad i = 1,2,3$
+
+%%
+% Specify the marginals (if no copula is given,
+% they are assumed independent): 
+for ii = 1:3
+    InputOpts.Marginals(ii).Type = 'Uniform';
+    InputOpts.Marginals(ii).Parameters = [-pi pi]; 
+end
+
+%%
+% Create an INPUT object based on the marginals:
+myInput = uq_createInput(InputOpts);
+
+%% 4 - COMPARISON OF THE MODEL RESPONSES FOR VARIOUS PARAMETER SETS
+
+%% 4.1 Using a fixed parameter set
+%
+% Create a validation sample of size $10^4$ from the input model
+% using latin hypercube sampling:
+X = uq_getSample(1e4,'LHS');
+
+%%
[…]

### 2019-09-25 sensitivity_analysis/UQLabCore_Rel1.0.0/Examples/Model/uq_Example_Model_03_MultipleOutputs.m

@@ -1,0 +1,169 @@
+%% MODEL MODULE: MULTIPLE OUTPUTS
+%
+% This example showcases the modeling of the deflection of a simply
+% supported beam subjected to a uniform random load at several points along
+% its length.
+
+%% 1 - INITIALIZE UQLAB
+%
+% Clear all variables from the workspace, set the random number generator
+% for reproducible results, and initialize the UQLab framework:
+clearvars
+rng(100,'twister')
+uqlab
+
+%% 2 - COMPUTATIONAL MODEL
+%
+% The simply supported beam problem is shown in the following figure:
+uq_figure('Position', [50 50 500 400])
+[I,~] = imread('SimplySupportedBeam.png');
+image(I)
+axis equal
+set(gca,'visible','off')
+
+%% 
+% The (negative) deflection of the beam at any longitudinal coordinate
+% $s$ is given by:
+%
+% $$V(s) = -\frac{p \,s (L^3 - 2\, s^2 L + s^3) }{2E b h^3}$$
+%
+% This computation is carried out by the function
+% |uq_SimplySupportedBeam9Points(X)| supplied with UQLab.
+% The function evaluates the inputs gathered in the $N \times M$
+% matrix |X|, where $N$ and $M$ are the numbers of realizations
+% and input variables, respectively.
+% The input variables are given in the following order:
+%
+% # $b$: beam width $(m)$
+% # $h$: beam height $(m)$
+% # $L$: beam length $(m)$
+% # $E$: Young's modulus $(Pa)$
+% # $p$: uniform load $(N/m)$
+%
+% The function returns the beam deflection $V(s_i)$
+% at nine equally-spaced points along the length
+% $s_i = i \cdot L/10, \; i=1,\ldots,9.$ 
+%
+% Create a MODEL object from the |uq_SimplySupportedBeam9points| function:
+ModelOpts.mFile = 'uq_SimplySupportedBeam9points';
+
+myModel = uq_createModel(ModelOpts);
+
+%% 3 - PROBABILISTIC INPUT MODEL
+%
+% The simply supported beam model has five independent inputs
+% modeled by lognormal random variables.
[…]

### 2019-09-25 sensitivity_analysis/UQLabCore_Rel1.0.0/Examples/Sensitivity/uq_Example_Sensitivity_01_Borehole_Model.m

@@ -1,0 +1,266 @@
+%% SENSITIVITY: BOREHOLE MODEL
+%
+% This example showcases the application of different sensitivity analysis
+% techniques available in UQLab to the 
+% <https://www.sfu.ca/~ssurjano/shortcol borehole function>.
+
+%% 1 - INITIALIZE UQLAB
+%
+% Clear all variables from the workspace, set the random number generator
+% for reproducible results, and initialize the UQLab framework:
+clearvars
+rng(100,'twister')
+uqlab
+
+%% 2 - COMPUTATIONAL MODEL
+%
+% The computational model is an $8$-dimensional analytical formula 
+% that is used to model the water flow through a borehole
+% The borehole function |uq_borehole| is supplied with UQLab.
+%
+% Create a MODEL object from the function file:
+ModelOpts.mFile = 'uq_borehole';
+myModel = uq_createModel(ModelOpts);
+
+%%
+% Type |help uq_borehole| for information on the model structure as well as
+% the description for each variable.
+
+%% 3 - PROBABILISTIC INPUT MODEL
+%
+% The probabilistic input model consists of eight independent random 
+% variables.
+%
+% Specify the marginals as follows:
+InputOpts.Marginals(1).Name = 'rw'; % Radius of the borehole
+InputOpts.Marginals(1).Type = 'Gaussian';
+InputOpts.Marginals(1).Parameters = [0.10 0.0161812]; % (m)
+
+InputOpts.Marginals(2).Name = 'r'; % Radius of influence
+InputOpts.Marginals(2).Type = 'Lognormal';
+InputOpts.Marginals(2).Parameters = [7.71 1.0056]; % (m)
+
+InputOpts.Marginals(3).Name = 'Tu'; % Transmissivity, upper aquifer
+InputOpts.Marginals(3).Type = 'Uniform';
+InputOpts.Marginals(3).Parameters = [63070 115600]; % (m^2/yr)
+
+InputOpts.Marginals(4).Name = 'Hu'; % Potentiometric head, upper aquifer
+InputOpts.Marginals(4).Type = 'Uniform';
+InputOpts.Marginals(4).Parameters = [990 1110]; % (m)
+
+InputOpts.Marginals(5).Name = 'Tl'; % Transmissivity, lower aquifer
+InputOpts.Marginals(5).Type = 'Uniform';
+InputOpts.Marginals(5).Parameters = [63.1 116]; % (m^2/yr)
+
[…]

### 2019-09-25 sensitivity_analysis/UQLabCore_Rel1.0.0/Examples/SimpleTestFunctions/uq_SimplySupportedBeamTwo.m

@@ -1,0 +1,18 @@
+function V = uq_SimplySupportedBeamTwo(X)
+% UQ_SIMPLYSUPPORTEDBEAM computes the midspan deflection of a simply 
+% supported beam under uniform loading at midspan and quarter span
+%
+%   Model with five input parameters  X= [b h L E p]
+%         b:   beam width
+%         h:   beam height
+%         L:   beam span
+%         E:   Young's modulus
+%         p:   uniform load
+%
+%   Output:  V = (5/32)*pL^4/(E*b*h^3)
+% 
+% See also: UQ_EXAMPLE_PCE_03_SIMPLYSUPPORTEDBEAM
+
+% Vectorized implementation
+V(:,1) = (5/32)*(X(:, 5).*X(:, 3).^4)./(X(:, 4).*X(:, 1).*X(:, 2).^3);
+V(:,2) = (57/512)*(X(:, 5).*X(:, 3).^4)./(X(:, 4).*X(:, 1).*X(:, 2).^3);

### 2019-09-25 sensitivity_analysis/UQLabCore_Rel1.0.0/core/interfaces/CLI/uq_createAnalysis.m

@@ -1,0 +1,22 @@
+% UQ_CREATEANALYSIS   create a UQLab ANALYSIS object.
+%    myAnalysis = UQ_CREATEANALYSIS(ANALYSISOPTS) creates and stores in the
+%    current UQLab session an object of type ANALYSIS based on the configuration
+%    options given in the ANALYSISOPTS structure. 
+%
+%    myAnalysis = UQ_CREATEANALYSIS(ANALYSISOPTS, '-private') creates an
+%    object of type ANALYSIS based on the configuration options given in
+%    the ANALYSISOPTS structure, without saving it in the UQLab session.
+%    This is useful, e.g., in parametric studies which create a large
+%    number of temporary ANALYSIS objects that do not need to be stored in
+%    memory. 
+%
+%    For more details about the available model types, please refer to
+%    analysis-specific initialization: 
+%       help <a href="matlab:uq_ReliabilityOptions">uq_ReliabilityOptions</a> - Reliability analysis (rare events estimation)
+%       help <a href="matlab:uq_SensitivityOptions">uq_SensitivityOptions</a> - Sensitivity analysis
+%       help <a href="matlab:uq_InversionOptions">uq_InversionOptions</a>   - Bayesian calibration options
+
+%
+%    See also: uq_getAnalysis, uq_listAnalyses, uq_selectAnalysis,
+%              uq_createInput, uq_createModel,  uq_doc
+%

### 2019-09-25 sensitivity_analysis/UQLabCore_Rel1.0.0/core/interfaces/CLI/uq_createInput.m

@@ -1,0 +1,32 @@
+% UQ_CREATEINPUT   create a UQLab INPUT object 
+%    myInput = UQ_CREATEINPUT(INPUTOPTS) creates and stores in the current
+%    UQLab session an object of type INPUT based on the configuration
+%    options given in the INPUTOPTS structure. 
+%
+%    myInput = UQ_CREATEINPUT(INPUTOPTS, '-private') creates an object of
+%    type INPUT based on the configuration options given in the INPUTOPTS
+%    structure, without saving it in the UQLab session. This is useful,
+%    e.g., in case of parametric studies which create a large number of
+%    temporary INPUT objects that do not need to be stored in memory.
+%
+%    Example: create a 2-dimensional INPUT object with independent
+%    uniform random variables in [-1,1]:
+%       INPUTOPTS.Marginals(1).Type = 'Uniform';
+%       INPUTOPTS.Marginals(1).Parameters = [-1 1];
+%       INPUTOPTS.Marginals(2).Type = 'Uniform';
+%       INPUTOPTS.Marginals(2).Parameters = [-1 1];
+%       myInput = UQ_CREATEINPUT(INPUTOPTS);
+%
+%    To draw N samples from the resulting object use:
+%       X = uq_getSample(N)
+%
+%    For additional information about the available INPUTOPTS configuration
+%    fields type: 
+%       help <a href="matlab:uq_InputOptions">uq_InputOptions</a>
+%
+%    To open the INPUT User Manual, type:
+%       uq_doc('Input')
+%
+%    See also: uq_getInput, uq_listInputs, uq_selectInput, uq_getSample,
+%              uq_createModel, uq_createAnalysis 
+%

### 2019-09-25 sensitivity_analysis/UQLabCore_Rel1.0.0/core/interfaces/CLI/uq_listModels.m

@@ -1,0 +1,18 @@
+% UQ_LISTMODELS   return all the MODEL objects available in the UQLab session
+%    UQ_LISTMODELS returns a list a MODEL objects that have been created by
+%    the <a href="matlab:help uq_createModel">uq_createModel</a> command. 
+%
+%    The MODEL that is currently selected (used by default in several UQLab
+%    commands like <a href="matlab:help uq_evalModel">uq_evalModel</a>) is highlighted by a '>' symbol to the left
+%    of its name.  
+%
+%    The MODEL objects listed by this command can be accessed from within any
+%    workspace (including functions) with the <a href="matlab:help uq_getModel">uq_getModel</a> function.
+%
+%    The name MODEL myModel created with 
+%       myModel = uq_createModel(MODELOPTS) 
+%    can be specified with the MODELOPTS.Name field.
+%
+%    See also: uq_createModel, uq_evalModel, uq_getModel, uq_selectModel,
+%              uq_listInputs, uq_listAnalyses 
+%

### 2019-09-25 sensitivity_analysis/UQLabCore_Rel1.0.0/lib/uq_display_cobweb.m

@@ -1,0 +1,266 @@
+function fighandle = uq_display_cobweb(X, Y, varargin)
+% disphand = UQ_DISPLAY_COBWEB(X, Y, varargin) Accepts a set of raw data
+% and a set of conditions and produces a 'cobweb' plot. X is an NxM array
+% and Y is an Nx1 array. In order to bring different sets of lines to the
+% front, use the scroll wheel on the plot.
+%
+% varargin can consist of the following Name-Value pairs:
+%   'YEdge', YEDGES  : YEDGES sets the values of Y where the cobweb lines
+%   change color. By default, the color changes at the mean.
+%
+%   'DataTags', TAGS : TAGS sets the names of the plotted data that will be
+%   shown on the X-axis. By default it's X1,X2,...,Y.
+%
+%
+%
+% Minimal working example:
+% uqlab
+% % Data
+% N_samp = 150;
+% X1 = randn(N_samp,1);
+% X2 = randn(N_samp,1);
+% X3 = 2.*X2;
+% X = [X1 X2 X3];
+% Y = X1+X2.^2-3.*X2;
+% Mean value to set edges
+% Ys = sort(Y);
+% Plot
+% uq_display_cobweb(X,Y,'YEdge',[Ys(floor(1/4*N_samp)),Ys(N_samp/2),Ys(ceil(3/4*N_samp))])
+%
+
+
+%% Parse varargin
+
+parse_keys = {'DataTags', 'YEdge'};
+parse_types = {'p','p'};
+
+[uq_cline, ~] = uq_simple_parser(varargin, parse_keys, parse_types);
+
+% check for data tags
+if ~strcmpi(uq_cline{1},'false')
+    dataTag_flag = true;
+    data_tags = uq_cline{1};
+else
+    dataTag_flag = false;
+end
+
+% check for provided Y edges
+if ~strcmpi(uq_cline{2},'false')
+    yEdge_flag = true;
+    yEdges = uq_cline{2};
+else
+    yEdge_flag = false;
+end
+
+
+
+%% Preparation
+%%
+% Concatenate data
+Xall = horzcat(X,Y);
+
+%%
+% Assign data_tags if not provided
+
+if ~dataTag_flag
+    data_tags = cell(1,size(Xall,2));
+    for kk = 1:size(Xall,2)
+        data_tags{kk} = sprintf('X_%i',kk);
+    end
+    data_tags{end} = 'Y';
+elseif dataTag_flag && length(data_tags)==size(Xall,2)-1
[…]

### 2019-09-25 sensitivity_analysis/UQLabCore_Rel1.0.0/lib/uq_plotConfidence.m

@@ -1,0 +1,60 @@
+function handles = uq_plotConfidence(X, Y, Conf, varargin)
+% UQ_PLOTCONFIDENCE plots the confidence bounds: X versus Y with confidence
+% CONF
+% 
+% See also: UQ_KRIGING_DISPLAY
+
+% Set to default the coloring options:
+Lcolor = uq_cmap(1);
+
+%gray
+BoundColor = [128,128,128]/255; 
+
+Tag = '';
+AlphaN = 0.3;
+Lwidth = 2;
+CreateFig = false;
+if nargin > 3
+    if mod(length(varargin), 2)
+        error('The number of options is incorrect');
+    end
+    for ii = 1:2:length(varargin) - 1
+        
+        switch(lower(varargin{ii}))
+            case 'linecolor'
+                Lcolor = varargin{ii + 1};
+            case {'tag', 'name'}
+                Tag = varargin{ii + 1};
+            case 'boundcolor'
+                BoundColor = varargin{ii + 1};
+            case 'alpha'
+                AlphaN = varargin{ii + 1};
+            case 'linewidth'
+                Lwidth = varargin{ii + 1};
+            case 'createfigure'
+                CreateFig = varargin{ii + 1};
+        end
+    end
+end
+if isempty(Tag)
+    MTag = '';
+else
+    MTag = ' - ';
+end
+if CreateFig
+    uq_figure('Name', sprintf('%s%sConfidence plot', MTag, Tag));
+end
+
+
+hold on
+
+% Plot the confidence bound:
+handles(2) = fill([X(:); flipud(X(:))],[Y(:) + Conf(:); flipud(Y(:) - Conf(:))],'r','LineStyle','none');
+handles(1) = plot(X, Y, 'Color', Lcolor, 'LineWidth', Lwidth);
+set(handles(2), 'FaceColor', BoundColor)
+alpha(AlphaN);
+box on
+
+
+% Remove the hold on:
+hold off

### 2019-09-25 sensitivity_analysis/UQLabCore_Rel1.0.0/lib/uq_printTable.m

@@ -1,0 +1,86 @@
+function varargout = uq_printTable(TableContents,Title)
+% function TSTRING = UQ_PRINTTABLE(COLTITLES,COLCONTENTS) print a table
+% with the provided titles and contents
+
+Tstr = [];
+
+%% Argument and consistency check
+
+if ~iscell(TableContents)
+    error('uq_printTable only accepts cell array inputs')
+end
+
+
+%% Retrieve table contents
+[NR, NC] = size(TableContents);
+
+%% Initialize table parameters based on the contents
+% identify maximum column width based on the cell contents
+maxColWidth = zeros(1,NC);
+for cc = 1:NC
+    for rr = 1:NR
+        % First convert numbers in strings
+        if isnumeric(TableContents{rr,cc})
+            TableContents{rr,cc} = sprintf('%.2g',TableContents{rr,cc});
+        end
+        % then check the maximum width of the corresponding column
+        maxColWidth(cc) = max(maxColWidth(cc), numel(TableContents{rr,cc}));
+    end
+end
+
+% If a title is given, check whether it's longer than the maximum length of
+% the columns. If it is, redistribute the extra space equally between them
+if exist('Title','var')
+   TotWidth = sum(maxColWidth)+4*(NC)+2;
+   TL = length(Title);
+   if sum(maxColWidth) < TL
+       DF = TL - TotWidth;
+       maxColWidth = maxColWidth + ceil(DF/NC);
+   end
+end
+
+%% Now let's use the format string for fixed width columns
+FStrings = cell(1,NC);
+for cc = 1:NC
+    FStrings{cc} = sprintf('%%-%ds',maxColWidth(cc));
+end
+
+%% Print the table
+TotWidth = sum(maxColWidth)+4*(NC)+2;
+hline = repmat('-',1,TotWidth);
+
+% Title first
+if exist('Title','var')
+    FTitle = sprintf('%%-%ds',TotWidth);
+    Tstr = [Tstr sprintf(FTitle,Title) '\n'];
+end
+
+for rr = 1:NR
+    if rr == 1 
+        Tstr = [Tstr sprintf(hline)];
+        Tstr = [Tstr sprintf('\n')];
+    end
+    % start each row with a vertical pipe
+    Tstr = [Tstr sprintf('|')];
+    for cc = 1:NC
+        Tstr = [Tstr sprintf('| ')];
[…]

### 2019-09-25 sensitivity_analysis/UQLabCore_Rel1.0.0/modules/uq_analysis/builtin/uq_inversion/MCMC/uq_HMC.m

@@ -1,0 +1,210 @@
+function [Sample,Accept,Time,ForwardModel,LogLikeliEval] = uq_HMC( Steps, ...
+                                        LeapfrogSteps, LeapfrogSize, Mass, ...
+                                        Seed, LogPrior, LogLikelihood, ...
+                                        Visualize, StoreModel)
+% UQ_HMC implements the Hamiltonian Monte Carlo MCMC algorithm 
+%   described in NEAL (2010) to sample from a posterior distribution.
+%
+%   SAMPLE = UQ_HMC(STEPS, LEAPFROGSTEPS, LEAPFROGSIZE, MASS, SEED, 
+%   LOGPRIOR, LOGLIKELIHOOD, VISUALIZE, STOREMODEL) samples from the posterior 
+%   distribution defined by the product: LOGPRIOR*LOGLIKELIHOOD. The
+%   algorithm tuning parameters are the number of leapfrog integrator steps
+%   LEAPFROGSTEPS, step size LEAPFROGSIZE and the mass matrix MASS. 
+%   STEPS is the number of undertaken MCMC iterations starting from the 
+%   SEED points. VISUALIZE is a structure containing options to show trace 
+%   plots of the sample chains at runtime. STOREMODEL is a logical variable 
+%   determining if the forward model evaluations are stored at runtime.
+%   SAMPLE is a 3D array containing the produced sample points.
+%   
+%   [SAMPLE, ACCEPT, TIME, FORWARDMODEL, LOGLIKELIEVAL] = UQ_HMC(STEPS, 
+%   LEAPFROGSTEPS, LEAPFROGSIZE, MASS, SEED, LOGPRIOR, LOGLIKELIHOOD, 
+%   VISUALIZE, STOREMODEL) additionally returns the acceptance rate ACCEPT 
+%   per chain, the total required execution time TIME, the forward model 
+%   runs FORWARDMODEL and the corresponding log-likelihood evaluations 
+%   LOGLIKELIEVAL.
+%
+%   References:
+%
+%   - R.M. Neal. MCMC using Hamiltonian dynamics. Handbook of Markov Chain
+%     Monte Carlo, 54:113-162, 2010
+%
+%   See also: UQ_INVERSION, UQ_PLOTMCMC, UQ_AM, UQ_MH, UQ_AIES                                    
+                                    
+% CHECK INPUT
[…]

### 2019-09-25 sensitivity_analysis/UQLabCore_Rel1.0.0/modules/uq_analysis/builtin/uq_inversion/MCMC/uq_MH.m

@@ -1,0 +1,166 @@
+function [Sample,Accept,Time,ForwardModel,LogLikeliEval] = uq_MH(  Steps,...
+                                        Seed, Proposal, LogPrior, ...
+                                        LogLikelihood, Visualize,...
+                                        StoreModel)                                    
+% UQ_MH implements the Metropolis-Hastings MCMC algorithm described in 
+%   HASTINGS (1970) to sample from a posterior distribution.
+%
+%   SAMPLE = UQ_MH(STEPS, SEED, PROPOSAL, LOGPRIOR, LOGLIKELIHOOD, 
+%   VISUALIZE, STOREMODEL) samples from the posterior distribution defined 
+%   by the product: LOGPRIOR*LOGLIKELIHOOD. A handle to the proposal 
+%   distribution is specified by PROPOSAL. STEPS is the number of 
+%   undertaken MCMC iterations starting from the SEED points. VISUALIZE is 
+%   a structure containing options to show trace plots of the sample chains
+%   at runtime. STOREMODEL is a logical variable determining if the forward 
+%   model evaluations are stored at runtime. SAMPLE is a 3D array 
+%   containing the produced sample points.
+%   
+%   [SAMPLE, ACCEPT, TIME, FORWARDMODEL, LOGLIKELIEVAL] = UQ_MH(STEPS, 
+%   SEED, PROPOSAL, LOGPRIOR, LOGLIKELIHOOD, VISUALIZE, STOREMODEL) 
+%   additionally returns the acceptance rate ACCEPT per chain, the total 
+%   required execution time TIME, the forward model runs FORWARDMODEL and
+%   the corresponding log-likelihood evaluations LOGLIKELIEVAL.
+%
+%   References:
+%
+%   - W.K. Hastings. Monte Carlo sampling methods using Markov chains and 
+%     their applications. Biometrika. 57:97-109, 1970
+%
+%   See also: UQ_INVERSION, UQ_PLOTMCMC, UQ_AIES, UQ_AM, UQ_HMC  
+
+% CHECK INPUT
+if or((numel(Steps) ~= 1),(mod(Steps,1) ~= 0))
+  error('The number of steps is not an integer!');
+end
+
+%%%%%%%%%%%%%%%%%%
+% INITIALIZATION %
+%%%%%%%%%%%%%%%%%%
+
+nPlotStep = Visualize.Interval; %plot steps  
[…]

### 2019-09-25 sensitivity_analysis/UQLabCore_Rel1.0.0/modules/uq_analysis/builtin/uq_inversion/MCMC/uq_getProposal.m

@@ -1,0 +1,51 @@
+function [x_cand, logCorrect] = uq_getProposal(Proposal,x_curr)
+% UQ_GETPROPOSAL samples from a proposal distribution
+%
+%   X_CAND = UQ_GETPROPOSAL(PROPOSAL, X_CURR) samples from the distribution 
+%   defined by the PROPOSAL structure conditioned on the sample point 
+%   X_CURR.
+%
+%   [X_CAND, LOGCORRECT] = UQ_GETPROPOSAL(PROPOSAL, X_CURR) additionally
+%   returns the logarithm of the metropolis-hastings correction factor.
+%
+%   See also: UQ_CREATECONDITINPUT, UQ_INVERSION
+
+% Initialize
+nChains = size(x_curr,1);
+
+if isfield(Proposal,'Cov')
+    % Supplied covariance matrix
+    % specify sampler and pdf handle
+    x_cand = mvnrnd(x_curr,Proposal.Cov,nChains);
+    logCorrect = zeros(nChains,1);
+else
+    % No supplied covariance matrix, use custom proposal
+    propDist = Proposal.Distribution;
+    propCond = Proposal.Conditioning;
+    % switch conditioning methods
+    switch lower(propCond)
+        case 'global' 
+            % no conditioning on current sample
+            x_cand = uq_getSample(propDist,nChains);
+            logCorrect = ...
+                uq_eval_logpdf(x_curr,propDist) - ...
+                uq_eval_logpdf(x_cand,propDist);
+        case 'previous'
+            % condition on current sample - create distribution objects
+            propDistCurr = uq_createConditInput(propDist,'mean',x_curr);
+            % sample
+            x_cand = zeros(size(x_curr));
+            for ii = 1:nChains
+                x_cand(ii,:) = uq_getSample(propDistCurr{ii},1);
+            end
+            % condition on candidate sample - create distribution objects
+            propDistCand = uq_createConditInput(propDist,'mean',x_cand);
+            % get correction factor
+            logCorrect = zeros(nChains,1);
+            for ii = 1:nChains
+                logCorrect(ii) = ...
+                    uq_eval_logpdf(x_curr(ii,:),propDistCand{ii}) - ...
+                    uq_eval_logpdf(x_cand(ii,:),propDistCurr{ii});
+            end
+    end
+end

### 2019-09-25 sensitivity_analysis/UQLabCore_Rel1.0.0/modules/uq_analysis/builtin/uq_inversion/Tests/uq_inversion_test_MCMC_MH_rwm.m

@@ -1,0 +1,44 @@
+function [success] = uq_inversion_test_MCMC_MH_rwm(level)
+% UQ_INVERSION_TEST_MCMC_MH_RWM tests the Metropolis-Hastings 
+%   algorithm for a simple conjugate Gaussian problem with a Gaussian
+%   proposal distribution.
+%
+%   See also: UQ_SELFTEST_UQ_INVERSION, UQ_MH
+
+%% START UQLAB
+uqlab('-nosplash');
+rng(8801)
+if nargin < 1
+    level = 'normal';
+end
+fprintf(['\nRunning: |' level '| uq_inversion_test_MCMC_MH_rwm...\n']);
+%% PROBLEM SETUP
+[y,Conj,PriorDist,ForwardModel,DiscrepancyOpt] = uq_inversion_test_MCMC_setup;
+
+%% SOLVER SETTINGS - Simple RWM
+Solver.Type = 'MCMC';
+Solver.MCMC.Sampler = 'MH';
+Solver.MCMC.Steps = 1e3;
+Solver.MCMC.NChains = 2;
+Solver.MCMC.Proposal.Cov = Conj.priorVariance;
+
+%% BAYESIAN MODEL
+BayesOpt.Type = 'Inversion';
+BayesOpt.Name = 'Bayesian model RWM';
+BayesOpt.Prior = PriorDist;
+BayesOpt.ForwardModel = ForwardModel;
+BayesOpt.Data.y = y;
+BayesOpt.Discrepancy = DiscrepancyOpt;
+BayesOpt.Solver = Solver;
+BayesianAnalysis = uq_createAnalysis(BayesOpt);
+
+
+%% SOME TESTING
+%Compute KL divergence between conjugate and samples
+KLDiv = uq_inversion_test_MCMC_computeKL(BayesianAnalysis,Conj);
+
+if KLDiv < 0.01
+    success = 1;
+else
+    success = 0;
+end

### 2019-09-25 sensitivity_analysis/UQLabCore_Rel1.0.0/modules/uq_analysis/builtin/uq_inversion/Tests/uq_inversion_test_func_AM.m

@@ -1,0 +1,68 @@
+function [success] = uq_inversion_test_func_AM(level)
+% UQ_INVERSION_TEST_FUNC_AM tests the functionality of the adaptive
+%   Metropolis algorithm.
+%
+%   See also: UQ_SELFTEST_UQ_INVERSION, UQ_AM
+
+%% START UQLAB
+uqlab('-nosplash');
+
+if nargin < 1
+    level = 'normal';
+end
+fprintf(['\nRunning: |' level '| uq_inversion_test_func_AM...\n']);
+
+%% PROBLEM SETUP
+load('uq_Example_BayesianLinearRegression');
+
+%% PRIOR DISTRIBUTION
+Prior.Name = 'Prior distribution';
+for i = 1:Model.M
+  Prior.Marginals(i).Name = sprintf('X%i',i);
+  Prior.Marginals(i).Type = 'Gaussian';
+  Prior.Marginals(i).Parameters = [0,1];
+end
+PriorDist = uq_createInput(Prior);
+
+%% FORWARD MODEL
+ModelOpt.Name = 'Forward model';
+ModelOpt.mHandle = @(x) x * Model.A;
+ModelOpt.isVectorized = true;
+ForwardModel.Model = uq_createModel(ModelOpt);
+
+%% UNKNOWN SIGMA
+DiscrepancyPriorOpt.Name = 'Prior of sigma';
+for ii = 1:size(Data,2)
+    DiscrepancyPriorOpt.Marginals(ii).Name = 'Sigma';
+    DiscrepancyPriorOpt.Marginals(ii).Type = 'Uniform';
+    DiscrepancyPriorOpt.Marginals(ii).Parameters = [0,20];
+end
+DiscrepancyPrior = uq_createInput(DiscrepancyPriorOpt);
+DiscrepancyOpt.Type = 'Gaussian';
+DiscrepancyOpt.Prior = DiscrepancyPrior;
+
+%% SOLVER SETTINGS
+% use MCMC with very few steps
+Solver.Type = 'MCMC';
+Solver.MCMC.Sampler = 'AM';
+Solver.MCMC.Steps = 100;
+Solver.MCMC.T0 = 10;
+Solver.MCMC.NChains = 2;
+
+%% BAYESIAN MODEL
+BayesOpt.Type = 'Inversion';
+BayesOpt.Name = 'Bayesian model';
+BayesOpt.Prior = PriorDist;
[…]

### 2019-09-25 sensitivity_analysis/UQLabCore_Rel1.0.0/modules/uq_analysis/builtin/uq_inversion/Tests/uq_inversion_test_func_DiscrepancyGroup.m

@@ -1,0 +1,73 @@
+function [success] = uq_inversion_test_func_DiscrepancyGroup(level)
+% UQ_INVERSION_TEST_FUNC_DISCREPANCYGROUP tests the functionality of the 
+%   Bayesian inversion module for multiple discrepancy/data groups.
+%
+%   See also: UQ_SELFTEST_UQ_INVERSION
+
+%% START UQLAB
+uqlab('-nosplash');
+
+%% PROBLEM SETUP
+load('uq_Example_BayesianLinearRegression');
+
+%% PRIOR DISTRIBUTION
+Prior.Name = 'Prior distribution';
+for i = 1:Model.M
+  Prior.Marginals(i).Name = sprintf('X%i',i);
+  Prior.Marginals(i).Type = 'Gaussian';
+  Prior.Marginals(i).Parameters = [0,1];
+end
+PriorDist = uq_createInput(Prior);
+
+%% FORWARD MODEL
+ModelOpt.Name = 'Forward model';
+ModelOpt.mHandle = @(x) x * Model.A;
+ModelOpt.isVectorized = true;
+ForwardModel.Model = uq_createModel(ModelOpt);
+
+%% DATA 
+myData(1).y = Data(1,1:12);
+myData(1).MOMap = (1:12);
+myData(1).Name = 'First Data';
+myData(2).y = Data(1,13:24);
+myData(2).MOMap = [ones(1,12);(13:24)];
+myData(2).Name = 'Second Data';
+
+%% DISCREPANCY MODEL
+DiscrepancyOpt(1).Type = 'Gaussian';
+DiscrepancyOpt(1).Parameters = 1;
+
+SigmaOpt.Name = 'Prior of sigma';
+SigmaOpt.Marginals(1).Name = 'Sigma';
+SigmaOpt.Marginals(1).Type = 'Uniform';
+SigmaOpt.Marginals(1).Parameters = [0,20];
+SigmaDist = uq_createInput(SigmaOpt);
+DiscrepancyOpt(2).Type = 'Gaussian';
+DiscrepancyOpt(2).Prior = SigmaDist;
+
+%% SOLVER SETTINGS
+% use MCMC with very few steps
+Solver.Type = 'MCMC';
+Solver.MCMC.Sampler = 'AM';
+Solver.MCMC.Steps = 1000;
+Solver.MCMC.NChains = 2;
+Solver.MCMC.Proposal.PriorScale = 1; 
[…]

### 2019-09-25 sensitivity_analysis/UQLabCore_Rel1.0.0/modules/uq_analysis/builtin/uq_inversion/Tests/uq_inversion_test_func_SigmaUnknownRow.m

@@ -1,0 +1,68 @@
+function [success] = uq_inversion_test_func_SigmaUnknownRow(level)
+% UQ_INVERSION_TEST_FUNC_SIGMAUNKNOWNROW tests the functionality of 
+%   Bayesian inversion module for unknown row discrepancy variance.
+%
+%   See also: UQ_SELFTEST_UQ_INVERSION
+
+%% START UQLAB
+uqlab('-nosplash');
+
+if nargin < 1
+    level = 'normal';
+end
+fprintf(['\nRunning: |' level '| uq_inversion_test_func_SigmaUnknownRow...\n']);
+%% PROBLEM SETUP
+load('uq_Example_BayesianLinearRegression');
+
+%% PRIOR DISTRIBUTION
+Prior.Name = 'Prior distribution';
+for i = 1:Model.M
+  Prior.Marginals(i).Name = sprintf('X%i',i);
+  Prior.Marginals(i).Type = 'Gaussian';
+  Prior.Marginals(i).Parameters = [0,1];
+end
+PriorDist = uq_createInput(Prior);
+
+%% FORWARD MODEL
+ModelOpt.Name = 'Forward model';
+ModelOpt.mHandle = @(x) x * Model.A;
+ModelOpt.isVectorized = true;
+ForwardModel.Model = uq_createModel(ModelOpt);
+
+%% UNKNOWN SIGMA
+DiscrepancyPriorOpt.Name = 'Prior of sigma';
+for ii = 1:size(Data,2)
+    DiscrepancyPriorOpt.Marginals(ii).Name = 'Sigma';
+    DiscrepancyPriorOpt.Marginals(ii).Type = 'Uniform';
+    DiscrepancyPriorOpt.Marginals(ii).Parameters = [0,20];
+end
+DiscrepancyPrior = uq_createInput(DiscrepancyPriorOpt);
+DiscrepancyOpt.Type = 'Gaussian';
+DiscrepancyOpt.Prior = DiscrepancyPrior;
+
+%% SOLVER SETTINGS
+% use MCMC with very few steps
+Solver.Type = 'MCMC';
+Solver.MCMC.Sampler = 'AM';
+Solver.MCMC.Steps = 100;
+Solver.MCMC.NChains = 2;
+Solver.MCMC.Proposal.PriorScale = 1; 
+Solver.MCMC.T0 = 1e3;
+
+%% BAYESIAN MODEL
+BayesOpt.Type = 'Inversion';
[…]

### 2019-09-25 sensitivity_analysis/UQLabCore_Rel1.0.0/modules/uq_analysis/builtin/uq_inversion/Tests/uq_selftest_uq_inversion.m

@@ -1,0 +1,90 @@
+function [pass] = uq_selftest_uq_inversion(level)
+% UQ_SELFTEST_UQ_INVERSION performs functionality and consistency
+%   tests for the Bayesian inversion module of UQLab. A number of tests are
+%   executed and their results are summarized and printed on screen.
+%
+% See also: UQ_INVERSION
+
+% start uqlab
+uqlab('-nosplash');
+
+% default level
+if nargin < 1
+    level = 'normal';
+end
+
+% tests to perform
+TestNames = {...
+    'uq_inversion_test_func_normpdf',...
+    'uq_inversion_test_func_lognormpdf',...
+    'uq_inversion_test_func_Visualize',...
+    'uq_inversion_test_func_InitOnly',...
+    'uq_inversion_test_func_noDiscrepancy',...
+    'uq_inversion_test_func_CustomLikelihood',...
+    'uq_inversion_test_func_SigmaKnownScalar',...
+    'uq_inversion_test_func_SigmaKnownScalar_WithConstant',...
+    'uq_inversion_test_func_SigmaKnownRow',...
+    'uq_inversion_test_func_SigmaKnownMatrix',...
+    'uq_inversion_test_func_SigmaUnknownScalar',...
+    'uq_inversion_test_func_SigmaUnknownRow',...
+    'uq_inversion_test_func_SigmaUnknownScalar_WithConstant',...
+    'uq_inversion_test_func_MultiModels',...
+    'uq_inversion_test_func_MultiModels_noDiscrepancy',...
+    'uq_inversion_test_func_DiscrepancyGroup',...
+    'uq_inversion_test_MCMC_MH_rwm',...
+    'uq_inversion_test_MCMC_MH_custProp',...
+    'uq_inversion_test_MCMC_AM_rwm',...
+    'uq_inversion_test_MCMC_AM_custProp',...
+    'uq_inversion_test_MCMC_AIES',...
+    'uq_inversion_test_MCMC_HMC',...
+    'uq_inversion_test_func_MH',...
+    'uq_inversion_test_func_AM',...
+    'uq_inversion_test_func_AIES',...
+    'uq_inversion_test_func_HMC',...
+    'uq_inversion_test_postProc',...
+    'uq_inversion_test_display',...
+    'uq_inversion_test_print'};
+
+% run the tests
+Ntests = length(TestNames);
+success = false(1, Ntests);
+Times = zeros(1, Ntests);
[…]

### 2019-09-25 sensitivity_analysis/UQLabCore_Rel1.0.0/modules/uq_analysis/builtin/uq_inversion/uq_display_uq_inversion.m

@@ -1,0 +1,752 @@
+function uq_display_uq_inversion(module, varargin)
+% UQ_DISPLAY_UQ_INVERSION graphically displays the results of an inverse
+%    analysis carried out with the Bayesian inversion module of UQLab.
+%
+%    UQ_DISPLAY_UQ_INVERSION(MODULE, NAME, VALUE) allows to choose
+%    more advanced plot functions by specifying Name/Value pairs:
+%
+%       Name               VALUE
+%       'acceptance'       Plots the acceptance rate per chain 
+%                          - Logical
+%                          default : false
+%       'scatterplot'      Plots a multi dimensional parameter scatter plot
+%                          of the parameters in the Results field of MODULE 
+%                          - Integer or 'all'
+%                          default : 'all'
+%       'predDist'         Plots the predictive distributions if available
+%                          from post processing (uq_postProcessInversion)
+%                          - Logical
+%                          default : true (if available)
+%       'meanConvergence'  Plots the convergence of the marginal mean for
+%                          the specified parameter averaged over all chains 
+%                          - Integer or 'all'
+%                          default : 0
+%       'trace'            Plots the trace plot of the marginal sample
+%                          points for the specified parameter 
+%                          - Integer or 'all'
+%                          default : 0
+%                          
+% See also: UQ_PRINT_UQ_INVERSION, UQ_POSTPROCESSINVERSION
+
+%% CONSISTENCY CHECKS
+if ~strcmp(module.Type, 'uq_inversion')
+   fprintf('uq_display_uq_inversion only operates on objects of type ''Inversion''') 
+end
+
+%check if MCMC Solver
+if or(~strcmp(module.Internal.Solver.Type, 'MCMC'),isempty(module.Results))
+    error('No results to display')
+end
+
+% switch if custom likelihood
+if module.Internal.customLikeli
+    CUSTOM_LIKELI = true;
+else
+    CUSTOM_LIKELI = false;
+end
+
+%% INITIALIZE
+% Check which post processed arrays are available
+if isfield(module.Results,'PostProc')
+    if isfield(module.Results.PostProc,'PriorPred')
+        priorPred_avail = true;
[…]

### 2019-09-25 sensitivity_analysis/UQLabCore_Rel1.0.0/modules/uq_analysis/builtin/uq_inversion/uq_eval_logpdf.m

@@ -1,0 +1,46 @@
+function value = uq_eval_logpdf(X,Input)
+% UQ_EVAL_LOGPDF calculates the logarithm of the multivariate probability 
+%     density function value that corresponds to samples of a random vector 
+%     collected in X with marginal distributions and copula specified by 
+%     the Input structure (or uq_default_input object) Input. Currently 
+%     only Gaussian copulas are implemented.
+%
+% See also UQ_EVAL_PDF, UQ_ALL_PDF, UQ_ALL_CDF, UQ_ALL_INVCDF
+
+marginals = Input.Marginals;
+copula    = Input.Copula;
+logXpdf   = log(uq_all_pdf(X,marginals));
+
+switch lower(copula.Type)
+
+    case 'independent'
+        % if the components of X are independent then the value of the log
+        % of the multivariate PDF is simply the sum of the univariate log PDFs
+        value = sum(logXpdf,2);
+
+    case 'gaussian'
+        % the components of X have dependency that is described by a
+        % Gaussian copula
+        R = copula.Parameters;
+        M = length(marginals);
+
+        % Specify the U marginals in the standard normal space
+        [U_marginals(1:M).Type]       = deal('Gaussian') ;
+        [U_marginals(1:M).Parameters] = deal([0,1]) ;
+
+        % Marginal CDFs
+        Xcdf = uq_all_cdf(X,marginals);
+
+        % Gaussian inverse CDF
+        invXcdf = uq_all_invcdf(Xcdf,U_marginals);
+        
+        % Finally calculate the multivariate PDF value based on the 
+        % logairthm of Sklar's theorem
+        % log(f_joint(x)) = log(c(F(x))) + sum(log(f(x_i)))
+        logDetR =  2 * sum(log(diag(chol(R))));
+        Cgauss =  (1/2)*logDetR -1/2*dot((invXcdf*(inv(R) - eye(M))),invXcdf, 2);
+        value = Cgauss + sum(logXpdf,2);
+
+    otherwise 
+        error('Copulas of type "%s" are not supported',copula.Type);
+end

### 2019-09-25 sensitivity_analysis/UQLabCore_Rel1.0.0/modules/uq_analysis/builtin/uq_inversion/uq_postProcessInversion.m

@@ -1,0 +1,556 @@
+function varargout = uq_postProcessInversion(module, varargin)
+% UQ_POSTPROCESSINVERSION post-processes an inverse analysis carried out
+%    with the Bayesian inversion module of UQLab.
+%
+%    UQ_DISPLAY_UQ_INVERSION(MODULE, NAME, VALUE) allows to choose
+%    more post processing options by specifying Name/Value pairs:
+%
+%       Name                  VALUE
+%       'burnIn'              Removes the burn in from the sample,
+%                             specified as a fraction (between 0 and 1) or
+%                             an Integer smaller than the number of
+%                             MCMC iterations
+%                             - Double or Integer
+%                             default : 0.5
+%       'badChains'           Removes chains specified by their id from the
+%                             sample
+%                             - Integer
+%                             default : 0
+%       'pointEstimate'       Computes a point estimate based on the
+%                             supplied sample.
+%                             - String : ('Mean','MAP')
+%                             default : 'mean'
+%       'percentiles'         Computes dimension-wise percentiles of 
+%                             supplied sample.
+%                             - Double
+%                             default : [0.05, 0.95]
+%       'gelmanRubin'         Computes the Rubin-Gelman scale reduction
+%                             factor
+%                             - Logical
+%                             default : false
+%       'prior'               Samples specified number of sample points 
+%                             from the prior distribution.
+%                             - Integer
+%                             default : 10,000
+%       'priorPredictive'     Samples specified number of sample points 
+%                             from the prior predictive distribution. This 
+%                             requires additional calls to the forward 
+%                             models
+%                             - Integer
+%                             default : 1,000
[…]

### 2019-09-25 sensitivity_analysis/UQLabCore_Rel1.0.0/modules/uq_analysis/builtin/uq_reliability/methods/uq_importance_sampling.m

@@ -1,0 +1,220 @@
+function [Results, Internal] = uq_importance_sampling(LimitStateFcn, SourceInput, InstrumentalInput, Options)
+% UQ_IMPORTANCE_SAMPLING conducts an importance sampling analysis based on
+% an instrumental density and a limit state function
+%
+% See also: UQ_SR_IMPORTANCE_SAMPLING
+
+% Now, select the instrumental input, for sampling during the method:
+%SourceMarginals = SourceInput.Marginals;
+%SourceCopula = SourceInput.Copula;
+
+if ~isfield(Options, 'Display') || ~isnumeric(Options.Display)
+    [~, Options] = uq_initialize_display(Options, Options);
+end
+Display = Options.Display;
+
+if nargin < 4
+    Options.Display = 1;
+    Options.Simulation.MaxSampleSize = 1e4;
+    Options.Simulation.BatchSize = 1e3;
+    Options.Simulation.Alpha = 0.05;
+end
+
+
+%% Get the input options 
+SimOpts = Options.Simulation;
+MaxSampleSize = SimOpts.MaxSampleSize;
+SampleSize = SimOpts.BatchSize;
+Alpha = SimOpts.Alpha;
+
+% Save evaluations or not:
+if isfield(Options, 'SaveEvaluations');
+    switch 1
+        case islogical(Options.SaveEvaluations) || isnumeric(Options.SaveEvaluations)
+            % These are both fine
+        case ischar(Options.SaveEvaluations)
+            % Parse the string:
+            switch lower(Options.SaveEvaluations)
+                case {'yes', 'true'}
+                    Options.SaveEvaluations = true;
+            end
+        otherwise
+            Options.SaveEvaluations = false;
+    end
+else
+    Options.SaveEvaluations = false;
+end
+
+% If there is no sampling strategy provided, stick to Monte Carlo:
+if ~isfield(SimOpts, 'Sampling') || ~ischar(SimOpts.Sampling);
+    SimOpts.Sampling = 'mc';
+end
+
+if isfield(SimOpts, 'TargetCoV') && ~isempty(SimOpts.TargetCoV);
+    % The method will stop when TargetCoV is reached
+    CoVThreshold = true; 
+    TargetCoV = SimOpts.TargetCoV;
+    
+else
+    % Do not check CoV convergence
+    CoVThreshold = false; 
+    
+end
+
+
+%% START IMPORTANCE SAMPLING:
[…]

### 2019-09-25 sensitivity_analysis/UQLabCore_Rel1.0.0/modules/uq_analysis/builtin/uq_reliability/tests/uq_Reliability_test_subsetsim_mc_constant.m

@@ -1,0 +1,119 @@
+function success = uq_Reliability_test_subsetsim_mc_constant(level)
+% SUCCESS = UQ_RELIABILITY_TEST_SUBSETSIM_MC_CONSTANT(LEVEL)
+%     This test checks whether the subset simulation returns exactly the 
+%     same value as Monte Carlo simulation in the case where the failure 
+%     probability is larger than p0=0.1 in the presence of constant inputs.
+%
+% See also UQ_SELFTEST_UQ_RELIABILITY
+
+%% Start test:
+uqlab('-nosplash');
+if nargin < 1
+    level = 'normal'; % TBD: Time that the tests will take
+end
+fprintf(['\nRunning: |' level '| ' mfilename '...\n']);
+
+
+%% set a seed
+seed = round(rand*100);
+
+
+%% threshold for numerical imprecision
+TH = 1e-8;
+
+
+%% create the input
+% Marginals:
+IOpts.Marginals(1).Name = 'R';
+IOpts.Marginals(1).Type = 'Gaussian';
+IOpts.Marginals(1).Moments = [1 0.8];
+
+IOpts.Marginals(2).Name = 'irrelevant';
+IOpts.Marginals(2).Type = 'constant';
+IOpts.Marginals(2).Parameters = rand(1);
+
+IOpts.Marginals(3).Name = 'S';
+IOpts.Marginals(3).Type = 'Gaussian';
+IOpts.Marginals(3).Moments = [0 0.6];
+
+IOpts.Marginals(4).Name = 'irrelevant';
+IOpts.Marginals(4).Type = 'constant';
+IOpts.Marginals(4).Parameters = 1.;
+
+% Copula definition:
+Rho = 0.15;
+IOpts.Copula.Type = 'Gaussian';
+copula_parameters = zeros(length(IOpts.Marginals));
+nonConst = ~ismember(lower({IOpts.Marginals.Type}),{'constant'});
+copula_parameters(find(nonConst),find(nonConst)) = [1, Rho; Rho, 1];
+IOpts.Copula.Parameters = copula_parameters;
+
+% Create the input:
+uq_createInput(IOpts);
+
+
+%% create the model
+MOpts.mString = '(X(:, 1) - X(:, 3)).*X(:,4) + X(:,1)';
+MOpts.isVectorized = true;
+uq_createModel(MOpts);
+
+
+%% MC analysis
+mopts.Type = 'reliability';
+mopts.Method = 'MC';
[…]

### 2019-09-25 sensitivity_analysis/UQLabCore_Rel1.0.0/modules/uq_analysis/builtin/uq_reliability/uq_initialize_uq_reliability.m

@@ -1,0 +1,576 @@
+function success = uq_initialize_uq_reliability(current_analysis)
+% success = UQ_INITIALIZE_UQ_RELIABILITY(current_analysis):
+%     initializes the structural reliability analyses defined in 
+%     CURRENT_ANALYSIS. All initializations are done in this file.
+% 
+% See also: UQ_RELIABILITY
+
+success = 0;
+
+
+%%
+% User options
+Options = current_analysis.Options; 
+
+% Actual options that the module will use
+Internal = current_analysis.Internal; 
+
+%% Check if the user has provided a Model and / or an Input
+[opt, Options] = uq_process_option(Options, 'Model', uq_getModel, 'uq_model');
+Internal.Model = opt.Value;
+
+[opt, Options] = uq_process_option(Options, 'Input', uq_getInput, 'uq_input');
+Internal.Input = opt.Value;
+
+%% Set the Method:
+% Depending on the method, some parts of the initialization script are
+% needed or not:
+InitSimulation = false;
+InitFORM = false;
+InitGradient = false;
+InitIS = false;
+InitSubsetSim = false;
+InitAKMCS = false;
+
+%% Reset the limitState count
+uq_evalLimitState('reset');
+
+%% Set the method specify defaults
+%process the method
+[Method, Options] = uq_process_option(Options, 'Method', 'mc', 'char');
+
+if Method.Missing || Method.Invalid
+    error('\nAn analysis method should be provided.\n');
+else
+    % Map the multiple methods description to a fixed set:
+    % mc, is, form, sorm
+    switch lower(Method.Value)
+        case {'mc','mcs','montecarlo', 'monte carlo'}
+            Internal.Method = 'mc';
+            % Needs to initialize the simulation options:
+            InitSimulation = true;
+            DEFAULTSaveEvaluations = 1;
+            DEFAULTBatchSize = 1e4;
+            
+        case {'importancesampling','is','importance sampling'}
+            Internal.Method = 'is';
+            % Needs to initialize the simulation options, FORM and the
+            % gradient:
+            InitSimulation = true;
+            InitFORM = true;
+            InitGradient = true;
+            InitIS = true;
+            DEFAULTSaveEvaluations = 1;
+            DEFAULTBatchSize = 1e2;
+            
+        case 'form'
[…]

### 2019-09-25 sensitivity_analysis/UQLabCore_Rel1.0.0/modules/uq_analysis/builtin/uq_sensitivity/Methods/Borgonovo_indices/uq_borgonovo_indices.m

@@ -1,0 +1,180 @@
+function Results = uq_borgonovo_indices(current_analysis)
+% RESULTS = UQ_SOBOL_BORGONOVO(ANALYSISOBJ): calculate the sample-based
+%     Borgonovo indices estimates for the problem specified in the analysis
+%     object ANALYSISOBJ. 
+%
+% See also: UQ_BORGONOVO_INDEX,UQ_SENSITIVITY
+
+%% Get the input options
+Options = current_analysis.Internal;
+
+% Initialize Cost
+Cost = 0;
+
+% Preallocation of Results
+Results = struct;
+
+% Verbosity
+Display = Options.Display;
+
+% Amount of variables
+M = Options.M;
+
+% Check if we need sampling
+Sampling = isfield(Options.Borgonovo,'Sampling');
+EDAvailable = isfield(Options.Borgonovo,'ExpDesign');
+
+
+%% SKIPPING UNUSED VARIABLES
+
+% The indices we want to compute:
+FactorIndex = Options.FactorIndex;
+
+% Number of factors that are considered:
+Mprime = sum(FactorIndex);
+
+% Collect the mean values of the unused factors to be used in the sampling
+if Mprime < M
+    Mu =  vertcat(Options.Input.Marginals.Moments);
+    Mu = Mu(:,1);
+end
+
+
+%% Check sampling
+if ~Sampling && ~EDAvailable
+    error(['You must either specify an experimental design or a model ',...
+        'and a sampling strategy in order to calculate Borgonovo indices.']);
+end
+% Sample if needed
+if Sampling
+    % Create a sample of X:
+    if Mprime < M
+        % Create an auxiliary input that sets the unused variables to constants
+        IOpts.Marginals = Options.Input.Marginals;
+        for jj = find(~FactorIndex)
+            IOpts.Marginals(jj).Type = 'Constant';
+            IOpts.Marginals(jj).Parameters = Mu(jj);
+        end
+        % Remove pre-calculated moments
+        IOpts.Marginals = rmfield(IOpts.Marginals,'Moments');
+        IOpts.Copula = Options.Input.Copula;
+        AuxInput = uq_createInput(IOpts, '-private');
+        % generate N samples
+        X_ED = uq_getSample(AuxInput, ...
+        Options.Borgonovo.SampleSize, ...
+        Options.Borgonovo.Sampling);
+    else
+        % generate N samples
[…]

### 2019-09-25 sensitivity_analysis/UQLabCore_Rel1.0.0/modules/uq_analysis/builtin/uq_sensitivity/Methods/Sobol_indices/uq_visualizeSobolIndices.m

@@ -1,0 +1,249 @@
+function uq_visualizeSobolIndices(Results, Outputs, varargin)
+% UQ_VISUALIZESOBOLINDICES(RESULTS,OUTIDX,VARARGIN): graphical
+%     representation of the Sobol' indices in the results structure
+%     RESULTS.
+%
+% See also: UQ_DISPLAY_UQ_SENSITIVITY
+
+%% Universal options for the Sobol' Indices Plots
+GlobalOptions.NIndices = 10; % max. number of Indices to plot (total and 1st order)
+GlobalOptions.NIndicesHigh = 5; % max number of indices to plot (higher orders)
+GlobalOptions.YTicks = 0:0.2:1; % Ticks of y-axis in plots
+GlobalOptions.AxisFontSize = 14; 
+GlobalOptions.NumOfColors = 6;
+GlobalOptions.IdxSorting = 'descend'; % How to sort high-order indices
+
+%% parse varargin options
+% initialization
+pie_flag = false;
+hist_flag = true;
+
+if nargin > 2
+    parse_keys = {'pie', 'hist'};
+    parse_types = {'f', 'f'};
+    [uq_cline, varargin] = uq_simple_parser(varargin, parse_keys, parse_types);
+    % 'coefficients' option additionally prints the coefficients
+    if strcmp(uq_cline{1}, 'true')
+        pie_flag = true;
+        hist_flag = false;
+    end
+    
+    % 'tolerance' option sets the default tolerance to plot coefficients
+    if strcmp(uq_cline{2}, 'true')
+        hist_flag = true;
+    end
+
+end
+
+%% Parse Input Arguments
+if ~exist('Outputs', 'var') || isempty(Outputs)
+    Outputs = 1;
+end
+
+Tag = '';
+
+if nargin < 3
+    if isfield(Results, 'Bootstrap') && ~isempty(Results.Bootstrap)
+        Bootstrap = 1;
+    else
+        Bootstrap = 0;
+    end
+else
+    Bootstrap = 0 ;
+end
+
+if nargin < 4 % Maybe some consistency checks...
+    MyColors = uq_cmap(GlobalOptions.NumOfColors);
+else
+    MyColors = varargin{2} ;
+end
+
+%% Produce plot(s)
+% Check how many outputs are there:
+NOuts = length(Outputs);
+OutputTag = '';
+for oo = Outputs
+    % Get the number of indices
+    NumOfIndices = length(Results.Total(:,oo));
[…]

### 2019-09-25 sensitivity_analysis/UQLabCore_Rel1.0.0/modules/uq_analysis/builtin/uq_sensitivity/Methods/uq_PCE_sobol_indices.m

@@ -1,0 +1,151 @@
+function Results = uq_PCE_sobol_indices(max_order, CurrentModel)
+% RESULTS = UQ_PCE_SOBOL_INDICES(MAX_ORDER,PCMODEL): analytically calculate Sobol'
+%     indices up to maximum order MAX_ORDER for the PCE model PCMODEL.
+%
+% See also: UQ_SOBOL_INDICES,UQ_SENSITIVITY
+%
+
+%% RETRIEVE INFORMATION FROM THE PCE MODEL
+
+% make sure the model is of "uq_metamodel" type.
+if ~strcmp(CurrentModel.Type, 'uq_metamodel')
+    error('Error, uq_getExpDesignSample is not defined for a model of type %s\n', CurrentModel.Type);
+end
+
+% Process Bootstrap, if available in the PCE model
+BootstrapFlag = false;
+if isfield(CurrentModel.Internal, 'Bootstrap') 
+    BOptions = CurrentModel.Internal.Bootstrap(1);
+    if isfield(BOptions,'BPCE') && isa(BOptions.BPCE, 'uq_model')
+       BootstrapFlag = true;
+    end
+end
+
+% Input/Output size, statistical moments of the PCE
+Nout = CurrentModel.Internal.Runtime.Nout;
+M = CurrentModel.Internal.Runtime.M;
+
+% if no max_order is requested, we choose the maximum order of the polynomials
+if ~exist('max_order', 'var')
+    max_order = 2;
+end
+
+% Variances
+for oo = 1:Nout
+    CurrentModel.Internal.Sensitivity.sobol_indices.sigma(oo) = norm(CurrentModel.PCE(oo).Coefficients(2:end),2);
+end
+
+%% Analytical Sobol's sensitivity indices
+%  Based on Sudret, 2008
+
+% Work only on non-constant variables
+nonConst = CurrentModel.Internal.Input.nonConst;
+
+% Initialize the (cell) array containing the (total) Sobol indices.
+sobol_cell_array = cell(max_order, 1) ;
+total_sobol_array = zeros(length(nonConst),Nout) ;
+
+% Initialize the cell to store the names of the variables of the Sobol Idx.
+VarIdx = cell(max_order,1);
+TotalVariance = zeros(1, Nout);
+
+% loop over the output variables
+for oo = 1:Nout
+    aoo = CurrentModel.PCE(oo).Coefficients ;
+    TotalVariance(oo) = sum(aoo(2:end).^2);
+    
[…]

### 2019-09-25 sensitivity_analysis/UQLabCore_Rel1.0.0/modules/uq_analysis/builtin/uq_sensitivity/Methods/uq_ancova_indices.m

@@ -1,0 +1,162 @@
+function Results = uq_ancova_indices(current_analysis)
+% RESULTS = UQ_ANCOVA_INDICES(ANALYSISOBJ) produces the ANCOVA importance
+%     indices for all variables in myInput using a provided PCE-Model
+
+% See also: UQ_SENSITIVITY,UQ_INITIALIZE_UQ_SENSITIVITY
+
+%% GET THE OPTIONS
+Options = current_analysis.Internal;
+
+% Verbosity
+Display = Options.Display;
+
+% Number of variables:
+M = Options.M;
+Mprime = sum(Options.FactorIndex);
+
+% Input object
+myInput = Options.Input;
+
+% Retrieve the model
+myModel = Options.Model;
+
+
+%% PRODUCE PCE METAMODEL IF NEEDED
+% a PCE will only include non constant input variables
+if ~Options.CustomPCE
+    if Display >1
+        fprintf('\nPlease note: A PCE of the provided model will be set up first.\n');
+    end
+    % Create the independent input
+    IndepInput = myInput.Options;
+    IndepInput.Copula.Type = 'Independent';
+    IndepInput.Copula.Parameters = eye(M);
+    myIndepInput = uq_createInput(IndepInput, '-private');
+    
+    % Add it to the PCE options
+    Options.PCEOpts.Input = myIndepInput;
+    Options.PCEOpts.FullModel = myModel;
+    
+    % Create the PCE metamodel from the options
+    myPCE = uq_createModel(Options.PCEOpts);
+    
+    % And save it in the Results
+    Results.PCE = myPCE;
+    
+    % Calculate the cost of the set up
+    if isfield(Options.PCEOpts.ExpDesign, 'NSamples')
+        Results.Cost = Options.PCEOpts.ExpDesign.NSamples; % input sample had to be evaluated
+    else
+        Results.Cost = 0;
+    end
+else
+    if Display > 1
+        fprintf('\nPlease note: The ANCOVA analysis will be done using the provided PCE metamodel.\n');
+    end
+    myPCE = myModel;
+    Results.Cost = 0;
+end
+
+%% Sampling and total variance
+% Get sample
+x_corr = uq_getSample(myInput,Options.ANCOVA.SampleSize,'lhs');
+
+% Produce PCE output variance and store it
+var_tot = var(uq_evalModel(myPCE,x_corr));
+Results.TotalVariance = var_tot;
+
+%% Identify subsets w, u & v for each variable
+
+% Storage preacllocation
[…]

### 2019-09-25 sensitivity_analysis/UQLabCore_Rel1.0.0/modules/uq_analysis/builtin/uq_sensitivity/Methods/uq_closed_sens_index.m

@@ -1,0 +1,156 @@
+function [Sclo,Cost,Options] = uq_closed_sens_index(Options,VariableSet)
+% [SCLO,COST,OPTIONS] = UQ_CLOSED_SENS_INDEX(OPTIONS,VARIABLESET)
+% produces the closed index of X_VARIABLESET, which corresponds to
+% Kucherenko's first order effect, and the associated cost.
+%   The OPTIONS are essentially the current analysis' options including a
+%   field named IndexOpts that is of type structure and contains
+%   .SampleSize, .Sampling, .Estimator and possibly samples .X1,.Y1,.X2,.Y2
+
+% See also: UQ_KUCHERENKO_INDICES, UQ_SHAPLEY_INDICES, UQ_TOTAL_SENS_INDEX
+
+%% GET THE INPUT ARGUMENTS
+
+% Verbosity
+Display = Options.Display;
+
+% Input object
+myInput = Options.Input;
+% Just another check to make sure the copula is Gaussian or Independent
+if  all(~strcmpi(myInput.Copula.Type,{'Independent','Gaussian'}))
+    error('Cannot compute sensitivity indices for copula of type "%s"', myInput.Copula.Type)
+end
+
+M = length(myInput.Marginals);
+
+% Get sampling specs
+N = Options.IndexOpts.SampleSize;
+Method = Options.IndexOpts.Sampling;
+
+% Kucherenko estimator
+Estimator = Options.IndexOpts.Estimator;
+
+% Determine which variables are not in the VariableSet
+noSet = myInput.nonConst(~ismember(myInput.nonConst,VariableSet));
+
+% Get the sample location
+Samples = Options.IndexOpts;
+
+% Turn variable indices into logicals
+% First check the format of the indices. We want them to be logical.
+if islogical(VariableSet)
+    % there should be M entries
+    if length(VariableSet) ~= M
+        fprintf('\n\nError: The conditioning indices are provided as logicals, \n but the length of the array is not equal to M!\n');
+        error('While initiating the conditional sampling')
+    end
+elseif isnumeric(VariableSet)
+    % maybe it's 1's and 0's and meant to be logical
+    if all(ismember(VariableSet,[0 1])) && length(VariableSet)==M
+        VariableSet = logical(VariableSet);
+    
+    % but maybe it's variable indices $\subset (1,...,M)$, turn into logical
+    elseif all(VariableSet < M+1) && length(unique(VariableSet)) == length(VariableSet)
+        logidx = false(1,M);
[…]

### 2019-09-25 sensitivity_analysis/UQLabCore_Rel1.0.0/modules/uq_analysis/builtin/uq_sensitivity/Methods/uq_correlation_indices.m

@@ -1,0 +1,63 @@
+function results = uq_correlation_indices(current_analysis)
+% RESULTS = UQ_CORRELATION_INDICES(ANALYSISOBJ): calculate the
+%     correlation-based sensitivity indices on the analysis object
+%     specified in ANALYSISOBJ. The results are stored in the RESULTS
+%     structure.
+%
+% See also: UQ_SENSITIVITY,UQ_INITIALIZE_UQ_SENSITIVITY
+
+Options = current_analysis.Internal;
+
+% Get the list of factors that need to be evaluated
+FactorIndex = Options.FactorIndex;
+
+%% Get a sample of points if the sample is not yet provided
+% figure out if we were already provided a sample or if we have to
+% calculate one.
+
+if isfield(Options.Correlation, 'Sample')
+    try 
+        Sample.X = Options.Correlation.Sample.X;
+        Sample.Y = Options.Correlation.Sample.Y;
+        NSamples = size(Sample.Y,1);
+    catch me
+        fprintf('The specified sample is not compatible with UQLab format: ')
+        Options.Correlation.Sample
+        rethrow(me);
+    end
+else
+    try % default to LHS sampling if not specified
+        NSamples = Options.Correlation.SampleSize;
+        if ~isfield(Options.Correlation, 'Sampling')
+            Sampling = 'LHS';
+        else 
+            Sampling = Options.Correlation.Sampling;
+        end
+        Sample.X = uq_getSample(Options.Input,NSamples, Sampling);
+        Sample.Y = uq_evalModel(Sample.X);
+    catch me
+        rethrow(me);
+    end
+end
+
+%% NOW EVALUATE THE CORRELATION COEFFICIENTS
+CorrIdx = corr(Sample.X, Sample.Y);
+RankCorrIdx = corr(Sample.X, Sample.Y, 'type', 'Spearman');
+
+% Remove the unwanted factors
+CorrIdx(~FactorIndex) = 0;
+RankCorrIdx(~FactorIndex) = 0;
+
+%% AND REPORT BACK THE RESULTS
+results.CorrIndices = CorrIdx;
+results.RankCorrIndices = RankCorrIdx;
+
+if Options.SaveEvaluations
+    results.ExpDesign = Sample;
+end
+
+if Options.Display > 0
+    fprintf('Correlation: Finished.\n');
+end
+
+results.Cost = NSamples;

### 2019-09-25 sensitivity_analysis/UQLabCore_Rel1.0.0/modules/uq_analysis/builtin/uq_sensitivity/Methods/uq_visualizeShapleyIndices.m

@@ -1,0 +1,141 @@
+function uq_visualizeShapleyIndices(Results, Outputs, varargin)
+% UQ_VISUALIZESHAPLEYINDICES(RESULTS,OUTIDX,VARARGIN): graphical
+%     representation of the Shapley indices in the results structure
+%     RESULTS.
+%
+% See also: UQ_DISPLAY_UQ_SENSITIVITY
+
+%% Universal options for the Sobol' Indices Plots
+GlobalOptions.NIndices = 10; % max. number of Indices to plot (total and 1st order)
+GlobalOptions.NIndicesHigh = 5; % max number of indices to plot (higher orders)
+GlobalOptions.YTicks = 0:0.2:1; % Ticks of y-axis in plots
+GlobalOptions.AxisFontSize = 14; 
+GlobalOptions.NumOfColors = 6;
+GlobalOptions.IdxSorting = 'descend'; % How to sort high-order indices
+
+%% parse varargin options
+% initialization
+pie_flag = false;
+hist_flag = true;
+
+if nargin > 2
+    parse_keys = {'pie', 'hist'};
+    parse_types = {'f', 'f'};
+    [uq_cline, varargin] = uq_simple_parser(varargin, parse_keys, parse_types);
+    % 'coefficients' option additionally prints the coefficients
+    if strcmp(uq_cline{1}, 'true')
+        pie_flag = true;
+        hist_flag = false;
+    end
+    
+    % 'tolerance' option sets the default tolerance to plot coefficients
+    if strcmp(uq_cline{2}, 'true')
+        hist_flag = true;
+    end
+
+end
+
+
+
+%% Parse Input Arguments
+if ~exist('Outputs', 'var') || isempty(Outputs)
+    Outputs = 1;
+end
+
+Tag = '';
+
+if nargin < 3
+    if isfield(Results, 'Bootstrap') && ~isempty(Results.Bootstrap)
+        Bootstrap = 1;
+    else
+        Bootstrap = 0;
+    end
+else
+    Bootstrap = 0 ;
+end
+
+if nargin < 4 % Maybe some consistency checks...
+    MyColors = uq_cmap(GlobalOptions.NumOfColors);
+else
+    MyColors = varargin{2} ;
+end
+%% Produce plot(s)
+% Check how many outputs are there:
+NOuts = length(Outputs);
+OutputTag = '';
+for oo = Outputs
+    % Get the number of indices
+    NumOfIndices = length(Results.Shapley(:,oo));
+    if NOuts > 1
[…]

### 2019-09-25 sensitivity_analysis/UQLabCore_Rel1.0.0/modules/uq_analysis/builtin/uq_sensitivity/Tests/uq_selftest_uq_sensitivity.m

@@ -1,0 +1,68 @@
+function pass = uq_selftest_uq_sensitivity(level)
+% PASS = UQ_SELFTEST_UQ_SENSITIVITY(LEVEL): non-regression and consistency
+%     test-suite for the Sensitivity module of UQLab. A number of tests are
+%     executed and their results are summarized and printed on screen.
+%
+% See also: UQ_SENSITIVITY
+
+uqlab('-nosplash');
+if nargin < 1
+    level = 'normal';
+end
+
+% Tests to perform:
+TestNames = {'uq_test_correlation',...
+    'uq_test_correlation_constant',...
+    'uq_test_SRC',...
+    'uq_test_perturbation',...
+    'uq_test_cotter_basic',...
+    'uq_test_sobol_indices',...
+    'uq_test_sobol_indices_constant',...
+    'uq_test_sobol_indices_PCE',...
+    'uq_test_sobol_indices_PCE_constant',...
+    'uq_test_morris',...
+    'uq_test_morris_high_order_interactions', ...
+    'uq_test_morris_linear_model', ...
+    'uq_test_morris_trajectory', ...
+    'uq_test_sobol_high_order_interactions', ...
+    'uq_test_cotter_high_order_interactions', ...
+    'uq_test_sensitivity_ishigami_outputs',...
+    'uq_test_sobol_indices_LRA',...
+    'uq_test_borgonovo_indices',...
+    'uq_test_ancova_indices',...
+    'uq_test_kucherenko_indices'...
+    };
+
+Ntests = length(TestNames);
+success = false(1, Ntests);
+Times = zeros(1, Ntests);
+for ii = 1:Ntests
+    TestTimer = tic;
+    success(ii) = eval([TestNames{ii} '(level);']);
+    Times(ii) = toc(TestTimer);
+end
+
+% Print out the results table and info:
+Result = {'ERROR','OK'};
+ResultChar = 60; % Character where the result of test is displayed
+MinusLine(1:ResultChar + 7) = deal('-');
+fprintf('\n%s\n',MinusLine);
+fprintf('UQ_SELFTEST_UQ_SENSITIVITY RESULTS');
+fprintf('\n%s\n',MinusLine);
+for ii = 1:Ntests
+    points(1:max(2,ResultChar-size(TestNames{ii},2))) = deal('.');
+    fprintf('%s %s %s @ %g sec.\n',TestNames{ii},points,Result{success(ii)+1},Times(ii));
[…]

### 2019-09-25 sensitivity_analysis/UQLabCore_Rel1.0.0/modules/uq_analysis/builtin/uq_sensitivity/Tests/uq_test_kucherenko_indices.m

@@ -1,0 +1,98 @@
+function success = uq_test_kucherenko_indices(level)
+% success = UQ_TEST_KUCHERENKO_INDICES(level) Test routine for Kucherenko indices.
+%     Evaluates the sensitivity for a portfolio function, and validate
+%     the results against analytical values. (Values from Kucherenko 2012)
+%
+% See also: UQ_SENSITIVITY, UQ_KUCHERENKO_INDICES
+
+success = 0;
+if nargin < 1
+    level = 'normal';
+end
+fprintf(['\nRunning: |', level,...
+    '| uq_test_kucherenko_indices...\n']);
+%% Start the framework:
+uqlab('-nosplash');
+
+%% Input
+M = 4;
+[Input.Marginals(1:M).Type] = deal('Gaussian');
+Input.Marginals(1).Parameters = [0 4];
+Input.Marginals(2).Parameters = [0 2];
+Input.Marginals(3).Parameters = [250 200];
+Input.Marginals(4).Parameters = [400 300];
+
+Input.Copula.Type = 'Gaussian';
+Input.Copula.Parameters = [ 1   0.3 0   0;...
+                            0.3 1   0   0;...
+                            0   0   1  -0.3;...
+                            0   0  -0.3 1];
+
+testInput = uq_createInput(Input, '-private');
+
+%% Ishigami model
+MOpts.mHandle = @(X) X(:,1).*X(:,3) + X(:,2).*X(:,4);
+MOpts.isVectorized = true;
+testModel = uq_createModel(MOpts, '-private');
+
+%% Reference indices
+
+reffirst = [0.507 0.399 0 0];
+reftotal = [0.492 0.3 0.192 0.108];
+
+%% Sensitivity
+% Initialization
+SOpts.Type = 'Sensitivity';
+SOpts.Input = testInput;
+SOpts.Model = testModel;
+SOpts.Method = 'Kucherenko';
+SOpts.Display = 'quiet';
+
+%% START THE TEST:
+
+% Absolute error allowed with respect to the analytical indices:
+AllowedError = 0.05;
+
+switch level
+    case 'slow'
+        SamplingStrat = {'mc', 'lhs', 'sobol', 'halton'};
+        
+    otherwise
+        SamplingStrat = {'mc'};
+        
+end
+
+% Further specs
+SOpts.Kucherenko.SampleSize = 15e3;
+SOpts.Kucherenko.Estimator = 'modified';
+
+for jj = 1:length(SamplingStrat)
+    % Loop on the sampling strategy
[…]

### 2019-09-25 sensitivity_analysis/UQLabCore_Rel1.0.0/modules/uq_analysis/builtin/uq_sensitivity/uq_initialize_uq_sensitivity.m

@@ -1,0 +1,845 @@
+function success = uq_initialize_uq_sensitivity(current_analysis)
+% SUCCESS = UQ_INITIALIZE_UQ_SENSITIVITY(ANALYSISOBJ): configure the
+%     analysis object ANALYSISOBJ based on the user specified options.
+%     Also the appropriate input/output structure of the ANALYSISOBJ object
+%     and run the analysis.
+%
+% See also: UQ_SENSITIVITY,UQ_CORRELATION_INDICES,UQ_SRC_INDICES,UQ_COTTER_INDICES,
+%           UQ_PERTURBATION_METHOD,UQ_MORRIS_INDICES,UQ_SOBOL_INDICES,
+%           UQ_ANCOVA_INDICES, UQ_KUCHERENKO_INDICES, UQ_SHAPLEY_INDICES
+
+
+%% INITIALIZATION AND RETRIEVAL OF THE REQUESTED OPTIONS
+% User-specified options as given in uq_createAnalysis(Options)
+Options = current_analysis.Options; 
+% Processed options that will be assigned to the object after
+% initialization
+Internal = current_analysis.Internal; 
+
+% Default to the full input distributions
+FactorDefaults.Boundaries = 1; 
+success = 0;
+
+
+
+%% CHECK FOR USER DEFINED MODEL/INPUT
+%  Retrieve the current ones if not defined
+
+% MODEL
+[opt, Options] = uq_process_option(Options, 'Model', uq_getModel);
+Internal.Model = uq_getModel(opt.Value);
+
+% INPUT
+[opt, Options] = uq_process_option(Options, 'Input', uq_getInput);
+Internal.Input = uq_getInput(opt.Value);
+CurrentInput = Internal.Input;
+
+% Determine the number of input variables from the following sources
+% (ordered by priority):
+% 1 - User-specified input object
+% 2 - Options.M
+
+% Define the methods, that do not need an input object
+methods_noinput = {'Correlation','SRC'};
+
+[opt, Options] = uq_process_option(Options, 'M', [], 'double');
+
+if ~isempty(opt.Value)
+    Internal.M = opt.Value;
+    M = Internal.M;
+    nonConst = 1:M;
+else
+    % throw error if input object is missing for non-allowed methods
+    if isempty(CurrentInput) && ~any(strcmpi(Options.Method,methods_noinput))
+        fprintf('\nError: It was not possible to determine the dimension of the problem.\n');
+        fprintf('\nYou can fix this by taking one of the following actions:');
[…]

### 2019-09-25 sensitivity_analysis/UQLabCore_Rel1.0.0/modules/uq_input/builtin/uq_default_input/copula/uq_NatafTransform.m

@@ -1,0 +1,58 @@
+function U = uq_NatafTransform( Xin, marginals, copula )
+% U = UQ_NATAFTRANSFORM(Xin, marginals, copula):
+%     performs (Generalised) Nataf Transformation of samples (X) 
+%     from the physical to the standard normal space (U)
+%
+% References:
+%
+%   Lebrun, R. and Dutfoy (2009), A., A generalization of the Nataf 
+%   transformation to distributions with elliptical copula, 
+%   Prob. Eng. Mech. 24(2), 172-178
+%
+% See also: UQ_INVNATAFTRANSFORM
+
+%% Get the non-constant components of the random vector
+% find the indices of the non-constant components
+Types = {marginals(:).Type};
+indConst = strcmpi(Types, 'constant'); 
+indNonConst = ~indConst ;
+% store non-constant components in X
+X = Xin(:,indNonConst) ;
+
+% NOTE: 
+% In the algorithm that follows the symbols from (Lebrun and Dutfoy, 2009)
+% are mostly adopted
+
+%% Do isoprobabilistic transformation from X to V
+% V denotes the samples in the standard elliptical space where covariance is still existent
+switch lower(copula.Type)
+    case 'gaussian'
+        [vMarginals(1:length(marginals(indNonConst))).Type] = deal('gaussian') ;
+        [vMarginals(1:length(marginals(indNonConst))).Parameters] = deal([0 1]) ;
+        V = uq_IsopTransform(X, marginals(indNonConst), vMarginals);
+    case 'student'
+        [vMarginals(1:length(marginals(indNonConst))).Type] = deal('student') ;
+        [vMarginals(1:length(marginals(indNonConst))).Parameters] = deal(1) ;
+        V = uq_IsopTransform(X, marginals(indNonConst), vMarginals);
+    otherwise
+        err('Error: Unknown type of elliptical copula!')
+        return;
+end
+%% Perform step 3: mapping from V to U
+% if the Cholesky decomposition of the correlation matrix already exists
+% use it, otherwise calculate it here (and store it)
+if isfield(copula,'cholR') && ~isempty(copula.cholR)
+    L = copula.cholR ;
+else
+    try
+        L = chol(copula.Parameters(indNonConst,indNonConst));
+        copula.cholR = L;
+    catch
+        error('Error: The copula correlation matrix is not positive definite or incorrectly defined!')
+    end
+end
+
+% Produce U (of non - constant marginals)
[…]

### 2019-09-25 sensitivity_analysis/UQLabCore_Rel1.0.0/modules/uq_input/builtin/uq_default_input/marginals/uq_IsopTransform.m

@@ -1,0 +1,89 @@
+function Y = uq_IsopTransform(X, X_marginals, Y_marginals)
+% Y = UQ_ISOPTRANSFORM(X, X_marginals, Y_marginals):
+%     Performs an isoprobabilistic transformation of samples from X to Y 
+%     where their corresponding marginal distributions are specified in the 
+%     structure arrays X_marginals and Y_marginals. In this function the 
+%     components of X and Y are assumed to be independent. 
+%
+%     For performing isoprobabilistic transformation of random vectors with
+%     dependent components check uq_GeneralIsopTransform
+%
+% See also UQ_GENERALISOPTRANSFORM
+
+% Check whether X and Y have the same marginal distributions. First check
+% only the first element to avaid the full comparison if they are different
+if strcmpi(X_marginals(1).Type,Y_marginals(1).Type) 
+    M = length(X_marginals);
+    [Xtypes{1:M}] = deal(X_marginals.Type);
+    [Ytypes{1:M}] = deal(Y_marginals.Type);
+    if all(strcmpi(Xtypes,Ytypes))
+        % All the types are the same, check moments
+        Xmoments = cell(1,M);
+        [Xmoments{:}] = deal(X_marginals(:).Parameters);
+        Ymoments = cell(1,M);
+        [Ymoments{:}] = deal(Y_marginals(:).Parameters);
+        % Handle cases where X or Y is bounded
+        Xbounds = cell(1,M);
+        Ybounds = cell(1,M);
+        if isfield(X_marginals, 'Bounds') 
+            [Xbounds{:}] = deal(X_marginals(:).Bounds);
+        else
+            [Xbounds{:}] = deal([-inf inf]);
+        end
+        if isfield(Y_marginals, 'Bounds') 
+            [Ybounds{:}] = deal(Y_marginals(:).Bounds);
+        else
+            [Ybounds{:}] = deal([-inf inf]);
+        end
+        if isequal(Xmoments,Ymoments) && isequal(Xbounds,Ybounds)
+            Y = X;
+            return;
+        end
+    end
+end
+
+%% Do the isoprobabilistic transformation
+cdfx = zeros(size(X));
+Y = zeros(size(X));
+for ii = 1 : length(X_marginals)
+    % if possible and there are no bounds defined, use a linear transform between Gaussians
+    if ~isfield(X_marginals(ii), 'Bounds') && ~isfield(Y_marginals(ii), 'Bounds') && ...
[…]

### 2019-09-25 sensitivity_analysis/UQLabCore_Rel1.0.0/modules/uq_input/builtin/uq_default_input/marginals/uq_all_invcdf.m

@@ -1,0 +1,59 @@
+function X = uq_all_invcdf(F, Marginals)
+% X = UQ_ALL_INVCDF(F, Marginals):
+%     Calculates the inverse Cumulative Distribution Function
+%     given CDF values of a random vector collected in F matrix. 
+%     Each column of F corresponds to a component of the random vector that 
+%     follows some distribution with some parameters as specified by the 
+%     Marginals structure array
+%
+%     For more information about available distributions and their 
+%     parameters please refer to the UQLab user manual: The Input module
+%
+% See also UQ_ALL_PDF, UQ_ALL_CDF
+
+X = zeros(size(F));
+for ii = 1:length(Marginals)
+    if strcmpi(Marginals(ii).Type, 'ks')
+        Marginals(ii).Parameters = Marginals(ii).KS;
+    end
+    % If bounds are given then calculate the bounded inverse CDF 
+    if isfield(Marginals(ii),'Bounds') && ...
+            ~isempty(Marginals(ii).Bounds)
+        if  ~isfield(Marginals(ii), 'Options')
+            Fbounds =  uq_cdfFun(Marginals(ii).Bounds(:), Marginals(ii).Type, ...
+                Marginals(ii).Parameters) ; 
+            Fa = Fbounds(1);
+            Fb = Fbounds(2) ; 
+            % 'squeeze' the inverse CDF so that it lies in [a,b] interval
+            X(:,ii) = uq_invcdfFun( F(:,ii)*(Fb-Fa) + Fa , Marginals(ii).Type, ...
+            Marginals(ii).Parameters);
+        else
+            if isfield(Marginals(ii), 'Options') && ~isempty(Marginals(ii).Options)
+                Fbounds =  uq_cdfFun(Marginals(ii).Bounds(:), Marginals(ii).Type, ...
+                    Marginals(ii).Parameters, Marginals(ii).Options) ;
+                Fa = Fbounds(1);
+                Fb = Fbounds(2) ;
+                X(:,ii) = uq_invcdfFun( F(:,ii)*(Fb-Fa) + Fa ,Marginals(ii).Type, ...
+                    Marginals(ii).Parameters, Marginals(ii).Options);
+            else
+                Fbounds =  uq_cdfFun(Marginals(ii).Bounds(:), Marginals(ii).Type, ...
+                    Marginals(ii).Parameters) ;
+                Fa = Fbounds(1);
+                Fb = Fbounds(2) ;
[…]

### 2019-09-25 sensitivity_analysis/UQLabCore_Rel1.0.0/modules/uq_input/builtin/uq_default_input/sampling/uq_sampleU.m

@@ -1,0 +1,78 @@
+function u = uq_sampleU(N, M, options)
+%UQ_SAMPLEU(N, M, options) returns N samples from the M-dimensional
+%uniform space [0,1]. The structure 'options' can be used to set some
+%additional options, namely:
+%
+%   Options.Method : allows to specify the sampling method. Possible
+%   values are 'MC' (default), 'grid', 'LHS', 'simpleLHS', 'Sobol',
+%   'Halton'
+%   Options.LHSiterations : allows to specify the number of different LHS
+%   designs that are going to be created (Options.Method = 'LHS')
+
+%% set some default values of not specified otherwise
+if ~exist('options', 'var')
+    options.Method = 'MC';
+    options.LHSiterations = 5;
+end
+
+switch lower(options.Method)
+    case {'regular', 'grid', 'meshgrid'}
+        % generate a regular mesh with N points in each direction
+        utmp = repmat(linspace(0,1,N).',1, M);
+        rhsstr = 'ndgrid(';
+        lhsstr = '[';
+        for ii=1:M
+            % to generate the final mesh
+            lsubstr = sprintf('tmpUi{%d},', ii);
+            lhsstr = sprintf('%s %s', lhsstr, lsubstr);
+            
+            rsubstr = sprintf('utmp(:, %d),', ii);
+            rhsstr = sprintf('%s %s',rhsstr, rsubstr);
+            
+        end
+        % remove the trailing comma and add the rest of the comvec command
+        rhsstr = [rhsstr(1:end-1) ');'];
+        lhsstr = [lhsstr(1:end-1) ']'];
+        eval([lhsstr ' = ' rhsstr]);
+        u = zeros(numel(tmpUi{1}), M);
+        for ii = 1:M
+            dimensions = ndims(tmpUi{ii}):-1:1;
+            tmpUi{ii} = permute(tmpUi{ii}, dimensions);
+            u(:,ii) = reshape(tmpUi{ii}, [],1);
+        end
+    case 'mc'
+        % standard Monte Carlo sampling
+        u = rand(N, M);
+    case 'lhs'
+        % Latin Hypercube Sampling
+        % add the LHSiterations if not provided
+        if ~isfield(options, 'LHSiterations') || isempty(options.LHSiterations)
+            options.LHSiterations = 5;
+        end
+        u = uq_lhs(N, M, 0, options.LHSiterations);
+    case 'simplelhs'
+        % Simple version of Latin Hypercube Sampling that does one
+        % iteration
+        u = uq_lhs(N, M,2);
[…]

### 2019-09-25 sensitivity_analysis/UQLabCore_Rel1.0.0/modules/uq_input/builtin/uq_default_input/tests/uq_default_input_test_DataMarginals.m

@@ -1,0 +1,61 @@
+function pass = uq_default_input_test_DataMarginals( level )
+% UQ_DEFAULT_INPUT_TEST_DATAMARGINALS validation test for data  
+% (kernel smoothing based) marginals of the default input module
+%
+% Summary:
+% Some marginal distributions are used to draw some samples that
+% are then used to define a kernel-smoothing-based random vector.
+% Samples are then drawn from that random vector and they are compared
+% in terms of their moments to the samples that were drawn from the
+% "true" random vector
+
+%% initialize test
+evalc('uqlab');
+if nargin < 1
+    level = 'normal'; 
+end
+fprintf(['\nRunning: |' level '| uq_default_input_test_DataMarginals...\n']);
+% fix the random seed
+rng(1001);
+
+%% parameters
+N0 = 5e3;
+N = 5e3;
+epsMean = 0.01; 
+epsStd  = 0.03;
+
+%% True Input
+inputOpts.Copula.Type = 'Gaussian';%% Generate the true input
+inputOpts.Marginals(1).Type = 'Gumbel';
+inputOpts.Marginals(1).Moments = [15 3];
+inputOpts.Marginals(2).Type = 'Beta';
+inputOpts.Marginals(2).Parameters = [2 3 0 7];
+inputOpts.Copula.Type = 'Gaussian';
+inputOpts.Copula.RankCorr = [1, 0.8; 0.8, 1];
+
+OrigInput = uq_createInput(inputOpts);
+
+% Get samples from the "true" input
+XOrig = uq_getSample(N0, 'LHS');
+
+%% Create the data-based input
+%  kernel smoothing-based marginals
+DataIN.Marginals(1).Type = 'ks';
+DataIN.Marginals(1).Parameters = XOrig(:,1);
+DataIN.Marginals(2).Type = 'ks';
+DataIN.Marginals(2).Parameters = XOrig(:,2);
+% Gaussian copula with appropriate spearman's correlation matrix
+DataIN.Copula.Type = 'Gaussian';
+DataIN.Copula.RankCorr = corr(XOrig, 'type', 'spearman');
+
+DataInput = uq_createInput(DataIN);
+
+% return the seed to original, so that the two sample sets are very close
+rng(1001);
+% Get samples from the data-based input
+XX = uq_getSample(N, 'LHS');
+
+%% validate results
[…]

### 2019-09-25 sensitivity_analysis/UQLabCore_Rel1.0.0/modules/uq_input/builtin/uq_default_input/uq_InputOptions.m

@@ -1,0 +1,38 @@
+% UQ_INPUTOPTIONS displays a helper for the main options needed to create an 
+% probabilistic input in UQLab.
+%    UQ_INPUTOPTIONS displays the main options needed by the command
+%    <a href="matlab:help uq_createInput">uq_createInput</a> to create an INPUT object in UQLab 
+%
+%    See also: uq_createInput, uq_getSample, uq_getInput, uq_listInputs,
+%              uq_selectInput 
+disp('Quickstart guide to the UQLab Input Module')
+disp('  ')
+disp('In the UQLab software, INPUT objects are created by the command:')
+disp('    myInput = uq_createInput(INPUTOPTIONS)')
+disp('The options are specified in the INPUTOPTIONS structure.')
+disp(' ');
+disp('Example: to create a 2-dimensional INPUT object that represents two random variables ')
+disp('in [-1,1] with covariance matrix C = [1 0.5; 0.5 1], type:')
+disp('    INPUTOPTIONS.Marginals(1).Type = ''Uniform'';')
+disp('    INPUTOPTIONS.Marginals(1).Parameters = [-1,1];')
+disp('    INPUTOPTIONS.Marginals(2).Type = ''Uniform'';')
+disp('    INPUTOPTIONS.Marginals(2).Parameters = [-1,1];')
+disp('    INPUTOPTIONS.Copula.Type = ''Gaussian'';')
+disp('    INPUTOPTIONS.Copula.Parameters = [1 0.5; 0.5 1];')
+disp('    myInput = uq_createInput(INPUTOPTIONS);')
+disp(' ');
+disp('To draw 100 samples from the resulting object, type:');
+disp('    X = uq_getSample(100)');
+disp(' ');
+AvailableMarginals = uq_getAvailableMarginals;
+disp('The following marginal distributions are available for use in UQLab:')
+for ii = 1:length(AvailableMarginals)
+fprintf('    - %s\n',regexprep(AvailableMarginals{ii},'(\<[a-z])','${upper($1)}'));
+end
+disp(' ');
+disp('Please refer to the Input User Manual (<a href="matlab:uq_doc(''Input'',''html'')">HTML</a>,<a href="matlab:uq_doc(''Input'',''pdf'')">PDF</a>)');
+disp('for more detailed information on the available features.')
+disp(' ');
+
+
+

### 2019-09-25 sensitivity_analysis/UQLabCore_Rel1.0.0/modules/uq_model/builtin/uq_metamodel/ExpDesign/uq_eval_ExpDesign.m

@@ -1,0 +1,63 @@
+function [Y] = uq_eval_ExpDesign(current_model, X)
+% Y = UQ_EVAL_EXPDESIGN(CURRENT_MODEL): evaluate the experimental desing as
+%     specified in the CURRENT_MODEL object
+%
+% Y = UQ_EVAL_EXPDESIGN(CURRENT_MODEL, X): evaluate the experimental desing as
+%     specified in the CURRENT_MODEL object on the points specified in X (which
+%     may differ from CURRENT_MODEL.ExpDesign.X;
+%
+% See also:
+% UQ_PCE_CALCULATE_COEFFICIENTS_REGRESSION,UQ_PCE_CALCULATE_COEFFICIENTS_PROJECTION
+
+
+%% Initialize common options and retrieve necessary information
+% Verbosity level
+DisplayLevel = current_model.Internal.Display;
+
+% If not specified in the command line, retrieve the points onto which to
+% calculate the experimental design from the CURRENT_MODEL object:
+if ~exist('X', 'var')
+    X = current_model.ExpDesign.X;
+end
+
+%% Calculate the experimental design only if it was not provided as an input
+switch lower(current_model.ExpDesign.Sampling)
+    case {'user','data'}
+        % do nothing: the ExpDesign is already calculated, just get it
+        Y = current_model.ExpDesign.Y;
+    otherwise % sampling methods
+        % get the full model
+        full_model = current_model.Internal.FullModel;
+        
+        if DisplayLevel > 1
+            fprintf('Evaluating the full model on the experimental design...\n')
+        end
+        
+        % evaluate the experimental design
+        Y = uq_evalModel(full_model,X);
+        
+        if DisplayLevel > 1
+            fprintf('Experimental design evaluation finished\n');
+        end
+end
+
+%% Preprocess the ED if configured
+if isfield(current_model.Internal,'ExpDesign') && isfield(current_model.Internal.ExpDesign,'PreprocY') && ~isempty(current_model.Internal.ExpDesign.PreprocY)
+    current_model.ExpDesign.Yraw = Y;
+    current_model.ExpDesign.Y = Y;
+    [current_model.ExpDesign, PostYPar] = uq_evalModel(current_model.Internal.ExpDesign.PreprocY,current_model.ExpDesign);
+    % re-associate the newly calculated exp design to Y;
+    Y = current_model.ExpDesign.Y;
[…]

### 2019-09-25 sensitivity_analysis/UQLabCore_Rel1.0.0/modules/uq_model/builtin/uq_metamodel/ExpDesign/uq_getExpDesignSample.m

@@ -1,0 +1,222 @@
+function [X, U, u, W] = uq_getExpDesignSample(current_model, varargin)
+% function [X, U, u, W] = UQ_GETEXPDESIGNSAMPLE(current_model,varargin):
+% get an experimental design and related information for the given
+% metamodelling type. The output arguments [X, U, u, W], when defined,
+% represent the ED in the original space X, the corresponding ED in the
+% reduced space U, the unit hypercube sample x and the quadrature weight
+% (if used) W.
+
+%% Argument and consistency checks
+
+% make sure the model is of "uq_metamodel" type.
+if ~strcmp(current_model.Type, 'uq_metamodel')
+    error('Error, uq_getExpDesignSample is not defined for a model of type %s\n', current_model.Type);
+end
+
+%% Retrieve of the experimental design options:
+EDOpts = current_model.ExpDesign;
+% Sampling scheme:
+sampling = EDOpts.Sampling;
+
+
+
+%% first retrieve the poly types. If not specified by the user as an option,
+% retrieve them as a function of the current_input module
+if strcmpi(current_model.MetaType,'pce') || strcmpi(current_model.MetaType,'lra')
+    switch lower(current_model.MetaType)
+        case 'pce'
+            Basis = current_model.PCE(1).Basis;
+        case 'lra'
+            Basis = current_model.LRA(1).Basis;
+    end
+    [PolyMarginals, PolyCopula] = uq_poly_marginals(Basis.PolyTypes, Basis.PolyTypesParams);
+end
+
+%% generation of the experimental design based both on the input and the metamodel
+% there are several different experimental design strategies, developed on the unit
+% hypercube of dimension M. Available methods, right now, are:
+% LHS - Latin Hypercube Sampling
+% MC  - Plain Monte Carlo
+% Sobol - Sobol pseudorandom series
+% Halton - Halton minimum discrepancy series
+% user - user-specified experimental design (directly by variables)
+% data - user-specified experimental design (loaded from .mat files)
+% gaussquad - standard gaussian quadrature
+% smolyakquad - Smolyak sparse grid with gaussian quadrature
+
+switch lower(sampling)
+    case {'data', 'user'}
+        % get the samples from the specified arrays
+        switch(lower(sampling))
+            case {'data','user'}
+                X = current_model.ExpDesign.X;
+                if isfield(current_model.ExpDesign, 'U')
+                    U = current_model.ExpDesign.U;
+                    u = U;
+                else
+                    U = X;
+                    u = U;
+                end
+        end
+        
[…]

### 2019-09-25 sensitivity_analysis/UQLabCore_Rel1.0.0/modules/uq_model/builtin/uq_metamodel/Kriging/eval_F/uq_Kriging_calc_auxMatrices.m

@@ -1,0 +1,84 @@
+function auxMatrices = uq_Kriging_calc_auxMatrices( R, F, Y, run_case )
+% UQ_KRIGING_CALC_AUXMATRICES: calculates some auxiliary matrices that can be
+% useful to have in order to speed up some calculations during training of a
+% Kriging metamodel as well as during predictions
+%
+% auxMatrices = UQ_KRIGING_CALC_AUXMATRICES( R, F, Y, run_case )
+% Given a correlation matrix R, and trend matrix F the following auxiliary
+% matrices are returned inside the auxMatrices structure:
+%   .cholR      : the Cholesky decomposition of R if exists otherwise NaN
+%   .Rinv       : the pseudo inverse of R in case the Cholesky decomposion
+%                 of it is not available due to R being ill conditioned. If
+%                 the Cholesky decomposition is available then this is
+%                 empty
+%   .FTRinv     : the quantity F' * R^(-1)
+%   .FTRinvF    : the quantity F' * R^(-1) * F
+%   .FTRinvF_inv: the pseudo-inverse (F' * R^(-1) * F)^(-1) in case 
+%                 F' * R^(-1) * F is ill-conditioned. Otherwise this field
+%                 is empty
+% ** aditional matrices that may be present in case of ML estimation
+% method:
+%   .Q          : (see documentation for description of this quantity)
+%   .G          : ( -//-)
+%   .Ytilde     : ( -//-)
+% the above fields will be empty in case of ill-conditioned R matrix.
+%
+% See also UQ_KRIGING_CALC_LEAVEKOUT, UQ_KRIGING_EVAL_J_OF_THETA_ML, 
+%  UQ_KRIGING_CALCULATE
+
+%1) try to calculate the Cholesky decomposition of R. This will speed up
+%   the calculation of the regression coefficients, beta, BUT it is only
+%   possible given that R is not ill-conditioned
+try
+    L = chol(R);
+    auxMatrices.cholR = L;
+    auxMatrices.Rinv = [];
+    FTRinv = (transpose(F) / L ) / transpose(L) ;
+    if any(strcmpi(run_case, {'ml_optimization', 'ml_estimation'}))
+        T = transpose(L);
+        Ytilde = T \ Y ;
+        [Q1,G] = qr(T \ F,0) ;
+        auxMatrices.Q1 = Q1;
+        auxMatrices.G = G;
+        auxMatrices.Ytilde = Ytilde;
+        auxMatrices.Ftilde = T \ F ;
[…]

### 2019-09-25 sensitivity_analysis/UQLabCore_Rel1.0.0/modules/uq_model/builtin/uq_metamodel/Kriging/eval_F/uq_Kriging_calc_sigmaSq.m

@@ -1,0 +1,38 @@
+function sigmaSq = uq_Kriging_calc_sigmaSq( parameters, estim_method )
+% UQ_KRIGING_CALC_SIGMASQ: returns the variance of the Guassian process
+% while trying to calculate it in the most efficient way depending on the
+% hyperparameters estimation method and in case of ML, the condition of the
+% correlation matrix R
+%
+% See also UQ_KRIGING_CALCULATE, UQ_KRIGING_EVAL_J_OF_THETA_ML
+
+switch lower(estim_method)
+    case 'cv'
+        sigmaSq =  sum(cell2mat(parameters.errors)./cell2mat(parameters.sigma_i_Sq)) ...
+            / parameters.CV_K ;
+    case 'ml_chol'
+          N = parameters.N ;
+          beta = parameters.beta ;
+          Ytilde = parameters.Ytilde;
+          Ftilde = parameters.Ftilde;
+          
+          z = Ytilde - Ftilde*beta ;
+          sigmaSq = transpose(z) * z / N; 
+    case 'ml_bypass_chol'
+        % bypass the calculation of the regression coefficients
+        N = parameters.N;
+        Q1 = parameters.Q1;
+        Ytilde = parameters.Ytilde;
+        
+        z = Ytilde - Q1*transpose(Q1)* Ytilde ;
+        sigmaSq = transpose(z) * z / N;
+     case {'ml_nochol', 'ml_bypass_nochol'}       
+        % calculation of beta cannot be avoided in that case
+        F = parameters.F;
+        Y = parameters.Y;
+        N = parameters.N ;
+        
+        beta = parameters.beta ;
+        z = Y - F* beta ;
+        sigmaSq = transpose(z) * parameters.Rinv * z  /N;
+end

### 2019-09-25 sensitivity_analysis/UQLabCore_Rel1.0.0/modules/uq_model/builtin/uq_metamodel/Kriging/eval_F/uq_Kriging_eval_F.m

@@ -1,0 +1,69 @@
+function F = uq_Kriging_eval_F( X, current_model )
+% F = UQ_KRIGING_EVAL_F(X, KRGMODEL): calculates the trend matrix (a.k.a. 
+% information matrix) F of the trend of a Kriging metamodel KRGMODEL on the 
+% points specified in X
+%
+% See also UQ_KRIGING_CALCULATE, UQ_KRIGING_EVAL, UQ_KRIGING_INITIALIZE, 
+% UQ_KRIGING_INITIALIZETREND
+
+M = current_model.Internal.Runtime.MnonConst;
+nonConstIdx = current_model.Internal.Runtime.nonConstIdx;
+N = size(X,1);
+
+% Obtain the current output
+current_output = current_model.Internal.Runtime.current_output ;
+
+%% IF the F function is given use it, otherwise proceed with polynomial F
+if isfield(current_model.Internal.Kriging(current_output).Trend,'CustomF') &&...
+        ~isempty(current_model.Internal.Kriging(current_output).Trend.CustomF)
+    % F is either numeric or cell array of handles
+    if isnumeric(current_model.Internal.Kriging(current_output).Trend.CustomF)
+        % F is numeric
+        F = repmat(current_model.Internal.Kriging(current_output).Trend.CustomF,N,1) ;
+    else
+        % This is the general case of custom trend
+        P = length(current_model.Internal.Kriging(current_output).Trend.CustomF);
+        F = zeros(N,P) ;
+        for ii = 1 : P
+            f_ii =  current_model.Internal.Kriging(current_output).Trend.CustomF{ii};
+            F(:,ii) = f_ii(X);
+        end
+    end
+else % the trend is not Custom
+    
+    PolyType = {current_model.Internal.Kriging(current_output).Trend.PolyTypes{nonConstIdx}};
+    P = current_model.Internal.Kriging(current_output).Trend.Degree;
+    
+    %% first get the polynomial basis index matrix
+    
+    % current maximum degree. It is set to the first value of Degree array, even for a non
+    % basis adaptive scheme
+    MaxDegree = current_model.Internal.Kriging(current_output).Trend.Degree(1) ;
+    
+    % if truncation is set to manual use the user-supplied polynomial indices, otherwise
+    % calculate them
+    
+    if isfield(current_model.Internal.Kriging(current_output).Trend.TruncOptions, 'Custom')% user-specified truncation
[…]

### 2019-09-25 sensitivity_analysis/UQLabCore_Rel1.0.0/modules/uq_model/builtin/uq_metamodel/Kriging/init/uq_Kriging_initialize_optimizer.m

@@ -1,0 +1,127 @@
+function optim_options = uq_Kriging_initialize_optimizer(current_model)
+% optim_options = UQ_KRIGING_INITIALIZE_OPTIMIZER(KRGModel) Parses various optimization 
+% options depending on the optimization method that is selected. It is assumed that 
+% for each output the same optimization method will be used
+%
+% See also: UQ_KRIGING_OPTIMIZER, UQ_KRIGING_CALCULATE
+
+
+nonConstIdx = current_model.Internal.Runtime.nonConstIdx;
+
+%% 
+if strcmpi(func2str(current_model.Internal.Kriging(1).GP.Corr.Handle),...
+        'uq_kriging_eval_r')
+    
+    %% filter out constants from the initial value and bounds definition
+    isIsotropic = current_model.Internal.Kriging(1).GP.Corr.Isotropic;
+    if isIsotropic
+        % do nothing about constants (assume that at least one variable is
+        % non-constant)
+    else
+        % anisotropic correlation function
+        if isfield(current_model.Internal.Kriging(1).Optim, 'InitialValue')
+            current_model.Internal.Kriging(1).Optim.InitialValue = ...
+             current_model.Internal.Kriging(1).Optim.InitialValue(nonConstIdx);
+        end
+
+        if isfield(current_model.Internal.Kriging(1).Optim, 'Bounds')
+            % Keep only  the bounds that correspond to the varying dimensions
+            % of the metamodel
+            current_model.Internal.Kriging(1).Optim.Bounds = ...
+                current_model.Internal.Kriging(1).Optim.Bounds(:,nonConstIdx);
+        end
+    end
+else
+   isIsotropic = false; 
+end
+
+%% parse optimization options
+switch lower(current_model.Internal.Kriging(1).Optim.Method)
+    case 'none'
+       optim_options = []; 
+       return
+    case {'gradbased','lbfgs', 'bfgs'}
+       optim_options = optimset(...
+    'Display', lower(current_model.Internal.Kriging(1).Optim.Display),...
+    'MaxIter', current_model.Internal.Kriging(1).Optim.MaxIter,...
+    'Algorithm','interior-point',...
+    'Hessian',{'lbfgs',current_model.Internal.Kriging(1).Optim.BFGS.nLM},...
+    'AlwaysHonorConstraints','none',...
+    'TolFun', current_model.Internal.Kriging(1).Optim.Tol );
[…]

### 2019-09-25 sensitivity_analysis/UQLabCore_Rel1.0.0/modules/uq_model/builtin/uq_metamodel/Kriging/optimizer/uq_Kriging_eval_J_of_theta_ML.m

@@ -1,0 +1,39 @@
+function J = uq_Kriging_eval_J_of_theta_ML(theta, parameters)
+% UQ_KRIGING_EVAL_J_OF_THETA_ML : Maximum Likelihood objective
+% function
+%
+% See also: UQ_KRIGING_OPTIMIZER, UQ_KRIGING_EVAL_J_OF_THETA_CV,
+% UQ_KRIGING_CALCULATE
+
+X = parameters.X;
+Y = parameters.Y;
+N = parameters.N;
+CorrOptions = parameters.CorrOptions;
+evalR_handle = CorrOptions.Handle ;
+F = parameters.F;
+trend_type = parameters.trend_type;
+
+try
+    % calculate R and auxiliary matrices
+    R = evalR_handle( X, X, theta, CorrOptions );
+    auxMatrices = uq_Kriging_calc_auxMatrices( R, F, Y, 'ml_optimization' );
+
+    if ~isnan(auxMatrices.cholR)
+        logDetR = 2* sum(log(diag(auxMatrices.cholR))) ;
+        parameters.Q1 = auxMatrices.Q1 ;
+        parameters.Ytilde = auxMatrices.Ytilde ;
+        sigmaSq = uq_Kriging_calc_sigmaSq( parameters, 'ml_bypass_chol' ) ;
+    else
+        eps = 1e-320; 
+        logDetR = log(max(det(R), eps));
+        
+        parameters.beta = uq_Kriging_calc_beta( ...
+                        F, trend_type, Y, 'standard', auxMatrices );
+        parameters.Rinv = auxMatrices.Rinv ;
+        sigmaSq = uq_Kriging_calc_sigmaSq( parameters, 'ml_bypass_NOchol' ) ;
+    end
+    
+    J = 0.5*(N * log(2*pi*sigmaSq) + logDetR + N) ;
+catch
+    J = realmax ;
+end

### 2019-09-25 sensitivity_analysis/UQLabCore_Rel1.0.0/modules/uq_model/builtin/uq_metamodel/Kriging/tests/uq_Kriging_selftest.m

@@ -1,0 +1,77 @@
+function pass = uq_Kriging_selftest(level)
+% PASS = UQ_KRIGING_SELFTEST(LEVEL): suite of non-regression and consistency
+%     checks for the Kriging module of UQLab
+%
+% See also: uq_selftest_uq_metamodel
+
+%% initialize test
+uqlab('-nosplash');
+
+if nargin < 1
+    level = 'normal'; % TBD: Time that the tests will take
+end   
+
+pass = 0;
+
+%% Are the required toolboxes available?
+% Required toolboxes for the Kriging module
+req_toolbox_names = {'Optimization Toolbox', 'Global Optimization Toolbox'};
+% Check 
+[ret_checks, ret_names] = uq_check_toolboxes();
+OPTIM_TOOLBOX_EXISTS = any(strcmpi(req_toolbox_names{1}, ret_names(ret_checks)));
+GOPTIM_TOOLBOX_EXISTS = any(strcmpi(req_toolbox_names{2}, ret_names(ret_checks)));
+
+%% Test Names are defined here
+TestNames = {'uq_Kriging_test_TrendTypes'};
+
+if OPTIM_TOOLBOX_EXISTS 
+    TestNames = [TestNames, {'uq_Kriging_test_ExpDesigns'}];
+else
+    warning('Skipping some tests due to absense of the Optimization Toolbox')
+end
+
+if OPTIM_TOOLBOX_EXISTS && GOPTIM_TOOLBOX_EXISTS
+    TestNames = [TestNames, {'uq_Kriging_test_OptimResult', ...
+        'uq_Kriging_test_CustomKriging','uq_Kriging_test_Constant'}];
+else
+    warning('Skipping some tests due to absense of the Global Optimization Toolbox and/or the Optimization Toolbox')
+end
+          
+%% Recursively execute each test defined in TestNames
+success = zeros(length(TestNames),1);
+Times = zeros(length(TestNames),1);
+TestTimer = tic;
+Tprev = 0;
+for iTest = 1 : length(TestNames)
+    % obtain the function handle of current test from its name
+    testFuncHandle = str2func(TestNames{iTest});
+    % run test
+    success(iTest) = testFuncHandle(level);
+    % calculate the time required from the current test to execute
+    Times(iTest) = toc(TestTimer) - Tprev ;
+    Tprev = Times(iTest);
+end
+
+
+%% Print out the results table and info:
[…]

### 2019-09-25 sensitivity_analysis/UQLabCore_Rel1.0.0/modules/uq_model/builtin/uq_metamodel/Kriging/tests/uq_Kriging_test_CustomKriging.m

@@ -1,0 +1,109 @@
+function pass = uq_Kriging_test_CustomKriging( level )
+% pass = UQ_KRIGING_TEST_CUSTOMKRIGING(LEVEL): non-regression test for the
+% custom Kriging (predictor-only) functionality in the Kriging module
+%
+% Summary:
+% Make sure that a user defined Kriging metamodel is working as expected
+
+%% initialize test
+pass = 1;
+if nargin < 1
+    level = 'normal';
+end
+evalc('uqlab');
+
+fprintf(['\nRunning: |' level '| uq_Kriging_test_CustomKriging...\n']);
+%% parameters
+eps = 1e-14 ;
+CorrFamilyHandle.Separable = @(X1,X2,th) max(0, 1 - abs(X1-X2)/th);
+CorrFamilyHandle.Ellipsoidal = @(h) max(0, 1 - h);
+if strcmpi(level, 'normal')
+    Families = {'matern-5_2',CorrFamilyHandle};    
+else
+    Families = {'matern-5_2','matern-3_2','gaussian','exponential',CorrFamilyHandle};    
+end
+
+%% input
+[inputopts.Marginals(1:3).Type] = deal('uniform') ;
+[inputopts.Marginals(1:3).Parameters] = deal([-pi pi]) ;
+% create the input
+myInput = uq_createInput( inputopts);
+
+%% model
+% Physical model: Ishigami function
+model.mFile = 'uq_ishigami_various_outputs' ;
+myModel = uq_createModel(model);
+
+%% first create a typical Kriging metamodel 
+if ~strcmpi(level,'normal')
+    Isotropy = {true, false};
+    CorrTypes = {'Separable','Ellipsoidal'};
+    % Get the indices of all possible combinations
+    combIdx = uq_findAllCombinations(Isotropy, CorrTypes, Families);
+    for ii  = 1 : length(combIdx)
+        % produce one different test-case for each combination
+        testCases(ii).Corr.Isotropic = Isotropy{combIdx(ii,1)};
+        testCases(ii).Corr.Type = CorrTypes{combIdx(ii,2)};
+        if isstruct(Families{combIdx(ii,3)})
+            testCases(ii).Corr.Family = ...
+                Families{combIdx(ii,3)}.(testCases(ii).Corr.Type);
+        else
[…]

### 2019-09-25 sensitivity_analysis/UQLabCore_Rel1.0.0/modules/uq_model/builtin/uq_metamodel/Kriging/tests/uq_Kriging_test_ExpDesigns.m

@@ -1,0 +1,128 @@
+function pass = uq_Kriging_test_ExpDesigns( level )
+% pass = UQ_KRIGING_TEST_EXPDESIGNS(LEVEL): non-regression test for the
+% various input scaling options in the Kriging module
+%
+% Summary:
+% Make sure that Kriging works with user defined experimental designs
+% regardless of scaling choice
+
+%% initialize test
+pass = 1;
+evalc('uqlab');
+if nargin < 1
+    level = 'normal'; 
+end
+fprintf(['\nRunning: |' level '| uq_Kriging_test_ExpDesigns...\n']);
+
+%% parameters
+Scaling_choices = [0,1];
+fname = 'KrigValTest.mat';
+N = 50;
+Nval = 500;
+eps = 1e-5;
+rng(10);
+
+%% Create input
+Input.Marginals.Type = 'Uniform' ;
+Input.Marginals.Parameters = [-1, 1] ;
+testInput = uq_createInput(Input);
+% Create Experimental Design
+model.Name = 'RungeFull';
+model.mFile = 'uq_runge' ;
+evalc('uq_createModel(model)');
+
+Xtrain = uq_getSample(N, 'Sobol');
+Xval = uq_getSample(Nval, 'Sobol');
+RungeFull = uq_selectModel(model.Name);
+
+Ytrain = uq_evalModel(RungeFull,Xtrain);
+Yval = uq_evalModel(RungeFull,Xval);
+
+% General kriging options
+metaopts.Type = 'Metamodel';
+metaopts.MetaType = 'Kriging';
+metaopts.Optim.InitialValue = 0.5;
+metaopts.Optim.Bounds = [0.01; 3] ;
+metaopts.Optim.Method = 'bfgs';
+metaopts.Optim.BFGS.nLM = 10;
+metaopts.Optim.MaxIter = 10;
+metaopts.Optim.Tol = 1e-4;
+metaopts.Trend.Type = 'polynomial';
+metaopts.Trend.Degree = 3;
+metaopts.Trend.PolyTypes = 'simple_poly';
+
+% Use ExpDesign of Sampling type 'user'
+metaopts.ExpDesign.X = Xtrain;
+metaopts.ExpDesign.Y = Ytrain;
+modelID = 0;
+for ii = 1 : length(Scaling_choices)
+    modelID = modelID + 1;
+    % 1) An input module has been defined
+    metaopts.Name = ['KrigModelTest',num2str(modelID)];
[…]

### 2019-09-25 sensitivity_analysis/UQLabCore_Rel1.0.0/modules/uq_model/builtin/uq_metamodel/Kriging/tests/uq_Kriging_test_OptimResult.m

@@ -1,0 +1,169 @@
+function pass = uq_Kriging_test_OptimResult( level )
+% UQ_KRIGING_TEST_OPTIMRESULT non-regression and validation test of the available
+% optimization methods for the Kriging module
+%
+% Summary:
+% Makes sure that different optimization methods using either ML or CV
+% estimation will converge to the correct minimum on some predefined
+% problems
+% 
+
+%% initialize test
+pass = 1;
+evalc('uqlab');
+
+if nargin < 1
+    level = 'normal';
+end
+fprintf(['\nRunning: |' level '| uq_Kriging_test_OptimResult...\n']);
+
+%% parameters
+eps_rel = 1e-2;
+rng(100);
+
+CorrFamilyHandle.Separable = @(X1,X2,th) max(0, 1 - abs(X1-X2)/th);
+CorrFamilyHandle.Ellipsoidal = @(h) max(0, 1 - h);
+CorrFamilies = {'exponential','linear','gaussian','matern-3_2','matern-5_2', 'handle'};
+CorrTypes = {'Separable','Ellipsoidal'};
+IsotropyCases = {true, false};
+optMethods = {'bfgs', 'hga','hsade'} ;
+estMethods = {'ml','cv'};
+avoid_BFGS_optim = [ 2 3 6]; % Corr. family cases to avoid gradient based optimisation
+Xtrain  = [ -1   -0.7143   -0.4286   -0.1429    0.1429    0.4286    0.7143 1]';
+Ytrain = [0.0385    0.0727    0.1788    0.6622    0.6622    0.1788    0.0727  0.0385]';
+
+% Get the indices of all possible combinations
+combIdx = uq_findAllCombinations(CorrFamilies, CorrTypes, IsotropyCases, optMethods, estMethods);
+% filter out the combinations that correspond to BFGS and families flaged
+% as "avoid_BFGS_optim"
+%accidx = ~(any(combIdx(:,1) == avoid_BFGS_optim,2) & combIdx(:,4) == 1);
+for ii = 1:length(avoid_BFGS_optim)
+   combIdx(combIdx(:,1)==avoid_BFGS_optim(ii) & combIdx(:,4) == 1,:) = []; 
+end
+%combIdx = combIdx(accidx,:);
+
+% for normal level randomly pick 30 cases
+if strcmpi(level, 'normal')
[…]

### 2019-09-25 sensitivity_analysis/UQLabCore_Rel1.0.0/modules/uq_model/builtin/uq_metamodel/Kriging/uq_Kriging_initialize.m

@@ -1,0 +1,1055 @@
+function success = uq_Kriging_initialize( current_model )
+% SUCCESS = UQ_KRIGING_INITIALIZE(KRGMODEL): Initialize a Kriging model based on the
+%     user-specified options.
+% 
+% See also: UQ_KRIGING_INITIALIZE_CUSTOM, UQ_PCE_INITIALIZE,
+% UQ_INITIALIZE_UQ_METAMODEL
+
+success = 0;
+
+M = current_model.Internal.Runtime.M;
+
+%% UNPROCESSED FIELDS
+skipFields = {'Type','Name','MetaType','Input','FullModel','ExpDesign', 'Display','ValidationSet'};
+
+%% DEFAULT VALUES
+% ExpDesign
+ScalingDefaults = 1 ;
+keepCache = 1; %flag to keep the cached matrices after the metamodel has been calculated
+
+% Trend
+TrendDefaults.Type = 'ordinary' ;
+TrendDefaults.Degree = 0 ;
+TrendDefaults.TruncOptions.qNorm = 1;
+TrendDefaults.CustomF = [];
+TrendDefaults.PolyTypes = cell(1,M);
+[TrendDefaults.PolyTypes{1:M}] = deal('simple_poly' );
+TrendDefaults.Handle = @uq_Kriging_eval_F;
+
+% Gaussian Process
+CorrDefaults.Family = 'matern-5_2';
+CorrDefaults.Type = 'ellipsoidal';
+CorrDefaults.Isotropic = false;
+CorrDefaults.Nugget = 1e-10;
+CorrDefaults.Handle = @uq_eval_Kernel;
+
+EstMethodDefaults.EstimMethod = 'CV' ;
+EstMethodDefaults.CV.LeaveKOut = 1;
+
+% Optimization
+OptimFieldsDatatypes.INITIALVALUE = 'double';
+OptimFieldsDatatypes.BOUNDS = 'double';
+OptimFieldsDatatypes.MAXITER = 'double';
+OptimFieldsDatatypes.TOL = 'double';
+OptimFieldsDatatypes.DISPLAY = 'char';
+OptimFieldsDatatypes.NLM = 'double';
+OptimFieldsDatatypes.NPOP = 'double';
+OptimFieldsDatatypes.NSTALL = 'double';
+OptimFieldsDatatypes.STRATEGIES = 'cell';
+OptimFieldsDatatypes.PSTR = 'double' ;
[…]

### 2019-09-25 sensitivity_analysis/UQLabCore_Rel1.0.0/modules/uq_model/builtin/uq_metamodel/Kriging/uq_Kriging_print.m

@@ -1,0 +1,219 @@
+function uq_Kriging_print(KRGModel, outArray, varargin)
+% UQ_KRIGING_PRINT(KRGMODEL,OUTARRAY,VARARGIN): pretty print information on the
+%     Kriging object in KRGMODEL for the specified set of output components
+%     OUTARRAY (default: OUTARRAY = 1).
+%
+% See also: UQ_KRIGING_DISPLAY,UQ_PCE_PRINT,UQ_PRINT_UQ_METAMODEL
+
+
+%% Consistency checks and command line parsing
+if ~exist('outArray', 'var') || isempty(outArray)
+    outArray = 1;
+    if length(KRGModel.Kriging) > 1;
+        warning('The selected Kriging metamodel has more than 1 output. Only the 1st output will be printed');
+        fprintf('You can specify the outputs you want to be displayed with the syntax:\n')
+        fprintf('uq_print(KRGModel, OUTARRAY)\nwhere OUTARRAY is the index of desired outputs, e.g. 1:3 for the first three\n\n')
+    end
+end
+if max(outArray) > length(KRGModel.Kriging)
+    error('Requested output range is too large') ;
+end
+
+%% parsing the residual command line
+% initialization
+if nargin > 2
+    parse_keys = {'beta', 'theta','F','optim','trend','gp','GP','R'};
+    parse_types = {'f', 'f','f', 'f','f', 'f','f','f'};
+    [uq_cline, varargin] = uq_simple_parser(varargin, parse_keys, parse_types);
+    % 'beta' option additionally prints beta
+    beta_flag = strcmp(uq_cline{1}, 'true');
+
+    % 'theta' option additionally prints beta
+    theta_flag = strcmp(uq_cline{2}, 'true');
+    
+    % 'F' option additionally prints the information matrix
+    F_flag = strcmp(uq_cline{3}, 'true');
+       
+    % 'optim' option additionally prints extensive optimization results
+    % report
+    optim_flag = strcmp(uq_cline{4}, 'true');
+    
+    % 'trend' option additionally prints extensive trend-related report
+    % report
+    trend_flag = strcmp(uq_cline{5}, 'true');
+    
+    % 'gp' option additionally prints extensive GP-related report
+    % report
[…]

### 2019-09-25 sensitivity_analysis/UQLabCore_Rel1.0.0/modules/uq_model/builtin/uq_metamodel/PCE/Display/uq_PCE_displayEE.m

@@ -1,0 +1,114 @@
+function uq_PCE_displayEE(myPCE, Order, CurOutput)
+% UQ_PCE_DISPLAYEE(PCEMODEL,ORDER,CURRENT_OUTPUT): plot the elementary effects
+%      for the specified PCEMODEL, up to interaction order ORDER for the
+%      output component CURRENT_OUTPUT.
+
+
+%% Command line parsing
+% default to first order
+if nargin < 2
+    Order = 1;
+end
+% default to 1st output
+if nargin < 3
+    CurOutput = 1;
+end
+
+if length(CurOutput) > 1
+    error('uq_PCE_displayEE only works for a single output conponent at a time');
+end
+
+alpha = 0.1;
+
+if Order > 2
+    error('uq_PCE_displayEE can only display elementary effects up to order 2');
+end
+
+
+%% Create the EEPCE
+myEEPCE = uq_PCE_createEEPCE(myPCE, Order, CurOutput);
+
+%% Collect all the necessary info for plotting
+% Input: will be used to create the grids
+myInput = myEEPCE.Internal.Input;
+
+% Variable names: will be used to properly name the plots
+VarNames = myEEPCE.Internal.VarNames;
+VarIdx = myEEPCE.Internal.VarIdx;
+
+
+%% Find plotting limits based on the input distributions
+Marginals = myInput.Marginals;
+M = length(Marginals);
+bounds = zeros(2,M);
+% Substitute the infinite bounds with the 95% interquantile range
+for ii = 1:M
+    bounds(1,ii) = uq_all_invcdf(0,Marginals(ii));
+    if isinf(bounds(:,1))
+        bounds(1,ii) = uq_all_invcdf(alpha/2,Marginals(ii));
+    end
+    bounds(2,ii) = uq_all_invcdf(1,Marginals(ii));
+    if isinf(bounds(2,ii))
+        bounds(2,ii) = uq_all_invcdf(1-alpha/2,Marginals(ii));
+    end
+end
+
+%% Now create the necessary plots and model evaluations
+% note: this works only in 1 or 2 dimensions
+switch Order
+    case 1
+        NGrid = 1000;
+        % create a 2x(ceil(M/2)) array of 1-dim subplots
+        D = ceil(M/2);
+        X = zeros(NGrid,M);
+        for ii = 1:M
+            X(:,ii) = linspace(bounds(1,ii),bounds(2,ii),NGrid)';
+        end
+        % evaluate the effects
[…]

### 2019-09-25 sensitivity_analysis/UQLabCore_Rel1.0.0/modules/uq_model/builtin/uq_metamodel/PCE/PolyBasis/uq_generate_basis_Apmj.m

@@ -1,0 +1,134 @@
+function [p_index, p_index_roots] = uq_generate_basis_Apmj(p, M, truncOpts)
+% BASIS = UQ_GENERATE_BASIS_APMJ(P, M, TRUNCOPTS): generate a polynomial chaos basis
+%     index to be used in evaluating the basis. P is the maximum polynomial order, M is the
+%     number of variables.
+%
+% Input Parameters:
+% 
+%  'p'            The order of the polynomials used
+%
+%  'M'            The dimensions along which to return the polynomial degree indices.
+%
+% Optional Parameters:
+%
+%  'truncOptions' Options relevant to the truncation of the basis.
+%
+% Return Values:
+%
+%  'BASIS'        A sparse matrix containing the indices that define the
+%                 basis
+%
+% See also UQ_GENERATE_JTUPLES_SUM_P
+
+
+
+% BASIS = UQ_GENERATE_BASIS_AMJP(P, M, TRUNCOPTS): generate a polynomial chaos basis
+%     index to be used in evaluating the basis. P is the maximum polynomial
+%     order, M is the number of variables. The optional structure TRUNCOPTS
+%     specifies the truncation strategy.
+%
+% See also: UQ_GENERATE_JTUPLES_SUM_P, UQ_PCE_CREATE_PSI
+
+%% parsing and checking the arguments for options and consistency
+
+% custom basis: the basis is already provided "as is", return it and do not
+% do anything
+if exist('truncOpts','var') && isfield(truncOpts, 'Custom')
+    % check the dimensionality and degree of the basis and w.r.t. the provided M
+    p_index = truncOpts.Custom;
+    p_index_roots = p_index;
+    if M ~= size(p_index, 2);
+       error('The custom basis dimension mismatch: %d provided, %d observed', M,size(p_index,2)); 
+    end
+    maxDeg = max(sum(p_index,2));
+    if maxDeg >max(p)
+       error('Custom basis max degree (%d) is higher than the specified max (%d)', maxDeg, p);
+    end
+    return;
+end
+
+% q-norm
+if exist('truncOpts', 'var') && isfield(truncOpts, 'qNorm')
+    truncation = 1;
+    q = truncOpts.qNorm;
+else
+    % default to no truncation 
+    truncation = 0;
+end
+
+% max interaction terms
+if exist('truncOpts', 'var') && isfield(truncOpts, 'MaxInteraction')
+    J = 1:min(M,truncOpts.MaxInteraction);
[…]

### 2019-09-25 sensitivity_analysis/UQLabCore_Rel1.0.0/modules/uq_model/builtin/uq_metamodel/PCE/uq_PCE_display.m

@@ -1,0 +1,85 @@
+function uq_PCE_display(PCEModel, outArray, varargin)
+% UQ_PCE_DISPLAY(PCMODEL,OUTARRAY,VARARGIN): plot the coefficient spectrum
+%     of the PCE in PCMODEL for output components specified in OUTARRAY
+%     (default OUTARRAY = 1). Colors and sizes of the circles correspond to 
+%     the total degree  of the corresponding polynomials
+% 
+% See also: UQ_KRIGING_DISPLAY, UQ_DISPLAY_UQ_METAMODEL
+
+if ~PCEModel.Internal.Runtime.isCalculated
+    fprintf('PCE module %s is not yet initialized!\nGiven Configuration Options:', PCEModel.Name);
+    PCEModel.Options
+    return;
+end
+
+
+%% Consistency checks and command line parsing
+if ~exist('outArray', 'var')
+    outArray = 1;
+    if length(PCEModel.PCE) > 1;
+        warning('The selected PCE has more than 1 output. Only the 1st output will be printed');
+        fprintf('You can specify the outputs you want to be displayed with the syntax:\n')
+        fprintf('uq_display(PCEModule, OUTARRAY)\nwhere OUTARRAY is the index of desired outputs, e.g. 1:3 for the first three\n\n')
+    end
+end
+if max(outArray) > length(PCEModel.PCE)
+    error('Requested output range is too large') ;
+end
+
+
+
+for ii = 1:length(outArray)
+    Coefs = PCEModel.PCE.Coefficients;
+    Basis = full(PCEModel.PCE(outArray(ii)).Basis.Indices);
+    
+    %  Get the non zero coefficients from the sparse storage
+    [Indices,~,NNZCoefs] = find(Coefs);   % Indices of non zero coefs and values
+    PsiDegrees=sum(Basis,2);     % degrees of all polynomial
+    PsiDegrees = PsiDegrees(Indices);      % degrees of non zero terms
+    logCoefs=log10(abs(NNZCoefs));          % log-scale of the coefs
+    
+    % Plot the coefficients
+    uq_figure('filename', 'PCECoeffSpectrum.fig',...
+        'Position', [50 50 500 400] );
+    try
+        MyColors = lines(6);
+    catch me
+        MyColors = jet(6);
+    end
+    
+    pp = [];
[…]

### 2019-09-25 sensitivity_analysis/UQLabCore_Rel1.0.0/modules/uq_model/builtin/uq_metamodel/PCE/uq_PCE_initialize_custom.m

@@ -1,0 +1,99 @@
+function success = uq_PCE_initialize_custom( current_model, Options)
+% success = UQ_PCE_INITIALIZE_CUSTOM(CURRENT_MODEL,OPTIONS) initialize a
+%     custom PCE (predictor only) in CURRENT_MODEL based on the options given
+%     in the OPTIONS structure
+%
+% See also: UQ_PCE_INITIALIZE
+
+% initialize return value to 0
+success = 0;
+if ~isfield(Options, 'PCE')
+    error('Custom PCE has been defined, but no PCE field is defined in the options!')
+end
+customPCE = Options.PCE;
+
+% check that the basis is is defined and retrieve it
+if ~isfield(customPCE, 'Basis')
+    error('Custom PCE has been defined, but no Basis was specified!')
+end
+Basis = customPCE.Basis;
+
+% retrieve the basis polynomial indices (\alpha vectors in the documentation)
+if ~isfield(Basis, 'Indices')
+    error('Custom PCE has been requested, but no basis indices are given!')
+end
+
+% retrieve the polynomial types
+if ~isfield(Basis, 'PolyTypes')
+    error('Custom PCE has been requested, but no polynomial types are given!')
+end
+
+% check that the coefficients are available
+if ~isfield(customPCE, 'Coefficients')
+    error('Custom PCE has been requested, but no polynomial coefficients are given!')
+end
+
+% Set the calculation method to 'custom' (no calculation)
+current_model.Internal.Method = 'custom';
+
+
+% Add the missing properties to the current_model object
+uq_addprop(current_model, 'PCE');
+current_model.PCE =  customPCE;
+
+
+% Don't allow custom PCE when no input has been defined
+if isfield(Options, 'Input') && ~isempty(Options.Input)
+    current_model.Internal.Input = uq_getInput(Options.Input);
+else
+    error('You have not specified an input module! Custom PCE is not allowed when an input module is not specifically defined.');
+end
+
+% loop over the output components and check that coefficients and basis
+% have the same dimensions. In case the recurrence terms of the orthogonal
+% polynomials have not  been calculated, calculate them and add them to the
+% Basis structure. 
+for oo = 1:length(current_model.PCE)
+    if length(current_model.PCE(oo).Coefficients) ~= size(current_model.PCE(oo).Basis.Indices,1)
[…]

### 2019-09-25 sensitivity_analysis/UQLabCore_Rel1.0.0/modules/uq_model/builtin/uq_metamodel/PCK/uq_PCK_display.m

@@ -1,0 +1,190 @@
+function uq_PCK_display(PCKRGModel, outArray, varargin)
+%Plots the mean and variance of a PC-Kriging predictor. Only works for 1- and 
+% 2-D inputs.
+
+% number of discretization points for the figures
+N1d = 500;
+N2d = 80;
+
+% check whether the PC-Kriging model has been calculated
+if ~PCKRGModel.Internal.Runtime.isCalculated
+    fprintf('PC-Kriging object %s is not yet initialized!\nGiven Configuration Options:', PCKRGModel.Name);
+    PCKRGModel.Options
+    return;
+end
+
+
+%% Consistency checks and command line parsing
+if ~exist('outArray', 'var')
+    outArray = 1;
+    if length(PCKRGModel.PCK) > 1;
+        warning('The selected PC-Kriging metamodel has more than 1 output. Only the 1st output will be printed');
+        fprintf('You can specify the outputs you want to be displayed with the syntax:\n')
+        fprintf('uq_print(PCKRGModule, OUTARRAY)\nwhere OUTARRAY is the index of desired outputs, e.g. 1:3 for the first three\n\n')
+    end
+end
+if max(outArray) > length(PCKRGModel.PCK)
+    error('Requested output range is too large') ;
+end
+
+
+%% parsing the residual command line
+% initialization
+if nargin > 2
+    parse_keys = {'R','nolegend'};
+    parse_types = {'f','f'};
+    [uq_cline, varargin] = uq_simple_parser(varargin, parse_keys, parse_types);
+    
+    % 'R' option additionally prints R matrix
+    R_flag = strcmp(uq_cline{1}, 'true');
+    noLegend_flag = strcmp(uq_cline{2}, 'true');
+        
+    flagWasSet = R_flag;
+else
+    flagWasSet = false;
+    noLegend_flag = false;
+end
+
+
+
+%% Produce non-default plots if any flag  was set
+if flagWasSet
+    if R_flag
+        for ii = 1:length(outArray)
+            current_output = outArray(ii);
+            uq_figure('name',sprintf('Output #%i', current_output), 'Position', [50 50 500 400])
+            imagesc(PCKRGModel.Internal.Kriging(current_output).Internal.Kriging.GP.R );
+            xlabel('Column index','FontSize', 14,'Interpreter','latex');
+            ylabel('Row index','FontSize', 14,'Interpreter','latex');
[…]

### 2019-09-25 sensitivity_analysis/UQLabCore_Rel1.0.0/modules/uq_model/builtin/uq_metamodel/SVR/uq_SVR_print.m

@@ -1,0 +1,157 @@
+function uq_SVR_print(SVRModel, outArray, varargin)
+
+%% Consistency checks and command line parsing
+if ~SVRModel.Internal.Runtime.isCalculated
+    fprintf('SVR object %s is not yet calculated!\nGiven Configuration Options:', SVRModel.Name);
+    SVRModel.Options
+    return;
+end
+if ~exist('outArray', 'var') || isempty(outArray)
+    outArray = 1;
+    if length(SVRModel.SVR) > 1
+        warning('The selected SVR metamodel has more than 1 output. Only the 1st output will be printed');
+        fprintf('You can specify the outputs you want to be displayed with the syntax:\n')
+        fprintf('uq_print(SVRModel, OUTARRAY)\nwhere OUTARRAY is the index of desired outputs, e.g. 1:3 for the first three\n\n')
+    end
+end
+if max(outArray) > length(SVRModel.SVR)
+    error('Requested output range is too large') ;
+end
+
+%% Produce the fixed header
+fprintf('\n%%--------------------------- SVR metamodel ---------------------------%%\n');
+fprintf('\tObject Name:\t\t%s\n', SVRModel.Name);
+
+%% Produce the default printout
+
+M = SVRModel.Internal.Runtime.M;
+fprintf('\tInput Dimension:\t%i\n', M);
+fprintf('\n\tExperimental Design\n')
+fprintf('\t\tSampling:\t%s\n', SVRModel.ExpDesign.Sampling)
+fprintf('\t\tX size:\t\t[%s]\n', [num2str(size(SVRModel.ExpDesign.X,1)),'x',num2str(size(SVRModel.ExpDesign.X,2))])
+fprintf('\t\tY size:\t\t[%s]\n', [num2str(size(SVRModel.ExpDesign.Y,1)),'x',num2str(size(SVRModel.ExpDesign.Y,2))])
+
+for ii =  1 : length(outArray)
+    % Get desired output
+    current_output = outArray(ii);
+    if length(outArray) > 1
+        fprintf('--- Output #%i:\n', current_output);
+    end
+    % Loss function
+    switch lower(SVRModel.Internal.SVR(current_output).Loss)
+        case 'l1-eps'
+            fprintf('\n\tLoss function:\t\t%s\n', 'L_1 epsilon-insensitive');
+        case 'l2-eps'
+            fprintf('\n\tLoss function:\t\t%s\n', 'L_2 epsilon-insensitive');
+        otherwise
[…]

### 2019-09-25 sensitivity_analysis/UQLabCore_Rel1.0.0/modules/uq_model/builtin/uq_metamodel/uq_calculateMetamodel.m

@@ -1,0 +1,35 @@
+function success = uq_calculateMetamodel(current_model, varargin)
+% SUCCESS = uq_calculateMetamodel(CURRENT_MODEL): entry point for the
+%     calculation of the metamodel object in CURRENT_MODEL. This function runs
+%     the required functions to calculate a metamodel based on their type
+%     (e.g., calculate coefficients for PCE, run hyperparameter optimization
+%     for Kriging, etc.). 
+%
+% See also UQ_PCE_CALCULATE_COEFFICIENTS, UQ_KRIGING_CALCULATE_COEFFICIENTS
+
+%% Argument retrieval and check
+if ~strcmp(current_model.Type, 'uq_metamodel')
+    error('Error: uq_metamodel cannot handle objects of type %s', current_model.Type);
+end
+
+%% Run the proper calculation function depending on the specified metamodel type
+switch lower(current_model.MetaType)
+    case 'pce'
+        success = uq_PCE_calculate_coefficients(current_model);
+    case 'lra'
+        success = uq_LRA_calculate_coefficients(current_model);
+    case 'kriging'
+        success = uq_Kriging_calculate(current_model);
+   case 'pck'
+        success = uq_PCK_calculate_coefficients(current_model);
+    case 'dicekriging'
+        success = uq_DiceKriging_calculate_coefficients(current_model);
+    case 'svr'
+        success = uq_SVR_calculate_coefficients(current_model);
+    case 'svc'
+        success = uq_SVC_calculate_coefficients(current_model);
+end
+
+if success
+   current_model.Internal.Runtime.isCalculated = 1; 
+end

### 2019-09-25 sensitivity_analysis/UQLabCore_Rel1.0.0/modules/uq_model/builtin/uq_metamodel/uq_initialize_uq_metamodel.m

@@ -1,0 +1,376 @@
+function success = uq_initialize_uq_metamodel(module)
+% UQ_INITIALIZE_UQ_METAMODEL: initializes the UQ metamodel infrastructure on the basis of
+% the specified module
+
+%% argument and consistency checks
+% if the module is not specified on the command line, retrieve the default
+if exist('module', 'var')
+    current_model = uq_getModel(module);
+else
+    current_model = uq_getModel;
+end
+
+% initialize the success status to 0
+success = 0;
+
+% retrieve the options (new structure)
+Options = current_model.Options;
+
+
+% some checks: the model must be of "uq_metamodel" type
+if ~strcmp(current_model.Type, 'uq_metamodel')
+    success = -1;
+    error('uq_initialize_uq_metamodel error: you must initialize a uq_metamodel type object, not a % one!!', module.Type);
+end
+% make sure that some metamodel type has been selected
+if ~isfield(Options,'MetaType') || isempty(Options.MetaType)
+    error('Could not initialize the metamodelling tool without defining a metamodel type!')
+end
+
+MetaTypeProp = uq_addprop(current_model, 'MetaType', Options.MetaType);
+MetaTypeProp.Hidden = true;
+
+%% VERBOSITY LEVEL
+DISPLAYDefaults = 1; % Only essential information
+
+[Options, current_model.Internal] = uq_initialize_display(Options, current_model.Internal);
+
+DisplayLevel = current_model.Internal.Display;
+if DisplayLevel > 1
+    fprintf('Initializing the metamodeling infrastructure... ');
+end
+
+%% VALIDATION SET
+[ValSet, Options] = uq_process_option(Options, 'ValidationSet');
+ValFlag = false;
+if ~ValSet.Missing
+    VSet = ValSet.Value;
+    if ~isfield(VSet(1),'X') || ~isfield(VSet(1),'Y')
+        error('The provided validation set is invalid');
+    else
+        ValFlag = true;
+        current_model.Internal.ValidationSet = VSet;
+    end
+end
+
+%% MANUAL DEFINITION OF PCE (predictor only)
+% in case we want to manually specify the PCE basis + coeff only and use
+% predictor without calculating, provide all the necessary information
+if isequal(lower(Options.MetaType), 'pce')
+    if isfield(Options, 'Method') && isequal(lower(Options.Method), 'custom')
+        uq_PCE_initialize_custom(current_model, Options);
+        return;
+    end
+end
+
[…]

### 2019-09-25 sensitivity_analysis/UQLabCore_Rel1.0.0/modules/uq_model/builtin/uq_uqlink/test/uq_uqlink_test_RecoverResume.m

@@ -1,0 +1,107 @@
+function pass = uq_uqlink_test_RecoverResume( level )
+% UQ_UQLINK_TEST_RECOVERRESUME tests the 'recover' and 'resume' options of
+% the uq_evalModel for uqlink objects
+%
+% See also: UQ_READ_SSBEAMDEFLECTION.m UQ_UQLINK_TEST_POSSIBLECASES.m
+
+% parameters
+pass = true ;
+eps_th = 1e-5 ;
+% X = [b h L E p]; Y = V = beam deflection
+X = [0.15 0.3 5 30000e6 1e4;
+     0.15 0.3 5 30000e6 1e4;
+     0.15 0.3 5 30000e6 1e4] ;
+ 
+Ytrue = 5 * X(:,5) .* X(:,3).^4 ./ (32 * X(:,4) .* X(:,1) .* X(:,2).^3) ;
+
+if nargin < 1
+    level = 'normal';
+end
+fprintf(['\nRunning: |' level '| uq_uqlink_test_ResumeRecover...\n']);
+
+uqlab('-nosplash')
+
+%% Executable name depends on architecture
+EXECBASENAME = 'myBeam';
+if ispc
+    EXECSUFFIX = 'win';
+elseif isunix
+    if ~ismac
+        EXECSUFFIX = 'linux';
+    else
+        EXECSUFFIX = 'mac';
+    end
+end
+EXECNAME = [EXECBASENAME '_' EXECSUFFIX];
+
+%% 
+Mopts.Type = 'UQLink' ;
+Mopts.Name = 'my Beam' ;
+Mopts.Command = [fullfile(uq_rootPath,'Examples','UQLink','C_SimplySupportedBeam', EXECNAME), ' SSB_Input.inp'] ;    
+Mopts.ExecutionPath = fullfile(uq_rootPath,'modules','uq_model','builtin','uq_uqlink','test') ;
+Mopts.Template = 'SSB_Input.inp.tpl' ;
+Mopts.Output.Parser= 'uq_read_SSBeamDeflection' ;
+Mopts.Output.FileName = 'SSB_Input.out' ;
+Mopts.Display = 'quiet' ;
+
+myModel = uq_createModel(Mopts) ;
+Yval = uq_evalModel(myModel,X);
+
+%% CASE 1: One of the output is NaN: It is found automatically and run again
+% Load Processed input/output and modify them
+load(fullfile(Mopts.ExecutionPath,'myBeam.mat'), 'uq_ProcessedX', 'uq_ProcessedY') ;
[…]

### 2019-09-25 sensitivity_analysis/cases/aero_module/aero_module.m

@@ -24,7 +24,7 @@ thickness = computeThickness(1, randVecThickness, P{31}, 0); % computeThickness 
 
 
 %% Get one random samples for YAWANGLE
-YAWANGLE = P{20} + P{35}.*X(end);
+YAWANGLE = X(end);
 %% Write the random data into the input file of AERO module
 writeAeroModuleInput(P{1},  P{2},  P{10}, P{3},  chord, ...
                      thickness./chord,  twist, P{7},  P{8},  P{9}, ... 

### 2019-09-25 sensitivity_analysis/testSensitivity.m

@@ -12,7 +12,7 @@ SobolOpts.Sobol.Order = 1;
 
 %% initialize UQlab
 % add path
-addpath(genpath('../../UQLabCore_Rel1.0.0/'));
+addpath(genpath([pwd,'/UQLabCore_Rel1.0.0/']));
 % start uqlab
 uqlab
 %% process input files
@@ -55,9 +55,10 @@ if (find(strcmp(methods,'MC')))
             
             disp(NsamplesMC(i));
             % get random samples ('experimental design')
-            X_ED = uq_getSample(NsamplesMC(i), 'MC') ;
+            X_ED = uq_getSample(NsamplesMC(i), 'MC');
+            
             % evaluate model at sample
-            Y_ED = uq_evalModel(myModel,X_ED) ;
+            Y_ED = uq_evalModel(myModel,X_ED);
             
             % moments of solution
             mean_MC(k,i) = mean(Y_ED);
@@ -68,10 +69,10 @@ if (find(strcmp(methods,'MC')))
             SobolAnalysis_MC           = uq_createAnalysis(SobolOpts);
             SobolResults_MC            = SobolAnalysis_MC.Results;
             Sobol_MC_FirstOrder(k,i,1:ndim) = SobolResults_MC.FirstOrder;
-            Sobol_MC_Total(k,i,1:ndim)      = SobolResults_MC.Total;
-            
+            Sobol_MC_Total(k,i,1:ndim)      = SobolResults_MC.Total;         
         end
-    end   
+    end
+    
     % take average over first dimension (multiple MC runs)
     AVG_Sobol_MC_FirstOrder = squeeze(mean(Sobol_MC_FirstOrder,1));
     AVG_Sobol_MC_Total      = squeeze(mean(Sobol_MC_Total,1));

### 2019-10-17 Paper/article.tex
some updates in article

@@
 \newcommand{\Lomd}{L^2(\Omega,\mathcal{D})}
-\newcommand{\myreferences}{../Bibliography/references}
+\newcommand{\myreferences}{./Bibliography/references}
 \bibliographystyle{elsarticle-num}
@@
 \usepackage{stix2}
-\usepackage{pgfplots}
 \usepackage{appendix}
 \usepackage[labelfont=bf]{caption}
-\pgfplotsset{compat=newest}
 \pagestyle{myheadings}
 %%%%%%%%%%%%%%%%%%%%%%%
-\title{A Global Sensitivity Analysis of Wind Turbine Aeroelastic Models}
+\title{A global sensitivity analysis of wind turbine aeroelastic models}
 \begin{document}
@@
 \begin{abstract}
-Wind turbines are a complex multi-physics systems.
-As the performance and the life span of a turbine is highly dependent on  manufacturing, meteorological and operational factors, it is very important to quantify the impact of each of these parameter in the output of interest.
+Wind turbines are a complex multi-physics systems whose performance and life span is highly dependent on manufacturing, meteorological and operational factors.
+It is therefore important to quantify the impact of these parameters on the turbine response.
+We perform comprehensive sensitivity analysis of aeroelastic models typically employed to compute turbine response.
+This paper also seeks to quantify sensitivities due to manufacturing tolerance in the turbine blades.
+We propose a non-uniform rational basis splines approach to parametrize the geometry of the blade and further utilize this for quantifying geometric sensitivities at different radial locations.
+[Some important results]
 \end{abstract}
@@
 In particular, BEM models are employed to predict turbine response such as the structural loads and power outputs.
-A number of parameters describing meteorological and operational conditions as well as manufacturing specifications are needed as input to simulate BEM models.
+A number of parameters describing meteorological and operational conditions as well as manufacturing specifications are needed as inputs for BEM simulations.
 In this work, we seek to quantify the sensitivities of these input parameters for different turbine responses.
+In the past a number of sensitivity studies have been performed to understand the influence of input parameters on different turbine response, for e.g.
+\cite{moriarty2002effect, eggers2003wind, McKay2014, dykes2014sensitivity, Robertson2018}.
+Many studies have confirmed that the wind parameters especially the wind speed and wind speed standard deviation have the most influence on aerodynamical performance of turbines.
+For example, in the of an upstream turbine, the wind speed is most sensitive to power production.
+Furthermore, wind speed in combination with wind speed standard deviation has a large influence on the power production of turbines operating in the wake of an upstream turbine.
+Among operational factors, the rotor RPM is the most sensitive to power production.
+[More on parameter sensitivities on structural loads]
+
+In this work, we also study the effect of manufacturing tolerances on the turbine response.
+Usually, there are some discrepancy between the  manufactured and nominally prescribed design of the turbine blade leading to a suboptimal performance.
+For BEM models, the turbine shape is described as a series of airfoils along the span of the blade where each airfoil shape is computed using the three quantities: chord length, thickness and twist.
+We perturb these three quantities to obtain a perturbed turbine blade and analyze with sections that have more influence on output quantities.
+
+To compute parameter sensitivities we use the Sobol expansion approach, which decomposes the total variance of the quantity of interest into contributions from individual parameters and their combinations.
+To perform Sobol analysis, we use Matlab-based general purpose uncertainty quantification toolbox \texttt{UQLab} \cite{uqlab}.
+\texttt{UQLab}'s modular structure allows for easy integration with available BEM codes.
+As our test wind turbine, we use the 2MW NM80 turbine (with an 80m rotor) from the DANAERO project \cite{DANAERO}.
 \section{Aeroelastic models}
@@
 \section{Global sensitivity analysis}
-The main objective of sensitivity analysis is to quantify the relative significance of individual inputs (or in combination) and how variations in input values affects the output of interest.
+The objective of sensitivity analysis is to quantify the relative significance of individual inputs (or in combination) and how variations in input values affects the output of interest.
 In engineering, sensitivity analysis can be employed for a number of reasons: to determine the stability and robustness of a computational model with respect to input parameters, for simplification of stochastic models by fixing the insensitive parameters, and to guide data acquisition campaigns and experimental design to refine the data on sensitive parameters.
-
-Sensitivity analysis techniques can be classified as local and global methods, see \cite{ralphSmith}.
-In local sensitivity analysis, individual parameters are perturbed around their nominal values --- allowing for the description of output variability only in a small neighbourhood of nominal input values.
+Sensitivity analysis techniques can be classified as local and global methods, see \cite{RSmith}.
+In local sensitivity analysis, individual parameters are perturbed around their nominal values allowing for the description of output variability only in a small neighbourhood of nominal input values.
 Although, local approaches are widely employed due to their ease of implementation and low computational cost, they are unable to quantify global behavior of nonlinearly parametrized models such as aeroelastic models.
@@
 \subsection{Sobol analysis}
-In this work, we employ variance-based Sobol decomposition to perform global sensitivity analysis.
-One of the advantages of this method is that the relative importance of input parameters is quantified on a scale of $[0,1]$ known as \emph{Sobol indices}.
-The main idea of Sobol analysis is to decompose the total variance of the output in term of contributions from individual parameters and their contribution.
+We employ variance-based Sobol decomposition to perform global sensitivity analysis.
+This approach allows for quantification of  the relative importance of the input parameters on a scale of $[0,1]$ known as \emph{Sobol indices}.
+
+The main idea of Sobol analysis is to express the total variance of the output in term of contributions from individual parameters and their combinations.
 For the ease of exposition, let us consider a nonlinear model:
@@
 \subsection{Geometric uncertainty}
-We use Non-Uniform Rational Basis Splines (NURBS) \cite{nurbs_book} to perturb the geometrical parameters of the turbine blade, such as the reference chord and twist curves.
+We use Non-Uniform Rational Basis Splines (NURBS) \cite{rogers2000} to perturb the geometrical parameters of the turbine blade, such as the reference chord and twist curves.
 The main advantage of using NURBS is that it provides great flexibility to approximate a large variety of curves with a limited number of control points.
@@
 \section*{Acknowledgements}
+\bibliography{\myreferences}
 \end{document}

### 2019-10-17 Paper/article.tex
improved postProcessing routine

@@
 
+[Sensitivity on the polars as novel element]
+
+[Mention this study is a part of IEA Task29 project]
+
 In this work, we also study the effect of manufacturing tolerances on the turbine response.
@@
 \texttt{UQLab}'s modular structure allows for easy integration with available BEM codes.
-As our test wind turbine, we use the 2MW NM80 turbine (with an 80m rotor) from the DANAERO project \cite{DANAERO}.
+For investigation the aeroelastic model of the  2MW NM80 turbine (with an 80m rotor) from the DANAERO project \cite{DANAERO} is utilized.
+
 \section{Aeroelastic models}

### 2019-10-17 sensitivity_analysis/cases/aero_module/turbineData.m

@@ -111,7 +111,6 @@ Input.Marginals(counter).Type = 'Gaussian';
 Input.Marginals(counter).Parameters = [PITCHANGLE, PITCHANGLE_Std];
 Input.Marginals(counter).Bounds = [PITCHANGLE_LB PITCHANGLE_UB];
 
-
 %% =======================RPM====================
 % Truncated Gaussian
 RPM_Std = 1;  % Standard deviation

### 2019-11-22 Torque_Abstract/main.tex
initial commit for Torque abstract

\documentclass[11pt]{article}
\usepackage[utf8]{inputenc}
\usepackage[parfill]{parskip}  % no indents
\usepackage{geometry}
 \geometry{
 a4paper,
 left=25mm,
 top=40mm,
 right=25mm,
 bottom=27mm
 }

\usepackage{natbib}  % references
\usepackage{graphicx}  % plots
\usepackage{scrextend}  % indent authors
\usepackage{amsmath}  % math typesetting

% typeset in sans serif to match the prompt
\usepackage{lmodern}
\renewcommand{\familydefault}{\sfdefault}
% correct the section heading size
\usepackage{titlesec}
\titleformat*{\section}{\normalsize\bfseries}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{document}
\raggedright  % don't stretch to right edge

% =============================================
\vspace*{80pt}
{\LARGE \textbf{Global sensitivity analysis of model parameters in aeroelastic wind-turbine models}}
\vspace*{28pt}

\begin{addmargin}[2.5cm]{0em}% 1em left, 2em right
\textbf{P. Kumar$^{\boldsymbol{\mathsf{1}}}$, B. Sanderse$^{\boldsymbol{\mathsf{1}}}$}

$^{\mathsf{1}}$ Centrum Wiskunde \& Informatica, Science Park 123, Amsterdam.

%$^{\mathsf{2}}$Address 2.
%\vspace{12pt}

Author contact email: b.sanderse@cwi.nl

Keywords: aero-elastic turbine model, sensitivity analysis
\end{addmargin}

% =============================================
\section{Introduction}
Aeroelastic models such as the Blade Element Momentum (BEM) models \cite{HandBook} continue to play a critical role in the design, development and optimization of modern wind turbines. In particular, BEM models are employed to predict turbine response such as the structural loads and power outputs.

A relatively large number of inputs are needed to perform BEM simulations, e.g. meteorological and operational conditions, geometric information of the, . In this work, we seek to quantify the sensitivities of these input parameters for different turbine responses. In the past a number of sensitivity studies have been performed to understand the influence of input parameters on different turbine response, for e.g. \cite{moriarty2002effect, eggers2003wind, McKay2014, dykes2014sensitivity, Robertson2018}. Many studies have confirmed that the wind parameters especially the wind speed and wind speed standard deviation have the most influence on aerodynamical performance of turbines. For example, in the of an upstream turbine, the wind speed is most sensitive to power production. Furthermore, wind speed in combination with wind speed standard deviation has a large influence on the power production of turbines operating in the wake of an upstream turbine. Among operational factors, the rotor RPM is the most sensitive to power production. [More on parameter sensitivities on structural loads]

Important work from Murcia et al. \cite{Murcia2018}.

[Sensitivity on the polars as novel element]

[Mention this study is a part of IEA Task29 project and WindTrue.

In this work, we also study the effect of manufacturing tolerances on the turbine response. Usually, there are some discrepancy between the  manufactured and nominally prescribed design of the turbine blade leading to a suboptimal performance. For BEM models, the turbine shape is described as a series of airfoils along the span of the blade where each airfoil shape is computed using the three quantities: chord length, thickness and twist. We perturb these three quantities to obtain a perturbed turbine blade and analyze with sections that have more influence on output quantities.

% =============================================
\section{Objectives}
The long-term objective of our study is to develop robust BEM models that give users an indication of the uncertainty associated with the predictions (loads, power, etc.) offered by the model. For this purpose, we will calibrate the model parameters present in BEM models. Examples of such model parameters are the time constant in dynamic stall models, wake correction factors, tip loss model parameters, and lift- and drag-polars. In order to limit the number of model parameters involved in the calibration process, \textit{the objective of the current study is to perform a global sensitivity study of the outputs of the BEM model towards the model parameters}.

% =============================================
\section{Methodology}
To compute parameter sensitivities we use a global sensitivity analysis based on the Sobol expansion approach, which decomposes the total variance of the quantity of interest (model output) into contributions from individual parameters and their combinations. To perform Sobol analysis, we use the uncertainty quantification toolbox \texttt{UQLab} \cite{uqlab}. \texttt{UQLab}'s modular structure allows for easy integration with available BEM codes. The aeroelastic code that we use is the \texttt{ECN Aero-Module} \cite{Boorsma2012} of the 2MW NM80 turbine from the DANAERO project \cite{Troldborg2013}.

Parameterization.

% =============================================
\section{Results}

% =============================================
\section{Conclusions}

% =============================================
\bibliographystyle{plain}
\bibliography{references}

\end{document}

### 2019-11-23 Paper/article.tex
added literature overview as appendix to paper

@@
 \documentclass[review]{elsarticle}
-\usepackage{lineno,hyperref}
-\modulolinenumbers[5]
+
 \journal{ }
+
+% packages
 \usepackage{microtype}
@@
 \usepackage[a4paper, total={6.2in, 9in}]{geometry}
-\usepackage[a4paper]{geometry}
 \renewcommand{\baselinestretch}{1.2} 
@@
 \usepackage{amsmath}
-\usepackage{tabularx,ragged2e,booktabs,caption}
-\renewcommand{\figurename}{{Figure}}
-\usepackage[labelfont=bf]{caption}
-\usepackage[font=footnotesize]{caption}
 \usepackage{amssymb,latexsym}
@@
 \usepackage{multirow}
-\usepackage[toc,page]{appendix}
+\usepackage{stix2}
+\usepackage[labelfont=bf,font=footnotesize]{caption}
+\usepackage{tabularx,ragged2e,booktabs}
+\usepackage{lineno}
+\modulolinenumbers[5]
+\usepackage[toc]{appendix}
+\usepackage{hyperref}
+
+%new commands
+\renewcommand{\figurename}{{Figure}}
 \newcolumntype{Y}{>{\centering\arraybackslash}X}
@@
 \numberwithin{equation}{section}
-\usepackage{tabularx,ragged2e,booktabs,caption}
 \newcommand{\footremember}[2]{%
@@
 }
-\usepackage[toc]{appendix}
+
 \DeclareMathAlphabet{\mathpzc}{OT1}{pzc}{m}{it}
@@
 \newcommand{\Lomd}{L^2(\Omega,\mathcal{D})}
-\newcommand{\myreferences}{./Bibliography/references}
+
+% bibliography
+\newcommand{\myreferences}{./Bibliography/references,./Bibliography/Mendeley_refs}
 \bibliographystyle{elsarticle-num}
+
 \numberwithin{equation}{section}
-\usepackage{tabularx,ragged2e,booktabs,caption}
 \graphicspath{{./images/}}
-\usepackage[toc]{appendix}
+
 \DeclareMathAlphabet{\mathpzc}{OT1}{pzc}{m}{it}
-\bibliographystyle{elsarticle-num}
-\usepackage{stix2}
-\usepackage{appendix}
-\usepackage[labelfont=bf]{caption}
+
 \pagestyle{myheadings}
+
 %%%%%%%%%%%%%%%%%%%%%%%
+\begin{document}
+
+\begin{frontmatter}
 \title{A global sensitivity analysis of wind turbine aeroelastic models}
-\begin{document}
-\begin{frontmatter}
+
 %\address[cwi]{CWI - Centrum Wiskunde \& Informatica, Amsterdam, The Netherlands}
@@
 %\author[cwi,aero]{Prashant~Kumar}
-\author[]{Prashant~Kumar}
-\ead{pkumar@cwi.nl}
+\author[]{Prashant~Kumar}\ead{pkumar@cwi.nl}
+
+\author{Benjamin~Sanderse}
+
 \begin{abstract}
@@
 \end{frontmatter}
+
 \linenumbers
+
 \section{Introduction}
@@
 
+Use of NURBS in wind-turbine blades (airfoils) optimization: \cite{Ribeiro2012,Bottasso2014}.
+
 \subsubsection{NURBS based perturbation}
@@
 \subsection{Wind velocity}
+
 \section{Description of test case}
+
 \section{Sensitivity analysis workflow}
+
 \section{Numerical experiment}
 \subsection{Interpretation of global sensitivity analysis result}
+
 \section{Conclusions}
+
 \section*{Acknowledgements}
 \bibliography{\myreferences}
+
+\appendix
+\section{Literature overview}
+List of BEM codes in \cite{Sayed2019,Vorpahl2013}.
+
+Uncertainties and corrections in BEM models \cite{Madsen2012}: complex inflow, e.g. sheared inflow.
+
+\cite{Eggers2003} study the effect of shear and turbulence on rotor loading, but not with a systematic (global) sensitivity analysis.
+
+\cite{McKay2014} perform a measurement data-only global sensitivity analysis using a neural network and a Fourier amplitude sensitivity test (a similar technique was used in \cite{Kusiak2010} for vibration analysis).
+The neural network is constructed based on measurement data.
+The output quantity of interest is the power production; several input parameters are used, including yaw angle, rotor speed, blade pitch angle, wind speed, ambient temperature, main bearing temperature, wind speed standard deviation and yaw angle standard deviation.
+
+\cite{Dykes2014} performed global sensitivity analysis of turbine costs with respect to key wind turbine configuration parameters including rotor diameter, rated power, hub height, and maximum tip speed.
+DAKOTA is used to calculate Sobol indices with Monte Carlo sampling.
+
+\cite{Rinker2016a} calculated the global sensitivity of wind turbine loads to wind inputs using polynomial response surfaces.
+She used Sobol indices, with quantity of interest the maximum blade root bending moment and as inputs four turbulence parameters: a reference mean wind speed, a reference turbulence intensity, the Kaimal length scale, and a parameter reflecting the nonstationarity.
+The turbine model studied is the WindPACT 5 MW reference model and the BEM model is FAST.
+
+\cite{Echeverria2017} employs a global sensitivity analysis to identify the design variables that affect wind turbine performance in order to reduce the number of variables in wind turbine design optimization.
+As input parameters, airfoil parameterization and chord and twist distributions are used; the outputs studied are annual energy production, maximum blade tip deflection, overall sound power level and blade mass.
+For example, Sobol indices are shown for the effect of chord and twist control points on the output quantities.
+
+\cite{Matthaus2017} performed uncertainty quantification including uncertainty in the aerodynamic properties; for example, the uncertainty in the lift coefficient at an airfoil section due to uncertain roughness (e.g.\ contamination) is achieved by interpolating between clean and fully rough state with a uniformly distributed random variable.
+Polynomial chaos and Kriging were used.
+See also \cite{Bortolotti2019}, in which the AVATAR turbine was analysed, using the Cp-Lambda aeroelastic model and DAKOTA as UQ toolbox.
+
+\cite{Murcia2018} similarly considered a global sensitivity analysis with Sobol indices computed by using sparse polynomial chaos expansion.
+The quantities of interest considered are the energy production and lifetime equivalent fatigue loads for the DTU 10 MW reference turbine.
+The uncertain input is given by the turbulent inflow field, with 4 parameters: mean hub height wind speed, std.
+dev.
+of hub height wind speed, shear exponent, yaw misalignment.
+The dependency between these parameters (as described by the Normal Turbulence Model) is taken into account by using a Rosenblatt transformation that transforms the dependent variables to a set of independent ones.
+The Chaospy package is used for the analysis.
+Seven different model outputs are considered: power, thrust, and several damage equivalent fatigue loads (EFL), computed using a rainflow counting algorithm.
+The sensitivity analysis shows that the turbulent inflow realization has a bigger impact on the total distribution of equivalent fatigue loads than the shear coefficient or yaw misalignment.
+
+\cite{Sayed2019} argue that BEM models are accurate mainly for small turbines but that for large turbines, when the tip deformations exceed 10\% of the blade radius, it is questionable if they can still be applied.
+They compare the results of engineering models to CFD-based aeroelastic simulations.
+It is shown that the 2D polars commonly used in BEM require 3D corrections, and that including more polars improves the results.
+`Turbine power and thrust predicted from the flexible rotor were reduced compared to the rigid rotor employing a BEM-based model.
+A reason behind the increase in power from the CFD-based model is the spanwise force component.
+It is not included in the BEM-based simulations as it is one of the main assumptions of its theory.
+This assumption could be used for small wind turbine simulations where the edgewise deformations were insignificant.
+But for such large wind turbines, the spanwise force distribution over the blade radius must be taken into account in the calculation of the power output due to the large edgewise deformations.
+To increase the accuracy of the engineering model, it is recommended to increase the number of polars calculated by 2D CFD simulations at different sections along the blade.
+This is because of the significant difference in the Reynolds number for the large blade.'
+
+\cite{Velarde2019} perform a global sensitivity analysis of fatigue loads with respect to structural, geotechnical and metocean parameters for a 5MW offshore wind turbine installed on a gravity based foundation.
+Linear regression of Monte Carlo simulations and Morris screening are performed for three design load cases.
+
+\cite{Robertson2018} assesses the sensitivity of different wind parameters on the loads of a wind turbine.
+They argue that the most common parameters considered normally in literature are the turbulence intensity variability, and the shear exponent, or wind profile, both having important influence on the turbine response.
+In the paper an assessment of which wind characteristics influence wind turbine structural loads is given.
+The correlated dependency between the individual parameters is not considered; rather, they focus on assessing the sensitivity of each parameter individually.
+An elementary-effects screening method (Morris screening) is used as global sensitivity method.
+The NREL 5MW turbine and FAST are used.
+
 \end{document}

### 2019-11-23 Torque_Abstract/main.tex
added literature overview as appendix to paper

@@
 
-A relatively large number of inputs are needed to perform BEM simulations, e.g.
-meteorological and operational conditions, geometric information of the, .
+A relatively large number of inputs are needed to perform BEM simulations, e.g. meteorological and operational conditions, geometric information of the ...
+
+`These tools use simplified methods such as BEM to find the unsteady aerodynamic loads and 1D structural models to determine the deformations.
+They are computationally cheap, but they are based on different corrections to account for the unsteadiness and the 3D effects.' \cite{Sayed2019}
+
 In this work, we seek to quantify the sensitivities of these input parameters for different turbine responses.

### 2019-11-25 Paper/article.tex
added more references to appendix in paper

@@
 \section*{Acknowledgements}
+
+\newpage
+
+\section{References}
 \bibliography{\myreferences}
 
+\newpage
+
 \appendix
-\section{Literature overview}
+\section{Literature overview of global sensitivity analysis in wind-turbine models}
 List of BEM codes in \cite{Sayed2019,Vorpahl2013}.
@@
 
+\cite{Fluck2018} develop an \textit{intrusive} stochastic BEM method: `a stochastic solution for the unsteady aerodynamic loads based on a projection of the unsteady Blade Element Momentum (BEM) equations onto a stochastic space spanned by chaos exponentials'.
+
+\cite{Hubler2017} used global sensitivity analysis (as part of a four-step approach) to reduce the parameter dimension for structural behaviour computations using FAST as aeroelastic code.
+The methods of expert knowledge, one-at-a-time variation, and regression based on Monte Carlo sampling are used as pre-processing steps to reduce the computational expense of global sensitivity analysis.
+
 \end{document}

### 2019-11-25 Torque_Abstract/main.tex
rewrote introduction, objective, methodology

--- 2019-11-22 Torque_Abstract/main.tex
@@
 \vspace*{80pt}
-{\LARGE \textbf{Global sensitivity analysis of model parameters in aeroelastic wind-turbine models}}
+{\LARGE \textbf{Global sensitivity analysis of model parameters in aeroelastic wind-turbine codes}}
 \vspace*{28pt}
@@
 
-Keywords: aero-elastic turbine model, sensitivity analysis
+Keywords: aero-elastic turbine model, BEM, sensitivity analysis, model uncertainty
 \end{addmargin}
@@
 \section{Introduction}
-Aeroelastic models such as the Blade Element Momentum (BEM) models \cite{HandBook} continue to play a critical role in the design, development and optimization of modern wind turbines.
-In particular, BEM models are employed to predict turbine response such as the structural loads and power outputs.
+Aeroelastic models such as the Blade Element Momentum (BEM) method \cite{HandBook} continue to play a critical role in the design, development and optimization of modern wind turbines.
+The accuracy of BEM predictions is affected to by uncertainties and inaccuracies, for example in the external conditions (wind parameters), in the turbine specification (geometric parameters), and in the BEM equations itself (model parameters).
+For the purpose of design, uncertainty quantification and optimization, is it crucial to limit the number of parameters, typically by performing sensitivity studies.
+Most studies focused on the effect of uncertainties in the external conditions and in the geometry, see e.g.\ \cite{Echeverria2017,Matthaus2017,Murcia2018,Robertson2018}.
 
-A relatively large number of inputs are needed to perform BEM simulations, e.g.
-meteorological and operational conditions, geometric information of the, .
-In this work, we seek to quantify the sensitivities of these input parameters for different turbine responses.
-In the past a number of sensitivity studies have been performed to understand the influence of input parameters on different turbine response, for e.g.
-\cite{moriarty2002effect, eggers2003wind, McKay2014, dykes2014sensitivity, Robertson2018}.
+\section{Objectives}
+The long-term objective of this study is to develop calibrated BEM models that give users an indication of the uncertainty associated with the predictions (loads, power, etc.) originating not only from external conditions and geometry, but also from the model formulation itself.
+For this purpose, we will calibrate the model parameters present in BEM models.
+Examples of such model parameters are the time constant in dynamic stall models, the wake correction factor, the tip loss model parameter, and the lift- and drag-polars \cite{Sayed2019}.
+In order to limit the number of model parameters involved in the calibration process, \textit{the objective of the current study is to perform a global sensitivity study of the outputs of the BEM model towards both geometric and model uncertainties}.
+
+%In the past a number of sensitivity studies have been performed to understand the influence of input parameters on different turbine response, e.g.\ \cite{moriarty2002effect, eggers2003wind, McKay2014, dykes2014sensitivity, Robertson2018}.
 Many studies have confirmed that the wind parameters especially the wind speed and wind speed standard deviation have the most influence on aerodynamical performance of turbines.
@@
 
-Important work from Murcia et al. \cite{Murcia2018}.
+% In particular, BEM models are employed to predict turbine response such as the structural loads and power outputs.
 
-[Sensitivity on the polars as novel element]
+%`These tools use simplified methods such as BEM to find the unsteady aerodynamic loads and 1D structural models to determine the deformations.
+They are computationally cheap, but they are based on different corrections to account for the unsteadiness and the 3D effects.' \cite{Sayed2019}
 
-[Mention this study is a part of IEA Task29 project and WindTrue.
+%[Sensitivity on the polars as novel element]
 
-In this work, we also study the effect of manufacturing tolerances on the turbine response.
+%[Mention this study is a part of IEA Task29 project and WindTrue.
+
+%In this work, we also study the effect of manufacturing tolerances on the turbine response.
 Usually, there are some discrepancy between the  manufactured and nominally prescribed design of the turbine blade leading to a suboptimal performance.
@@
 % =============================================
-\section{Objectives}
-The long-term objective of our study is to develop robust BEM models that give users an indication of the uncertainty associated with the predictions (loads, power, etc.) offered by the model.
-For this purpose, we will calibrate the model parameters present in BEM models.
-Examples of such model parameters are the time constant in dynamic stall models, wake correction factors, tip loss model parameters, and lift- and drag-polars.
-In order to limit the number of model parameters involved in the calibration process, \textit{the objective of the current study is to perform a global sensitivity study of the outputs of the BEM model towards the model parameters}.
 
@@
 \section{Methodology}
-To compute parameter sensitivities we use a global sensitivity analysis based on the Sobol expansion approach, which decomposes the total variance of the quantity of interest (model output) into contributions from individual parameters and their combinations.
-To perform Sobol analysis, we use the uncertainty quantification toolbox \texttt{UQLab} \cite{uqlab}.
+To compute parameter sensitivities we use a global sensitivity analysis based on the Sobol expansion approach, which decomposes the total variance of the quantity of interest (model output) into contributions from individual parameters and their combinations, similar to \cite{Echeverria2017,Murcia2018,Rinker2016}.
+We employ the uncertainty quantification toolbox \texttt{UQLab} \cite{uqlab}, which computes the Sobol indices from a sparse polynomial chaos expansion.
 \texttt{UQLab}'s modular structure allows for easy integration with available BEM codes.
-The aeroelastic code that we use is the \texttt{ECN Aero-Module} \cite{Boorsma2012} of the 2MW NM80 turbine from the DANAERO project \cite{Troldborg2013}.
 
-Parameterization.
+The geometric uncertainties currently considered are chord and twist distribution, whereas the model uncertainty enters via uncertainty in lift- and drag-polars.
+In order to express the chord, twist, lift and drag distributions along the turbine blade, an efficient parameterization is needed that gives flexible control over the prescribed uncertainty while limit the number of required parameters.
+We have chosen to use NURBS curves for this purpose, similar to \cite{Echeverria2017}.
 
@@
 \section{Results}
+The aeroelastic code that we use is the \texttt{ECN Aero-Module} \cite{Boorsma2012} and the turbine is the 2MW NM80 turbine from the DANAERO project \cite{Troldborg2013}.
 
@@
 \bibliographystyle{plain}
-\bibliography{references}
+\bibliography{references,Mendeley_refs}
 

### 2019-11-29 Torque_Abstract/main.tex
SA for power torque abstract run number 2. Also some clean of was performed

@@
  }
-
 \usepackage{natbib}  % references
@@
 \usepackage{amsmath}  % math typesetting
+\usepackage{graphicx}
+\usepackage{subcaption}
 
@@
 In order to express the chord, twist, lift and drag distributions along the turbine blade, an efficient parameterization is needed that gives flexible control over the prescribed uncertainty while limit the number of required parameters.
-We have chosen to use NURBS curves for this purpose, similar to \cite{Echeverria2017}.
+We have chosen to use Non-Uniform Rational Basis Splines (NURBS) for this purpose, similar to \cite{Echeverria2017}.
 
@@
 \section{Results}
-The aeroelastic code that we use is the \texttt{ECN Aero-Module} \cite{Boorsma2012} and the turbine is the 2MW NM80 turbine from the DANAERO project \cite{Troldborg2013}.
+The aeroelastic code that we use is the \texttt{ECN Aero-Module} \cite{Boorsma2012} and the turbine is the 2MW NM80 turbine from the DANAERO project \cite{Troldborg2013} with a blade radius of 38.8m.
+The data for lift ($Cl$) and drag ($Cd$) polars are available at four locations along the blade radius at 11.87m, 17.82m, 28.97m and 35.53m.
+The reference value of polars are obtained from wind-tunnel experiment with 3D corrections.
+Random samples of chord, twist, lift- and drag-polars are obtained by perturbing the control points with a uniformly distributed random variable.
+For chord and twist, we independently perturb control points at different locations as shown in Fig.
+\ref{fig:samples} (a) - (b).
+For $Cl$ and $Cd$, all control points are perturbed using a same random number.
 
+In Fig. \ref{fig:Sobol}, we show the total order Sobol indices as a measure of sensitivity for different geometric and model parameters on power production. The control points are named
+
+\begin{figure}[h!]
+  \centering
+  \begin{subfigure}[b]{0.4\linewidth}
+    \includegraphics[width=\linewidth]{Chord_samples_torque.eps}
+   % \caption{Coffee.}
+  \end{subfigure}
+  \begin{subfigure}[b]{0.4\linewidth}
+    \includegraphics[width=\linewidth]{Twist_samples_torque.eps}
+   % \caption{More coffee.}
+  \end{subfigure}
+
+  \begin{subfigure}[b]{0.4\linewidth}
+    \includegraphics[width=\linewidth]{Cl_samples_torque.eps}
+    %\caption{Coffee.}
+  \end{subfigure}
+  \begin{subfigure}[b]{0.4\linewidth}
+    \includegraphics[width=\linewidth]{Cd_samples_torque}
+    %\caption{More coffee.}
+  \end{subfigure}
+  \caption{Random realization of twist, chord, lift- and drag-polars.}
+  \label{fig:samples}
+\end{figure}
+\begin{figure}[h!]
+  \centering
+    \includegraphics[width=\linewidth]{SA_Power_Chord_Twist_CD_CL_torque.eps}
+  \caption{Total order Sobol indices for the sensitivity analysis of geometrical and model parameters.}
+   \label{fig:Sobol}
+\end{figure}
 % =============================================
 \section{Conclusions}
-
+Global sensitivity analysis is a powerful tool to identify important input parameters associated with aeroelastic models.
+Sobol indices computed using sparse polynomial expansion is very efficient and can be used to analyze large number of input parameters.
+In the full paper, we will present a sensitivity assessment with a more model parameters.
 % =============================================

### 2019-11-29 Paper/article.tex
updated section 1 in Torque paper
- moved literature overview into separate file so it is easily commented out

--- 2019-10-17 Paper/article.tex
@@
 \newcommand{\Lomd}{L^2(\Omega,\mathcal{D})}
-\newcommand{\myreferences}{./Bibliography/references}
+
+% bibliography
+\newcommand{\myreferences}{./Bibliography/references,./Bibliography/Mendeley_refs}
 \bibliographystyle{elsarticle-num}
+
 \numberwithin{equation}{section}
-\usepackage{tabularx,ragged2e,booktabs,caption}
 \graphicspath{{./images/}}
-\usepackage[toc]{appendix}
+
 \DeclareMathAlphabet{\mathpzc}{OT1}{pzc}{m}{it}
-\bibliographystyle{elsarticle-num}
-\usepackage{stix2}
-\usepackage{appendix}
-\usepackage[labelfont=bf]{caption}
+
 \pagestyle{myheadings}
+
 %%%%%%%%%%%%%%%%%%%%%%%
-\title{A global sensitivity analysis of wind turbine aeroelastic models}
 \begin{document}
+
 \begin{frontmatter}
-%\address[cwi]{CWI - Centrum Wiskunde \& Informatica, Amsterdam, The Netherlands}
+\title{A global sensitivity analysis of model uncertainty in aeroelastic wind turbine models}
+
+\address[cwi]{Centrum Wiskunde \& Informatica (CWI), Amsterdam, The Netherlands}
 %\address[aero]{Faculty of Aerospace Engineering, Delft University of Technology, Delft, The Netherlands.}
 %\author[cwi,aero]{Prashant~Kumar}
-\author[]{Prashant~Kumar}
-\ead{pkumar@cwi.nl}
+\author[]{Prashant~Kumar}\ead{pkumar@cwi.nl}
+
+\author{Benjamin~Sanderse}
+
 \begin{abstract}
@@
 \begin{keyword}
-Global sensitivity analysis, UQ, Sobol indices, aeroelastic models
+BEM, global sensitivity analysis, model uncertainty, Sobol indices
 \end{keyword}
 \end{frontmatter}
+
 \linenumbers
+
 \section{Introduction}
 Aeroelastic models such as the Blade Element Momentum (BEM) models \cite{HandBook} play a critical role in the design, development and optimization of modern wind turbines.
-In particular, BEM models are employed to predict turbine response such as the structural loads and power outputs.
-A number of parameters describing meteorological and operational conditions as well as manufacturing specifications are needed as inputs for BEM simulations.
+A large number of BEM models have been developed to predict turbine responses such as the structural loads and power output \cite{Vorpahl2013}.
+
+As a consequence of the strong model assumptions at the basis of BEM theory, the results from BEM codes can be subject to significant inaccuracies or uncertainties.
+For example, the effect of sheared inflow \cite{Madsen2012} is not naturally accounted for in the theory, and needs to be incorporated via correction terms.
+Other major model uncertainties in BEM models are for example the time constant in dynamic stall models, the wake correction factor, the tip loss model parameter, and the lift- and drag-polars used to compute local aerodynamic forces.
+Especially for increasing turbine sizes, these model parameters  are foreseen to be not sufficiently accurate \cite{Sayed2019}.
+In other words, the uncertainty associated to the output of currently employed BEM models is rather large.
+% iIn order to arrive at robust BEM models that include an uncertainty estimate associated
+
+Recently, several papers have addressed the uncertainty in BEM model output by performing forward uncertainty propagation or sensitivity studies, e.g.\ \cite{Bos2019a,Echeverria2017,Matthaus2017,Murcia2018,Robertson2018}.
+In these studies, the focus in mainly on uncertainties in the external conditions (wind parameters) and/or uncertainty in the turbine specification (geometric parameters).
+Apart from understanding how uncertainty in the output is related to the different uncertain inputs, such sensitivity studies are very useful to reduce the number of parameters as needed for example in design optimization \cite{Echeverria2017}.
+However, the uncertainty in BEM model output as caused by uncertainty in the model formulation itself, e.g.\ through the values chosen for model parameters, has been given little attention (exceptions being the effect of aerodynamic properties studied in \cite{Matthaus2017,Bortolotti2019}).
+
+The goal of this work is to perform a systematic assessment of the uncertainty in model parameters in BEM models.
+We approach this by performing a global sensitivity analysis based on the Sobol expansion approach, which decomposes the total variance of the quantity of interest (model output) into contributions from individual parameters and their combinations, similar to \cite{Echeverria2017,Murcia2018,Rinker2016}.
+We will employ the uncertainty quantification toolbox \texttt{UQLab} \cite{uqlab}, which computes the Sobol indices based on a sparse polynomial chaos expansion.
+
+Clearly, the main novelty of the work lies in the study of the effect of model uncertainties rather than external conditions or geometric uncertainties.
+The current sensitivity analysis is part of the so-called WindTrue project, in which the long-term goal is the development of calibrated BEM models, that possess a quantified level of uncertainty.
+Anticipating on this goal, we are using the NM80 wind turbine model from the Tjaereborg wind farm, for which extensive measurement data is available through the DanAero project \cite{Troldborg2013} and IEA Task 29.
+The measurement data will enable the calibration step, and our sensitivity analysis will make clear which of the model parameters should be used in the calibration step.
+
+This paper is structured as follows.
+First, in section \ref{sec:model_description} we give a short description of the BEM model and associated uncertainties.
+In section \ref{sec:parameterization} the parameterization of the uncertainties is described, and in section \ref{sec:GSA} the global sensitivity analysis methodology.
+Section \ref{sec:results} discusses the results of the sensitivity analysis for the NM80 turbine test case, and conclusions follow in section \ref{sec:conclusions}.
+
+%A number of parameters describing meteorological and operational conditions as well as manufacturing specifications are needed as inputs for BEM simulations.
 In this work, we seek to quantify the sensitivities of these input parameters for different turbine responses.
@@
 \cite{moriarty2002effect, eggers2003wind, McKay2014, dykes2014sensitivity, Robertson2018}.
-Many studies have confirmed that the wind parameters especially the wind speed and wind speed standard deviation have the most influence on aerodynamical performance of turbines.
-For example, in the of an upstream turbine, the wind speed is most sensitive to power production.
+Many studies have confirmed that the wind parameters especially the wind speed and wind speed standard deviation have the most influence on aerodynamic performance of turbines.
+For example, in the case of an upstream turbine, the wind speed is most sensitive to power production.
 Furthermore, wind speed in combination with wind speed standard deviation has a large influence on the power production of turbines operating in the wake of an upstream turbine.
@@
 [More on parameter sensitivities on structural loads]
-
-[Sensitivity on the polars as novel element]
-
-[Mention this study is a part of IEA Task29 project]
-
-In this work, we also study the effect of manufacturing tolerances on the turbine response.
+%
+%[Sensitivity on the polars as novel element]
+%
+%[Mention this study is a part of IEA Task29 project]
+%
+%In this work, we also study the effect of manufacturing tolerances on the turbine response.
 Usually, there are some discrepancy between the  manufactured and nominally prescribed design of the turbine blade leading to a suboptimal performance.
@@
 We perturb these three quantities to obtain a perturbed turbine blade and analyze with sections that have more influence on output quantities.
-
-To compute parameter sensitivities we use the Sobol expansion approach, which decomposes the total variance of the quantity of interest into contributions from individual parameters and their combinations.
+%
+%To compute parameter sensitivities we use the Sobol expansion approach, which decomposes the total variance of the quantity of interest into contributions from individual parameters and their combinations.
 To perform Sobol analysis, we use Matlab-based general purpose uncertainty quantification toolbox \texttt{UQLab} \cite{uqlab}.
@@
 
-\section{Aeroelastic models}
+\section{Aeroelastic models}\label{sec:model_description}
 \subsection{BEM models}
 \subsection{Input parameters of aeroelastic models}
-\section{Global sensitivity analysis}
-The objective of sensitivity analysis is to quantify the relative significance of individual inputs (or in combination) and how variations in input values affects the output of interest.
-In engineering, sensitivity analysis can be employed for a number of reasons: to determine the stability and robustness of a computational model with respect to input parameters, for simplification of stochastic models by fixing the insensitive parameters, and to guide data acquisition campaigns and experimental design to refine the data on sensitive parameters.
-Sensitivity analysis techniques can be classified as local and global methods, see \cite{RSmith}.
-In local sensitivity analysis, individual parameters are perturbed around their nominal values allowing for the description of output variability only in a small neighbourhood of nominal input values.
-Although, local approaches are widely employed due to their ease of implementation and low computational cost, they are unable to quantify global behavior of nonlinearly parametrized models such as aeroelastic models.
-Global sensitivity approaches, on the other hand, consider the entire range of input values to compute output sensitivities.
-Therefore, global sensitivity analysis is more suitable for aeroelastic models considered in this work.
 
-\subsection{Sobol analysis}
-We employ variance-based Sobol decomposition to perform global sensitivity analysis.
-This approach allows for quantification of  the relative importance of the input parameters on a scale of $[0,1]$ known as \emph{Sobol indices}.
+\section{Parametrization of uncertain inputs in the BEM model}\label{sec:parameterization}
 
-The main idea of Sobol analysis is to express the total variance of the output in term of contributions from individual parameters and their combinations.
-For the ease of exposition, let us consider a nonlinear model:
-\begin{equation}\label{nonlinear_model}
-Y = f(\mathbf{Z}),
-\end{equation}
-where $\mathbf{Z} = [z_1, z_2, ..., z_M]\in \mathcal{D}_{\mathbf{Z}}\in \mathbb{R}^M$ is the input vector.
-For simplicity, we assume input parameters are uniformly distributed i.e.
-$z_i \sim \mathcal{U}(0,1)$ and the support of input set is $\mathcal{D}_\mathbf{Z}  =  [0,1]^M$ where $M$ is the total number input parameters.
-The Sobol decomposition is defined as:
-\begin{equation}\label{sobol_decomp}
-f(z_1, z_2, ..., z_M) = f_0+\sum_{i=1}^M f_i(z_i) + \sum_{1\leq<i<j\leq M} f_{ij}(z_i,z_j) + ... + f_{1, 2, ..., M}(z_1, z_2, ..., z_M),
-\end{equation}
-The above decomposition is only valid for independent input parameters. 
-\subsubsection{PCE Ordinary Least Square (PCE\_OLS)}
-\subsubsection{PCE Least Angle Regression (PCE\_LAR)}
-\section{Parametrization of uncertain inputs in the BEM model}
 \subsection{Geometric uncertainty}
@@
 Further, the set of control points and knots can be directly manipulated to control the smoothness and curvature.
+
+Use of NURBS in wind-turbine blades (airfoils) optimization: \cite{Ribeiro2012,Bottasso2014}.
 
@@
 \subsection{Wind velocity}
-\section{Description of test case}
-\section{Sensitivity analysis workflow}
-\section{Numerical experiment}
-\subsection{Interpretation of global sensitivity analysis result}
-\section{Conclusions}
+
+\section{Global sensitivity analysis}\label{sec:GSA}
+The objective of sensitivity analysis is to quantify the relative significance of individual inputs (or in combination) and how variations in input values affects the output of interest.
+In engineering, sensitivity analysis can be employed for a number of reasons: to determine the stability and robustness of a computational model with respect to input parameters, for simplification of stochastic models by fixing the insensitive parameters, and to guide data acquisition campaigns and experimental design to refine the data on sensitive parameters.
+Sensitivity analysis techniques can be classified as local and global methods, see \cite{RSmith}.
+In local sensitivity analysis, individual parameters are perturbed around their nominal values allowing for the description of output variability only in a small neighbourhood of nominal input values.
+Although, local approaches are widely employed due to their ease of implementation and low computational cost, they are unable to quantify global behavior of nonlinearly parametrized models such as aeroelastic models.
+Global sensitivity approaches, on the other hand, consider the entire range of input values to compute output sensitivities.
+Therefore, global sensitivity analysis is more suitable for aeroelastic models considered in this work.
+
+\subsection{Sobol analysis}
+We employ variance-based Sobol decomposition to perform global sensitivity analysis.
+This approach allows for quantification of  the relative importance of the input parameters on a scale of $[0,1]$ known as \emph{Sobol indices}.
+
+The main idea of Sobol analysis is to express the total variance of the output in term of contributions from individual parameters and their combinations.
+For the ease of exposition, let us consider a nonlinear model:
+\begin{equation}\label{nonlinear_model}
+Y = f(\mathbf{Z}),
+\end{equation}
+where $\mathbf{Z} = [z_1, z_2, ..., z_M]\in \mathcal{D}_{\mathbf{Z}}\in \mathbb{R}^M$ is the input vector.
+For simplicity, we assume input parameters are uniformly distributed i.e.
+$z_i \sim \mathcal{U}(0,1)$ and the support of input set is $\mathcal{D}_\mathbf{Z}  =  [0,1]^M$ where $M$ is the total number input parameters.
+The Sobol decomposition is defined as:
+\begin{equation}\label{sobol_decomp}
+f(z_1, z_2, ..., z_M) = f_0+\sum_{i=1}^M f_i(z_i) + \sum_{1\leq<i<j\leq M} f_{ij}(z_i,z_j) + ... + f_{1, 2, ..., M}(z_1, z_2, ..., z_M),
+\end{equation}
+The above decomposition is only valid for independent input parameters. 
+\subsubsection{PCE Ordinary Least Square (PCE\_OLS)}
+\subsubsection{PCE Least Angle Regression (PCE\_LAR)}
+
+%\section{Description of test case}
+
+%\section{Sensitivity analysis workflow}
+
+\section{Results}\label{sec:results}
+
+\subsection{NM80}
+\subsubsection{Verification of GSA method}
+Convergence study of UQLab with increasing number of samples for Quad, OLS, LARS.
+
+\subsection{AVATAR}
+
+\section{Conclusions}\label{sec:conclusions}
+
 \section*{Acknowledgements}
+
+\newpage
+
+\section{References}
 \bibliography{\myreferences}
+
+\newpage
+
+\appendix
+\input{literature_overview}
+
 \end{document}

### 2019-11-29 Paper/literature_overview.tex
updated section 1 in Torque paper
- moved literature overview into separate file so it is easily commented out

\section{Literature overview of global sensitivity analysis in wind-turbine models}
List of BEM codes in \cite{Sayed2019,Vorpahl2013}.

Uncertainties and corrections in BEM models \cite{Madsen2012}: complex inflow, e.g. sheared inflow.

\cite{Eggers2003} study the effect of shear and turbulence on rotor loading, but not with a systematic (global) sensitivity analysis.

\cite{McKay2014} perform a measurement data-only global sensitivity analysis using a neural network and a Fourier amplitude sensitivity test (a similar technique was used in \cite{Kusiak2010} for vibration analysis). The neural network is constructed based on measurement data. The output quantity of interest is the power production; several input parameters are used, including yaw angle, rotor speed, blade pitch angle, wind speed, ambient temperature, main bearing temperature, wind speed standard deviation and yaw angle standard deviation.

\cite{Dykes2014} performed global sensitivity analysis of turbine costs with respect to key wind turbine configuration parameters including rotor diameter, rated power, hub height, and maximum tip speed. DAKOTA is used to calculate Sobol indices with Monte Carlo sampling.

\cite{Rinker2016a} calculated the global sensitivity of wind turbine loads to wind inputs using polynomial response surfaces. She used Sobol indices, with quantity of interest the maximum blade root bending moment and as inputs four turbulence parameters: a reference mean wind speed, a reference turbulence intensity, the Kaimal length scale, and a parameter reflecting the nonstationarity. The turbine model studied is the WindPACT 5 MW reference model and the BEM model is FAST.

\cite{Echeverria2017} employs a global sensitivity analysis to identify the design variables that affect wind turbine performance in order to reduce the number of variables in wind turbine design optimization. As input parameters, airfoil parameterization and chord and twist distributions are used; the outputs studied are annual energy production, maximum blade tip deflection, overall sound power level and blade mass. For example, Sobol indices are shown for the effect of chord and twist control points on the output quantities.

\cite{Matthaus2017} performed uncertainty quantification including uncertainty in the aerodynamic properties; for example, the uncertainty in the lift coefficient at an airfoil section due to uncertain roughness (e.g.\ contamination) is achieved by interpolating between clean and fully rough state with a uniformly distributed random variable. Polynomial chaos and Kriging were used. See also \cite{Bortolotti2019}, in which the AVATAR turbine was analysed, using the Cp-Lambda aeroelastic model and DAKOTA as UQ toolbox.

\cite{Murcia2018} similarly considered a global sensitivity analysis with Sobol indices computed by using sparse polynomial chaos expansion. The quantities of interest considered are the energy production and lifetime equivalent fatigue loads for the DTU 10 MW reference turbine. The uncertain input is given by the turbulent inflow field, with 4 parameters: mean hub height wind speed, std. dev. of hub height wind speed, shear exponent, yaw misalignment. The dependency between these parameters (as described by the Normal Turbulence Model) is taken into account by using a Rosenblatt transformation that transforms the dependent variables to a set of independent ones. The Chaospy package is used for the analysis. Seven different model outputs are considered: power, thrust, and several damage equivalent fatigue loads (EFL), computed using a rainflow counting algorithm. The sensitivity analysis shows that the turbulent inflow realization has a bigger impact on the total distribution of equivalent fatigue loads than the shear coefficient or yaw misalignment.

\cite{Sayed2019} argue that BEM models are accurate mainly for small turbines but that for large turbines, when the tip deformations exceed 10\% of the blade radius, it is questionable if they can still be applied. They compare the results of engineering models to CFD-based aeroelastic simulations. It is shown that the 2D polars commonly used in BEM require 3D corrections, and that including more polars improves the results. `Turbine power and thrust predicted from the flexible rotor were reduced compared to the rigid rotor employing a BEM-based model. A reason behind the increase in power from the CFD-based model is the spanwise force component. It is not included in the BEM-based simulations as it is one of the main assumptions of its theory. This assumption could be used for small wind turbine simulations where the edgewise deformations were insignificant. But for such large wind turbines, the spanwise force distribution over the blade radius must be taken into account in the calculation of the power output due to the large edgewise deformations. To increase the accuracy of the engineering model, it is recommended to increase the number of polars calculated by 2D CFD simulations at different sections along the blade. This is because of the significant difference in the Reynolds number for the large blade.'

\cite{Velarde2019} perform a global sensitivity analysis of fatigue loads with respect to structural, geotechnical and metocean parameters for a 5MW offshore wind turbine installed on a gravity based foundation. Linear regression of Monte Carlo simulations and Morris screening are performed for three design load cases.

\cite{Robertson2018} assesses the sensitivity of different wind parameters on the loads of a wind turbine. They argue that the most common parameters considered normally in literature are the turbulence intensity variability, and the shear exponent, or wind profile, both having important influence on the turbine response. In the paper an assessment of which wind characteristics influence wind turbine structural loads is given. The correlated dependency between the individual parameters is not considered; rather, they focus on assessing the sensitivity of each parameter individually. An elementary-effects screening method (Morris screening) is used as global sensitivity method. The NREL 5MW turbine and FAST are used.

\cite{Fluck2018} develop an \textit{intrusive} stochastic BEM method: `a stochastic solution for the unsteady aerodynamic loads based on a projection of the unsteady Blade Element Momentum (BEM) equations onto a stochastic space spanned by chaos exponentials'.

\cite{Hubler2017} used global sensitivity analysis (as part of a four-step approach) to reduce the parameter dimension for structural behaviour computations using FAST as aeroelastic code. The methods of expert knowledge, one-at-a-time variation, and regression based on Monte Carlo sampling are used as pre-processing steps to reduce the computational expense of global sensitivity analysis.

### 2019-11-29 Papers/Torque_Paper/article.tex
made reference file shared between the different papers
added section on model uncertainties

--- 2019-11-29 Paper/article.tex
@@
 % bibliography
-\newcommand{\myreferences}{./Bibliography/references,./Bibliography/Mendeley_refs}
-\bibliographystyle{elsarticle-num}
+\newcommand{\myreferences}{../references,../Mendeley_refs}
+\bibliographystyle{abbrv} %elsarticle-num}
 
@@
 \section{Aeroelastic models}\label{sec:model_description}
-\subsection{BEM models}
-\subsection{Input parameters of aeroelastic models}
+\subsection{BEM model uncertainties}
+State-of-the-art wind-turbine models simulate the aeroelastic behavior of wind turbines by combining the concept of momentum conservation of the flow (BEM) with the equations of motion for the structure, possibly extended with hydrodynamics of the sea and control algorithms \cite{Vorpahl2013}.
+In this work, we only concentrate on the first aspect, namely the prediction of flow and blade forces as given by the BEM method.
+The particular BEM code that we use is  ECN's \texttt{Aero-Module} \cite{Boorsma2012}.
+
+For this case of a rigid turbine, with a uniform inflow field, the main model uncertainties that arise in the BEM formulation that we will discuss in this work are the following \cite{Hansen1993}:
+\begin{itemize}
+\item The use of (2D) \textit{airfoil polars}; uncertainty arises not only because the actual flow along the blade is 3D, but also because the 2D polars itself can be inaccurate, either when obtained from measurement or from 2D codes like Xfoil;
+%\item Sheared inflow
+\item The assumption of locally 2D flow along the blade is also invalid at the tip; this is typically corrected via \textit{a tip loss model};
+\item Three-dimensional flow effects also cause stall delay at the root; this is typically modelled via \textit{a dynamic stall model }and time constant;
+\item Yawed inflow is not naturally included in BEM and typically incorporated via \textit{a wake correction}.
+
+%\item Dynamic wake - Oye time lag constant
+%\item Tower shadow constants – Potential flow / Downwind Empirical model
+
+\end{itemize}
+
+%\subsection{Input parameters of aeroelastic models}
+%The main
 
@@
 \subsection{NM80}
+ and the turbine is the 2MW NM80 turbine from the DANAERO project \cite{Troldborg2013} with a blade radius of 38.8m.
+The data for lift ($Cl$) and drag ($Cd$) polars are available at four locations along the blade radius at 11.87m, 17.82m, 28.97m and 35.53m.
+The reference value of polars are obtained from wind-tunnel experiment with 3D corrections.
+Random samples of chord, twist, lift- and drag-polars are obtained by perturbing the control points with a uniformly distributed random variable.
+For chord and twist, we independently perturb control points at different locations as shown in Fig.
+\ref{fig:samples} (a) - (b).
+For $Cl$ and $Cd$, all control points are perturbed using a same random number.
+
 \subsubsection{Verification of GSA method}

### 2019-12-02 Torque_Abstract/main.tex
added draft of torque abstract

--- 2019-11-25 Torque_Abstract/main.tex
@@
  }
-
 \usepackage{natbib} % references
@@
 \usepackage{amsmath} % math typesetting
-
+\usepackage{graphicx}
+\usepackage{subcaption}
+\usepackage{microtype}
 % typeset in sans serif to match the prompt
@@
 Aeroelastic models such as the Blade Element Momentum (BEM) method \cite{HandBook} continue to play a critical role in the design, development and optimization of modern wind turbines.
-The accuracy of BEM predictions is affected to by uncertainties and inaccuracies, for example in the external conditions (wind parameters), in the turbine specification (geometric parameters), and in the BEM equations itself (model parameters).
-For the purpose of design, uncertainty quantification and optimization, is it crucial to limit the number of parameters, typically by performing sensitivity studies.
-Most studies focused on the effect of uncertainties in the external conditions and in the geometry, see e.g.\ \cite{Echeverria2017,Matthaus2017,Murcia2018,Robertson2018}.
-
+The accuracy of BEM predictions is affected by uncertainties and inaccuracies, for e.g., in the external conditions (wind parameters), in the turbine specifications (geometric parameters), and in the BEM equations itself (model parameters).
+For the purpose of design, uncertainty quantification and optimization, it is crucial to limit the number of parameters, typically by performing sensitivity studies.
+Most studies focused on the effect of uncertainties in the external conditions and geometry, see e.g.
+\cite{Echeverria2017,Robertson2018}.
+% removed reference Matthaus2017,Murcia2018
 \section{Objectives}
-The long-term objective of this study is to develop calibrated BEM models that give users an indication of the uncertainty associated with the predictions (loads, power, etc.) originating not only from external conditions and geometry, but also from the model formulation itself.
+The long-term objective of this study is to develop calibrated BEM models that give users an indication of the uncertainty associated with the predictions (loads, power, etc.) originating not only from external conditions and geometry but also from the model formulation itself.
 For this purpose, we will calibrate the model parameters present in BEM models.
 Examples of such model parameters are the time constant in dynamic stall models, the wake correction factor, the tip loss model parameter, and the lift- and drag-polars \cite{Sayed2019}.
-In order to limit the number of model parameters involved in the calibration process, \textit{the objective of the current study is to perform a global sensitivity study of the outputs of the BEM model towards both geometric and model uncertainties}.
+To limit the number of model parameters involved in the calibration process, \textit{the objective of the current study is to perform a global sensitivity study of the outputs of the BEM model towards both geometric and model uncertainties}.
 
-%In the past a number of sensitivity studies have been performed to understand the influence of input parameters on different turbine response, e.g.\ \cite{moriarty2002effect, eggers2003wind, McKay2014, dykes2014sensitivity, Robertson2018}.
-Many studies have confirmed that the wind parameters especially the wind speed and wind speed standard deviation have the most influence on aerodynamical performance of turbines.
-For example, in the of an upstream turbine, the wind speed is most sensitive to power production.
+%In the past several sensitivity studies have been performed to understand the influence of input parameters on different turbine responses, e.g.\ \cite{moriarty2002effect, eggers2003wind, McKay2014, dykes2014sensitivity, Robertson2018}.
+Many studies have confirmed that wind parameters especially wind speed and wind speed standard deviation have the most influence on the aerodynamical performance of turbines.
+For example, in an upstream turbine, the wind speed is most sensitive to power production.
 Furthermore, wind speed in combination with wind speed standard deviation has a large influence on the power production of turbines operating in the wake of an upstream turbine.
@@
 %In this work, we also study the effect of manufacturing tolerances on the turbine response.
-Usually, there are some discrepancy between the  manufactured and nominally prescribed design of the turbine blade leading to a suboptimal performance.
-For BEM models, the turbine shape is described as a series of airfoils along the span of the blade where each airfoil shape is computed using the three quantities: chord length, thickness and twist.
+Usually, there is some discrepancy between the manufactured and nominally prescribed design of the turbine blade leading to suboptimal performance.
+For BEM models, the turbine shape is described as a series of airfoils along the span of the blade where each airfoil shape is computed using the three quantities: chord length, thickness, and twist.
 We perturb these three quantities to obtain a perturbed turbine blade and analyze with sections that have more influence on output quantities.
@@
 \section{Methodology}
-To compute parameter sensitivities we use a global sensitivity analysis based on the Sobol expansion approach, which decomposes the total variance of the quantity of interest (model output) into contributions from individual parameters and their combinations, similar to \cite{Echeverria2017,Murcia2018,Rinker2016}.
+To compute parameter sensitivities we use a global sensitivity analysis (GSA) based on the Sobol expansion approach, which decomposes the total variance of the quantity of interest (model output) into contributions from individual parameters and their combinations, similar to \cite{Echeverria2017,Rinker2016}.
 We employ the uncertainty quantification toolbox \texttt{UQLab} \cite{uqlab}, which computes the Sobol indices from a sparse polynomial chaos expansion.
@@
 The geometric uncertainties currently considered are chord and twist distribution, whereas the model uncertainty enters via uncertainty in lift- and drag-polars.
-In order to express the chord, twist, lift and drag distributions along the turbine blade, an efficient parameterization is needed that gives flexible control over the prescribed uncertainty while limit the number of required parameters.
-We have chosen to use NURBS curves for this purpose, similar to \cite{Echeverria2017}.
+To express the chord, twist, lift and drag distributions along the turbine blade, an efficient parameterization is needed that gives flexible control over the prescribed uncertainty while limiting the number of required parameters.
+We have chosen to use Non-Uniform Rational Basis Splines (NURBS) for this purpose, similar to \cite{Echeverria2017}.
 
@@
 \section{Results}
-The aeroelastic code that we use is the \texttt{ECN Aero-Module} \cite{Boorsma2012} and the turbine is the 2MW NM80 turbine from the DANAERO project \cite{Troldborg2013}.
-
+The aeroelastic code that we use is the \texttt{ECN Aero-Module} \cite{Boorsma2012} and the turbine is the 2MW NM80 turbine (blade radius of 38.8m) from the DANAERO project \cite{Troldborg2013}.
+The data for airfoil lift- and drag-polars are available at four locations along the blade at 11.87m, 17.82m, 28.97m and 35.53m.
+%The reference value of polars are obtained from wind-tunnel experiment with 3D corrections.
+The lift variables are numbered from Cl1 - Cl4 and drag from Cd1 - Cd4 along the blade radius.
+Random samples of chord, twist, lift- and drag-polars are obtained by perturbing the control points with a uniformly distributed random number.
+Each of the control points for chord (Ch1 - Ch5) and twist (Tw1 - Tw6)  are perturbed independently.
+Thus, by locally perturbing the chord and twist curves, we can identify locations along the blade that are sensitive to turbine response.
+For a given lift and drag curve, all control points are perturbed (globally) using the same random number.
+Here, we are interested in knowing at what locations the inaccuracies in polars may lead to the most variations in the turbine output.
+Random samples for these input parameters are presented in Fig.
+\ref{fig:samples}.
+In Fig.
+\ref{fig:Sobol}, we show the total order Sobol indices as a measure of sensitivity for different geometric and model parameters on the power output.
+We observe that the model parameters are relatively more sensitive than the geometric parameters.
+Within geometric parameters, chord distribution shows a significantly higher sensitivity compared to the twist variables.
+\begin{figure}[h!]
+ \centering
+ \includegraphics[trim={0 1.6cm 0 0},clip, scale=0.65]{figure1.pdf}
+ \caption{Random realization of chord, twist, lift- (Cl2) and drag-polars (Cd2).}
+ \label{fig:samples}
+\end{figure}
+\begin{figure}[h!]
+ \centering
+  \includegraphics[width=\linewidth]{SA_Power_chord_twist_Cl_Cd.eps}
+ \caption{Total order Sobol indices for the sensitivity analysis of geometrical and model parameters.}
+  \label{fig:Sobol}
+\end{figure}
 % =============================================
 \section{Conclusions}
-
+We have shown that Sobol indices computed using sparse polynomial expansion is highly suitable for high-dimensional GSA.
+In the full paper, we will present a detailed sensitivity assessment that will include more model parameters.
+The identified sensitive parameters will further be utilized to develop calibrated BEM models.
 % =============================================

### 2019-12-02 sensitivity_analysis/AEROmoduleWrapper/writeAeroModuleInput.m

@@ -287,13 +287,13 @@ for i = 1:P{31}{1} % Loop over the polar files
 end
 
 %% Uncomment to plot the random samples of chord, twist, CL, CD
-% figure(1)
-% hold on
-% plot(P{3}, chord, 'color','k','linestyle','--','linewidth',0.5,'HandleVisibility','off')
-% 
-% figure(2)
-% hold on 
-% plot(P{3}, twist, 'color','k','linestyle','--','linewidth',0.5,'HandleVisibility','off')
+figure(1)
+hold on
+plot(P{3}, chord, 'color','k','linestyle','--','linewidth',0.5,'HandleVisibility','off')
+
+figure(2)
+hold on 
+plot(P{3}, twist, 'color','k','linestyle','--','linewidth',0.5,'HandleVisibility','off')
 % 
 % figure(3)
 % hold on

### 2019-12-02 sensitivity_analysis/cases/aero_module/initialize.m

@@ -29,7 +29,7 @@ NsamplesOLS = [8]; % if not specified, the number of samples from Quad is taken
 OLS_repeat = 1; % like MC_repeat
  
 % for PCE-LARS:
-NsamplesLARS = [1024]; % if not specified, the number of samples from Quad is taken
+NsamplesLARS = [20]; % if not specified, the number of samples from Quad is taken
 LARS_repeat = 1; % like MC_repeat
 
 %% Assemble the Input.Marginal for sensitivity analysis by text comparison

### 2019-12-02 sensitivity_analysis/testSensitivity.m

@@ -1,6 +1,6 @@
 %% Sobol indices computed with Monte Carlo and PCE-type methods
 clc
-close all
+% close all
 clearvars 
 
 caseName = 'aero_module'; % 'airfoil_lift';

### 2019-12-02 Papers/Torque_Abstract/main.tex
made reference file shared between the different papers
added section on model uncertainties

--- 2019-11-25 Torque_Abstract/main.tex
@@
 \section{Results}
-The aeroelastic code that we use is the \texttt{ECN Aero-Module} \cite{Boorsma2012} and the turbine is the 2MW NM80 turbine from the DANAERO project \cite{Troldborg2013}.
-
+The aeroelastic code that we use is the \texttt{ECN Aero-Module} \cite{Boorsma2012} and the turbine is the 2MW NM80 turbine (blade radius of 38.8m) from the DANAERO project \cite{Troldborg2013}.
+The data for airfoil lift- and drag-polars are available at four locations along the blade at 11.87m, 17.82m, 28.97m and 35.53m.
+%The reference value of polars are obtained from wind-tunnel experiment with 3D corrections.
+The lift variables are numbered from Cl1 - Cl4 and drag from Cd1 - Cd4 along the blade radius.
+Random samples of chord, twist, lift- and drag-polars are obtained by perturbing the control points with a uniformly distributed random number.
+Each of the control points for chord (Ch1 - Ch5) and twist (Tw1 - Tw6)  are perturbed independently.
+Thus, by locally perturbing the chord and twist curves, we can identify locations along the blade that are sensitive to turbine response.
+For a given lift and drag curve, all control points are perturbed (globally) using the same random number.
+Here, we are interested in knowing at what locations the inaccuracies in polars may lead to the most variations in the turbine output.
+Random samples for these input parameters are presented in Fig.
+\ref{fig:samples}.
+In Fig.
+\ref{fig:Sobol}, we show the total order Sobol indices as a measure of sensitivity for different geometric and model parameters on the power output.
+We observe that the model parameters are relatively more sensitive than the geometric parameters.
+Within geometric parameters, chord distribution shows a significantly higher sensitivity compared to the twist variables.
+\begin{figure}[h!]
+ \centering
+ \includegraphics[trim={0 1.6cm 0 0},clip, scale=0.65]{figure1.pdf}
+ \caption{Random realization of chord, twist, lift- (Cl2) and drag-polars (Cd2).}
+ \label{fig:samples}
+\end{figure}
+\begin{figure}[h!]
+ \centering
+  \includegraphics[width=\linewidth]{SA_Power_chord_twist_Cl_Cd.eps}
+ \caption{Total order Sobol indices for the sensitivity analysis of geometrical and model parameters.}
+  \label{fig:Sobol}
+\end{figure}
 % =============================================
 \section{Conclusions}
-
+We have shown that Sobol indices computed using sparse polynomial expansion is highly suitable for high-dimensional GSA.
+In the full paper, we will present a detailed sensitivity assessment that will include more model parameters.
+The identified sensitive parameters will further be utilized to develop calibrated BEM models.
 % =============================================
 \bibliographystyle{plain}
-\bibliography{references,Mendeley_refs}
+\bibliography{../references,../Mendeley_refs}
 

### 2019-12-03 sensitivity_analysis/cases/airfoil_lift/postProcessing.m
Added the workshop presentation

@@ -69,11 +69,11 @@ if (find(strcmp(methods,'MC')))
     semilogx(NsamplesMC', AVG_Sobol_MC_Total(:, end), 'x-','Linewidth', 2, 'Color', cmap(1,:));  
 end
 if (find(strcmp(methods,'PCE_Quad')))
-    if(size(AVG_Sobol_Quad_Total,2)>1)
-        semilogx(NsamplesQuad', AVG_Sobol_Quad_Total(:, 1:end-1), 's-','Linewidth', 2,'Color', cmap(2,:), 'HandleVisibility','off');
+    if(size(Sobol_Quad_Total,2)>1)
+        semilogx(NsamplesQuad', Sobol_Quad_Total(:, 1:end-1), 's-','Linewidth', 2,'Color', cmap(2,:), 'HandleVisibility','off');
         hold on
     end
-    semilogx(NsamplesQuad', AVG_Sobol_Quad_Total(:, end), 's-','Linewidth', 2,'Color', cmap(2,:));
+    semilogx(NsamplesQuad', Sobol_Quad_Total(:, end), 's-','Linewidth', 2,'Color', cmap(2,:));
 end
 if (find(strcmp(methods,'PCE_OLS')))
     if(size(AVG_Sobol_OLS_Total,2)>1)
@@ -114,7 +114,7 @@ if (find(strcmp(methods,'MC')))
 end
 
 if (find(strcmp(methods,'PCE_Quad')))
-    uq_bar((1:ndim)+ coords(k), AVG_Sobol_Quad_Total(end,:), bar_width, 'FaceColor', cmap(k,:), 'EdgeColor', 'none')
+    uq_bar((1:ndim)+ coords(k), Sobol_Quad_Total(end,:), bar_width, 'FaceColor', cmap(k,:), 'EdgeColor', 'none')
     k = k+1;
 end
 
@@ -151,94 +151,94 @@ legend(methods, 'Interpreter', 'none')
 ylabel('Total order Sobol index');
 ylim([0 1])
 
-%% plot the polynomial response surface for PCE quadrature-based
-
-if (find(strcmp(methods,'PCE_Quad')))
-    
-    n_inputs = length(Input.Marginals);
-    
-    if (n_inputs == 1)
-        Xsamples  = myPCE_Quad.ExpDesign.X;
-        Ysamples  = myPCE_Quad.ExpDesign.Y;
-        
-        N = 100;
-        domain = getMarginalBounds(myInput.Marginals(1));
[…]

### 2019-12-06 Papers/Torque_Abstract/main.tex
Added the workshop presentation
font size increase for figure

@@
 \titleformat*{\section}{\normalsize\bfseries}
-
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 %The reference value of polars are obtained from wind-tunnel experiment with 3D corrections.
-The lift variables are numbered from Cl1 - Cl4 and drag from Cd1 - Cd4 along the blade radius.
-Random samples of chord, twist, lift- and drag-polars are obtained by perturbing the control points with a uniformly distributed random number.
-Each of the control points for chord (Ch1 - Ch5) and twist (Tw1 - Tw6)  are perturbed independently.
-Thus, by locally perturbing the chord and twist curves, we can identify locations along the blade that are sensitive to turbine response.
+The lift and drag variables at these sections are numbered Cl1 - Cl4 and Cd1 - Cd4 respectively.
+Uncertainty enters by multiplying the reference lift and drag curves by a random variable (see Fig.
+\ref{fig:samples}).
+The chord and twist are perturbed via a number of control points that define the NURBS representation; we use 5 chord (Ch1 - Ch5) and 6 twist (Tw1 - Tw6) control points and perturb each value randomly following a uniformly distribution.
+Examples of random samples for these input parameters are presented in Fig.~\ref{fig:samples}.
+
+%Thus, by locally perturbing the chord and twist curves, we can identify locations along the blade that are sensitive to turbine response.
 For a given lift and drag curve, all control points are perturbed (globally) using the same random number.
 Here, we are interested in knowing at what locations the inaccuracies in polars may lead to the most variations in the turbine output.
-Random samples for these input parameters are presented in Fig.
-\ref{fig:samples}.
-In Fig.
-\ref{fig:Sobol}, we show the total order Sobol indices as a measure of sensitivity for different geometric and model parameters on the power output.
-We observe that the model parameters are relatively more sensitive than the geometric parameters.
-Within geometric parameters, chord distribution shows a significantly higher sensitivity compared to the twist variables.
+In Fig.~\ref{fig:Sobol}, we show the total order Sobol indices as a measure of the sensitivity of the different geometric and model parameters on the power output.
+In this case, we observe the important result that the uncertainty in model parameters is at least as important as the uncertainty in geometric parameters.
+Furthermore, within the geometric parameters, the chord distribution shows a significantly higher sensitivity compared to the twist variables.
+
 \begin{figure}[h!]
  \centering
- \includegraphics[trim={0 1.6cm 0 0},clip, scale=0.65]{figure1.pdf}
+ \includegraphics[trim={0 1.6cm 0 1.5cm},clip, scale=0.67]{figure1.pdf}
  \caption{Random realization of chord, twist, lift- (Cl2) and drag-polars (Cd2).}
@@
 \end{figure}
+
 \begin{figure}[h!]
@@
   \includegraphics[width=\linewidth]{SA_Power_chord_twist_Cl_Cd.eps}
- \caption{Total order Sobol indices for the sensitivity analysis of geometrical and model parameters.}
+ \caption{Sensitivity of power output towards geometric and model parameters, expressed by Sobol indices.}
   \label{fig:Sobol}
@@
 \section{Conclusions}
-We have shown that Sobol indices computed using sparse polynomial expansion is highly suitable for high-dimensional GSA.
-In the full paper, we will present a detailed sensitivity assessment that will include more model parameters.
-The identified sensitive parameters will further be utilized to develop calibrated BEM models.
+We have shown how Sobol indices computed using sparse adaptive polynomial expansion can be used for high-dimensional global sensitivity analysis of both geometric and model uncertainties.
+In the full paper, we will extend the sensitivity assessment to include more model uncertainties, such as parameters in dynamic stall and tip loss models.
+The identified sensitive model parameters will be utilized in future work to develop calibrated BEM models with built-in uncertainty estimates as part of the \textit{WindTrue} project, and is connected to IEA Task 29.
+
 % =============================================
-\bibliographystyle{plain}
+\bibliographystyle{abbrv}
 \bibliography{../references,../Mendeley_refs}

### 2019-12-17 Papers/Torque_Paper/article.tex
updated article text

@@
 \begin{itemize}
-\item The use of (2D) \textit{airfoil polars}; uncertainty arises not only because the actual flow along the blade is 3D, but also because the 2D polars itself can be inaccurate, either when obtained from measurement or from 2D codes like Xfoil;
+\item The use of (2D) \textit{airfoil polars}; uncertainty arises not only because the actual flow along the blade is 3D, but also because the 2D polars itself can be inaccurate, either when obtained from measurement or from 2D codes like \texttt{XFOIL};
 %\item Sheared inflow
@@
 \subsection{Geometric uncertainty}
-We use Non-Uniform Rational Basis Splines (NURBS) \cite{rogers2000} to perturb the geometrical parameters of the turbine blade, such as the reference chord and twist curves.
-The main advantage of using NURBS is that it provides great flexibility to approximate a large variety of curves with a limited number of control points.
-Further, the set of control points and knots can be directly manipulated to control the smoothness and curvature.
-
-Use of NURBS in wind-turbine blades (airfoils) optimization: \cite{Ribeiro2012,Bottasso2014}.
-
+We use Non-Uniform Rational Basis Splines (NURBS) \cite{rogers2000} to perturb the geometrical parameters of the turbine blade, such as the reference chord, twist or experimentally obtained drag-, lift-polar curves.
+The main advantage of using NURBS is that it provides great flexibility to approximate a large variety of curves with a limited number of control points which can be useful in reducing the number of uncertain parameter.
+%Further, the set of control points and knots can be directly manipulated to control the smoothness and curvature.
+Use of NURBS has already been exploited in many aerodynamics applications, for instance, design optimization of wind-turbine blades, see for e.g.
+\cite{Ribeiro2012,Bottasso2014} and to parametrize chord and twist curves \cite{Echeverria2017}.
+In the following, we briefly outline the procedure used to parametrize curves using NURBS.
 \subsubsection{NURBS based perturbation}
-The value of NURBS curve at each location $x$ is computed using a weighted sum of $N$ basis functions (or B-splines):
+The value of NURBS curve at each location $x$ is computed using a weighted sum of $n$ basis functions (or B-splines):
 \begin{equation}\label{NURB_curve}
-S(x)  = \sum_{i=1}^{N} c_i B_{i,p}(x),
+S(x)  = \sum_{i=1}^{n} c_i B_{i,p}(x),
 \end{equation}
-where $S(x)$ is the value of the curve at location $x, c_i$ is the weight of control point $i$ and $B_{i,p}(x)$ is the value of B-spline corresponding to the $i$-th control point at $x$.
-The subscript $p$ denote the polynomial degree of the NURBS curve.
-The total number of control points $N = m+p+1$ where $m$ is the number of knots  and the degree of NURBS curve.
-The above definition can be easily extended to higher dimensions.
-
-The B-splines are recursive in polynomial degree, for example, we can derive quadratic B-splines ($p=2$) using linear B-splines ($p=1$), cubic ($p=3$) from quadratic B-splines and so on.
-Given $m$ knot locations $t_1, t_2, ..., t_{m}$, the B-spline of degree 0 is defined as:
+where $S(x)$ is the value of the NURBS curve at location $x$; $c_i$ is the weight of the $i$-th control point and $B_{i,p}(x)$ is the value of B-spline corresponding to the $i$-th control point at $x$ with the subscript $p$ denoting its polynomial degree.
+A sequence of non-decreasing parameters $X = \{x_1,x_2,...,x_m\}$ called the \textit{knot vector} determines the domain in which any control point is active with the number of elements in the knot vector is given by $m = n+p+1$.
+B-splines are non-zero only within a knot interval $[x_i,x_{i+1})$ for $i = 1, 2, ..., m-1$ and are zero elsewhere.
+Therefore, only a few nearby control points are active when computing the NURBS curve \eqref{NURB_curve}.
+For a curve of polynomial degree p, there are $p+1$ active control points at any location $x$, for e.g., for a linear curve ($p=1$) only two adjacent control points are active at any location.
+Also, note that the number of B-splines are equal to the number of elements in $X$.
+Another useful property of the B-splines is that they are recursive in polynomial degree thus we can derive quadratic B-splines using linear B-splines, cubic from quadratic B-splines and so on.
+Starting with B-spline of degree 0 defined as
 \begin{equation}\label{linearBspline}
@@
 \begin{cases}
-1\quad t_i\leq x < t_{i+1},\quad i = 1,2,...,m,\\
-0\quad\text{elsewhere.}
+1\quad x_i\leq x < x_{i+1},\quad i = 1,2,...,m,\\
+0\quad\text{elsewhere,}
 \end{cases} 
 \end{equation}
-Higher order B-splines can then be derived using the recurrence relation \cite{deBoor}:
+we can derive higher order B-splines in the knot interval $[x_{i+p},x_{i+p+1})$ using the recurrence relation \cite{deBoor}:
 \begin{equation}\label{NURBS_recurrence}
-B_{i,p}(x) := \frac{x-t_i}{t_{i+p} - t_i}B_{i,p-1}(x) + \frac{t_{i+p+1}  -  x}{t_{i+p+1}  -  t_{i+1}}B_{i+1,p-1}(x),\quad p\geq1, i = 1,2,...,m.
+B_{i,p}(x) := \frac{x - x_i}{x_{i+p} - x_i}B_{i,p-1}(x) + \frac{x_{i+p+1}  -  x}{x_{i+p+1}  -  x_{i+1}}B_{i+1,p-1}(x),\quad p\geq1.
 \end{equation} 
-Something about padding...
-These B-splines can be constructed efficiently using the De Boor's algorithm \cite{deBoor}.
+These B-splines can be constructed efficiently using the de Boor's algorithm \cite{deBoor}.
+Note that to perform above computation, we need to pad the knot vector $X$ by repeating the first and last elements $p$ times each.
 In Fig.
-\ref{B-splines}, we show linear, quadratic and cubic splines for $x\in[0,1]$.
-At the interval boundaries these B-splines go to zero smoothly.
-The domain of influence of a given control point depends on the respective B-spline (rephrase)??
+\ref{B-splines}, we show linear, quadratic and cubic splines for $x\in[0,1]$ [Also show the padding].
 
 Next, we describe steps to generate perturbed samples of chord from a given reference chord, $S_{ref}(x)$, using NURBS based parametrization.
-The first step is to approximate the given reference chord using a NURBS curve with a fixed degree $p$.
-For this, we need to sample $S_{ref}(x)$ at $N$ locations $\{x_j\}_{j=1}^N$ and compute $B_{i,p}(x_j), \text{ for }i, j = 1,2, ..., N$, such that we can compute the set of control points $\mathbf{c}=\{c_i\}_{i=0}^N$ by solving the following linear system:
+The first step is to approximate the given reference chord using a NURBS curve with a fixed degree $p$ and a limited number of knots $X =\{x_j\}_{j=0}^n$.
+The knots can be uniformly spaced or chosen heuristically, more advanced approaches for knot selection can be found in, for e.g.
+\cite{razdan1999knot,LI2005791}.
+Next we sample $S_{ref}(x)$ at $X$ and also compute $B_{i,p}(x_j), \text{ for }i, j = 1,2, ..., n$.
+Now we can formulate a linear system to solve for a set of control points $\mathbf{c}=\{c_i\}_{i=0}^n$ as:
 \begin{equation}\label{nurbs_inversion}
-\mathbf{B}\mathbf{c} = \mathbf{S},
+\mathbf{B}\mathbf{c} = \mathbf{S}_{ref},
 \end{equation}
-where $\mathbf{S}\in \mathbb{R}^{N}$ is a vector containing sampled values of the reference curve and $\mathbf{B}\in \mathbb{R}^{N\times N}$ is a matrix with $j$-th row consisting of B-splines values at $x_j$ locations, i.e., $B_{i,p}(x_j), i= 1, 2,..., N$.
-Once the control points are obtained, we can derive the approximate reference curve $S_N(x)$ using \eqref{NURB_curve}.
-Note that the accuracy of the approximated curve is dependent on the sample locations $x_j$ as well as the degree $p$ and can be set heuristically \cite{adaptiveNurbs}.
-The number of sampled locations $N$ can be adaptively increased until the following tolerance criteria is met:
-\begin{equation}
-\frac{||S_N - S_{ref}||}{||S_{ref}||} <\varepsilon.
-\end{equation}
-More advanced approaches, monotonicity preserving methods, etc??
-%\begin{remark}
-%How to chose $x$?
-%\end{remark}
-%The final step is to introduce uncertainty in the above computed control points by associating a probability distribution.
-%
-%The steps to obtained perturbed samples from a given reference curve is outlined in Fig. \ref{}.
+where $\mathbf{S}_{ref}\in\mathbb{R}^{n}$ is a vector containing sampled values of $S_{ref}(x)$ and $\mathbf{B}\in \mathbb{R}^{n\times n}$ is a matrix with $j$-th row consisting of values of $n$ B-splines sampled at the location $x_j$.
+Once the control points are obtained, we can derive the approximate reference curve $S_n(x)$ using \eqref{NURB_curve}.
+In this work, we use uniform random variable to perturb the baseline control points $\mathbf{c}$.
+
 \subsection{Model uncertainty}
 \subsection{Wind velocity}
-
 \section{Global sensitivity analysis}\label{sec:GSA}
@@
 This approach allows for quantification of  the relative importance of the input parameters on a scale of $[0,1]$ known as \emph{Sobol indices}.
-
-The main idea of Sobol analysis is to express the total variance of the output in term of contributions from individual parameters and their combinations.
+In a Sobol analysis, we express the total variance of the output in term of contributions from individual parameters and their combinations.
 For the ease of exposition, let us consider a nonlinear model:
 \begin{equation}\label{nonlinear_model}
-Y = f(\mathbf{Z}),
+Y = f({z}),
 \end{equation}
-where $\mathbf{Z} = [z_1, z_2, ..., z_M]\in \mathcal{D}_{\mathbf{Z}}\in \mathbb{R}^M$ is the input vector.
+where ${z} = [z_1, z_2, ..., z_M]\in \mathcal{D}_{z}\in \mathbb{R}^M$ is a vector of random  inputs.
 For simplicity, we assume input parameters are uniformly distributed i.e.
-$z_i \sim \mathcal{U}(0,1)$ and the support of input set is $\mathcal{D}_\mathbf{Z}  =  [0,1]^M$ where $M$ is the total number input parameters.
-The Sobol decomposition is defined as:
+$z_i \sim \mathcal{U}(0,1)$ and the support of input set is $\mathcal{D}_{z}  =  [0,1]^M$ where $M$ is the total number input parameters.
+The Sobol decomposition is based on the hierarchical representation \cite{RSmith}:
 \begin{equation}\label{sobol_decomp}
@@
 \end{equation}
-The above decomposition is only valid for independent input parameters. 
-\subsubsection{PCE Ordinary Least Square (PCE\_OLS)}
-\subsubsection{PCE Least Angle Regression (PCE\_LAR)}
+where the zeroth-order function $f_0$ is the mean response of $f$, the first-order univariate functions $f_i(z_i)$ quantify independent contributions due to each random input, second-order functions $f_{i,j}(z_i,z_j)$ represent the effect of interaction between $z_i$ and $z_j$ on the response $f$.
+Note that the \eqref{sobol_decomp} is only valid for independent input parameters.
+Higher-order terms are interpreted in a similar manner.
+Formally, the zeroth-, first- and second-order terms are defined as:
+\begin{align}\label{sobol_terms}
+f_0 &= \int_{\mathcal{D}_{z}}f(z)dz,\\
+f_i(z_i) &= \int_{{\mathcal{D}_{z}}^{M-1}}f(z)dz_{\sim \{i\}} - f_0,\\
+f_{ij}(z_i,z_j) &= \int_{{\mathcal{D}_{z}}^{M-2}}f(z)dz_{\sim \{i,j\}} - f_i(z_i)  - f_j(z_j) - f_0,
+\end{align}
+where ${\mathcal{D}_{z}}^{M-1} = [0,1]^{M-1}$ and the notation $z_{\sim \{i,j\}}$ denotes the vector having all the components of ${z}$ except in the set $\{i,j\}$.
+We only consider second order expansion of \eqref{sobol_decomp}, as for a number of applications, the second order expansion is usually sufficient and higher-order terms have negligible effect on the output response.
+In case of wind turbine, suppose $y$ represent the power output that depends on say 3 random inputs:  windspeed $(z_1)$, wind standard deviation $(z_2)$ and RPM $(z_3)$, then $f_0$ represents the mean power output considering all the three random inputs, $f_1(z_1)$ represent the independent contribution of windspeed on the power and $f_{1,2}(z_1,z_2)$ quantify the interactions of wind speed and wind standard deviation on power.
+Further, the total variance of $f(z)$ is defined as:
+\begin{equation}\label{tot_var}
+D = \text{Var}[f(z)] = \int_{\mathcal{D}_{z}} f^2(z)dz - f_0^2
+\end{equation}
+As the expansion in \eqref{sobol_decomp} is not unique, some orthogonality conditions must be imposed, see \cite{Rabitz1999,SOBOL2001271} which allows to express the total variance as
+\begin{equation}
+D = \sum_{i=1}^M D_i + \sum_{1\leq i<j\leq M} D_{ij},
+\end{equation}
+where the first- and second-order partial variances are defined as
+\begin{align}
+D_i &= \int_{0}^{1} f^2_i(z_i)dz_i, \label{partial_variance1}\\
+D_{ij} &= \int_{0}^{1} \int_{0}^{1} f^2_{ij}(z_i,z_j)dz_idz_j \label{partial_variance2}. 
+\end{align}
+The first- and second-order Sobol indices are computed as
+\begin{equation}\label{sobol_ind}
+S_i = \frac{D_i}{D}, \quad  S_{ij} = \frac{D_{ij}}{D}, \quad i,j=1,2, ..., M,
+\end{equation}
+and the sum of all indices satisfy
+\begin{equation}
+\sum_{i=1}^M S_i + \sum_{1\leq i<j\leq M} S_{ij} = 1.
+\end{equation}
+Finally, the total effect of the parameter $z_i$ on the output response $Y$ is quantified using the total sensitivity indices measure
+\begin{equation}
+S_{T_i} = S_i + \sum_{j=1}^M S_{ij}.
+\end{equation}
+The total sensitivity indices can be interpreted as an importance measure for the parameter $z_i$, therefore a large $S_{T_i}$ implies that $z_i$ has a strong influence on $Y$.
+The computation of $S_{T_i}$ requires the approximation of partial variances defined in \eqref{partial_variance1} - \eqref{partial_variance2}.
+Using the Monte Carlo method for computing these variance can be prohibitive for computationally expensive models $Y$.
+Polynomial Chaos Expansions (PCE) based approaches provides a more efficient alternative for sensitivity analysis.
+As shown in  \cite{SUDRET2008964}, one can compute the Sobol indices analytically by post-processing the PCE coefficients.
+\subsection{PCE based Sobol indices computation}
 
+\subsubsection{Ordinary Least Square (PCE\_OLS)}
+\subsubsection{Least Angle Regression (PCE\_LAR)}
 %\section{Description of test case}
-
 %\section{Sensitivity analysis workflow}
-
 \section{Results}\label{sec:results}
@@
 \newpage
-
-\section{References}
 \bibliography{\myreferences}

### 2019-12-17 sensitivity_analysis/cases/aero_module/initialize.m

@@ -29,7 +29,7 @@ NsamplesOLS = [8 16]; % if not specified, the number of samples from Quad is tak
 OLS_repeat = 1; % like MC_repeat
  
 % for PCE-LARS:
-NsamplesLARS = [32]; % if not specified, the number of samples from Quad is taken
+NsamplesLARS = [64]; % if not specified, the number of samples from Quad is taken
 LARS_repeat = 1; % like MC_repeat
 
 %% Assemble the Input.Marginal for sensitivity analysis by text comparison

### 2019-12-17 Papers/Torque_Abstract/main.tex
small corrections to the Torque abstract

@@
 \begin{addmargin}[2.5cm]{0em}% 1em left, 2em right
-\textbf{P. Kumar$^{\boldsymbol{\mathsf{1}}}$, B. Sanderse$^{\boldsymbol{\mathsf{1}}}$}
+\textbf{P. Kumar$^{\boldsymbol{\mathsf{1}}}$, B. Sanderse$^{\boldsymbol{\mathsf{1}}}$, K. Boorsma$^{\boldsymbol{\mathsf{2}}}$, M. Caboni$^{\boldsymbol{\mathsf{2}}}$}
 
-$^{\mathsf{1}}$ Centrum Wiskunde \& Informatica, Science Park 123, Amsterdam.
+$^{\mathsf{1}}$ Centrum Wiskunde \& Informatica, Science Park 123, Amsterdam, The Netherlands.
+$^{\mathsf{2}}$ TNO, Westerduinweg 3, Petten, The Netherlands.
 
@@
 For the purpose of design, uncertainty quantification and optimization, it is crucial to limit the number of parameters, typically by performing sensitivity studies.
-Most studies focused on the effect of uncertainties in the external conditions and geometry, see e.g.
-\cite{Echeverria2017,Robertson2018}.
+Most studies so far focused on the effect of uncertainties in the external conditions and in geometry, see e.g.~\cite{Echeverria2017,Robertson2018}.
 % removed reference Matthaus2017,Murcia2018
@@
 \section{Methodology}
-To compute parameter sensitivities we use a global sensitivity analysis (GSA) based on the Sobol expansion approach, which decomposes the total variance of the quantity of interest (model output) into contributions from individual parameters and their combinations, similar to \cite{Echeverria2017,Rinker2016}.
+To compute parameter sensitivities we use a global sensitivity analysis (GSA) based on the Sobol expansion approach, which decomposes the total variance of the quantity of interest (model output) into contributions from individual parameters and their combinations, similar to \cite{Echeverria2017,Rinker2016a}.
 We employ the uncertainty quantification toolbox \texttt{UQLab} \cite{uqlab}, which computes the Sobol indices from a sparse polynomial chaos expansion.
@@
 \section{Results}
-The aeroelastic code that we use is the \texttt{ECN Aero-Module} \cite{Boorsma2012} and the turbine is the 2MW NM80 turbine (blade radius of 38.8m) from the DANAERO project \cite{Troldborg2013}.
-The data for airfoil lift- and drag-polars are available at four locations along the blade at 11.87m, 17.82m, 28.97m and 35.53m.
+The aeroelastic code that we use is the BEM model of the \texttt{ECN Aero-Module} \cite{Boorsma2012} and the turbine is the 2MW NM80 turbine (blade radius of 38.8m) from the DANAERO project \cite{Madsen2010a}.
+The data for the airfoil lift- and drag-polars is available at four locations along the blade: at 11.87m, 17.82m, 28.97m and 35.53m.
 %The reference value of polars are obtained from wind-tunnel experiment with 3D corrections.
@@
 \ref{fig:samples}).
-The chord and twist are perturbed via a number of control points that define the NURBS representation; we use 5 chord (Ch1 - Ch5) and 6 twist (Tw1 - Tw6) control points and perturb each value randomly following a uniformly distribution.
+The chord and twist are perturbed via a number of control points that define the NURBS representation; we use 5 chord (Ch1 - Ch5) and 6 twist (Tw1 - Tw6) control points and perturb each value randomly following a uniform distribution.
 Examples of random samples for these input parameters are presented in Fig.~\ref{fig:samples}.

### 2019-12-17 Papers/Torque_Paper/article.tex
minor changes to paper

--- 2019-11-29 Papers/Torque_Paper/article.tex
@@
 
-\section{Aeroelastic models}\label{sec:model_description}
-\subsection{BEM model uncertainties}
+\section{Uncertainties in BEM models}\label{sec:model_description}
+\subsection{Model uncertainties}
 State-of-the-art wind-turbine models simulate the aeroelastic behavior of wind turbines by combining the concept of momentum conservation of the flow (BEM) with the equations of motion for the structure, possibly extended with hydrodynamics of the sea and control algorithms \cite{Vorpahl2013}.
@@
 \item The assumption of locally 2D flow along the blade is also invalid at the tip; this is typically corrected via \textit{a tip loss model};
-\item Three-dimensional flow effects also cause stall delay at the root; this is typically modelled via \textit{a dynamic stall model }and time constant;
-\item Yawed inflow is not naturally included in BEM and typically incorporated via \textit{a wake correction}.
+\item Three-dimensional flow effects also cause stall delay, mainly at the root; this is typically modelled via \textit{a dynamic stall model }and time constant;
+\item Yawed inflow is not naturally included in BEM and typically incorporated via \textit{a skew correction function}.
 
@@
 \end{itemize}
+
+\subsection{Geometric uncertainties}
 

### 2020-01-07 Papers/Torque_Paper/article.tex
added discrete routines

--- 2019-12-17 Papers/Torque_Paper/article.tex
@@
 
-Recently, several papers have addressed the uncertainty in BEM model output by performing forward uncertainty propagation or sensitivity studies, e.g.\ \cite{Bos2019a,Echeverria2017,Matthaus2017,Murcia2018,Robertson2018}.
+Recently, several papers have addressed the uncertainty in BEM model output by performing forward uncertainty propagation or sensitivity studies, e.g.\ \cite{Echeverria2017,Matthaus2017,Murcia2018,Robertson2018,Bos2019a}.
 In these studies, the focus in mainly on uncertainties in the external conditions (wind parameters) and/or uncertainty in the turbine specification (geometric parameters).
@@
 Therefore, only a few nearby control points are active when computing the NURBS curve \eqref{NURB_curve}.
-For a curve of polynomial degree p, there are $p+1$ active control points at any location $x$, for e.g., for a linear curve ($p=1$) only two adjacent control points are active at any location.
+For a curve of polynomial degree $p$, there are $p+1$ active control points at any location $x$, for e.g., for a linear curve ($p=1$) only two adjacent control points are active at any location.
 Also, note that the number of B-splines are equal to the number of elements in $X$.
@@
 
-\subsection{Model uncertainty}
-\subsection{Wind velocity}
+There are two ways to obtain perturbed curves either by independently perturbing each of the control points (local perturbation) or using the same random number to perturb all the control points (global perturbation).
+We use local perturbation for chord control points Ch1-Ch5 and twist control points Tw1-Tw6 control points, see Fig.
+\ref{perturbed_samples} (a) - (b), respectively.
+Each control points are perturbed using a uniform random number with bounds given by $\pm5\%$ of the magnitude of control point.
+For lift and drag curve, all control points are perturbed using same uniform random number resulting in samples of the lift and drag curves as shown in Fig.
+\ref{perturbed_samples} (c) - (d), respectively.
+For simplicity, we only consider uniform distributions for perturbing the control points, however, we can also use other probability distributions for this.
+Also note that when using a uniform distribution with bounds given by relative magnitude of control points, we tend to obtain small perturbations for control points with values close to zero.
+\begin{figure}[h!]
+\centering
+\includegraphics[trim={0 1.6cm 0 1.5cm},clip, scale=0.67]{figure1.pdf}
+\caption{Random realization of chord, twist, lift- (Cl2) and drag-polars (Cd2).}
+\label{perturbed_samples}
+\end{figure}
+
 \section{Global sensitivity analysis}\label{sec:GSA}
@@
 
-\subsubsection{Ordinary Least Square (PCE\_OLS)}
+Quadrature based integration methods suffer from curse of dimensionality where the number of integration points rises exponentially with an increase in the number of dimensions.
+For applications with a fixed computational budget methods that minimizes the approximation error with a given number of samples are usually more practical.
+Next, we briefly describe two such methods to compute PCE coefficients based on a given number of model evaluations.
+
+\subsubsection{Ordinary Least Square}
+The 
 \subsubsection{Least Angle Regression (PCE\_LAR)}
+\subsection{Sensitivity analysis with discrete random variable}
 %\section{Description of test case}

### 2020-01-07 sensitivity_analysis/cases/aero_module/NM80.m

@@ -268,11 +268,15 @@ Input.Marginals(counter).Type = 'Uniform';
 Input.Marginals(counter).Parameters = [-0.5 0.5];
 Input.Marginals(counter).Bounds = [-0.5 0.5];
 
+
+
 %% Specify uncertain parameters to be considered in the sensitivity analysis
 % The parameter should be defined in the following format {name,index,rel_perturbation} where
 % rel_pertubation defines the amount of relative perturbation for B-spline
-% curves. This parameter may not be required for other scalar random
-% variables 
+% curves. For example, {'Twist',2,0.2} defines the uncertainty in the
+% second control point of Twist curve and 0.2 means a relative uncertainty of plus minus 10% for this  control point. 
+% This parameter may not be required for other scalar random
+% variables. 
 
 % uncertain_params = {{'Twist',2,0.2},{'Twist',3,0.2},{'Twist',4,0.2},{'Twist',5,0.2},{'Twist',6,0.2},{'Twist',7,0.2},...
 %                    {'Chord',2,0.2},{'Chord',4,0.2},{'Chord',6,0.2},{'Chord',8,0.2}, ...
@@ -283,6 +287,6 @@ Input.Marginals(counter).Bounds = [-0.5 0.5];
 %                    {'CM',1, 0.2}, {'CM',2, 0.2}, {'CM',3, 0.2},{'CM',4,0.2}
 %                    {'DYNSTALLTYPE','',''}, {'CORR3DTYPE','',''}};
 
-uncertain_params = {{'YAW','',''}, {'DYNSTALLTYPE','',''},{'CORR3DTYPE','',''}};
+uncertain_params = {{'YAW','',''},{'PITCHANGLE','',''}};
 % Specify quantity of interest
 QoI = 'Power'; % 'Axial_Force' or 'Power'

### 2020-01-07 sensitivity_analysis/cases/aero_module/initialize.m

@@ -1,5 +1,5 @@
 % Name of Matlab file representing the turbine data
-turbineName = 'NM80'; % 'NM80', 'AVATAR'
+turbineName = 'NM80_discrete'; % 'NM80', 'AVATAR'
 
 %% model description 
 % Name of Matlab file representing the model
@@ -19,22 +19,25 @@ methods = {'PCE_LARS'}; % {'MC','PCE_Quad','PCE_OLS','PCE_LARS'};
 % graphs
 MC_repeat = 1;
 % number of samples with MC
-NsamplesMC = [8 16 32];
+NsamplesMC = [8];
 
-% for PCE-Quad, specify the polynomial degrees to be tested
+% for PCE_Quad, specify the polynomial degrees to be tested
 DegreesQuad = 1:3; %[1 2 3 4 5 6];
 
 % % for PCE-OLS:
-NsamplesOLS = [8 16]; % if not specified, the number of samples from Quad is taken
+NsamplesOLS = [8 16 32]; % if not specified, the number of samples from Quad is taken
 OLS_repeat = 1; % like MC_repeat
  
 % for PCE-LARS:
-NsamplesLARS = [64]; % if not specified, the number of samples from Quad is taken
+NsamplesLARS = [8]; % if not specified, the number of samples from Quad is taken
 LARS_repeat = 1; % like MC_repeat
 
 %% Assemble the Input.Marginal for sensitivity analysis by text comparison
 ndim = length(P{26});
 ntot = length(P{25}.Marginals); 
+discrete_index = [];
+cont_index = [];
+discrete_param_vals = [];
 for i=1:ndim    
     for j = 1:ntot
         if(strcmp([P{25}.Marginals(j).Name,num2str(P{25}.Marginals(j).Index)],[P{26}{i}{1},num2str(P{26}{i}{2})]))
@@ -42,8 +45,15 @@ for i=1:ndim
             Input.Marginals(i).Type = P{25}.Marginals(j).Type; 
             Input.Marginals(i).Parameters = P{25}.Marginals(j).Parameters;
             Input.Marginals(i).Bounds = P{25}.Marginals(j).Bounds;
+            if(strcmp(Input.Marginals(i).Type,'Constant')) % Get the index and parameter of discrete variable
+                discrete_index = [discrete_index i];
+                discrete_param_vals = [discrete_param_vals Input.Marginals(i).Parameters];
+            else
+                cont_index = [cont_index i];
+            end
             break;
         end
     end
 end
- 
+
+

### 2020-01-07 sensitivity_analysis/testSensitivityDiscrete.m

@@ -1,0 +1,351 @@
+%% Sobol indices computed with Monte Carlo and PCE-type methods
+clc
+close all
+clearvars
+
+caseName = 'aero_module'; % 'airfoil_lift','aero_module', etc;
+input_file = caseName; % specify directory which contains test case settings and model
+
+%% Sobol options
+SobolOpts.Type        = 'Sensitivity';
+SobolOpts.Method      = 'Sobol';
+SobolOpts.Sobol.Order = 1;
+
+%% Add paths for dependent routines located in the directories 'NURBS','AEROmoduleWrapper' and 'Geometry'
+addpath([pwd,'/AEROmoduleWrapper/']);
+addpath([pwd,'/NURBS/']);
+addpath([pwd,'/Geometry/']);
+
+%% initialize UQlab
+
+% add path
+run('config.m');
+addpath(genpath(UQLab_path));
+
+% start uqlab
+uqlab
+
+%% process input files
+run(['cases/' input_file '/initialize.m']);
+
+if (exist('mean_exact','var'))
+    compare_mean = 1;
+else
+    compare_mean = 0;
+end
+if (exist('std_exact','var'))
+    compare_std = 1;
+else
+    compare_std = 0;
+end
+
+% create and add the model to UQLab
+myModel = uq_createModel(Model);
+% create input object with UQLab
+myInput = uq_createInput(Input) ;
+
+% display input properties
+uq_print(myInput);
+uq_display(myInput);
+
+%% Monte-Carlo
+if(find(strcmp(methods,'MC')))
+    
+    disp('=========MC==========');
+    N_MC     = length(NsamplesMC);
+    mean_MC  = zeros(MC_repeat,N_MC);
+    std_MC   = zeros(MC_repeat,N_MC);
+    Sobol_MC_FirstOrder = zeros(MC_repeat,N_MC,ndim);
+    Sobol_MC_Total      = zeros(MC_repeat,N_MC,ndim);
+    
+    % perform multiple runs to decrease effect of random sampling
+    for k = 1:MC_repeat
+        for i=1:N_MC
+            
+            %disp(NsamplesMC(i));
+            % get random samples ('experimental design')
+            X_ED = uq_getSample(NsamplesMC(i), 'MC');
+ 
+            if(size(discrete_index,1)>=1)
+                X_ED(:,discrete_index) = []; % get rid of the constant columns
+                X_temp  = allcomb(discrete_param_vals); % Use ndgrids to make a grid of discrete variables
[…]

### 2020-01-13 Papers/Torque_Paper/article.tex

@@
 Apart from understanding how uncertainty in the output is related to the different uncertain inputs, such sensitivity studies are very useful to reduce the number of parameters as needed for example in design optimization \cite{Echeverria2017}.
-However, the uncertainty in BEM model output as caused by uncertainty in the model formulation itself, e.g.\ through the values chosen for model parameters, has been given little attention (exceptions being the effect of aerodynamic properties studied in \cite{Matthaus2017,Bortolotti2019}).
+However, the uncertainty in BEM model output as caused by uncertainty in the model formulation itself, e.g.\ through the values chosen for model parameters, has been given little attention (exceptions being the effect of aerodynamic properties studied in \cite{Bortolotti2019,Matthaus2017}).
 
@@
 \begin{equation}\label{nonlinear_model}
-Y = f({z}),
+Y = f({\mathbf{z}}),
 \end{equation}
-where ${z} = [z_1, z_2, ..., z_M]\in \mathcal{D}_{z}\in \mathbb{R}^M$ is a vector of random  inputs.
+where ${\mathbf{z}} = [z_1, z_2, ..., z_M]\in \mathcal{D}_{\mathbf{z}}\in \mathbb{R}^M$ is a vector of random  inputs.
 For simplicity, we assume input parameters are uniformly distributed i.e.
-$z_i \sim \mathcal{U}(0,1)$ and the support of input set is $\mathcal{D}_{z}  =  [0,1]^M$ where $M$ is the total number input parameters.
+$z_i \sim \mathcal{U}(0,1)$ and the support of input set is $\mathcal{D}_{\mathbf{z}}  =  [0,1]^M$ where $M$ is the total number input parameters.
 The Sobol decomposition is based on the hierarchical representation \cite{RSmith}:
@@
 \begin{align}\label{sobol_terms}
-f_0 &= \int_{\mathcal{D}_{z}}f(z)dz,\\
-f_i(z_i) &= \int_{{\mathcal{D}_{z}}^{M-1}}f(z)dz_{\sim \{i\}} - f_0,\\
-f_{ij}(z_i,z_j) &= \int_{{\mathcal{D}_{z}}^{M-2}}f(z)dz_{\sim \{i,j\}} - f_i(z_i)  - f_j(z_j) - f_0,
+f_0 &= \int_{\mathcal{D}_{\mathbf{z}}}f(\mathbf{z})d\mathbf{z},\\
+f_i(z_i) &= \int_{{\mathcal{D}_{\mathbf{z}}}^{M-1}}f(\mathbf{z})d\mathbf{z}_{\sim \{i\}} - f_0,\\
+f_{ij}(z_i,z_j) &= \int_{{\mathcal{D}_{\mathbf{z}}}^{M-2}}f(\mathbf{z})d\mathbf{z}_{\sim \{i,j\}} - f_i(z_i)  - f_j(z_j) - f_0,
 \end{align}
-where ${\mathcal{D}_{z}}^{M-1} = [0,1]^{M-1}$ and the notation $z_{\sim \{i,j\}}$ denotes the vector having all the components of ${z}$ except in the set $\{i,j\}$.
+where ${\mathcal{D}_{\mathbf{z}}}^{M-1} = [0,1]^{M-1}$ and the notation $\mathbf{z}_{\sim \{i,j\}}$ denotes the vector having all the components of $\mathbf{z}$ except in the set $\{i,j\}$.
 We only consider second order expansion of \eqref{sobol_decomp}, as for a number of applications, the second order expansion is usually sufficient and higher-order terms have negligible effect on the output response.
 In case of wind turbine, suppose $y$ represent the power output that depends on say 3 random inputs:  windspeed $(z_1)$, wind standard deviation $(z_2)$ and RPM $(z_3)$, then $f_0$ represents the mean power output considering all the three random inputs, $f_1(z_1)$ represent the independent contribution of windspeed on the power and $f_{1,2}(z_1,z_2)$ quantify the interactions of wind speed and wind standard deviation on power.
-Further, the total variance of $f(z)$ is defined as:
+Further, the total variance of $f(\mathbf{z})$ is defined as:
 \begin{equation}\label{tot_var}
-D = \text{Var}[f(z)] = \int_{\mathcal{D}_{z}} f^2(z)dz - f_0^2
+D = \text{Var}[f(\mathbf{z})] = \int_{\mathcal{D}_{\mathbf{z}}} f^2(\mathbf{z})d\mathbf{z} - f_0^2
 \end{equation}
@@
 \subsection{PCE based Sobol indices computation}
+The PCE $f^{K}(\mathbf{z})$ of the computational model $Y$ is defined as a weighted sum of multivariate polynomials in $\mathbf{z}$ \cite{RSmith}
+\begin{equation}
+Y = f^{K}(\mathbf{z}) = \sum_{|\mathbf{k}| = 0}^K y_{\mathbf{k}}\Psi_{\mathbf{k}}(\mathbf{z}),
+\end{equation}
+where $\mathbf{k}\in  \mathbb{N}_0^M$ is a $M$-dimensional multi-index with the magnitude $|\mathbf{k}| = k_1+k_2 + ...
++ k_M$, whereas $\Psi_{\mathbf{k}}(\mathbf{z})$ is the multivariate orthogonal polynomials that is computed using a tensor product of univariate polynomials $\psi_{k_i}^{(i)}(z_i)$
+\begin{equation}
+\Psi_{\mathbf{k}}(\mathbf{z}) := \prod_{i=1}^M\psi_{k_i}^{(i)}(z_i).
+\end{equation}
+The choice of the univariate orthogonal polynomial $\psi_{k_i}$ depends on the type of random variable, $z_i$.
+For example, if for uniformly distributed random variable, the Legendre family of polynomials is used, see  \cite{Xiu2002} for details.
+Projection methods are typically employed for the calculation
 

### 2020-01-22 Papers/Torque_Paper/article.tex
Ran simulation for paper

@@
 \begin{abstract}
-Wind turbines are a complex multi-physics systems whose performance and life span is highly dependent on manufacturing, meteorological and operational factors.
-It is therefore important to quantify the impact of these parameters on the turbine response.
-We perform comprehensive sensitivity analysis of aeroelastic models typically employed to compute turbine response.
-This paper also seeks to quantify sensitivities due to manufacturing tolerance in the turbine blades.
-We propose a non-uniform rational basis splines approach to parametrize the geometry of the blade and further utilize this for quantifying geometric sensitivities at different radial locations.
-[Some important results]
+Sensitivity analysis of model parameters associated with the Blade element momentum models is studied. Sobol based 
 \end{abstract}
@@
 \section{Introduction}
-Aeroelastic models such as the Blade Element Momentum (BEM) models \cite{HandBook} play a critical role in the design, development and optimization of modern wind turbines.
+Aeroelastic models such as the Blade Element Momentum (BEM) models \cite{HandBook} play a critical role in the design, development, and optimization of modern wind turbines.
 A large number of BEM models have been developed to predict turbine responses such as the structural loads and power output \cite{Vorpahl2013}.
@@
 As a consequence of the strong model assumptions at the basis of BEM theory, the results from BEM codes can be subject to significant inaccuracies or uncertainties.
-For example, the effect of sheared inflow \cite{Madsen2012} is not naturally accounted for in the theory, and needs to be incorporated via correction terms.
+For example, the effect of sheared inflow \cite{Madsen2012} is not naturally accounted for in the theory and needs to be incorporated via correction terms.
 Other major model uncertainties in BEM models are for example the time constant in dynamic stall models, the wake correction factor, the tip loss model parameter, and the lift- and drag-polars used to compute local aerodynamic forces.
 Especially for increasing turbine sizes, these model parameters are foreseen to be not sufficiently accurate \cite{Sayed2019}.
-In other words, the uncertainty associated to the output of currently employed BEM models is rather large.
-% iIn order to arrive at robust BEM models that include an uncertainty estimate associated
+In other words, the uncertainty associated with the output of currently employed BEM models is rather large.
+% to arrive at robust BEM models that include an uncertainty estimate associated
 
@@
 
-Clearly, the main novelty of the work lies in the study of the effect of model uncertainties rather than external conditions or geometric uncertainties.
+The main novelty of the work lies in the study of the effect of model uncertainties rather than external conditions or geometric uncertainties.
 The current sensitivity analysis is part of the so-called WindTrue project, in which the long-term goal is the development of calibrated BEM models, that possess a quantified level of uncertainty.
@@
 
-%A number of parameters describing meteorological and operational conditions as well as manufacturing specifications are needed as inputs for BEM simulations.
+% Several parameters describing meteorological and operational conditions, as well as manufacturing specifications, are needed as inputs for BEM simulations.
 In this work, we seek to quantify the sensitivities of these input parameters for different turbine responses.
-In the past a number of sensitivity studies have been performed to understand the influence of input parameters on different turbine response, for e.g.
+In the past, several sensitivity studies have been performed to understand the influence of input parameters on different turbine response, e.g.
 \cite{moriarty2002effect, eggers2003wind, McKay2014, dykes2014sensitivity, Robertson2018}.
-Many studies have confirmed that the wind parameters especially the wind speed and wind speed standard deviation have the most influence on aerodynamic performance of turbines.
+Many studies have confirmed that wind parameters especially wind speed and wind speed standard deviation have the most influence on the aerodynamic performance of turbines.
 For example, in the case of an upstream turbine, the wind speed is most sensitive to power production.
@@
 %To compute parameter sensitivities we use the Sobol expansion approach, which decomposes the total variance of the quantity of interest into contributions from individual parameters and their combinations.
-To perform Sobol analysis, we use Matlab-based general purpose uncertainty quantification toolbox \texttt{UQLab} \cite{uqlab}.
+To perform Sobol analysis, we use Matlab-based general-purpose uncertainty quantification toolbox \texttt{UQLab} \cite{uqlab}.
 \texttt{UQLab}'s modular structure allows for easy integration with available BEM codes.
-For investigation the aeroelastic model of the  2MW NM80 turbine (with an 80m rotor) from the DANAERO project \cite{DANAERO} is utilized.
+For investigation, the aeroelastic model of the 2MW NM80 turbine (with an 80m rotor) from the DANAERO project \cite{DANAERO} is utilized.
 
@@
 \subsection{BEM model uncertainties}
-State-of-the-art wind-turbine models simulate the aeroelastic behavior of wind turbines by combining the concept of momentum conservation of the flow (BEM) with the equations of motion for the structure, possibly extended with hydrodynamics of the sea and control algorithms \cite{Vorpahl2013}.
+State-of-the-art wind-turbine models simulate the aeroelastic behavior of wind turbines by combining the concept of momentum conservation of the flow (BEM) with the equations of motion for the structure, possibly extended with the hydrodynamics of the sea and control algorithms \cite{Vorpahl2013}.
 In this work, we only concentrate on the first aspect, namely the prediction of flow and blade forces as given by the BEM method.
@@
 \begin{itemize}
-\item The use of (2D) \textit{airfoil polars}; uncertainty arises not only because the actual flow along the blade is 3D, but also because the 2D polars itself can be inaccurate, either when obtained from measurement or from 2D codes like \texttt{XFOIL};
+\item The use of (2D) \textit{airfoil polars}; uncertainty arises not only because the actual flow along the blade is 3D, but also because the 2D polars itself can be inaccurate, either when obtained from measurement or 2D codes like \texttt{XFOIL};
 %\item Sheared inflow
 \item The assumption of locally 2D flow along the blade is also invalid at the tip; this is typically corrected via \textit{a tip loss model};
-\item Three-dimensional flow effects also cause stall delay at the root; this is typically modelled via \textit{a dynamic stall model }and time constant;
+\item Three-dimensional flow effects also cause stall delay at the root; this is typically modeled via \textit{a dynamic stall model }and time constant;
 \item Yawed inflow is not naturally included in BEM and typically incorporated via \textit{a wake correction}.
@@
 %\item Dynamic wake - Oye time lag constant
-%\item Tower shadow constants – Potential flow / Downwind Empirical model
+%\item Tower shadow constants " Potential flow / Downwind Empirical model
 
@@
 We use Non-Uniform Rational Basis Splines (NURBS) \cite{rogers2000} to perturb the geometrical parameters of the turbine blade, such as the reference chord, twist or experimentally obtained drag-, lift-polar curves.
-The main advantage of using NURBS is that it provides great flexibility to approximate a large variety of curves with a limited number of control points which can be useful in reducing the number of uncertain parameter.
+The main advantage of using NURBS is that it provides great flexibility to approximate a large variety of curves with a limited number of control points which can be useful in reducing the number of uncertain parameters.
 %Further, the set of control points and knots can be directly manipulated to control the smoothness and curvature.
 Use of NURBS has already been exploited in many aerodynamics applications, for instance, design optimization of wind-turbine blades, see for e.g.
-\cite{Ribeiro2012,Bottasso2014} and to parametrize chord and twist curves \cite{Echeverria2017}.
+\cite{Bottasso2014,Ribeiro2012} and to parametrize chord and twist curves \cite{Echeverria2017}.
 In the following, we briefly outline the procedure used to parametrize curves using NURBS.
 \subsubsection{NURBS based perturbation}
-The value of NURBS curve at each location $x$ is computed using a weighted sum of $n$ basis functions (or B-splines):
+The value of NURBS curve $S(x)$ is computed using a weighted sum of $n$ basis functions (or B-splines):
 \begin{equation}\label{NURB_curve}
@@
 \end{equation}
-where $S(x)$ is the value of the NURBS curve at location $x$; $c_i$ is the weight of the $i$-th control point and $B_{i,p}(x)$ is the value of B-spline corresponding to the $i$-th control point at $x$ with the subscript $p$ denoting its polynomial degree.
+where $c_i$ is the weight of the $i$-th control point and $B_{i,p}(x)$ is the value of B-spline corresponding to the $i$-th control point at $x$ with the subscript $p$ denoting its polynomial degree.
 A sequence of non-decreasing parameters $X = \{x_1,x_2,...,x_m\}$ called the \textit{knot vector} determines the domain in which any control point is active with the number of elements in the knot vector is given by $m = n+p+1$.
-B-splines are non-zero only within a knot interval $[x_i,x_{i+1})$ for $i = 1, 2, ..., m-1$ and are zero elsewhere.
+A B-spline $B_{i,p}(x)$ is non-zero only within a knot interval $[x_i,x_{i+1})$ for $i = 1, 2, ..., m-1$ and are zero elsewhere.
 Therefore, only a few nearby control points are active when computing the NURBS curve \eqref{NURB_curve}.
@@
 The knots can be uniformly spaced or chosen heuristically, more advanced approaches for knot selection can be found in, for e.g.
-\cite{razdan1999knot,LI2005791}.
+\cite{LI200579,razdan1999knot}.
 Next we sample $S_{ref}(x)$ at $X$ and also compute $B_{i,p}(x_j), \text{ for }i, j = 1,2, ..., n$.
@@
 \ref{perturbed_samples} (a) - (b), respectively.
-Each control points are perturbed using a uniform random number with bounds given by $\pm5\%$ of the magnitude of control point.
-For lift and drag curve, all control points are perturbed using same uniform random number resulting in samples of the lift and drag curves as shown in Fig.
+Each control point are perturbed using a uniform random number with bounds given by $\pm5\%$ of the magnitude of a control point.
+For lift and drag curves, all control points are perturbed using the same uniform random number resulting in samples of the lift and drag curves as shown in Fig.
 \ref{perturbed_samples} (c) - (d), respectively.
 For simplicity, we only consider uniform distributions for perturbing the control points, however, we can also use other probability distributions for this.
-Also note that when using a uniform distribution with bounds given by relative magnitude of control points, we tend to obtain small perturbations for control points with values close to zero.
+Also note that when using a uniform distribution with bounds given by the relative magnitude of control points, we tend to obtain small perturbations for control points with values close to zero.
 \begin{figure}[h!]
 \centering
-\includegraphics[trim={0 1.6cm 0 1.5cm},clip, scale=0.67]{figure1.pdf}
-\caption{Random realization of chord, twist, lift- (Cl2) and drag-polars (Cd2).}
+\includegraphics[trim={0 0cm 0 0cm},clip, scale=.75]{basis_example.eps}
+\caption{}
 \label{perturbed_samples}
@@
 \section{Global sensitivity analysis}\label{sec:GSA}
-The objective of sensitivity analysis is to quantify the relative significance of individual inputs (or in combination) and how variations in input values affects the output of interest.
-In engineering, sensitivity analysis can be employed for a number of reasons: to determine the stability and robustness of a computational model with respect to input parameters, for simplification of stochastic models by fixing the insensitive parameters, and to guide data acquisition campaigns and experimental design to refine the data on sensitive parameters.
+The objective of sensitivity analysis is to quantify the relative significance of individual inputs (or in combination) and how variations in input values affect the output of interest.
+In engineering, sensitivity analysis can be employed for several reasons: to determine the stability and robustness of a computational model for input parameters, for simplification of stochastic models by fixing the insensitive parameters, and to guide data acquisition campaigns and experimental design to refine the data on sensitive parameters.
 Sensitivity analysis techniques can be classified as local and global methods, see \cite{RSmith}.
-In local sensitivity analysis, individual parameters are perturbed around their nominal values allowing for the description of output variability only in a small neighbourhood of nominal input values.
-Although, local approaches are widely employed due to their ease of implementation and low computational cost, they are unable to quantify global behavior of nonlinearly parametrized models such as aeroelastic models.
+In a local sensitivity analysis, individual parameters are perturbed around their nominal values allowing for the description of output variability only in a small neighborhood of nominal input values.
+Although, local approaches are widely employed due to their ease of implementation and low computational cost, they are unable to quantify the global behavior of nonlinearly parametrized models such as aeroelastic models.
 Global sensitivity approaches, on the other hand, consider the entire range of input values to compute output sensitivities.
@@
 \subsection{PCE based Sobol indices computation}
-The PCE $f^{K}(\mathbf{z})$ of the computational model $Y$ is defined as a weighted sum of multivariate polynomials in $\mathbf{z}$ \cite{RSmith}
-\begin{equation}
-Y = f^{K}(\mathbf{z}) = \sum_{|\mathbf{k}| = 0}^K y_{\mathbf{k}}\Psi_{\mathbf{k}}(\mathbf{z}),
+The PCE approximation $f^{K}(\mathbf{z})$ of the computational model $Y$ can be defined as a weighted sum of multivariate polynomials in $\mathbf{z}$ \cite{RSmith}
+\begin{equation}\label{eq:PCE}
+f(\mathbf{z}) \approx f^{K}(\mathbf{z}) = \sum_{|\mathbf{k}| = 0}^K w_{\mathbf{k}}\Psi_{\mathbf{k}}(\mathbf{z}),
 \end{equation}
 where $\mathbf{k}\in \mathbb{N}_0^M$ is a $M$-dimensional multi-index with the magnitude $|\mathbf{k}| = k_1+k_2 + ...
-+ k_M$, whereas $\Psi_{\mathbf{k}}(\mathbf{z})$ is the multivariate orthogonal polynomials that is computed using a tensor product of univariate polynomials $\psi_{k_i}^{(i)}(z_i)$
++ k_M$.
+The total number of terms $K$ in \eqref{eq:PCE} can be based on a standard truncation scheme where we use all polynomials in $M$ random variable with order less than or equal to $p$.
+More advanced truncation schemes such as the maximum interaction or a hyperbolic scheme can be employed depending on the application \cite{BlatmanThesis}.
+The multivariate polynomials $\Psi_{\mathbf{k}}(\mathbf{z})$ is orthogonal with respect to the joint probability distribution of random vector $\mathbf{z}$.
+The multivariate distribution is computed using a tensor product of univariate polynomials $\psi_{k_i}^{(i)}(z_i)$
 \begin{equation}
@@
 The choice of the univariate orthogonal polynomial $\psi_{k_i}$ depends on the type of random variable, $z_i$.
-For example, if for uniformly distributed random variable, the Legendre family of polynomials is used, see  \cite{Xiu2002} for details.
-Projection methods are typically employed for the calculation
+For example, for uniform random variable the Legendre family of polynomials is used, see \cite{Xiu2002} for details.
+Exploiting the orthogonality of $\Psi_{\mathbf{k}}(\mathbf{z})$, the coefficients can be computed using the projection approach:
+\begin{equation}\label{eq:projection}
+w_{\mathbf{k}} = \mathbb{E}[\Psi_{\mathbf{k}}(\mathbf{z})\cdot f^{K}(\mathbf{z})]
+\end{equation} 
+The above expectation can be computed using a quadrature based integration methods such as the Gaussian quadrature, sparse quadrature rules (for e.g., Smolyak's tensor grid), etc.
+Quadrature based integration methods suffer from curse of dimensionality where the number of integration points rises exponentially with an increase in the number of dimensions.
+In the current work where we have a fixed computational budget, methods that minimizes the approximation error for a given number of samples are more practical.
+Next, we briefly describe two such methods to compute PCE coefficients based on a fixed number of model evaluations.
+\subsubsection{Ordinary Least Square}
+Instead of using projection based method one can also use least squares based approaches to compute coefficients \cite{Berveiller2006}.
+The PCE expansion \eqref{eq:PCE} can be expressed as a vector-matrix product
+\begin{equation}\label{eq:LSM}
+Y = f^{K}(\mathbf{z}) +\varepsilon_K = \sum_{|\mathbf{k}| = 0}^K w_{\mathbf{k}}\Psi_{\mathbf{k}}(\mathbf{z}) + \varepsilon_K = \mathbf{w}^T\boldsymbol{\Psi}(\mathbf{z}) +\varepsilon_K,
+\end{equation} 
+where $\mathbf{w} = \{w_0,w_1,...,w_K\}^T$ is the coefficient vector, $\boldsymbol{\Psi}(\mathbf{z}) = \{\Psi_0(\mathbf{z}), \Psi_1(\mathbf{z}), ..., \Psi_K(\mathbf{z})\}^T$ is the matrix containing multivariate polynomials in $\mathbf{z}$ and $\varepsilon_K$ is the approximation error.
+The least-square minimization problem can then be formulated as:
+\begin{equation}\label{eq:PCE_OLS}
+\mathbf{\hat{w}} = \text{arg min }\mathbb{E}\left[\left(\mathbf{w}^T \boldsymbol{\Psi}(\mathbf{z}) - f(\mathbf{z})\right)^2\right]
+ \end{equation}
+The coefficient vector $\mathbf{\hat{w}}$ can be computed using the Ordinary Least Square (OLS) method.
+Given $N$ model evaluations $\mathbf{Y} = \{f(\mathbf{z}_1), f(\mathbf{z}_2), ..., f(\mathbf{z}_N)\}$, we can obtain the coefficients as:
+\begin{equation}\label{eq:OLS}
+\hat{\mathbf{w}} = (\mathbf{X}^T\mathbf{X})^{-1}\mathbf{X}^T\mathbf{Y},
+\end{equation}
+where $\mathbf{X}\in \mathbb{R}^{N\times K}$ is the design matrix where $\mathbf{X}_{i,j} = \Psi_{j}(\mathbf{z}_i)$, $i = 1,2, ..., N$ and $j = 0,1,2,..., K$.
+Although the OLS method allows to compute the coefficients using a limited number of model evaluations one the main drawback of this approach is that it may result in a PCE model consisting of high-order interaction terms that may lead to overfitting.
+For many engineering problems, low-order interactions between random input variables are more important.
+Therefore, the least square form in \eqref{eq:PCE_OLS} can be regularized such that a the minimization results in a low order sparse PCE model.
 
-Quadrature based integration methods suffer from curse of dimensionality where the number of integration points rises exponentially with an increase in the number of dimensions.
-For applications with a fixed computational budget methods that minimizes the approximation error with a given number of samples are usually more practical.
-Next, we briefly describe two such methods to compute PCE coefficients based on a given number of model evaluations.
+\subsubsection{Least Angle Regression}
+Least angle regression (LARS) is one of the several methods can be applied to compute sparse PCE models \cite{BlatmanThesis,BLATMAN2011}.
 
-\subsubsection{Ordinary Least Square}
-The 
-\subsubsection{Least Angle Regression (PCE\_LAR)}
+The LARS algorithm can be summarized in the following steps:
+
 \subsection{Sensitivity analysis with discrete random variable}
-%\section{Description of test case}
-%\section{Sensitivity analysis workflow}
+
 \section{Results}\label{sec:results}
@@
 \subsection{NM80}
- and the turbine is the 2MW NM80 turbine from the DANAERO project \cite{Troldborg2013} with a blade radius of 38.8m.
-The data for lift ($Cl$) and drag ($Cd$) polars are available at four locations along the blade radius at 11.87m, 17.82m, 28.97m and 35.53m.
-The reference value of polars are obtained from wind-tunnel experiment with 3D corrections.
+ and the turbine is the 2MW NM80 turbine from the DANAERO project \cite{Troldborg2013} with a blade radius of 38.8 m.
+The data for lift ($Cl$) and drag ($Cd$) polars are available at four locations along the blade radius at 11.87 m, 17.82 m, 28.97 m, and 35.53 m.
+The reference value of polars is obtained from the wind-tunnel experiment with 3D corrections.
 Random samples of chord, twist, lift- and drag-polars are obtained by perturbing the control points with a uniformly distributed random variable.
@@
 \ref{fig:samples} (a) - (b).
-For $Cl$ and $Cd$, all control points are perturbed using a same random number.
+For $Cl$ and $Cd$, all control points are perturbed using the same random number.
 
 \subsubsection{Verification of GSA method}
-Convergence study of UQLab with increasing number of samples for Quad, OLS, LARS.
+Convergence study of UQLab with an increasing number of samples for OLS and LARS.
+\begin{figure}[h!]
+\centering
+\includegraphics[trim={0cm 0cm 0 0cm},clip, scale=1.3]{nurbs_approx.eps}
+\caption{Random realization of chord, twist, lift- (Cl2) and drag-polars (Cd2).}
+\label{perturbed_samples}
+\end{figure}
+
+\begin{figure*}[h!]
+    \centering
+    \begin{subfigure}[t]{0.32\textwidth}
+        \centering
+        \includegraphics[trim={1cm 0cm 0 0cm},clip, scale=0.22]{convergence_chord1_4.eps}{a}
+        \caption{$M=4$}
+    \end{subfigure}%
+    ~ 
+    \begin{subfigure}[t]{0.32\textwidth}
+        \centering
+        \includegraphics[trim={1cm 0cm 0 0cm},clip, scale=0.22]{convergence_chord1_6.eps}
+        \caption{$M=6$}
+    \end{subfigure}
+      ~ 
+    \begin{subfigure}[t]{0.32\textwidth}
+        \centering
+        \includegraphics[trim={1cm 0cm 0 0cm},clip, scale=0.22]{convergence_chord1_8.eps}
+        \caption{ $M=8$}
+ \end{subfigure}
+    \caption{ Convergence of $PCE\_OLS$ and $PCE\_LARS$ with increasing number of uncertain inputs $M$ and number of samples $N$}
+\end{figure*}
+
+\begin{figure}[h!]
+\centering
+\includegraphics[trim={1cm 0cm 0 0cm},clip, scale=0.67]{figure1.pdf}
+\caption{Random realization of chord, twist, lift- (Cl2) and drag-polars (Cd2).}
+\label{perturbed_samples}
+\end{figure}
+
+\begin{figure}[h!]
+\centering
+\includegraphics[trim={1cm 0cm 0 0cm},clip, scale=0.4]{SA_Power_chord_twist_Cl_Cd.eps}
+\caption{}
+\label{perturbed_samples}
+\end{figure}
 

### 2020-01-22 sensitivity_analysis/cases/aero_module/initialize.m

@@ -1,5 +1,5 @@
 % Name of Matlab file representing the turbine data
-turbineName = 'AVATAR'; % 'NM80', 'AVATAR'
+turbineName = 'NM80'; % 'NM80', 'AVATAR'
 
 %% model description 
 % Name of Matlab file representing the model
@@ -25,11 +25,11 @@ NsamplesMC = [8 16 32];
 DegreesQuad = 1:3; %[1 2 3 4 5 6];
 
 % % for PCE-OLS:
-NsamplesOLS = [8 16]; % if not specified, the number of samples from Quad is taken
+NsamplesOLS = [8 16 32 64 128]; % if not specified, the number of samples from Quad is taken
 OLS_repeat = 1; % like MC_repeat
  
 % for PCE-LARS:
-NsamplesLARS = [8 16]; % if not specified, the number of samples from Quad is taken
+NsamplesLARS = [8 16 32 64 128]; % if not specified, the number of samples from Quad is taken
 LARS_repeat = 1; % like MC_repeat
 
 %% Assemble the Input.Marginal for sensitivity analysis by text comparison

### 2020-01-29 Papers/Torque_Paper/article.tex

@@
 \begin{abstract}
-Sensitivity analysis of model parameters associated with the Blade element momentum models is studied. Sobol based 
+A global sensitivity analysis of model parameters associated with the Blade Element Momentum (BEM) models is performed.
+The Sobol indices based on a sparse polynomial  expansion are used as a measure of global sensitivities.
+The sensitivity analysis workflow is developed using the uncertainty quantification toolbox \texttt{UQLab} that is integrated with the ECN's \texttt{Aero-Module} aeroelastic code.
+Sensitivity studies are performed on the NM80 wind turbine model from the DANAERO project.
 \end{abstract}
@@
 The goal of this work is to perform a systematic assessment of the uncertainty in model parameters in BEM models.
-We approach this by performing a global sensitivity analysis based on the Sobol expansion approach, which decomposes the total variance of the quantity of interest (model output) into contributions from individual parameters and their combinations, similar to \cite{Echeverria2017,Murcia2018,Rinker2016}.
+We approach this by performing a global sensitivity analysis based on the Sobol expansion approach, which decomposes the total variance of the quantity of interest (model output) into contributions from individual parameters and their combinations, similar to \cite{Echeverria2017,Murcia2018, Rinker2016a}.
 We will employ the uncertainty quantification toolbox \texttt{UQLab} \cite{uqlab}, which computes the Sobol indices based on a sparse polynomial chaos expansion.
@@
 \cite{Bottasso2014,Ribeiro2012} and to parametrize chord and twist curves \cite{Echeverria2017}.
-In the following, we briefly outline the procedure used to parametrize curves using NURBS.
+In the following, we briefly outline the procedure used to parametrize a curve using NURBS.
 \subsubsection{NURBS based perturbation}
@@
 where $c_i$ is the weight of the $i$-th control point and $B_{i,p}(x)$ is the value of B-spline corresponding to the $i$-th control point at $x$ with the subscript $p$ denoting its polynomial degree.
-A sequence of non-decreasing parameters $X = \{x_1,x_2,...,x_m\}$ called the \textit{knot vector} determines the domain in which any control point is active with the number of elements in the knot vector is given by $m = n+p+1$.
-A B-spline $B_{i,p}(x)$ is non-zero only within a knot interval $[x_i,x_{i+1})$ for $i = 1, 2, ..., m-1$ and are zero elsewhere.
-Therefore, only a few nearby control points are active when computing the NURBS curve \eqref{NURB_curve}.
-For a curve of polynomial degree $p$, there are $p+1$ active control points at any location $x$, for e.g., for a linear curve ($p=1$) only two adjacent control points are active at any location.
-Also, note that the number of B-splines are equal to the number of elements in $X$.
-Another useful property of the B-splines is that they are recursive in polynomial degree thus we can derive quadratic B-splines using linear B-splines, cubic from quadratic B-splines and so on.
+A sequence of non-decreasing parameters $X = \{x_1,x_2,...,x_m\}$ called the \textit{knot vector} determines the domain in which any control point is active.
+A control point $c_i$ is said to be active at a location when the corresponding B-spline $B_{i,p}(x)$ is non-zero at $x$.
+B-splines are recursive in polynomial degree thus we can derive quadratic B-splines using linear B-splines, cubic from quadratic B-splines and so on.
 Starting with B-spline of degree 0 defined as
@@
 \end{equation}
-we can derive higher order B-splines in the knot interval $[x_{i+p},x_{i+p+1})$ using the recurrence relation \cite{deBoor}:
+we can derive higher-order B-splines using the recurrence relation \cite{deBoor}:
 \begin{equation}\label{NURBS_recurrence}
@@
 \end{equation}
+Note that with increasing $p$ the interval in which a B-spline is non-zero also become larger.
+For a curve of polynomial degree $p$, there are $p+1$ active control points at any location $x$, for e.g., for a linear curve ($p=1$) only two adjacent control points are active at any location.
+Therefore, only a few nearby control points are active when computing the NURBS curve \eqref{NURB_curve}.
 These B-splines can be constructed efficiently using the de Boor's algorithm \cite{deBoor}.
-Note that to perform above computation, we need to pad the knot vector $X$ by repeating the first and last elements $p$ times each.
+To compute high-order B-splines in \eqref{NURBS_recurrence}, we need to pad the knot vector $X$ by repeating the first and last elements $p$ times each.
 In Fig.
-\ref{B-splines}, we show linear, quadratic and cubic splines for $x\in[0,1]$ [Also show the padding].
+\ref{basis_examples}, we show linear, quadratic and cubic splines for $x\in[0,1]$ with original knot vector $X = \{0,1/3,2/3,1\}$, padded accordingly with the curve order, for e.g., for linear B-splines, we use the knot vector $X = \{0,0,1/3,2/3,1,1\}$.
+\begin{figure}[h!]
+\centering
+\includegraphics[trim={0 0cm 0 0cm},clip, scale=.75]{basis_example.eps}
+\caption{Examples of B-splines with original knot vector $X = \{0,1/3,2/3,1\}$.}
+\label{basis_examples}
+\end{figure}
 
 Next, we describe steps to generate perturbed samples of chord from a given reference chord, $S_{ref}(x)$, using NURBS based parametrization.
-The first step is to approximate the given reference chord using a NURBS curve with a fixed degree $p$ and a limited number of knots $X =\{x_j\}_{j=0}^n$.
-The knots can be uniformly spaced or chosen heuristically, more advanced approaches for knot selection can be found in, for e.g.
-\cite{LI200579,razdan1999knot}.
+The first step is to approximate the given reference chord using a NURBS curve with a fixed degree $p$ and a limited number of knots $X =\{x_j\}_{j=1}^n$.
+The knots can be uniformly spaced or chosen heuristically, more advanced approaches for knot selection can be found in \cite{LI2005791,razdan1999knot}.
 Next we sample $S_{ref}(x)$ at $X$ and also compute $B_{i,p}(x_j), \text{ for }i, j = 1,2, ..., n$.
-Now we can formulate a linear system to solve for a set of control points $\mathbf{c}=\{c_i\}_{i=0}^n$ as:
+Now we can formulate a linear system to solve for a set of control points $\mathbf{c}=\{c_i\}_{i=1}^n$ as:
 \begin{equation}\label{nurbs_inversion}
@@
 \end{equation}
-where $\mathbf{S}_{ref}\in\mathbb{R}^{n}$ is a vector containing sampled values of $S_{ref}(x)$ and $\mathbf{B}\in \mathbb{R}^{n\times n}$ is a matrix with $j$-th row consisting of values of $n$ B-splines sampled at the location $x_j$.
+where $\mathbf{S}_{ref}\in\mathbb{R}^{n}$ is a vector containing sampled values of $S_{ref}(x)$ and $\mathbf{B}\in \mathbb{R}^{n\times n}$ is a matrix with $j$-th row consisting of values of $n$ B-splines sampled at location $x_j$.
 Once the control points are obtained, we can derive the approximate reference curve $S_n(x)$ using \eqref{NURB_curve}.
 In this work, we use uniform random variable to perturb the baseline control points $\mathbf{c}$.
-
-There are two ways to obtain perturbed curves either by independently perturbing each of the control points (local perturbation) or using the same random number to perturb all the control points (global perturbation).
-We use local perturbation for chord control points Ch1-Ch5 and twist control points Tw1-Tw6 control points, see Fig.
-\ref{perturbed_samples} (a) - (b), respectively.
-Each control point are perturbed using a uniform random number with bounds given by $\pm5\%$ of the magnitude of a control point.
-For lift and drag curves, all control points are perturbed using the same uniform random number resulting in samples of the lift and drag curves as shown in Fig.
-\ref{perturbed_samples} (c) - (d), respectively.
-For simplicity, we only consider uniform distributions for perturbing the control points, however, we can also use other probability distributions for this.
-Also note that when using a uniform distribution with bounds given by the relative magnitude of control points, we tend to obtain small perturbations for control points with values close to zero.
+In Fig.
+\ref{nurbs_approx}, we show a parameterization of  chord curve using 9 control points and second-order B-splines.
 \begin{figure}[h!]
 \centering
-\includegraphics[trim={0 0cm 0 0cm},clip, scale=.75]{basis_example.eps}
-\caption{}
-\label{perturbed_samples}
+\includegraphics[trim={0cm 0cm 0 0cm},clip, scale=1.3]{nurbs_approx.eps}
+\caption{Parametrization of chord using 9 control points and second-order B-splines.}
+\label{nurbs_approx}
 \end{figure}
+
+Finally, to obtain random samples of the chord curve, first we associate a probability distribution for each control point.
+And then we independently sample each control point and use them in \eqref{NURB_curve} to compute a random samples of chord curve.
 
@@
 Sensitivity analysis techniques can be classified as local and global methods, see \cite{RSmith}.
-In a local sensitivity analysis, individual parameters are perturbed around their nominal values allowing for the description of output variability only in a small neighborhood of nominal input values.
+In a local sensitivity analysis, individual parameters are perturbed around their nominal values allowing for the description of output variability only in a small neighbourhood of nominal input values.
 Although, local approaches are widely employed due to their ease of implementation and low computational cost, they are unable to quantify the global behavior of nonlinearly parametrized models such as aeroelastic models.
 Global sensitivity approaches, on the other hand, consider the entire range of input values to compute output sensitivities.
-Therefore, global sensitivity analysis is more suitable for aeroelastic models considered in this work.
+Therefore, a global sensitivity analysis is more suitable for aeroelastic models considered in this work.
 
@@
 \begin{equation}\label{sobol_decomp}
-f(z_1, z_2, ..., z_M) = f_0+\sum_{i=1}^M f_i(z_i) + \sum_{1\leq<i<j\leq M} f_{ij}(z_i,z_j) + ... + f_{1, 2, ..., M}(z_1, z_2, ..., z_M),
+f(z_1, z_2, ..., z_M) = f_0+\sum_{i=1}^M f_i(z_i) + \sum_{1\leq i<j\leq M} f_{ij}(z_i,z_j) + ... + f_{1, 2, ..., M}(z_1, z_2, ..., z_M),
 \end{equation}
 where the zeroth-order function $f_0$ is the mean response of $f$, the first-order univariate functions $f_i(z_i)$ quantify independent contributions due to each random input, second-order functions $f_{i,j}(z_i,z_j)$ represent the effect of interaction between $z_i$ and $z_j$ on the response $f$.
+Higher-order terms are interpreted in a similar manner.
 Note that the \eqref{sobol_decomp} is only valid for independent input parameters.
-Higher-order terms are interpreted in a similar manner.
+
 Formally, the zeroth-, first- and second-order terms are defined as:
@@
 where ${\mathcal{D}_{\mathbf{z}}}^{M-1} = [0,1]^{M-1}$ and the notation $\mathbf{z}_{\sim \{i,j\}}$ denotes the vector having all the components of $\mathbf{z}$ except in the set $\{i,j\}$.
-We only consider second order expansion of \eqref{sobol_decomp}, as for a number of applications, the second order expansion is usually sufficient and higher-order terms have negligible effect on the output response.
-In case of wind turbine, suppose $y$ represent the power output that depends on say 3 random inputs: windspeed $(z_1)$, wind standard deviation $(z_2)$ and RPM $(z_3)$, then $f_0$ represents the mean power output considering all the three random inputs, $f_1(z_1)$ represent the independent contribution of windspeed on the power and $f_{1,2}(z_1,z_2)$ quantify the interactions of wind speed and wind standard deviation on power.
+We only consider second-order expansion of \eqref{sobol_decomp}, as for a number of applications, the second-order expansion is usually sufficient and higher-order terms have negligible effect on the output response.
+In the case of wind turbine, suppose $y$ represent the power output that depends on say 3 random inputs: windspeed $(z_1)$, wind standard deviation $(z_2)$ and RPM $(z_3)$, then $f_0$ represents the mean power output considering all the three random inputs, $f_1(z_1)$ represent the independent contribution of windspeed on the power and $f_{1,2}(z_1,z_2)$ quantify the interactions of wind speed and wind standard deviation on power.
 Further, the total variance of $f(\mathbf{z})$ is defined as:
@@
 \end{equation}
-As the expansion in \eqref{sobol_decomp} is not unique, some orthogonality conditions must be imposed, see \cite{Rabitz1999,SOBOL2001271} which allows to express the total variance as
+As the expansion in \eqref{sobol_decomp} is not unique, some orthogonality conditions must be imposed  \cite{Rabitz1999,SOBOL2001271} which allows to express the total variance as
 \begin{equation}
@@
 \end{align}
-The first- and second-order Sobol indices are computed as
+The first- and second-order Sobol indices are then computed as
 \begin{equation}\label{sobol_ind}
@@
 \begin{equation}\label{eq:PCE}
-f(\mathbf{z}) \approx f^{K}(\mathbf{z}) = \sum_{|\mathbf{k}| = 0}^K w_{\mathbf{k}}\Psi_{\mathbf{k}}(\mathbf{z}),
+Y= f(\mathbf{z}) \approx f^{K}(\mathbf{z}) = \sum_{|\mathbf{k}| = 0}^K w_{\mathbf{k}}\Psi_{\mathbf{k}}(\mathbf{z}),
 \end{equation}
 where $\mathbf{k}\in \mathbb{N}_0^M$ is a $M$-dimensional multi-index with the magnitude $|\mathbf{k}| = k_1+k_2 + ...
-+ k_M$.
-The total number of terms $K$ in \eqref{eq:PCE} can be based on a standard truncation scheme where we use all polynomials in $M$ random variable with order less than or equal to $p$.
-More advanced truncation schemes such as the maximum interaction or a hyperbolic scheme can be employed depending on the application \cite{BlatmanThesis}.
-The multivariate polynomials $\Psi_{\mathbf{k}}(\mathbf{z})$ is orthogonal with respect to the joint probability distribution of random vector $\mathbf{z}$.
-The multivariate distribution is computed using a tensor product of univariate polynomials $\psi_{k_i}^{(i)}(z_i)$
++ k_M$ and $\Psi_{\mathbf{k}}(\mathbf{z})$ multivariate polynomial is computed using a tensor product of univariate polynomials $\psi_{k_i}^{(i)}(z_i)$
 \begin{equation}
@@
 \end{equation}
+where $k_i$ is the order of univariate polynomial.
+The multivariate polynomials $\Psi_{\mathbf{k}}(\mathbf{z})$ is orthogonal with respect to the joint probability distribution of random vector $\mathbf{z}$.
 The choice of the univariate orthogonal polynomial $\psi_{k_i}$ depends on the type of random variable, $z_i$.
 For example, for uniform random variable the Legendre family of polynomials is used, see \cite{Xiu2002} for details.
-Exploiting the orthogonality of $\Psi_{\mathbf{k}}(\mathbf{z})$, the coefficients can be computed using the projection approach:
+
+The total number of terms $K$ in \eqref{eq:PCE} can be based on a standard truncation scheme where we use all polynomials in $M$ random variable with order less than or equal to $p$.
+More advanced truncation schemes such as the maximum interaction or a hyperbolic scheme can also be employed depending on the application, see \cite{BlatmanThesis}.
+
+The PCE coefficient $w_{\mathbf{k}}$ corresponding to $\Psi_{\mathbf{k}}(\mathbf{z})$ can be computed using the projection approach:
 \begin{equation}\label{eq:projection}
-w_{\mathbf{k}} = \mathbb{E}[\Psi_{\mathbf{k}}(\mathbf{z})\cdot f^{K}(\mathbf{z})]
+w_{\mathbf{k}} = \mathbb{E}[\Psi_{\mathbf{k}}(\mathbf{z})\cdot f^{K}(\mathbf{z})].
 \end{equation} 
-The above expectation can be computed using a quadrature based integration methods such as the Gaussian quadrature, sparse quadrature rules (for e.g., Smolyak's tensor grid), etc.
+The above expectation can be computed using a quadrature based integration methods such as the Gaussian quadrature, sparse quadrature rules (e.g.
+Smolyak's tensor grid), etc.
 Quadrature based integration methods suffer from curse of dimensionality where the number of integration points rises exponentially with an increase in the number of dimensions.
@@
 where $\mathbf{w} = \{w_0,w_1,...,w_K\}^T$ is the coefficient vector, $\boldsymbol{\Psi}(\mathbf{z}) = \{\Psi_0(\mathbf{z}), \Psi_1(\mathbf{z}), ..., \Psi_K(\mathbf{z})\}^T$ is the matrix containing multivariate polynomials in $\mathbf{z}$ and $\varepsilon_K$ is the approximation error.
-The least-square minimization problem can then be formulated as:
+The least-squares minimization problem can then be formulated as:
 \begin{equation}\label{eq:PCE_OLS}
-\mathbf{\hat{w}} = \text{arg min }\mathbb{E}\left[\left(\mathbf{w}^T \boldsymbol{\Psi}(\mathbf{z}) - f(\mathbf{z})\right)^2\right]
+\mathbf{\overline{w}} = \text{arg min }\mathbb{E}\left[\left(\mathbf{w}^T \boldsymbol{\Psi}(\mathbf{z}) - f(\mathbf{z})\right)^2\right]
  \end{equation}
-The coefficient vector $\mathbf{\hat{w}}$ can be computed using the Ordinary Least Square (OLS) method.
+The coefficient vector $\mathbf{\overline{w}}$ can be computed using the Ordinary Least Squares (OLS) method.
 Given $N$ model evaluations $\mathbf{Y} = \{f(\mathbf{z}_1), f(\mathbf{z}_2), ..., f(\mathbf{z}_N)\}$, we can obtain the coefficients as:
 \begin{equation}\label{eq:OLS}
-\hat{\mathbf{w}} = (\mathbf{X}^T\mathbf{X})^{-1}\mathbf{X}^T\mathbf{Y},
+\overline{\mathbf{w}} = (\mathbf{X}^T\mathbf{X})^{-1}\mathbf{X}^T\mathbf{Y},
 \end{equation}
 where $\mathbf{X}\in \mathbb{R}^{N\times K}$ is the design matrix where $\mathbf{X}_{i,j} = \Psi_{j}(\mathbf{z}_i)$, $i = 1,2, ..., N$ and $j = 0,1,2,..., K$.
-Although the OLS method allows to compute the coefficients using a limited number of model evaluations one the main drawback of this approach is that it may result in a PCE model consisting of high-order interaction terms that may lead to overfitting.
-For many engineering problems, low-order interactions between random input variables are more important.
-Therefore, the least square form in \eqref{eq:PCE_OLS} can be regularized such that a the minimization results in a low order sparse PCE model.
+Although the OLS method allows to compute the coefficients using a limited number of model evaluations, one of the main drawback of this approach is that it may result in a PCE model that consists of high-order interaction terms that may lead to overfitting.
+For many engineering problems, low-order interactions between the input variables are more important.
+Therefore, the least squares form in \eqref{eq:PCE_OLS} can be regularized such that the minimization results is a low-order sparse PCE model.
 
 \subsubsection{Least Angle Regression}
-Least angle regression (LARS) is one of the several methods can be applied to compute sparse PCE models \cite{BlatmanThesis,BLATMAN2011}.
+Least angle regression (LARS) algorithm \cite{efron2004} is one of the several methods can be applied to compute sparse PCE models \cite{BlatmanThesis, BLATMAN2011}.
+The key idea is to find the best set of polynomials from a given candidate set that minimizes the regularized form of \eqref{eq:PCE_OLS}:
+\begin{equation}\label{regularized_LS}
+\mathbf{\overline{w}} = \text{arg min }\mathbb{E}\left[\left(\mathbf{w}^T \boldsymbol{\Psi}(\mathbf{z}) - f(\mathbf{z})\right)^2\right] + \lambda ||\mathbf{w}||_1.
+\end{equation}
+The LARS algorithm in the context of PCE can be summarized as follows.
+We start with all PCE coefficients for the candidate polynomials set to zero, and find the polynomial, say $\Psi_1(\mathbf{w})$ most correlated with the model evaluations.
+This polynomial enters the active set.
+Next, we take the largest step in the direction of $\Psi_1(\mathbf{w})$ until some other polynomial $\Psi_2(\mathbf{w})$ from the candidate set, has as much correlation with the current residual.
+The second polynomial is now also added to the active set.
+In the next step, we move both the coefficients in the active set towards their least-square values until a third polynomial from candidate set shows as much correlation as the current PCE model based on two polynomials show with the model evaluations.
+The number of iteration is given by the minimum of $\{K,N\}$.
+After every iteration a posteriori error, for instance, Leave-One-Out (LOO) cross-validation error is computed.
+The active set with the smallest LOO is then chosen as the best sparse PCE
+model.
 
-The LARS algorithm can be summarized in the following steps:
+One of the main advantage of the LARS algorithm is that it works well in cases when $K>>N$, i.e.
+the number of polynomials in the candidate set is much more larger than the number of model evaluations.
 
 \subsection{Sensitivity analysis with discrete random variable}
-
 \section{Results}\label{sec:results}
@@
 \subsection{NM80}
- and the turbine is the 2MW NM80 turbine from the DANAERO project \cite{Troldborg2013} with a blade radius of 38.8 m.
+As a reference wind turbine, we consider the 2MW NM80 turbine from the DANAERO project \cite{Troldborg2013} with a blade radius of 38.8 m.
 The data for lift ($Cl$) and drag ($Cd$) polars are available at four locations along the blade radius at 11.87 m, 17.82 m, 28.97 m, and 35.53 m.
+The lift and drag variables at these sections are numbered Cl1 - Cl4 and Cd1 - Cd4 respectively.
 The reference value of polars is obtained from the wind-tunnel experiment with 3D corrections.
 Random samples of chord, twist, lift- and drag-polars are obtained by perturbing the control points with a uniformly distributed random variable.
-For chord and twist, we independently perturb control points at different locations as shown in Fig.
-\ref{fig:samples} (a) - (b).
-For $Cl$ and $Cd$, all control points are perturbed using the same random number.
 
-\subsubsection{Verification of GSA method}
-Convergence study of UQLab with an increasing number of samples for OLS and LARS.
+There are two ways to obtain perturbed curves either by independently perturbing each of the control points (local perturbation) or using the same random number to perturb all the control points (global perturbation).
+We use local perturbation for chord control points Ch3 - Ch7 and twist control points Tw2 - Tw7 control points, see Fig.
+\ref{perturbed_samples} (a) - (b), respectively.
+Each control point are perturbed using a uniform random number with bounds given by $\pm5\%$ of the magnitude of a control point.
+For lift and drag curves, all control points are perturbed using the same uniform random number resulting in samples of the lift and drag curves as shown in Fig.
+\ref{perturbed_samples} (c) - (d), respectively.
+For simplicity, we only consider uniform distributions for perturbing the control points, however, we can also use other probability distributions for this purpose.
+Also note that when using a uniform distribution with bounds given by the relative magnitude of control points, we tend to obtain small perturbations for control points with values close to zero.
 \begin{figure}[h!]
 \centering
-\includegraphics[trim={0cm 0cm 0 0cm},clip, scale=1.3]{nurbs_approx.eps}
+\includegraphics[trim={0.5cm 0cm 0 0cm},clip, scale=0.67]{figure1.pdf}
 \caption{Random realization of chord, twist, lift- (Cl2) and drag-polars (Cd2).}
@@
 \end{figure}
+
+\subsubsection{Verification of GSA method}
+We perform a convergence study for OLS and LARS to determine which method converges fastest with increasing dimension.
+For this, we use the parametrized chord from Fig.
+\ref{perturbed_samples}(a) with 9 control points.
+In Fig.
+\ref{convergence_dim}, we show the convergence behavior of the two methods with increasing number of uncertain inputs, consisting of 4, 6 and 8 control points and for different number of model evaluations.
+We see that both methods scales well with increasing dimensions, however, $PCE\_LARS$ exhibits faster convergence than $PCE\_OLS$ in high dimensions.
 
@@
         \centering
-        \includegraphics[trim={1cm 0cm 0 0cm},clip, scale=0.22]{convergence_chord1_4.eps}{a}
-        \caption{$M=4$}
+        \includegraphics[trim={1cm 0cm 0 0cm}, clip, scale = 0.22]{convergence_chord1_4.eps}{a}
+        \caption{$M=4, Ch1-Ch4$}
     \end{subfigure}%
@@
         \includegraphics[trim={1cm 0cm 0 0cm},clip, scale=0.22]{convergence_chord1_6.eps}
-        \caption{$M=6$}
+        \caption{$M=6, Ch1-Ch6$}
     \end{subfigure}
@@
         \includegraphics[trim={1cm 0cm 0 0cm},clip, scale=0.22]{convergence_chord1_8.eps}
-        \caption{ $M=8$}
+        \caption{ $M=8, Ch1-Ch8$}
  \end{subfigure}
-    \caption{ Convergence of $PCE\_OLS$ and $PCE\_LARS$ with increasing number of uncertain inputs $M$ and number of samples $N$}
+    \caption{Convergence of $PCE\_OLS$ and $PCE\_LARS$ with increasing number of uncertain inputs $M$ and samples $N$.}
+\label{convergence_dim}
 \end{figure*}
-
-\begin{figure}[h!]
-\centering
-\includegraphics[trim={1cm 0cm 0 0cm},clip, scale=0.67]{figure1.pdf}
-\caption{Random realization of chord, twist, lift- (Cl2) and drag-polars (Cd2).}
-\label{perturbed_samples}
-\end{figure}
+\subsubsection{Sensitivity analysis results}
+In Fig.~\ref{sensitivity_analysis}, we show the total order Sobol indices as a measure of the sensitivity of the different geometric and model parameters on the power output.
+In this case, we observe the important result that the uncertainty in model parameters is at least as important as the uncertainty in geometric parameters.
+Furthermore, within the geometric parameters, the chord distribution shows a significantly higher sensitivity compared to the twist variables
 
@@
 \caption{}
-\label{perturbed_samples}
+\label{sensitivity_analysis}
 \end{figure}
@@
 \section{Conclusions}\label{sec:conclusions}
-
+We have shown how Sobol indices computed using sparse adaptive polynomial expansion can be used for high-dimensional global sensitivity analysis of both geometric and model uncertainties.
+The identified sensitive model parameters will be utilized in future work to develop calibrated BEM models with built-in uncertainty estimates as part of the \textit{WindTrue} project, and is connected to IEA Task 29.
 \section*{Acknowledgements}

### 2020-02-10 Papers/Torque_Paper/article.tex
Torque paper in IOP format
* changed Latex style
* updated sections 1-3

--- 2020-01-13 Papers/Torque_Paper/article.tex
@@
 \begin{abstract}
-Wind turbines are a complex multi-physics systems whose performance and life span is highly dependent on manufacturing, meteorological and operational factors.
-It is therefore important to quantify the impact of these parameters on the turbine response.
-We perform comprehensive sensitivity analysis of aeroelastic models typically employed to compute turbine response.
-This paper also seeks to quantify sensitivities due to manufacturing tolerance in the turbine blades.
-We propose a non-uniform rational basis splines approach to parametrize the geometry of the blade and further utilize this for quantifying geometric sensitivities at different radial locations.
-[Some important results]
+A global sensitivity analysis of model parameters associated with the Blade Element Momentum (BEM) models is performed.
+The Sobol indices based on a sparse polynomial  expansion are used as a measure of global sensitivities.
+The sensitivity analysis workflow is developed using the uncertainty quantification toolbox \texttt{UQLab} that is integrated with the ECN's \texttt{Aero-Module} aeroelastic code.
+Sensitivity studies are performed on the NM80 wind turbine model from the DANAERO project.
 \end{abstract}
@@
 
-\section{Introduction}
-Aeroelastic models such as the Blade Element Momentum (BEM) models \cite{HandBook} play a critical role in the design, development and optimization of modern wind turbines.
-A large number of BEM models have been developed to predict turbine responses such as the structural loads and power output \cite{Vorpahl2013}.
-
-As a consequence of the strong model assumptions at the basis of BEM theory, the results from BEM codes can be subject to significant inaccuracies or uncertainties.
-For example, the effect of sheared inflow \cite{Madsen2012} is not naturally accounted for in the theory, and needs to be incorporated via correction terms.
-Other major model uncertainties in BEM models are for example the time constant in dynamic stall models, the wake correction factor, the tip loss model parameter, and the lift- and drag-polars used to compute local aerodynamic forces.
-Especially for increasing turbine sizes, these model parameters  are foreseen to be not sufficiently accurate \cite{Sayed2019}.
-In other words, the uncertainty associated to the output of currently employed BEM models is rather large.
-% iIn order to arrive at robust BEM models that include an uncertainty estimate associated
-
-Recently, several papers have addressed the uncertainty in BEM model output by performing forward uncertainty propagation or sensitivity studies, e.g.\ \cite{Echeverria2017,Matthaus2017,Murcia2018,Robertson2018,Bos2019a}.
-In these studies, the focus in mainly on uncertainties in the external conditions (wind parameters) and/or uncertainty in the turbine specification (geometric parameters).
-Apart from understanding how uncertainty in the output is related to the different uncertain inputs, such sensitivity studies are very useful to reduce the number of parameters as needed for example in design optimization \cite{Echeverria2017}.
-However, the uncertainty in BEM model output as caused by uncertainty in the model formulation itself, e.g.\ through the values chosen for model parameters, has been given little attention (exceptions being the effect of aerodynamic properties studied in \cite{Bortolotti2019,Matthaus2017}).
-
-The goal of this work is to perform a systematic assessment of the uncertainty in model parameters in BEM models.
-We approach this by performing a global sensitivity analysis based on the Sobol expansion approach, which decomposes the total variance of the quantity of interest (model output) into contributions from individual parameters and their combinations, similar to \cite{Echeverria2017,Murcia2018,Rinker2016}.
-We will employ the uncertainty quantification toolbox \texttt{UQLab} \cite{uqlab}, which computes the Sobol indices based on a sparse polynomial chaos expansion.
-
-Clearly, the main novelty of the work lies in the study of the effect of model uncertainties rather than external conditions or geometric uncertainties.
-The current sensitivity analysis is part of the so-called WindTrue project, in which the long-term goal is the development of calibrated BEM models, that possess a quantified level of uncertainty.
-Anticipating on this goal, we are using the NM80 wind turbine model from the Tjaereborg wind farm, for which extensive measurement data is available through the DanAero project \cite{Troldborg2013} and IEA Task 29.
-The measurement data will enable the calibration step, and our sensitivity analysis will make clear which of the model parameters should be used in the calibration step.
-
-This paper is structured as follows.
-First, in section \ref{sec:model_description} we give a short description of the BEM model and associated uncertainties.
-In section \ref{sec:parameterization} the parameterization of the uncertainties is described, and in section \ref{sec:GSA} the global sensitivity analysis methodology.
-Section \ref{sec:results} discusses the results of the sensitivity analysis for the NM80 turbine test case, and conclusions follow in section \ref{sec:conclusions}.
-
-%A number of parameters describing meteorological and operational conditions as well as manufacturing specifications are needed as inputs for BEM simulations.
-In this work, we seek to quantify the sensitivities of these input parameters for different turbine responses.
-In the past a number of sensitivity studies have been performed to understand the influence of input parameters on different turbine response, for e.g.
-\cite{moriarty2002effect, eggers2003wind, McKay2014, dykes2014sensitivity, Robertson2018}.
-Many studies have confirmed that the wind parameters especially the wind speed and wind speed standard deviation have the most influence on aerodynamic performance of turbines.
-For example, in the case of an upstream turbine, the wind speed is most sensitive to power production.
-Furthermore, wind speed in combination with wind speed standard deviation has a large influence on the power production of turbines operating in the wake of an upstream turbine.
-Among operational factors, the rotor RPM is the most sensitive to power production.
-[More on parameter sensitivities on structural loads]
-%
-%[Sensitivity on the polars as novel element]
-%
-%[Mention this study is a part of IEA Task29 project]
-%
-%In this work, we also study the effect of manufacturing tolerances on the turbine response.
-Usually, there are some discrepancy between the  manufactured and nominally prescribed design of the turbine blade leading to a suboptimal performance.
-For BEM models, the turbine shape is described as a series of airfoils along the span of the blade where each airfoil shape is computed using the three quantities: chord length, thickness and twist.
-We perturb these three quantities to obtain a perturbed turbine blade and analyze with sections that have more influence on output quantities.
-%
-%To compute parameter sensitivities we use the Sobol expansion approach, which decomposes the total variance of the quantity of interest into contributions from individual parameters and their combinations.
-To perform Sobol analysis, we use Matlab-based general purpose uncertainty quantification toolbox \texttt{UQLab} \cite{uqlab}.
-\texttt{UQLab}'s modular structure allows for easy integration with available BEM codes.
-For investigation the aeroelastic model of the  2MW NM80 turbine (with an 80m rotor) from the DANAERO project \cite{DANAERO} is utilized.
-
-\section{Aeroelastic models}\label{sec:model_description}
-\subsection{BEM model uncertainties}
-State-of-the-art wind-turbine models simulate the aeroelastic behavior of wind turbines by combining the concept of momentum conservation of the flow (BEM) with the equations of motion for the structure, possibly extended with hydrodynamics of the sea and control algorithms \cite{Vorpahl2013}.
-In this work, we only concentrate on the first aspect, namely the prediction of flow and blade forces as given by the BEM method.
-The particular BEM code that we use is  ECN's \texttt{Aero-Module} \cite{Boorsma2012}.
-
-For this case of a rigid turbine, with a uniform inflow field, the main model uncertainties that arise in the BEM formulation that we will discuss in this work are the following \cite{Hansen1993}:
-\begin{itemize}
-\item The use of (2D) \textit{airfoil polars}; uncertainty arises not only because the actual flow along the blade is 3D, but also because the 2D polars itself can be inaccurate, either when obtained from measurement or from 2D codes like \texttt{XFOIL};
-%\item Sheared inflow
-\item The assumption of locally 2D flow along the blade is also invalid at the tip; this is typically corrected via \textit{a tip loss model};
-\item Three-dimensional flow effects also cause stall delay at the root; this is typically modelled via \textit{a dynamic stall model }and time constant;
-\item Yawed inflow is not naturally included in BEM and typically incorporated via \textit{a wake correction}.
-
-%\item Dynamic wake - Oye time lag constant
-%\item Tower shadow constants – Potential flow / Downwind Empirical model
-
-\end{itemize}
-
-%\subsection{Input parameters of aeroelastic models}
-%The main
-
-\section{Parametrization of uncertain inputs in the BEM model}\label{sec:parameterization}
-
-\subsection{Geometric uncertainty}
-We use Non-Uniform Rational Basis Splines (NURBS) \cite{rogers2000} to perturb the geometrical parameters of the turbine blade, such as the reference chord, twist or experimentally obtained drag-, lift-polar curves.
-The main advantage of using NURBS is that it provides great flexibility to approximate a large variety of curves with a limited number of control points which can be useful in reducing the number of uncertain parameter.
-%Further, the set of control points and knots can be directly manipulated to control the smoothness and curvature.
-Use of NURBS has already been exploited in many aerodynamics applications, for instance, design optimization of wind-turbine blades, see for e.g.
-\cite{Ribeiro2012,Bottasso2014} and to parametrize chord and twist curves \cite{Echeverria2017}.
-In the following, we briefly outline the procedure used to parametrize curves using NURBS.
-\subsubsection{NURBS based perturbation}
-The value of NURBS curve at each location $x$ is computed using a weighted sum of $n$ basis functions (or B-splines):
-\begin{equation}\label{NURB_curve}
-S(x)  = \sum_{i=1}^{n} c_i B_{i,p}(x),
-\end{equation}
-where $S(x)$ is the value of the NURBS curve at location $x$; $c_i$ is the weight of the $i$-th control point and $B_{i,p}(x)$ is the value of B-spline corresponding to the $i$-th control point at $x$ with the subscript $p$ denoting its polynomial degree.
-A sequence of non-decreasing parameters $X = \{x_1,x_2,...,x_m\}$ called the \textit{knot vector} determines the domain in which any control point is active with the number of elements in the knot vector is given by $m = n+p+1$.
-B-splines are non-zero only within a knot interval $[x_i,x_{i+1})$ for $i = 1, 2, ..., m-1$ and are zero elsewhere.
-Therefore, only a few nearby control points are active when computing the NURBS curve \eqref{NURB_curve}.
-For a curve of polynomial degree $p$, there are $p+1$ active control points at any location $x$, for e.g., for a linear curve ($p=1$) only two adjacent control points are active at any location.
-Also, note that the number of B-splines are equal to the number of elements in $X$.
-Another useful property of the B-splines is that they are recursive in polynomial degree thus we can derive quadratic B-splines using linear B-splines, cubic from quadratic B-splines and so on.
-Starting with B-spline of degree 0 defined as
-\begin{equation}\label{linearBspline}
-B_{i,0}(x) :=
-\begin{cases}
-1\quad x_i\leq x < x_{i+1},\quad i = 1,2,...,m,\\
-0\quad\text{elsewhere,}
-\end{cases} 
-\end{equation}
-we can derive higher order B-splines in the knot interval $[x_{i+p},x_{i+p+1})$ using the recurrence relation \cite{deBoor}:
-\begin{equation}\label{NURBS_recurrence}
-B_{i,p}(x) := \frac{x - x_i}{x_{i+p} - x_i}B_{i,p-1}(x) + \frac{x_{i+p+1}  -  x}{x_{i+p+1}  -  x_{i+1}}B_{i+1,p-1}(x),\quad p\geq1.
-\end{equation} 
-These B-splines can be constructed efficiently using the de Boor's algorithm \cite{deBoor}.
-Note that to perform above computation, we need to pad the knot vector $X$ by repeating the first and last elements $p$ times each.
-In Fig.
-\ref{B-splines}, we show linear, quadratic and cubic splines for $x\in[0,1]$ [Also show the padding].
-
-Next, we describe steps to generate perturbed samples of chord from a given reference chord, $S_{ref}(x)$, using NURBS based parametrization.
-The first step is to approximate the given reference chord using a NURBS curve with a fixed degree $p$ and a limited number of knots $X =\{x_j\}_{j=0}^n$.
-The knots can be uniformly spaced or chosen heuristically, more advanced approaches for knot selection can be found in, for e.g.
-\cite{razdan1999knot,LI2005791}.
-Next we sample $S_{ref}(x)$ at $X$ and also compute $B_{i,p}(x_j), \text{ for }i, j = 1,2, ..., n$.
-Now we can formulate a linear system to solve for a set of control points $\mathbf{c}=\{c_i\}_{i=0}^n$ as:
-\begin{equation}\label{nurbs_inversion}
-\mathbf{B}\mathbf{c} = \mathbf{S}_{ref},
-\end{equation}
-where $\mathbf{S}_{ref}\in\mathbb{R}^{n}$ is a vector containing sampled values of $S_{ref}(x)$ and $\mathbf{B}\in \mathbb{R}^{n\times n}$ is a matrix with $j$-th row consisting of values of $n$ B-splines sampled at the location $x_j$.
-Once the control points are obtained, we can derive the approximate reference curve $S_n(x)$ using \eqref{NURB_curve}.
-In this work, we use uniform random variable to perturb the baseline control points $\mathbf{c}$.
-
-There are two ways to obtain perturbed curves either by independently perturbing each of the control points (local perturbation) or using the same random number to perturb all the control points (global perturbation).
-We use local perturbation for chord control points Ch1-Ch5 and twist control points Tw1-Tw6 control points, see Fig.
-\ref{perturbed_samples} (a) - (b), respectively.
-Each control points are perturbed using a uniform random number with bounds given by $\pm5\%$ of the magnitude of control point.
-For lift and drag curve, all control points are perturbed using same uniform random number resulting in samples of the lift and drag curves as shown in Fig.
-\ref{perturbed_samples} (c) - (d), respectively.
-For simplicity, we only consider uniform distributions for perturbing the control points, however, we can also use other probability distributions for this.
-Also note that when using a uniform distribution with bounds given by relative magnitude of control points, we tend to obtain small perturbations for control points with values close to zero.
-\begin{figure}[h!]
-\centering
-\includegraphics[trim={0 1.6cm 0 1.5cm},clip, scale=0.67]{figure1.pdf}
-\caption{Random realization of chord, twist, lift- (Cl2) and drag-polars (Cd2).}
-\label{perturbed_samples}
-\end{figure}
-
-\section{Global sensitivity analysis}\label{sec:GSA}
-The objective of sensitivity analysis is to quantify the relative significance of individual inputs (or in combination) and how variations in input values affects the output of interest.
-In engineering, sensitivity analysis can be employed for a number of reasons: to determine the stability and robustness of a computational model with respect to input parameters, for simplification of stochastic models by fixing the insensitive parameters, and to guide data acquisition campaigns and experimental design to refine the data on sensitive parameters.
-Sensitivity analysis techniques can be classified as local and global methods, see \cite{RSmith}.
-In local sensitivity analysis, individual parameters are perturbed around their nominal values allowing for the description of output variability only in a small neighbourhood of nominal input values.
-Although, local approaches are widely employed due to their ease of implementation and low computational cost, they are unable to quantify global behavior of nonlinearly parametrized models such as aeroelastic models.
-Global sensitivity approaches, on the other hand, consider the entire range of input values to compute output sensitivities.
-Therefore, global sensitivity analysis is more suitable for aeroelastic models considered in this work.
-
-\subsection{Sobol analysis}
-We employ variance-based Sobol decomposition to perform global sensitivity analysis.
-This approach allows for quantification of  the relative importance of the input parameters on a scale of $[0,1]$ known as \emph{Sobol indices}.
-In a Sobol analysis, we express the total variance of the output in term of contributions from individual parameters and their combinations.
-For the ease of exposition, let us consider a nonlinear model:
-\begin{equation}\label{nonlinear_model}
-Y = f({\mathbf{z}}),
-\end{equation}
-where ${\mathbf{z}} = [z_1, z_2, ..., z_M]\in \mathcal{D}_{\mathbf{z}}\in \mathbb{R}^M$ is a vector of random  inputs.
-For simplicity, we assume input parameters are uniformly distributed i.e.
-$z_i \sim \mathcal{U}(0,1)$ and the support of input set is $\mathcal{D}_{\mathbf{z}}  =  [0,1]^M$ where $M$ is the total number input parameters.
-The Sobol decomposition is based on the hierarchical representation \cite{RSmith}:
-\begin{equation}\label{sobol_decomp}
-f(z_1, z_2, ..., z_M) = f_0+\sum_{i=1}^M f_i(z_i) + \sum_{1\leq<i<j\leq M} f_{ij}(z_i,z_j) + ... + f_{1, 2, ..., M}(z_1, z_2, ..., z_M),
-\end{equation}
-where the zeroth-order function $f_0$ is the mean response of $f$, the first-order univariate functions $f_i(z_i)$ quantify independent contributions due to each random input, second-order functions $f_{i,j}(z_i,z_j)$ represent the effect of interaction between $z_i$ and $z_j$ on the response $f$.
-Note that the \eqref{sobol_decomp} is only valid for independent input parameters.
-Higher-order terms are interpreted in a similar manner.
-Formally, the zeroth-, first- and second-order terms are defined as:
-\begin{align}\label{sobol_terms}
-f_0 &= \int_{\mathcal{D}_{\mathbf{z}}}f(\mathbf{z})d\mathbf{z},\\
-f_i(z_i) &= \int_{{\mathcal{D}_{\mathbf{z}}}^{M-1}}f(\mathbf{z})d\mathbf{z}_{\sim \{i\}} - f_0,\\
-f_{ij}(z_i,z_j) &= \int_{{\mathcal{D}_{\mathbf{z}}}^{M-2}}f(\mathbf{z})d\mathbf{z}_{\sim \{i,j\}} - f_i(z_i)  - f_j(z_j) - f_0,
-\end{align}
-where ${\mathcal{D}_{\mathbf{z}}}^{M-1} = [0,1]^{M-1}$ and the notation $\mathbf{z}_{\sim \{i,j\}}$ denotes the vector having all the components of $\mathbf{z}$ except in the set $\{i,j\}$.
-We only consider second order expansion of \eqref{sobol_decomp}, as for a number of applications, the second order expansion is usually sufficient and higher-order terms have negligible effect on the output response.
-In case of wind turbine, suppose $y$ represent the power output that depends on say 3 random inputs:  windspeed $(z_1)$, wind standard deviation $(z_2)$ and RPM $(z_3)$, then $f_0$ represents the mean power output considering all the three random inputs, $f_1(z_1)$ represent the independent contribution of windspeed on the power and $f_{1,2}(z_1,z_2)$ quantify the interactions of wind speed and wind standard deviation on power.
-Further, the total variance of $f(\mathbf{z})$ is defined as:
-\begin{equation}\label{tot_var}
-D = \text{Var}[f(\mathbf{z})] = \int_{\mathcal{D}_{\mathbf{z}}} f^2(\mathbf{z})d\mathbf{z} - f_0^2
-\end{equation}
-As the expansion in \eqref{sobol_decomp} is not unique, some orthogonality conditions must be imposed, see \cite{Rabitz1999,SOBOL2001271} which allows to express the total variance as
-\begin{equation}
-D = \sum_{i=1}^M D_i + \sum_{1\leq i<j\leq M} D_{ij},
-\end{equation}
-where the first- and second-order partial variances are defined as
-\begin{align}
-D_i &= \int_{0}^{1} f^2_i(z_i)dz_i, \label{partial_variance1}\\
-D_{ij} &= \int_{0}^{1} \int_{0}^{1} f^2_{ij}(z_i,z_j)dz_idz_j \label{partial_variance2}. 
-\end{align}
-The first- and second-order Sobol indices are computed as
-\begin{equation}\label{sobol_ind}
-S_i = \frac{D_i}{D}, \quad  S_{ij} = \frac{D_{ij}}{D}, \quad i,j=1,2, ..., M,
-\end{equation}
-and the sum of all indices satisfy
-\begin{equation}
-\sum_{i=1}^M S_i + \sum_{1\leq i<j\leq M} S_{ij} = 1.
-\end{equation}
-Finally, the total effect of the parameter $z_i$ on the output response $Y$ is quantified using the total sensitivity indices measure
-\begin{equation}
-S_{T_i} = S_i + \sum_{j=1}^M S_{ij}.
-\end{equation}
-The total sensitivity indices can be interpreted as an importance measure for the parameter $z_i$, therefore a large $S_{T_i}$ implies that $z_i$ has a strong influence on $Y$.
-The computation of $S_{T_i}$ requires the approximation of partial variances defined in \eqref{partial_variance1} - \eqref{partial_variance2}.
-Using the Monte Carlo method for computing these variance can be prohibitive for computationally expensive models $Y$.
-Polynomial Chaos Expansions (PCE) based approaches provides a more efficient alternative for sensitivity analysis.
-As shown in \cite{SUDRET2008964}, one can compute the Sobol indices analytically by post-processing the PCE coefficients.
-\subsection{PCE based Sobol indices computation}
-The PCE $f^{K}(\mathbf{z})$ of the computational model $Y$ is defined as a weighted sum of multivariate polynomials in $\mathbf{z}$ \cite{RSmith}
-\begin{equation}
-Y = f^{K}(\mathbf{z}) = \sum_{|\mathbf{k}| = 0}^K y_{\mathbf{k}}\Psi_{\mathbf{k}}(\mathbf{z}),
-\end{equation}
-where $\mathbf{k}\in  \mathbb{N}_0^M$ is a $M$-dimensional multi-index with the magnitude $|\mathbf{k}| = k_1+k_2 + ...
-+ k_M$, whereas $\Psi_{\mathbf{k}}(\mathbf{z})$ is the multivariate orthogonal polynomials that is computed using a tensor product of univariate polynomials $\psi_{k_i}^{(i)}(z_i)$
-\begin{equation}
-\Psi_{\mathbf{k}}(\mathbf{z}) := \prod_{i=1}^M\psi_{k_i}^{(i)}(z_i).
-\end{equation}
-The choice of the univariate orthogonal polynomial $\psi_{k_i}$ depends on the type of random variable, $z_i$.
-For example, if for uniformly distributed random variable, the Legendre family of polynomials is used, see  \cite{Xiu2002} for details.
-Projection methods are typically employed for the calculation
-
-Quadrature based integration methods suffer from curse of dimensionality where the number of integration points rises exponentially with an increase in the number of dimensions.
-For applications with a fixed computational budget methods that minimizes the approximation error with a given number of samples are usually more practical.
-Next, we briefly describe two such methods to compute PCE coefficients based on a given number of model evaluations.
-
-\subsubsection{Ordinary Least Square}
-The 
-\subsubsection{Least Angle Regression (PCE\_LAR)}
-\subsection{Sensitivity analysis with discrete random variable}
-%\section{Description of test case}
-%\section{Sensitivity analysis workflow}
-\section{Results}\label{sec:results}
-
-\subsection{NM80}
- and the turbine is the 2MW NM80 turbine from the DANAERO project \cite{Troldborg2013} with a blade radius of 38.8m.
-The data for lift ($Cl$) and drag ($Cd$) polars are available at four locations along the blade radius at 11.87m, 17.82m, 28.97m and 35.53m.
-The reference value of polars are obtained from wind-tunnel experiment with 3D corrections.
-Random samples of chord, twist, lift- and drag-polars are obtained by perturbing the control points with a uniformly distributed random variable.
-For chord and twist, we independently perturb control points at different locations as shown in Fig.
-\ref{fig:samples} (a) - (b).
-For $Cl$ and $Cd$, all control points are perturbed using a same random number.
-
-\subsubsection{Verification of GSA method}
-Convergence study of UQLab with increasing number of samples for Quad, OLS, LARS.
-
-\subsection{AVATAR}
-
-\section{Conclusions}\label{sec:conclusions}
+\input{content}
 

### 2020-02-10 Papers/Torque_Paper/article_IOP.tex
Torque paper in IOP format
* changed Latex style
* updated sections 1-3

--- 2020-02-10 Papers/Torque_Paper/article.tex
@@
-\documentclass[review]{elsarticle}
+\documentclass[a4paper]{jpconf}
 
-\journal{ }
+%\journal{ }
 
 % packages
-\usepackage{microtype}
-\usepackage{algorithm}
-\usepackage[noend]{algpseudocode}
-\usepackage[a4paper, total={6.2in, 9in}]{geometry}
-\renewcommand{\baselinestretch}{1.2} 
-\usepackage{color}
-\usepackage[x11names]{xcolor}
-\usepackage[utf8]{inputenc}
-\usepackage{amsmath}
-\usepackage{amssymb,latexsym}
+%\usepackage{microtype}
+%\usepackage{algorithm}
+%\usepackage[noend]{algpseudocode}
+%\usepackage[a4paper, total={6.2in, 9in}]{geometry}
+%\renewcommand{\baselinestretch}{1.2} 
+%\usepackage{color}
+%\usepackage[x11names]{xcolor}
+%\usepackage[utf8]{inputenc}
+%\usepackage{amsmath}
+%\usepackage{amssymb}
+\usepackage{latexsym}
 \usepackage{graphicx}
@@
 \usepackage{amsfonts}
-\usepackage{tikz}
-\usepackage{array}
+%\usepackage{tikz}
+% \usepackage{array}
 \usepackage{mathtools}
 \usepackage{relsize}
-\usepackage[T1]{fontenc}
+%\usepackage[T1]{fontenc}
 \usepackage{cases}
@@
 \usepackage{multirow}
-\usepackage{stix2}
-\usepackage[labelfont=bf,font=footnotesize]{caption}
+%\usepackage{stix2}
+%\usepackage[labelfont=bf,font=footnotesize]{caption}
 \usepackage{tabularx,ragged2e,booktabs}
-\usepackage{lineno}
-\modulolinenumbers[5]
-\usepackage[toc]{appendix}
+\usepackage{todonotes}
+%\usepackage{lineno}
+%\modulolinenumbers[5]
+%\usepackage[toc]{appendix}
 \usepackage{hyperref}
 
-%new commands
-\renewcommand{\figurename}{{Figure}}
-\newcolumntype{Y}{>{\centering\arraybackslash}X}
-\newtheorem{theorem}{Theorem}
-\newtheorem{remark}{Remark}[section]
-\numberwithin{equation}{section}
-\newcommand{\footremember}[2]{%
-  \footnote{#2}
-  \newcounter{#1}
-  \setcounter{#1}{\value{footnote}}%
-}
-\newcommand{\footrecall}[1]{%
-  \footnotemark[\value{#1}]%
-}
-
-\DeclareMathAlphabet{\mathpzc}{OT1}{pzc}{m}{it}
+%\DeclareMathAlphabet{\mathpzc}{OT1}{pzc}{m}{it}
 \newcommand{\lnorm}{\left|\left|}
@@
 \newcommand{\myreferences}{../references,../Mendeley_refs}
-\bibliographystyle{abbrv} %elsarticle-num}
+\bibliographystyle{iopart-num} %elsarticle-num}
 
@@
 
-\DeclareMathAlphabet{\mathpzc}{OT1}{pzc}{m}{it}
+%\DeclareMathAlphabet{\mathpzc}{OT1}{pzc}{m}{it}
 
@@
 
-\begin{frontmatter}
+%\author{Jacky Mucklow}
+
+%\address{Production Editor, \jpcs, \iopp, Dirac House, Temple Back, Bristol BS1~6BE, UK}
+
+%\ead{jacky.mucklow@iop.org}
+
 \title{A global sensitivity analysis of model uncertainty in aeroelastic wind turbine models}
 
-\address[cwi]{Centrum Wiskunde \& Informatica (CWI), Amsterdam, The Netherlands}
-%\address[aero]{Faculty of Aerospace Engineering, Delft University of Technology, Delft, The Netherlands.}
-%\author[cwi,aero]{Prashant~Kumar}
-\author[]{Prashant~Kumar}\ead{pkumar@cwi.nl}
+%$^{\mathsf{1}}$ Centrum Wiskunde \& Informatica, Science Park 123, Amsterdam, The Netherlands.
 
-\author{Benjamin~Sanderse}
+\author{P.~Kumar$^{1}$, B.~Sanderse$^{1}$, K.~Boorsma$^{2}$, M.~Caboni$^{2}$} \ead{b.sanderse@cwi.nl}
+\address{$^{1}$ Centrum Wiskunde \& Informatica (CWI), Amsterdam, The Netherlands}
+\address{$^{2}$ TNO, Westerduinweg 3, Petten, The Netherlands}
+
+%\author{Benjamin~Sanderse}
+%\textsl{•}
 
@@
 \end{abstract}
-\begin{keyword}
-BEM, global sensitivity analysis, model uncertainty, Sobol indices
-\end{keyword}
-\end{frontmatter}
 
-\linenumbers
+%\begin{keyword}
+%BEM, global sensitivity analysis, model uncertainty, Sobol indices
+%\end{keyword}
+%\end{frontmatter}
+
+%\linenumbers
 
@@
 
-\section*{Acknowledgements}
+%\section*{Acknowledgements}
 
-\newpage
+%\newpage
 \bibliography{\myreferences}
 
-\newpage
+%\newpage
 
-\appendix
-\input{literature_overview}
+%\appendix
+%\input{literature_overview}
 

### 2020-02-10 Papers/Torque_Paper/content.tex
Torque paper in IOP format
* changed Latex style
* updated sections 1-3

\section{Introduction}
Aeroelastic models such as the Blade Element Momentum (BEM) models \cite{HandBook} play a critical role in the design, development, and optimization of modern wind turbines. A large number of BEM models have been developed to predict turbine responses such as the structural loads and power output \cite{Vorpahl2013}.

As a consequence of the strong model assumptions at the basis of BEM theory, the results from BEM codes can be subject to significant inaccuracies or uncertainties. For example, the effect of sheared inflow \cite{Madsen2012} is not naturally accounted for in the theory and needs to be incorporated via correction terms. Other major model uncertainties in BEM models are for example the time constant in dynamic stall models, the wake correction factor, the tip loss model parameter, and the lift- and drag-polars used to compute local aerodynamic forces. Especially for increasing turbine sizes, these model parameters are foreseen to be not sufficiently accurate \cite{Sayed2019}. In other words, the uncertainty associated with the output of currently employed BEM models is rather large. 
% to arrive at robust BEM models that include an uncertainty estimate associated

Recently, several papers have addressed the uncertainty in BEM model output by performing forward uncertainty propagation or sensitivity studies, e.g.\ \cite{Echeverria2017,Matthaus2017,Murcia2018,Robertson2018,Bos2019a}. In these studies, the focus in mainly on uncertainties in the external conditions (wind parameters) and/or uncertainty in the turbine specification (geometric parameters). Apart from understanding how uncertainty in the output is related to the different uncertain inputs, such sensitivity studies are very useful to reduce the number of parameters as needed for example in design optimization \cite{Echeverria2017}. However, the uncertainty in BEM model output as caused by uncertainty in the model formulation itself, e.g.\ through the values chosen for model parameters, has been given little attention (exceptions being the effect of aerodynamic properties studied in \cite{Bortolotti2019,Matthaus2017})\todo{this is also what we do}.

The goal of this work is to perform a systematic assessment of the uncertainty in model parameters in BEM models. We approach this by performing a global sensitivity analysis based on the Sobol expansion approach, which decomposes the total variance of the quantity of interest (model output) into contributions from individual parameters and their combinations, similar to \cite{Echeverria2017,Murcia2018, Rinker2016a}. We will employ the uncertainty quantification toolbox \texttt{UQLab} \cite{uqlab}, which computes the Sobol indices based on a sparse polynomial chaos expansion.

The main novelty of the work lies in the study of the effect of model uncertainties rather than external conditions or geometric uncertainties. As such, the current sensitivity analysis is part of the so-called WindTrue project, in which the long-term goal is the development of calibrated BEM models, that possess a quantified level of uncertainty. Anticipating on this goal, we are using the NM80 wind turbine model from the Tjaereborg wind farm, for which extensive measurement data is available through the DanAero project \cite{Troldborg2013} and IEA Task 29. The measurement data will enable the calibration step, and our sensitivity analysis will make clear which of the model parameters should be used in this calibration step.

This paper is structured as follows. First, in section \ref{sec:model_description} we give a short description of the BEM model and associated geometric and model uncertainties. In section \ref{sec:parameterization} the parameterization of the uncertainties is described, and in section \ref{sec:GSA} the global sensitivity analysis methodology. Section \ref{sec:results} discusses the results of the sensitivity analysis for the NM80 turbine test case, and conclusions follow in section \ref{sec:conclusions}.

% Several parameters describing meteorological and operational conditions, as well as manufacturing specifications, are needed as inputs for BEM simulations. In this work, we seek to quantify the sensitivities of these input parameters for different turbine responses. In the past, several sensitivity studies have been performed to understand the influence of input parameters on different turbine response, e.g. \cite{moriarty2002effect, eggers2003wind, McKay2014, dykes2014sensitivity, Robertson2018}. Many studies have confirmed that wind parameters especially wind speed and wind speed standard deviation have the most influence on the aerodynamic performance of turbines. For example, in the case of an upstream turbine, the wind speed is most sensitive to power production. Furthermore, wind speed in combination with wind speed standard deviation has a large influence on the power production of turbines operating in the wake of an upstream turbine. Among operational factors, the rotor RPM is the most sensitive to power production. [More on parameter sensitivities on structural loads]
%
%[Sensitivity on the polars as novel element]
%
%[Mention this study is a part of IEA Task29 project]
%
%In this work, we also study the effect of manufacturing tolerances on the turbine response. Usually, there is some discrepancy between the manufactured and nominally prescribed design of the turbine blade leading to suboptimal performance. For BEM models, the turbine shape is described as a series of airfoils along the span of the blade where each airfoil shape is computed using the three quantities: chord length, thickness, and twist. We perturb these three quantities to obtain a perturbed turbine blade and analyze with sections that have more influence on output quantities. 
%
%To compute parameter sensitivities we use the Sobol expansion approach, which decomposes the total variance of the quantity of interest into contributions from individual parameters and their combinations. To perform Sobol analysis, we use Matlab-based general-purpose uncertainty quantification toolbox \texttt{UQLab} \cite{uqlab}. \texttt{UQLab}'s modular structure allows for easy integration with available BEM codes. For investigation, the aeroelastic model of the 2MW NM80 turbine (with an 80m rotor) from the DANAERO project \cite{DANAERO} is utilized.

\section{Aeroelastic model uncertainties}\label{sec:model_description}
%\subsection{BEM model uncertainties}
State-of-the-art wind-turbine models simulate the aeroelastic behavior of wind turbines by combining the concept of momentum conservation of the flow (BEM theory) with the equations of motion for the structure, possibly extended with the hydrodynamics of the sea and control algorithms \cite{Vorpahl2013}. In this work, we only concentrate on the first aspect, namely the prediction of flow and blade forces as given by the BEM method. The particular BEM code that we use is ECN's \texttt{Aero-Module} \cite{Boorsma2012}.

For the case of a rigid turbine, with a uniform inflow field, the main model uncertainties that arise in the BEM formulation are the following \cite{Hansen1993}:
\begin{itemize}
\item The use of (2D) \textit{airfoil polars}; uncertainty arises not only because the actual flow along the blade is 3D, but also because the 2D polars itself can be inaccurate, either when obtained from measurements or 2D codes like \texttt{XFOIL};
%\item Sheared inflow
\item The assumption of locally 2D flow along the blade is also invalid at the tip; this is typically corrected via \textit{a tip loss model};
\item Three-dimensional flow effects also cause stall delay at the root; this is typically modeled via \textit{a dynamic stall model} and associated time constant;
\item Yawed inflow is not naturally included in BEM and typically incorporated via \textit{a wake correction}.
%\item Dynamic wake - Oye time lag constant
%\item Tower shadow constants " Potential flow / Downwind Empirical model
\end{itemize}
Next to these model uncertainties we will consider the chord and twist distribution along the blade to be uncertain (geometric uncertainties).

%\subsection{Parameters selected in this study}
%The main

\section{Parametrization of uncertain inputs}\label{sec:parameterization}

%\subsection{Geometric uncertainties with NURBS curves}
The chord, twist, lift coefficient, and drag coefficient are functions of the radial distance along the blade, leading to a very high-dimensional number of uncertain parameters. In order to reduce this number, we parametrize the variation along the blade by using Non-Uniform Rational Basis Splines (NURBS) \cite{rogers2000}. In this way only a limited number of control points is needed to approximate a large variety of curves, so that the resulting number of uncertain parameters is relatively small. 
%To perturb the geometrical parameters of the turbine blade, such as the reference chord or twist. This parameterization will also be used to perturb lift-and drag polars. %Further, the set of control points and knots can be directly manipulated to control the smoothness and curvature.
NURBS have already been exploited in several aerodynamics applications, for instance, design optimization of wind-turbine blades, e.g.~\cite{Bottasso2014,Ribeiro2012}, and to parametrize chord and twist curves \cite{Echeverria2017}. In the following, we briefly outline the procedure used to parametrize a curve using NURBS.

%The uncertainty in a variable, e.g.\ the chord, is then expressed through uncertainties in the value of the control points.

\subsection{NURBS representation}
A NURBS curve $S(x)$ is expressed using a weighted sum of $n$ basis functions (or B-splines):
\begin{equation}\label{NURB_curve}
S(x) = \sum_{i=1}^{n} c_i B_{i,p}(x),
\end{equation}
where $c_i$ is the weight of the $i$-th control point and $B_{i,p}(x)$ is the value of the B-spline corresponding to the $i$-th control point at $x$, with the subscript $p$ denoting its polynomial degree.  In our case, $S(x)$ represent for example the chord as a function of the radial location along the blade. A sequence of non-decreasing parameters $X = \{x_1,x_2,...,x_m\}$, called the \textit{knot vector}, determines the domain in which any control point is active. A control point $c_i$ is said to be active at a location when the corresponding B-spline $B_{i,p}(x)$ is non-zero at $x$. B-splines are recursive in polynomial degree, thus we can derive quadratic B-splines using linear B-splines, cubic from quadratic B-splines and so on. Starting with a B-spline of degree 0 defined as 
\begin{equation}\label{linearBspline}
B_{i,0}(x) :=
\begin{cases}
1\quad x_i\leq x < x_{i+1},\quad i = 1,2,...,m,\\
0\quad\text{elsewhere,}
\end{cases} 
\end{equation}
we can derive higher-order B-splines using the recurrence relation \cite{deBoor}:
\begin{equation}\label{NURBS_recurrence}
B_{i,p}(x) := \frac{x - x_i}{x_{i+p} - x_i}B_{i,p-1}(x) + \frac{x_{i+p+1} - x}{x_{i+p+1} - x_{i+1}}B_{i+1,p-1}(x),\quad p\geq1.
\end{equation}
Note that with increasing $p$ the interval in which a B-spline is non-zero also become larger. For a curve of polynomial degree $p$, there are $p+1$ active control points at any location $x$. For example, for a linear curve ($p=1$) only two adjacent control points are active at any location. Therefore, only a few nearby control points are active when computing the NURBS curve \eqref{NURB_curve}. These B-splines can be constructed efficiently using the de Boor's algorithm \cite{deBoor}. To compute high-order B-splines in \eqref{NURBS_recurrence}, we need to pad the knot vector $X$ by repeating the first and last elements $p$ times each. In Fig. \ref{basis_examples}, we show linear, quadratic and cubic splines for $x\in[0,1]$ with original knot vector $X = \{0,1/3,2/3,1\}$, padded accordingly with the curve order, for e.g., for linear B-splines, we use the knot vector $X = \{0,0,1/3,2/3,1,1\}$.
\begin{figure}[h!]
\centering
\includegraphics[trim={0 0cm 0 0cm},clip, scale=.75]{basis_example.eps}
\caption{Examples of B-splines with original knot vector $X = \{0,1/3,2/3,1\}$.}
\label{basis_examples}
\end{figure}

In order to parameterize a given curve, e.g. the blade reference chord, we choose a degree $p$ and a number of knots $X =\{x_j\}_{j=1}^n$. The knots can be uniformly spaced or chosen heuristically (for advanced approaches for knot selection see\cite{LI2005791,razdan1999knot}). Next we sample $S_{ref}(x)$ at $X$ and also compute $B_{i,p}(x_j), \text{ for }i, j = 1,2, ..., n$. We then formulate a linear system and solve for the set of control points $\mathbf{c}=\{c_i\}_{i=1}^n$ as:
\begin{equation}\label{nurbs_inversion}
\mathbf{B}\mathbf{c} = \mathbf{S}_{ref},
\end{equation}
where $\mathbf{S}_{ref}\in\mathbb{R}^{n}$ is a vector containing sampled values of $S_{ref}(x)$ and $\mathbf{B}\in \mathbb{R}^{n\times n}$ is a matrix with the $j$-th row consisting of values of $n$ B-splines sampled at location $x_j$. Once the control points are obtained, the approximate reference curve $S_n(x)$ follows from \eqref{NURB_curve}. Fig.~\ref{nurbs_approx} shows a parameterization of the chord curve using $n=9$ control points and second-order ($p=2$) B-splines.
\begin{figure}[h!]
\centering
\includegraphics[trim={0cm 0cm 0 0cm},clip, scale=1.3]{nurbs_approx.eps}
\caption{Parametrization of chord using 9 control points and second-order B-splines.}
\label{nurbs_approx}
\end{figure}

\subsection{Perturbing the NURBS curves}
Given the reference values of $\mathbf{c}$, it becomes straightforward to generate perturbed curves: we sample a uniform random variable to perturb the baseline control points and use these to compute a perturbed curve $\tilde{S}(x)$:
\begin{equation}
\tilde{S} (x) = \sum_{i=1}^{n} \tilde{c}_{i} B_{i,p}(x),
\end{equation}
where
\begin{equation}
\tilde{c}_{i} = c_{i} (1 + \Delta_{i}),
\end{equation}
and the perturbation is randomly sampled as $\Delta_{i} \sim \mathcal{U}[-0.05,0.05]$.

\subsection{Model uncertainties}
The above framework is directly applicable to express the geometric uncertainty in twist- and chord distribution, with $c$ denoting either the value of the twist control points or of the chord control points. For model uncertainties, i.e.\ lift- and drag polars, \todo{Insert mathematical expression}

%
%Finally, to obtain random samples of the chord curve, first we associate a probability distribution for each control point. And then we independently sample each control point and use them in \eqref{NURB_curve} to compute a random samples of chord curve.

% Next, we describe steps to generate perturbed samples of chord from a given reference chord, $S_{ref}(x)$, using NURBS based parametrization.

\section{Global sensitivity analysis}\label{sec:GSA}
The objective of sensitivity analysis is to quantify the relative significance of individual inputs (or their combination) and how variations in input values affect the output of interest. In engineering, sensitivity analysis can be employed for several reasons: to determine the stability and robustness of a computational model for input parameters, for simplification of stochastic models by fixing the insensitive parameters, and to guide data acquisition campaigns and experimental design to refine the data on sensitive parameters. Sensitivity analysis techniques can be classified as local and global methods, see \cite{RSmith}. In a local sensitivity analysis, individual parameters are perturbed around their nominal values allowing for the description of output variability only in a small neighbourhood of nominal input values. Although, local approaches are widely employed due to their ease of implementation and low computational cost, they are unable to quantify the global behavior of nonlinearly parametrized models such as aeroelastic models. Global sensitivity approaches, on the other hand, consider the entire range of input values to compute output sensitivities. Therefore, a global sensitivity analysis is more suitable for aeroelastic models considered in this work.

\subsection{Sobol analysis}
We employ variance-based Sobol decomposition to perform global sensitivity analysis. This approach allows for quantification of the relative importance of the input parameters on a scale of $[0,1]$ known as \emph{Sobol indices}. In a Sobol analysis, we express the total variance of the output in term of contributions from individual parameters and their combinations. For the ease of exposition, let us consider a nonlinear model:
\begin{equation}\label{nonlinear_model}
Y = f({\mathbf{z}}),
\end{equation}
where ${\mathbf{z}} = [z_1, z_2, ..., z_M]\in \mathcal{D}_{\mathbf{z}}\in \mathbb{R}^M$ is a vector of random inputs. For simplicity, we assume input parameters are uniformly distributed i.e. $z_i \sim \mathcal{U}(0,1)$ and the support of input set is $\mathcal{D}_{\mathbf{z}} = [0,1]^M$ where $M$ is the total number input parameters. The Sobol decomposition is based on the hierarchical representation \cite{RSmith}:
\begin{equation}\label{sobol_decomp}
f(z_1, z_2, ..., z_M) = f_0+\sum_{i=1}^M f_i(z_i) + \sum_{1\leq i<j\leq M} f_{ij}(z_i,z_j) + ... + f_{1, 2, ..., M}(z_1, z_2, ..., z_M),
\end{equation}
where the zeroth-order function $f_0$ is the mean response of $f$, the first-order univariate functions $f_i(z_i)$ quantify independent contributions due to each random input, second-order functions $f_{i,j}(z_i,z_j)$ represent the effect of interaction between $z_i$ and $z_j$ on the response $f$.  Higher-order terms are interpreted in a similar manner. Note that the \eqref{sobol_decomp} is only valid for independent input parameters.

Formally, the zeroth-, first- and second-order terms are defined as:
\begin{align}\label{sobol_terms}
f_0 &= \int_{\mathcal{D}_{\mathbf{z}}}f(\mathbf{z})d\mathbf{z},\\
f_i(z_i) &= \int_{{\mathcal{D}_{\mathbf{z}}}^{M-1}}f(\mathbf{z})d\mathbf{z}_{\sim \{i\}} - f_0,\\
f_{ij}(z_i,z_j) &= \int_{{\mathcal{D}_{\mathbf{z}}}^{M-2}}f(\mathbf{z})d\mathbf{z}_{\sim \{i,j\}} - f_i(z_i) - f_j(z_j) - f_0,
\end{align}
where ${\mathcal{D}_{\mathbf{z}}}^{M-1} = [0,1]^{M-1}$ and the notation $\mathbf{z}_{\sim \{i,j\}}$ denotes the vector having all the components of $\mathbf{z}$ except in the set $\{i,j\}$. We only consider second-order expansion of \eqref{sobol_decomp}, as for a number of applications, the second-order expansion is usually sufficient and higher-order terms have negligible effect on the output response. In the case of wind turbine, suppose $y$ represent the power output that depends on say 3 random inputs: windspeed $(z_1)$, wind standard deviation $(z_2)$ and RPM $(z_3)$, then $f_0$ represents the mean power output considering all the three random inputs, $f_1(z_1)$ represent the independent contribution of windspeed on the power and $f_{1,2}(z_1,z_2)$ quantify the interactions of wind speed and wind standard deviation on power. Further, the total variance of $f(\mathbf{z})$ is defined as:
\begin{equation}\label{tot_var}
D = \text{Var}[f(\mathbf{z})] = \int_{\mathcal{D}_{\mathbf{z}}} f^2(\mathbf{z})d\mathbf{z} - f_0^2
\end{equation}
As the expansion in \eqref{sobol_decomp} is not unique, some orthogonality conditions must be imposed  \cite{Rabitz1999,SOBOL2001271} which allows to express the total variance as
\begin{equation}
D = \sum_{i=1}^M D_i + \sum_{1\leq i<j\leq M} D_{ij},
\end{equation}
where the first- and second-order partial variances are defined as
\begin{align}
D_i &= \int_{0}^{1} f^2_i(z_i)dz_i, \label{partial_variance1}\\
D_{ij} &= \int_{0}^{1} \int_{0}^{1} f^2_{ij}(z_i,z_j)dz_idz_j \label{partial_variance2}. 
\end{align}
The first- and second-order Sobol indices are then computed as
\begin{equation}\label{sobol_ind}
S_i = \frac{D_i}{D}, \quad S_{ij} = \frac{D_{ij}}{D}, \quad i,j=1,2, ..., M,
\end{equation}
and the sum of all indices satisfy
\begin{equation}
\sum_{i=1}^M S_i + \sum_{1\leq i<j\leq M} S_{ij} = 1.
\end{equation}
Finally, the total effect of the parameter $z_i$ on the output response $Y$ is quantified using the total sensitivity indices measure
\begin{equation}
S_{T_i} = S_i + \sum_{j=1}^M S_{ij}.
\end{equation}
The total sensitivity indices can be interpreted as an importance measure for the parameter $z_i$, therefore a large $S_{T_i}$ implies that $z_i$ has a strong influence on $Y$. The computation of $S_{T_i}$ requires the approximation of partial variances defined in \eqref{partial_variance1} - \eqref{partial_variance2}. Using the Monte Carlo method for computing these variance can be prohibitive for computationally expensive models $Y$. Polynomial Chaos Expansions (PCE) based approaches provides a more efficient alternative for sensitivity analysis. As shown in \cite{SUDRET2008964}, one can compute the Sobol indices analytically by post-processing the PCE coefficients.
\subsection{PCE based Sobol indices computation}
The PCE approximation $f^{K}(\mathbf{z})$ of the computational model $Y$ can be defined as a weighted sum of multivariate polynomials in $\mathbf{z}$ \cite{RSmith}
\begin{equation}\label{eq:PCE}
Y= f(\mathbf{z}) \approx f^{K}(\mathbf{z}) = \sum_{|\mathbf{k}| = 0}^K w_{\mathbf{k}}\Psi_{\mathbf{k}}(\mathbf{z}),
\end{equation}
where $\mathbf{k}\in \mathbb{N}_0^M$ is a $M$-dimensional multi-index with the magnitude $|\mathbf{k}| = k_1+k_2 + ... + k_M$ and $\Psi_{\mathbf{k}}(\mathbf{z})$ multivariate polynomial is computed using a tensor product of univariate polynomials $\psi_{k_i}^{(i)}(z_i)$
\begin{equation}
\Psi_{\mathbf{k}}(\mathbf{z}) := \prod_{i=1}^M\psi_{k_i}^{(i)}(z_i).
\end{equation}
where $k_i$ is the order of univariate polynomial. The multivariate polynomials $\Psi_{\mathbf{k}}(\mathbf{z})$ is orthogonal with respect to the joint probability distribution of random vector $\mathbf{z}$. The choice of the univariate orthogonal polynomial $\psi_{k_i}$ depends on the type of random variable, $z_i$. For example, for uniform random variable the Legendre family of polynomials is used, see \cite{Xiu2002} for details.

The total number of terms $K$ in \eqref{eq:PCE} can be based on a standard truncation scheme where we use all polynomials in $M$ random variable with order less than or equal to $p$. More advanced truncation schemes such as the maximum interaction or a hyperbolic scheme can also be employed depending on the application, see \cite{BlatmanThesis}.

The PCE coefficient $w_{\mathbf{k}}$ corresponding to $\Psi_{\mathbf{k}}(\mathbf{z})$ can be computed using the projection approach:
\begin{equation}\label{eq:projection}
w_{\mathbf{k}} = \mathbb{E}[\Psi_{\mathbf{k}}(\mathbf{z})\cdot f^{K}(\mathbf{z})].
\end{equation} 
The above expectation can be computed using a quadrature based integration methods such as the Gaussian quadrature, sparse quadrature rules (e.g. Smolyak's tensor grid), etc. Quadrature based integration methods suffer from curse of dimensionality where the number of integration points rises exponentially with an increase in the number of dimensions. In the current work where we have a fixed computational budget, methods that minimizes the approximation error for a given number of samples are more practical. Next, we briefly describe two such methods to compute PCE coefficients based on a fixed number of model evaluations.
\subsubsection{Ordinary Least Square}
Instead of using projection based method one can also use least squares based approaches to compute coefficients \cite{Berveiller2006}. The PCE expansion \eqref{eq:PCE} can be expressed as a vector-matrix product
\begin{equation}\label{eq:LSM}
Y = f^{K}(\mathbf{z}) +\varepsilon_K = \sum_{|\mathbf{k}| = 0}^K w_{\mathbf{k}}\Psi_{\mathbf{k}}(\mathbf{z}) + \varepsilon_K = \mathbf{w}^T\boldsymbol{\Psi}(\mathbf{z}) +\varepsilon_K,
\end{equation} 
where $\mathbf{w} = \{w_0,w_1,...,w_K\}^T$ is the coefficient vector, $\boldsymbol{\Psi}(\mathbf{z}) = \{\Psi_0(\mathbf{z}), \Psi_1(\mathbf{z}), ..., \Psi_K(\mathbf{z})\}^T$ is the matrix containing multivariate polynomials in $\mathbf{z}$ and $\varepsilon_K$ is the approximation error. The least-squares minimization problem can then be formulated as:
\begin{equation}\label{eq:PCE_OLS}
\mathbf{\overline{w}} = \text{arg min }\mathbb{E}\left[\left(\mathbf{w}^T \boldsymbol{\Psi}(\mathbf{z}) - f(\mathbf{z})\right)^2\right]
 \end{equation}
The coefficient vector $\mathbf{\overline{w}}$ can be computed using the Ordinary Least Squares (OLS) method. Given $N$ model evaluations $\mathbf{Y} = \{f(\mathbf{z}_1), f(\mathbf{z}_2), ..., f(\mathbf{z}_N)\}$, we can obtain the coefficients as:
\begin{equation}\label{eq:OLS}
\overline{\mathbf{w}} = (\mathbf{X}^T\mathbf{X})^{-1}\mathbf{X}^T\mathbf{Y},
\end{equation}
where $\mathbf{X}\in \mathbb{R}^{N\times K}$ is the design matrix where $\mathbf{X}_{i,j} = \Psi_{j}(\mathbf{z}_i)$, $i = 1,2, ..., N$ and $j = 0,1,2,..., K$. Although the OLS method allows to compute the coefficients using a limited number of model evaluations, one of the main drawback of this approach is that it may result in a PCE model that consists of high-order interaction terms that may lead to overfitting. For many engineering problems, low-order interactions between the input variables are more important. Therefore, the least squares form in \eqref{eq:PCE_OLS} can be regularized such that the minimization results is a low-order sparse PCE model.

\subsubsection{Least Angle Regression}
Least angle regression (LARS) algorithm \cite{efron2004} is one of the several methods can be applied to compute sparse PCE models \cite{BlatmanThesis, BLATMAN2011}. The key idea is to find the best set of polynomials from a given candidate set that minimizes the regularized form of \eqref{eq:PCE_OLS}:
\begin{equation}\label{regularized_LS}
\mathbf{\overline{w}} = \text{arg min }\mathbb{E}\left[\left(\mathbf{w}^T \boldsymbol{\Psi}(\mathbf{z}) - f(\mathbf{z})\right)^2\right] + \lambda ||\mathbf{w}||_1.
\end{equation}
The LARS algorithm in the context of PCE can be summarized as follows. We start with all PCE coefficients for the candidate polynomials set to zero, and find the polynomial, say $\Psi_1(\mathbf{w})$ most correlated with the model evaluations. This polynomial enters the active set. Next, we take the largest step in the direction of $\Psi_1(\mathbf{w})$ until some other polynomial $\Psi_2(\mathbf{w})$ from the candidate set, has as much correlation with the current residual. The second polynomial is now also added to the active set. In the next step, we move both the coefficients in the active set towards their least-square values until a third polynomial from candidate set shows as much correlation as the current PCE model based on two polynomials show with the model evaluations. The number of iteration is given by the minimum of $\{K,N\}$. After every iteration a posteriori error, for instance, Leave-One-Out (LOO) cross-validation error is computed. The active set with the smallest LOO is then chosen as the best sparse PCE
model.

One of the main advantage of the LARS algorithm is that it works well in cases when $K>>N$, i.e. the number of polynomials in the candidate set is much more larger than the number of model evaluations.

\subsection{Sensitivity analysis with discrete random variable}
\section{Results}\label{sec:results}

\subsection{NM80}
As a reference wind turbine, we consider the 2MW NM80 turbine from the DANAERO project \cite{Troldborg2013} with a blade radius of 38.8 m. The data for lift ($Cl$) and drag ($Cd$) polars are available at four locations along the blade radius at 11.87 m, 17.82 m, 28.97 m, and 35.53 m. The lift and drag variables at these sections are numbered Cl1 - Cl4 and Cd1 - Cd4 respectively. The reference value of polars is obtained from the wind-tunnel experiment with 3D corrections. Random samples of chord, twist, lift- and drag-polars are obtained by perturbing the control points with a uniformly distributed random variable.

There are two ways to obtain perturbed curves either by independently perturbing each of the control points (local perturbation) or using the same random number to perturb all the control points (global perturbation). We use local perturbation for chord control points Ch3 - Ch7 and twist control points Tw2 - Tw7 control points, see Fig. \ref{perturbed_samples} (a) - (b), respectively. Each control point are perturbed using a uniform random number with bounds given by $\pm5\%$ of the magnitude of a control point. For lift and drag curves, all control points are perturbed using the same uniform random number resulting in samples of the lift and drag curves as shown in Fig. \ref{perturbed_samples} (c) - (d), respectively. For simplicity, we only consider uniform distributions for perturbing the control points, however, we can also use other probability distributions for this purpose. Also note that when using a uniform distribution with bounds given by the relative magnitude of control points, we tend to obtain small perturbations for control points with values close to zero. 
\begin{figure}[h!]
\centering
\includegraphics[trim={0.5cm 0cm 0 0cm},clip, scale=0.67]{figure1.pdf}
\caption{Random realization of chord, twist, lift- (Cl2) and drag-polars (Cd2).}
\label{perturbed_samples}
\end{figure}

\subsubsection{Verification of GSA method}
We perform a convergence study for OLS and LARS to determine which method converges fastest with increasing dimension. For this, we use the parametrized chord from Fig. \ref{perturbed_samples}(a) with 9 control points. In Fig. \ref{convergence_dim}, we show the convergence behavior of the two methods with increasing number of uncertain inputs, consisting of 4, 6 and 8 control points and for different number of model evaluations. We see that both methods scales well with increasing dimensions, however, $PCE\_LARS$ exhibits faster convergence than $PCE\_OLS$ in high dimensions.

\begin{figure*}[h!]
    \centering
    \begin{subfigure}[t]{0.32\textwidth}
        \centering
        \includegraphics[trim={1cm 0cm 0 0cm}, clip, scale = 0.22]{convergence_chord1_4.eps}{a}
        \caption{$M=4, Ch1-Ch4$}
    \end{subfigure}%
    ~ 
    \begin{subfigure}[t]{0.32\textwidth}
        \centering
        \includegraphics[trim={1cm 0cm 0 0cm},clip, scale=0.22]{convergence_chord1_6.eps}
        \caption{$M=6, Ch1-Ch6$}
    \end{subfigure}
      ~ 
    \begin{subfigure}[t]{0.32\textwidth}
        \centering
        \includegraphics[trim={1cm 0cm 0 0cm},clip, scale=0.22]{convergence_chord1_8.eps}
        \caption{ $M=8, Ch1-Ch8$}
 \end{subfigure}
    \caption{Convergence of $PCE\_OLS$ and $PCE\_LARS$ with increasing number of uncertain inputs $M$ and samples $N$.}
\label{convergence_dim}
\end{figure*}
\subsubsection{Sensitivity analysis results}
In Fig.~\ref{sensitivity_analysis}, we show the total order Sobol indices as a measure of the sensitivity of the different geometric and model parameters on the power output. In this case, we observe the important result that the uncertainty in model parameters is at least as important as the uncertainty in geometric parameters. Furthermore, within the geometric parameters, the chord distribution shows a significantly higher sensitivity compared to the twist variables

\begin{figure}[h!]
\centering
\includegraphics[trim={1cm 0cm 0 0cm},clip, scale=0.4]{SA_Power_chord_twist_Cl_Cd.eps}
\caption{}
\label{sensitivity_analysis}
\end{figure}

\subsection{AVATAR}

\section{Conclusions}\label{sec:conclusions}
We have shown how Sobol indices computed using sparse adaptive polynomial expansion can be used for high-dimensional global sensitivity analysis of both geometric and model uncertainties. The identified sensitive model parameters will be utilized in future work to develop calibrated BEM models with built-in uncertainty estimates as part of the \textit{WindTrue} project, and is connected to IEA Task 29.

### 2020-02-10 Papers/Torque_Paper/content.tex
updates section 4-5

@@
 where
-\begin{equation}
+\begin{equation}\label{eqn:perturb_c}
 \tilde{c}_{i} = c_{i} (1 + \Delta_{i}),
@@
 The objective of sensitivity analysis is to quantify the relative significance of individual inputs (or their combination) and how variations in input values affect the output of interest.
-In engineering, sensitivity analysis can be employed for several reasons: to determine the stability and robustness of a computational model for input parameters, for simplification of stochastic models by fixing the insensitive parameters, and to guide data acquisition campaigns and experimental design to refine the data on sensitive parameters.
-Sensitivity analysis techniques can be classified as local and global methods, see \cite{RSmith}.
-In a local sensitivity analysis, individual parameters are perturbed around their nominal values allowing for the description of output variability only in a small neighbourhood of nominal input values.
-Although, local approaches are widely employed due to their ease of implementation and low computational cost, they are unable to quantify the global behavior of nonlinearly parametrized models such as aeroelastic models.
+Sensitivity analysis techniques can be classified as local and global \cite{RSmith}.
+In a local sensitivity analysis, individual parameters are perturbed around their nominal values allowing for the description of output variability only in a small neighbourhood around the nominal input values.
+Although local approaches are widely employed due to their ease of implementation and low computational cost, they are unable to quantify the global behaviour of nonlinearly parametrized models such as aeroelastic models.
 Global sensitivity approaches, on the other hand, consider the entire range of input values to compute output sensitivities.
-Therefore, a global sensitivity analysis is more suitable for aeroelastic models considered in this work.
+Therefore, a global sensitivity analysis is more suitable for the aeroelastic models considered in this work.
+%In engineering, sensitivity analysis can be employed for several reasons: to determine the stability and robustness of a computational model for input parameters, for simplification of stochastic models by fixing the insensitive parameters, and to guide data acquisition campaigns and experimental design to refine the data on sensitive parameters.
 
 \subsection{Sobol analysis}
-We employ variance-based Sobol decomposition to perform global sensitivity analysis.
+We employ the variance-based Sobol decomposition to perform global sensitivity analysis.
 This approach allows for quantification of the relative importance of the input parameters on a scale of $[0,1]$ known as \emph{Sobol indices}.
 In a Sobol analysis, we express the total variance of the output in term of contributions from individual parameters and their combinations.
-For the ease of exposition, let us consider a nonlinear model:
+Consider a generic BEM model $f(\mathbf{z})$, depending on uncertain parameters $\mathbf{z}$ and returning a quantity of interest $Y$ (e.g.\ power, loads, moments):
 \begin{equation}\label{nonlinear_model}
@@
 \end{equation}
-where ${\mathbf{z}} = [z_1, z_2, ..., z_M]\in \mathcal{D}_{\mathbf{z}}\in \mathbb{R}^M$ is a vector of random inputs.
-For simplicity, we assume input parameters are uniformly distributed i.e.
-$z_i \sim \mathcal{U}(0,1)$ and the support of input set is $\mathcal{D}_{\mathbf{z}} = [0,1]^M$ where $M$ is the total number input parameters.
-The Sobol decomposition is based on the hierarchical representation \cite{RSmith}:
+where ${\mathbf{z}} = [z_1, z_2, ..., z_M]\in \mathcal{D}_{\mathbf{z}}\in \mathbb{R}^M$ is a vector of random variables (parameters).
+For ease of notation, we assume these input parameters to be uniformly distributed i.e.
+$z_i \sim \mathcal{U}(0,1)$ and the support of the input set is $\mathcal{D}_{\mathbf{z}} = [0,1]^M$ where $M$ is the total number of input parameters.
+The Sobol decomposition is based on the following hierarchical representation \cite{RSmith}:
 \begin{equation}\label{sobol_decomp}
-f(z_1, z_2, ..., z_M) = f_0+\sum_{i=1}^M f_i(z_i) + \sum_{1\leq i<j\leq M} f_{ij}(z_i,z_j) + ... + f_{1, 2, ..., M}(z_1, z_2, ..., z_M),
+f(z_1, z_2, ..., z_M) = f_0+\sum_{i=1}^M f_i(z_i) + \sum_{1\leq i<j\leq M} f_{ij}(z_i,z_j) + ... + f_{1, 2, ..., M}(z_1, z_2, ..., z_M).
 \end{equation}
-where the zeroth-order function $f_0$ is the mean response of $f$, the first-order univariate functions $f_i(z_i)$ quantify independent contributions due to each random input, second-order functions $f_{i,j}(z_i,z_j)$ represent the effect of interaction between $z_i$ and $z_j$ on the response $f$.
-Higher-order terms are interpreted in a similar manner.
-Note that the \eqref{sobol_decomp} is only valid for independent input parameters.
-
+Here the zeroth-order function $f_0$ is the mean response of $f$, the first-order univariate functions $f_i(z_i)$ quantify independent contributions due to each random input, and second-order functions $f_{i,j}(z_i,z_j)$ represent the effect of interaction between $z_i$ and $z_j$ on the response $f$.
 Formally, the zeroth-, first- and second-order terms are defined as:
@@
 where ${\mathcal{D}_{\mathbf{z}}}^{M-1} = [0,1]^{M-1}$ and the notation $\mathbf{z}_{\sim \{i,j\}}$ denotes the vector having all the components of $\mathbf{z}$ except in the set $\{i,j\}$.
-We only consider second-order expansion of \eqref{sobol_decomp}, as for a number of applications, the second-order expansion is usually sufficient and higher-order terms have negligible effect on the output response.
-In the case of wind turbine, suppose $y$ represent the power output that depends on say 3 random inputs: windspeed $(z_1)$, wind standard deviation $(z_2)$ and RPM $(z_3)$, then $f_0$ represents the mean power output considering all the three random inputs, $f_1(z_1)$ represent the independent contribution of windspeed on the power and $f_{1,2}(z_1,z_2)$ quantify the interactions of wind speed and wind standard deviation on power.
-Further, the total variance of $f(\mathbf{z})$ is defined as:
+Higher-order terms are interpreted in a similar manner.
+Note that \eqref{sobol_decomp} is only valid for independent input parameters.
+We only consider a second-order expansion of \eqref{sobol_decomp}, as for many applications this is sufficient because higher-order terms have negligible effect on the output response.
+
+As an example, $Y$ can represent the power output of a wind turbine, with random inputs windspeed $(z_1)$, wind standard deviation $(z_2)$ and rotational speed $(z_3)$.
+$f_0$ then represents the mean power output considering all the three random inputs, $f_1(z_1)$ represent the independent contribution of windspeed on the power and $f_{1,2}(z_1,z_2)$ quantifies the interactions of wind speed and wind standard deviation on power.
+
+The definition of the variance is:
 \begin{equation}\label{tot_var}
-D = \text{Var}[f(\mathbf{z})] = \int_{\mathcal{D}_{\mathbf{z}}} f^2(\mathbf{z})d\mathbf{z} - f_0^2
+D = \text{Var}[f(\mathbf{z})] = \int_{\mathcal{D}_{\mathbf{z}}} f^2(\mathbf{z})d\mathbf{z} - f_0^2.
 \end{equation}
@@
 \end{equation}
-where the first- and second-order partial variances are defined as
+where the first- and second-order partial variances are defined via the Sobol decomposition as
 \begin{align}
-D_i &= \int_{0}^{1} f^2_i(z_i)dz_i, \label{partial_variance1}\\
-D_{ij} &= \int_{0}^{1} \int_{0}^{1} f^2_{ij}(z_i,z_j)dz_idz_j \label{partial_variance2}. 
+D_i &= \int_{0}^{1} f^2_i(z_i)dz_i, & D_{ij} &= \int_{0}^{1} \int_{0}^{1} f^2_{ij}(z_i,z_j)dz_idz_j \label{partial_variance2}. 
 \end{align}
@@
 \begin{equation}\label{sobol_ind}
-S_i = \frac{D_i}{D}, \quad S_{ij} = \frac{D_{ij}}{D}, \quad i,j=1,2, ..., M,
+S_i = \frac{D_i}{D}, \quad S_{ij} = \frac{D_{ij}}{D}, \quad i,j=1,2, ..., M.
 \end{equation}
-and the sum of all indices satisfy
-\begin{equation}
-\sum_{i=1}^M S_i + \sum_{1\leq i<j\leq M} S_{ij} = 1.
-\end{equation}
-Finally, the total effect of the parameter $z_i$ on the output response $Y$ is quantified using the total sensitivity indices measure
+%and the sum of all indices satisfy
+%\begin{equation}
+%\sum_{i=1}^M S_i + \sum_{1\leq i<j\leq M} S_{ij} = 1.
+%\end{equation}
+The total effect of $z_i$ on the output $Y$ can be quantified using the total sensitivity index
 \begin{equation}
@@
 The total sensitivity indices can be interpreted as an importance measure for the parameter $z_i$, therefore a large $S_{T_i}$ implies that $z_i$ has a strong influence on $Y$.
-The computation of $S_{T_i}$ requires the approximation of partial variances defined in \eqref{partial_variance1} - \eqref{partial_variance2}.
-Using the Monte Carlo method for computing these variance can be prohibitive for computationally expensive models $Y$.
-Polynomial Chaos Expansions (PCE) based approaches provides a more efficient alternative for sensitivity analysis.
+The computation of $S_{T_i}$ requires the approximation of partial variances defined in \eqref{partial_variance2}.
+Using the Monte Carlo method for computing these variances can be prohibitive for computationally expensive models $Y$.
+On the other hand, Polynomial Chaos Expansions (PCE) based approaches provide a more efficient alternative for sensitivity analysis, and will be used here.
 As shown in \cite{SUDRET2008964}, one can compute the Sobol indices analytically by post-processing the PCE coefficients.
-\subsection{PCE based Sobol indices computation}
+
+\subsection{PCE-based Sobol indices computation}
 The PCE approximation $f^{K}(\mathbf{z})$ of the computational model $Y$ can be defined as a weighted sum of multivariate polynomials in $\mathbf{z}$ \cite{RSmith}
@@
 \end{equation}
-where $\mathbf{k}\in \mathbb{N}_0^M$ is a $M$-dimensional multi-index with the magnitude $|\mathbf{k}| = k_1+k_2 + ...
-+ k_M$ and $\Psi_{\mathbf{k}}(\mathbf{z})$ multivariate polynomial is computed using a tensor product of univariate polynomials $\psi_{k_i}^{(i)}(z_i)$
+where $\mathbf{k}\in \mathbb{N}_0^M$ is a $M$-dimensional multi-index with magnitude $|\mathbf{k}| = k_1+k_2 + ...
++ k_M$ and $\Psi_{\mathbf{k}}(\mathbf{z})$ a multivariate polynomial basis, which is computed using a tensor product of univariate polynomials $\psi_{k_i}^{(i)}(z_i)$
 \begin{equation}
-\Psi_{\mathbf{k}}(\mathbf{z}) := \prod_{i=1}^M\psi_{k_i}^{(i)}(z_i).
+\Psi_{\mathbf{k}}(\mathbf{z}) := \prod_{i=1}^M\psi_{k_i}^{(i)}(z_i),
 \end{equation}
-where $k_i$ is the order of univariate polynomial.
-The multivariate polynomials $\Psi_{\mathbf{k}}(\mathbf{z})$ is orthogonal with respect to the joint probability distribution of random vector $\mathbf{z}$.
+where $k_i$ is the order of the univariate polynomial.
 The choice of the univariate orthogonal polynomial $\psi_{k_i}$ depends on the type of random variable, $z_i$.
-For example, for uniform random variable the Legendre family of polynomials is used, see \cite{Xiu2002} for details.
+For example, for uniform random variables the Legendre family of polynomials is used \cite{Xiu2002}.
+%The multivariate polynomials $\Psi_{\mathbf{k}}(\mathbf{z})$ is orthogonal with respect to the joint probability distribution of random vector $\mathbf{z}$. 
+The total number of terms $K$ in \eqref{eq:PCE} can be based for example on a standard truncation scheme in which all polynomials in $M$ random variable with order less than or equal to $p$ are used.
+More advanced truncation schemes such as the maximum interaction or a hyperbolic scheme can also be employed depending on the application, see \cite{BlatmanThesis}.
+\todo{which one is used?}
 
-The total number of terms $K$ in \eqref{eq:PCE} can be based on a standard truncation scheme where we use all polynomials in $M$ random variable with order less than or equal to $p$.
-More advanced truncation schemes such as the maximum interaction or a hyperbolic scheme can also be employed depending on the application, see \cite{BlatmanThesis}.
+The PCE coefficients $w_{\mathbf{k}}$ corresponding to $\Psi_{\mathbf{k}}(\mathbf{z})$ can be computed using a pseudo-spectral approach by using quadrature methods.
+For a small number of dimensions this is an efficient approach.
+However, in the current work we will encounter more than 10 dimensions and quadrature methods are typically too expensive as they suffer strongly from the curse of dimensionality.
+We therefore focus on methods in which the sampling points can be chosen more freely: Ordinary Least Squares (OLS) and Least-Angle Regression (LARS).
+%In the current work where we have a limited computational budget, methods that minimizes the approximation error for a given number of samples are more practical.
+Next, we briefly describe two such methods to compute PCE coefficients based on a fixed number of model evaluations.
 
-The PCE coefficient $w_{\mathbf{k}}$ corresponding to $\Psi_{\mathbf{k}}(\mathbf{z})$ can be computed using the projection approach:
-\begin{equation}\label{eq:projection}
-w_{\mathbf{k}} = \mathbb{E}[\Psi_{\mathbf{k}}(\mathbf{z})\cdot f^{K}(\mathbf{z})].
-\end{equation} 
-The above expectation can be computed using a quadrature based integration methods such as the Gaussian quadrature, sparse quadrature rules (e.g.
+%\subsubsection{Quadrature}
+%The PCE coefficients $w_{\mathbf{k}}$ corresponding to $\Psi_{\mathbf{k}}(\mathbf{z})$ can be computed using the projection approach:
+%\begin{equation}\label{eq:projection}
+%w_{\mathbf{k}} = \mathbb{E}[\Psi_{\mathbf{k}}(\mathbf{z})\cdot f^{K}(\mathbf{z})].
+%\end{equation} 
+%The above expectation can be computed using a quadrature based integration methods such as the Gaussian quadrature, sparse quadrature rules (e.g.
 Smolyak's tensor grid), etc.
-Quadrature based integration methods suffer from curse of dimensionality where the number of integration points rises exponentially with an increase in the number of dimensions.
-In the current work where we have a fixed computational budget, methods that minimizes the approximation error for a given number of samples are more practical.
+Quadrature based integration methods typically suffer from the curse of dimensionality because the number of integration points rises exponentially with an increase in the number of dimensions.
+In the current work where we have a limited computational budget, methods that minimizes the approximation error for a given number of samples are more practical.
 Next, we briefly describe two such methods to compute PCE coefficients based on a fixed number of model evaluations.
-\subsubsection{Ordinary Least Square}
-Instead of using projection based method one can also use least squares based approaches to compute coefficients \cite{Berveiller2006}.
+
+\subsubsection{Ordinary Least Squares}
+Instead of using a projection-based method one can also use least squares based approaches to compute coefficients \cite{Berveiller2006}.
 The PCE expansion \eqref{eq:PCE} can be expressed as a vector-matrix product
@@
 \end{equation} 
-where $\mathbf{w} = \{w_0,w_1,...,w_K\}^T$ is the coefficient vector, $\boldsymbol{\Psi}(\mathbf{z}) = \{\Psi_0(\mathbf{z}), \Psi_1(\mathbf{z}), ..., \Psi_K(\mathbf{z})\}^T$ is the matrix containing multivariate polynomials in $\mathbf{z}$ and $\varepsilon_K$ is the approximation error.
+where $\mathbf{w} = \{w_0,w_1,...,w_K\}^T$ is the coefficient vector, $\boldsymbol{\Psi}(\mathbf{z}) = \{\Psi_0(\mathbf{z}), \Psi_1(\mathbf{z}), ..., \Psi_K(\mathbf{z})\}^T$ is the matrix containing multivariate polynomials in $\mathbf{z}$, and $\varepsilon_K$ is the approximation error.
 The least-squares minimization problem can then be formulated as:
 \begin{equation}\label{eq:PCE_OLS}
-\mathbf{\overline{w}} = \text{arg min }\mathbb{E}\left[\left(\mathbf{w}^T \boldsymbol{\Psi}(\mathbf{z}) - f(\mathbf{z})\right)^2\right]
+\mathbf{\overline{w}} = \argmin_{\mathbf{w}} \mathbb{E}\left[\left(\mathbf{w}^T \boldsymbol{\Psi}(\mathbf{z}) - f(\mathbf{z})\right)^2\right]
  \end{equation}
@@
 where $\mathbf{X}\in \mathbb{R}^{N\times K}$ is the design matrix where $\mathbf{X}_{i,j} = \Psi_{j}(\mathbf{z}_i)$, $i = 1,2, ..., N$ and $j = 0,1,2,..., K$.
-Although the OLS method allows to compute the coefficients using a limited number of model evaluations, one of the main drawback of this approach is that it may result in a PCE model that consists of high-order interaction terms that may lead to overfitting.
-For many engineering problems, low-order interactions between the input variables are more important.
-Therefore, the least squares form in \eqref{eq:PCE_OLS} can be regularized such that the minimization results is a low-order sparse PCE model.
+Although the OLS method allows to compute the coefficients using a limited number of model evaluations, one of the main drawbacks of this approach is that it may result in a PCE model that consists of high-order interaction terms that may lead to overfitting.
+For many engineering problems, low-order interactions between the input variables are most important.
+Therefore, the least squares form in \eqref{eq:PCE_OLS} can be regularized such that the minimization results is a low-order sparse PCE model, as discussed next.
 
 \subsubsection{Least Angle Regression}
-Least angle regression (LARS) algorithm \cite{efron2004} is one of the several methods can be applied to compute sparse PCE models \cite{BlatmanThesis, BLATMAN2011}.
+The LARS algorithm \cite{efron2004} is one of the methods that can be applied to compute sparse PCE models \cite{BlatmanThesis, BLATMAN2011}.
 The key idea is to find the best set of polynomials from a given candidate set that minimizes the regularized form of \eqref{eq:PCE_OLS}:
 \begin{equation}\label{regularized_LS}
-\mathbf{\overline{w}} = \text{arg min }\mathbb{E}\left[\left(\mathbf{w}^T \boldsymbol{\Psi}(\mathbf{z}) - f(\mathbf{z})\right)^2\right] + \lambda ||\mathbf{w}||_1.
+\mathbf{\overline{w}} = \argmin_{\mathbf{w}} \mathbb{E}\left[\left(\mathbf{w}^T \boldsymbol{\Psi}(\mathbf{z}) - f(\mathbf{z})\right)^2\right] + \lambda ||\mathbf{w}||_1.
 \end{equation}
@@
 The second polynomial is now also added to the active set.
-In the next step, we move both the coefficients in the active set towards their least-square values until a third polynomial from candidate set shows as much correlation as the current PCE model based on two polynomials show with the model evaluations.
-The number of iteration is given by the minimum of $\{K,N\}$.
-After every iteration a posteriori error, for instance, Leave-One-Out (LOO) cross-validation error is computed.
-The active set with the smallest LOO is then chosen as the best sparse PCE
-model.
+In the next step, we move both the coefficients in the active set towards their least-square values until a third polynomial from the candidate set shows as much correlation as the current PCE model based on two polynomials shows with the model evaluations.
+The number of iterations is given by the minimum of $\{K,N\}$.
+After every iteration a posteriori error, for instance the Leave-One-Out (LOO) cross-validation error, is computed.
+The active set with the smallest LOO is then chosen as the best sparse PCE model.
 
-One of the main advantage of the LARS algorithm is that it works well in cases when $K>>N$, i.e.
-the number of polynomials in the candidate set is much more larger than the number of model evaluations.
+One of the main advantage of the LARS algorithm is that it works well in cases when $K>>N$, i.e. the number of polynomials in the candidate set is much larger than the number of model evaluations.
 
-\subsection{Sensitivity analysis with discrete random variable}
+\subsection{Sobol index computation}
+
+%\subsection{Sensitivity analysis with discrete random variable}
+
 \section{Results}\label{sec:results}
@@
 \subsection{NM80}
-As a reference wind turbine, we consider the 2MW NM80 turbine from the DANAERO project \cite{Troldborg2013} with a blade radius of 38.8 m.
-The data for lift ($Cl$) and drag ($Cd$) polars are available at four locations along the blade radius at 11.87 m, 17.82 m, 28.97 m, and 35.53 m.
+\subsubsection{Operating conditions}
+As a reference wind turbine, we consider the 2MW NM80 turbine from the DANAERO project \cite{Troldborg2013} with a blade radius of 38.8 m. \todo{Add operating conditions}
+
+\subsubsection{Choice of uncertain inputs}
+The data for lift ($C_l$) and drag ($C_d$) polars are available at four locations along the blade radius at 11.87 m, 17.82 m, 28.97 m, and 35.53 m.
 The lift and drag variables at these sections are numbered Cl1 - Cl4 and Cd1 - Cd4 respectively.
-The reference value of polars is obtained from the wind-tunnel experiment with 3D corrections.
+The reference value of these polars is obtained from the wind-tunnel experiment with 3D corrections.
 Random samples of chord, twist, lift- and drag-polars are obtained by perturbing the control points with a uniformly distributed random variable.
 
-There are two ways to obtain perturbed curves either by independently perturbing each of the control points (local perturbation) or using the same random number to perturb all the control points (global perturbation).
-We use local perturbation for chord control points Ch3 - Ch7 and twist control points Tw2 - Tw7 control points, see Fig.
-\ref{perturbed_samples} (a) - (b), respectively.
-Each control point are perturbed using a uniform random number with bounds given by $\pm5\%$ of the magnitude of a control point.
-For lift and drag curves, all control points are perturbed using the same uniform random number resulting in samples of the lift and drag curves as shown in Fig.
-\ref{perturbed_samples} (c) - (d), respectively.
+There are two ways to obtain perturbed curves: independently perturbing each of the control points (local perturbation) or using a single random number to perturb all the control points (global perturbation).
+We use local perturbation for chord control points Ch3 - Ch7 and twist control points Tw2 - Tw7 control points, see Fig.\ \ref{perturbed_samples} (a) - (b), respectively.
+Each control point are perturbed using a uniform random number with bounds given by $\pm5\%$ of the magnitude of a control point (see equation \eqref{eqn:perturb_c}).
+For the lift and drag curves, all control points are perturbed using the same uniform random number resulting in samples of the lift and drag curves as shown in Fig.\ \ref{perturbed_samples} (c) - (d), respectively.
 For simplicity, we only consider uniform distributions for perturbing the control points, however, we can also use other probability distributions for this purpose.
@@
 
-\subsubsection{Verification of GSA method}
-We perform a convergence study for OLS and LARS to determine which method converges fastest with increasing dimension.
-For this, we use the parametrized chord from Fig.
-\ref{perturbed_samples}(a) with 9 control points.
-In Fig.
-\ref{convergence_dim}, we show the convergence behavior of the two methods with increasing number of uncertain inputs, consisting of 4, 6 and 8 control points and for different number of model evaluations.
-We see that both methods scales well with increasing dimensions, however, $PCE\_LARS$ exhibits faster convergence than $PCE\_OLS$ in high dimensions.
+\subsubsection{Verification of sensitivity analysis method}
+We perform a convergence study for both OLS and LARS in order to determine which method converges fastest with increasing number of uncertain parameters and number of model runs.
+For this purpose we only include uncertainty in the chord.
+The NURBS representation is defined by 9 control points, of which a subset is taken uncertain (see Fig.\ \ref{perturbed_samples}(a)).
+In Fig.\ \ref{convergence_dim}, we show for both OLS and LARS the convergence behaviour of the total Sobol indices of the power output with increasing number of model evaluations $N$ (runs of the AeroModule), and with increasing number of uncertain inputs $M$ (4, 6 and 8 control points).
+We see that both methods scale well with increasing dimensions, however, LARS exhibits faster convergence than OLS in high dimensions due to its adaptive nature, and will be the method of choice in the subsequent tests.
+\todo{what is a reasonable choice of $N$, depending on $M$?}
 
@@
  \end{subfigure}
-    \caption{Convergence of $PCE\_OLS$ and $PCE\_LARS$ with increasing number of uncertain inputs $M$ and samples $N$.}
+    \caption{Convergence of OLS and LARS with increasing number of uncertain inputs $M$ and model evaluations $N$.}
 \label{convergence_dim}
 \end{figure*}
+
 \subsubsection{Sensitivity analysis results}

### 2020-02-13 Papers/Torque_Paper/content.tex
updated torque article

@@
 More advanced truncation schemes such as the maximum interaction or a hyperbolic scheme can also be employed depending on the application, see \cite{BlatmanThesis}.
-\todo{which one is used?}
+\textcolor{red}{In this work, we use the hyperbolic truncation scheme with a $qnorm = 0.75$.} \todo{which one is used?}
 
@@
 \subsubsection{Operating conditions}
-As a reference wind turbine, we consider the 2MW NM80 turbine from the DANAERO project \cite{Troldborg2013} with a blade radius of 38.8 m. \todo{Add operating conditions}
+As a reference wind turbine, we consider the 2MW NM80 onshore turbine from the DANAERO project \cite{Troldborg2013} with a blade radius of 38.8 m and \textcolor{red}{a hub height of 57.19 m with the mean wind speed at hub height set as 6.1m/s.
+Also, we use the following mean values for: RPM = 12.3, Pitch angle = 0.15 and Yaw  = 0.}
+\todo{Add operating conditions}
 
@@
 We perform a convergence study for both OLS and LARS in order to determine which method converges fastest with increasing number of uncertain parameters and number of model runs.
-For this purpose we only include uncertainty in the chord.
+For this purpose we only include uncertainty in the chord and \textcolor{red}{set power production as the quantity of interest}.
 The NURBS representation is defined by 9 control points, of which a subset is taken uncertain (see Fig.\ \ref{perturbed_samples}(a)).
@@
 We see that both methods scale well with increasing dimensions, however, LARS exhibits faster convergence than OLS in high dimensions due to its adaptive nature, and will be the method of choice in the subsequent tests.
-\todo{what is a reasonable choice of $N$, depending on $M$?}
+\textcolor{red}{Ideally, for spectral methods, $N$ should be increased exponentially with increasing $M$.
+In the numerical experiments, for both OLS and LARS, we start with a small $N$ and double it with every iteration until converged values of Sobol indices are obtained.}  \todo{what is a reasonable choice of $N$, depending on $M$?}
 
@@
 \subsubsection{Sensitivity analysis results}
-In Fig.~\ref{sensitivity_analysis}, we show the total order Sobol indices as a measure of the sensitivity of the different geometric and model parameters on the power output.
+In Fig.~\ref{sensitivity_analysis}, we show the total order Sobol indices as a measure of the sensitivity of the different geometric and model parameters on the power output \textcolor{red}{computed using LARS method with the number of uncertain parameter $M = 19$ chosen from Fig.~\ref{perturbed_samples}.
+We found that 1024 model evaluations were sufficient to achieve converged indices}.
 In this case, we observe the important result that the uncertainty in model parameters is at least as important as the uncertainty in geometric parameters.
-Furthermore, within the geometric parameters, the chord distribution shows a significantly higher sensitivity compared to the twist variables
+Furthermore, within the geometric parameters, the chord distribution shows a significantly higher sensitivity compared to the twist variables.
 

### 2020-02-20 Papers/Torque_Paper/content.tex
minor edits and accepting Prashant's changes to Torque article

--- 2020-02-10 Papers/Torque_Paper/content.tex
@@
 The total number of terms $K$ in \eqref{eq:PCE} can be based for example on a standard truncation scheme in which all polynomials in $M$ random variable with order less than or equal to $p$ are used.
-More advanced truncation schemes such as the maximum interaction or a hyperbolic scheme can also be employed depending on the application, see \cite{BlatmanThesis}.
-\todo{which one is used?}
+In this work, a more advanced truncation schemes known as the hyperbolic truncation scheme is used \cite{BlatmanThesis}.
+%\textcolor{red}{In this work, we use the hyperbolic truncation scheme with a $qnorm = 0.75$.}
 
@@
 \subsubsection{Operating conditions}
-As a reference wind turbine, we consider the 2MW NM80 turbine from the DANAERO project \cite{Troldborg2013} with a blade radius of 38.8 m. \todo{Add operating conditions}
+As a reference wind turbine, we consider the 2MW NM80 onshore turbine from the DANAERO project \cite{Troldborg2013} with a blade radius of 38.8 m and a hub height of 57.19 m with the mean wind speed at hub height set as 6.1m/s.
+Furthermore, the RPM is set at 12.3$/\text{min.}$, the blade pitch angle at 0.15 $\deg$ and the turbine yaw angle is zero.
 
@@
 We perform a convergence study for both OLS and LARS in order to determine which method converges fastest with increasing number of uncertain parameters and number of model runs.
-For this purpose we only include uncertainty in the chord.
+For this purpose we only include uncertainty in the chord and set the power production as the quantity of interest.
 The NURBS representation is defined by 9 control points, of which a subset is taken uncertain (see Fig.\ \ref{perturbed_samples}(a)).
@@
 We see that both methods scale well with increasing dimensions, however, LARS exhibits faster convergence than OLS in high dimensions due to its adaptive nature, and will be the method of choice in the subsequent tests.
-\todo{what is a reasonable choice of $N$, depending on $M$?}
+For both OLS and LARS, we start with a small $N$ and double it with every iteration until converged values of Sobol indices are obtained.
+%}  \todo{what is a reasonable choice of $N$, depending on $M$?}
 
@@
 \subsubsection{Sensitivity analysis results}
-In Fig.~\ref{sensitivity_analysis}, we show the total order Sobol indices as a measure of the sensitivity of the different geometric and model parameters on the power output.
+In Fig.~\ref{sensitivity_analysis}, we show the total order Sobol indices as a measure of the sensitivity of the different geometric and model parameters on the power output computed using LARS with $M=19$ uncertain parameter chosen from Fig.~\ref{perturbed_samples}.
+We found that $N=1024$ model evaluations were sufficient to achieve converged indices.
 In this case, we observe the important result that the uncertainty in model parameters is at least as important as the uncertainty in geometric parameters.
-Furthermore, within the geometric parameters, the chord distribution shows a significantly higher sensitivity compared to the twist variables
+Furthermore, within the geometric parameters, the chord distribution shows a significantly higher sensitivity compared to the twist variables.
 
@@
 
-\subsection{AVATAR}
-
 \section{Conclusions}\label{sec:conclusions}

### 2020-02-20 Papers/Torque_Paper/content.tex
minor updates to paper
* edited conclusion
* changed x to r
* description of RVs

@@
 
-As a consequence of the strong model assumptions at the basis of BEM theory, the results from BEM codes can be subject to significant inaccuracies or uncertainties.
+As a consequence of the relatively strong model assumptions at the basis of BEM theory, the results from BEM codes can be subject to significant inaccuracies or uncertainties.
 For example, the effect of sheared inflow \cite{Madsen2012} is not naturally accounted for in the theory and needs to be incorporated via correction terms.
@@
 \subsection{NURBS representation}
-A NURBS curve $S(x)$ is expressed using a weighted sum of $n$ basis functions (or B-splines):
+A NURBS curve $S(r)$ is expressed using a weighted sum of $n$ basis functions (or B-splines):
 \begin{equation}\label{NURB_curve}
-S(x) = \sum_{i=1}^{n} c_i B_{i,p}(x),
+S(r) = \sum_{i=1}^{n} c_i B_{i,p}(r),
 \end{equation}
-where $c_i$ is the weight of the $i$-th control point and $B_{i,p}(x)$ is the value of the B-spline corresponding to the $i$-th control point at $x$, with the subscript $p$ denoting its polynomial degree.
-In our case, $S(x)$ represent for example the chord as a function of the radial location along the blade.
-A sequence of non-decreasing parameters $X = \{x_1,x_2,...,x_m\}$, called the \textit{knot vector}, determines the domain in which any control point is active.
-A control point $c_i$ is said to be active at a location when the corresponding B-spline $B_{i,p}(x)$ is non-zero at $x$.
+where $c_i$ is the weight of the $i$-th control point and $B_{i,p}(r)$ is the value of the B-spline corresponding to the $i$-th control point at $x$, with the subscript $p$ denoting its polynomial degree.
+In our case, $S(r)$ represent for example the chord as a function of the radial location along the blade.
+A sequence of non-decreasing parameters $X = \{r_1,r_2,...,r_m\}$, called the \textit{knot vector}, determines the domain in which any control point is active.
+A control point $c_i$ is said to be active at a location when the corresponding B-spline $B_{i,p}(r)$ is non-zero at $r$.
 B-splines are recursive in polynomial degree, thus we can derive quadratic B-splines using linear B-splines, cubic from quadratic B-splines and so on.
@@
 \begin{equation}\label{linearBspline}
-B_{i,0}(x) :=
+B_{i,0}(r) :=
 \begin{cases}
-1\quad x_i\leq x < x_{i+1},\quad i = 1,2,...,m,\\
+1\quad r_i\leq r < r_{i+1},\quad i = 1,2,...,m,\\
 0\quad\text{elsewhere,}
@@
 \begin{equation}\label{NURBS_recurrence}
-B_{i,p}(x) := \frac{x - x_i}{x_{i+p} - x_i}B_{i,p-1}(x) + \frac{x_{i+p+1} - x}{x_{i+p+1} - x_{i+1}}B_{i+1,p-1}(x),\quad p\geq1.
+B_{i,p}(r) := \frac{r - r_i}{r_{i+p} - r_i}B_{i,p-1}(r) + \frac{r_{i+p+1} - r}{r_{i+p+1} - r_{i+1}}B_{i+1,p-1}(r),\quad p\geq1.
 \end{equation}
 Note that with increasing $p$ the interval in which a B-spline is non-zero also become larger.
-For a curve of polynomial degree $p$, there are $p+1$ active control points at any location $x$.
+For a curve of polynomial degree $p$, there are $p+1$ active control points at any location $r$.
 For example, for a linear curve ($p=1$) only two adjacent control points are active at any location.
@@
 In Fig.
-\ref{basis_examples}, we show linear, quadratic and cubic splines for $x\in[0,1]$ with original knot vector $X = \{0,1/3,2/3,1\}$, padded accordingly with the curve order, for e.g., for linear B-splines, we use the knot vector $X = \{0,0,1/3,2/3,1,1\}$.
+\ref{basis_examples}, we show linear, quadratic and cubic splines for $r\in[0,1]$ with original knot vector $X = \{0,1/3,2/3,1\}$, padded accordingly with the curve order, for e.g., for linear B-splines, we use the knot vector $X = \{0,0,1/3,2/3,1,1\}$.
 \begin{figure}[h!]
@@
 In order to parameterize a given curve, e.g.
-the blade reference chord, we choose a degree $p$ and a number of knots $X =\{x_j\}_{j=1}^n$.
+the blade reference chord, we choose a degree $p$ and a number of knots $X =\{r_j\}_{j=1}^n$.
 The knots can be uniformly spaced or chosen heuristically (for advanced approaches for knot selection see\cite{LI2005791,razdan1999knot}).
-Next we sample $S_{ref}(x)$ at $X$ and also compute $B_{i,p}(x_j), \text{ for }i, j = 1,2, ..., n$.
+Next we sample $S_{ref}(r)$ at $X$ and also compute $B_{i,p}(r_j), \text{ for }i, j = 1,2, ..., n$.
 We then formulate a linear system and solve for the set of control points $\mathbf{c}=\{c_i\}_{i=1}^n$ as:
@@
 \end{equation}
-where $\mathbf{S}_{ref}\in\mathbb{R}^{n}$ is a vector containing sampled values of $S_{ref}(x)$ and $\mathbf{B}\in \mathbb{R}^{n\times n}$ is a matrix with the $j$-th row consisting of values of $n$ B-splines sampled at location $x_j$.
-Once the control points are obtained, the approximate reference curve $S_n(x)$ follows from \eqref{NURB_curve}.
+where $\mathbf{S}_{ref}\in\mathbb{R}^{n}$ is a vector containing sampled values of $S_{ref}(r)$ and $\mathbf{B}\in \mathbb{R}^{n\times n}$ is a matrix with the $j$-th row consisting of values of $n$ B-splines sampled at location $r_j$.
+Once the control points are obtained, the approximate reference curve $S_n(r)$ follows from \eqref{NURB_curve}.
 Fig.~\ref{nurbs_approx} shows a parameterization of the chord curve using $n=9$ control points and second-order ($p=2$) B-splines.
@@
 \subsection{Perturbing the NURBS curves}
-Given the reference values of $\mathbf{c}$, it becomes straightforward to generate perturbed curves: we sample a uniform random variable to perturb the baseline control points and use these to compute a perturbed curve $\tilde{S}(x)$:
-\begin{equation}
-\tilde{S} (x) = \sum_{i=1}^{n} \tilde{c}_{i} B_{i,p}(x),
-\end{equation}
-where
+Given the reference values of $\mathbf{c}$, it becomes straightforward to generate perturbed curves: we sample a uniform random variable to perturb the baseline control points and use these to compute a perturbed curve $\tilde{S}(r)$:
 \begin{equation}\label{eqn:perturb_c}
-\tilde{c}_{i} = c_{i} (1 + \Delta_{i}),
+\tilde{S} (r) = \sum_{i=1}^{n} \tilde{c}_{i} B_{i,p}(r) \qquad \text{where} \qquad \tilde{c}_{i} = c_{i} (1 + \Delta_{i}),
 \end{equation}
 and the perturbation is randomly sampled as $\Delta_{i} \sim \mathcal{U}[-0.05,0.05]$.
-
-\subsection{Model uncertainties}
-The above framework is directly applicable to express the geometric uncertainty in twist- and chord distribution, with $c$ denoting either the value of the twist control points or of the chord control points.
-For model uncertainties, i.e.\ lift- and drag polars, \todo{Insert mathematical expression}
+This equation is applicable to both the twist- and chord distribution, where $\tilde{c}_{i}$ can denote either the value of the twist control points, $\text{Tw}_i$, or of the chord control points, $\text{Ch}_{i}$.
+For the lift- and drag polars, we use the following parametrization:
+\begin{equation}\label{eqn:perturb_cl}
+\tilde{S} (r,\alpha) = \sum_{i=1}^{n} \tilde{c}_{i}(\alpha) B_{i,p}(r) \qquad \text{where} \qquad \tilde{c}_{i}(\alpha) = c_{\text{ref},i} (\alpha) (1 + \Delta_{i}),
+\end{equation}
+where $c_{i}(\alpha)$ can denote either the lift coefficient $\text{Cl}_{i}$ or drag coefficient $\text{Cd}_{i}$, and the subscript $\text{ref}$ denotes the reference curve.
+% $c_{\text{ref},i}(\alpha)$ can denote either the reference lift coefficient $C_{l,\text{ref},i}(\alpha)$ or drag coefficient $C_{d,\text{ref},i}(\alpha)$ at section $i$.
 
@@
 \subsubsection{Choice of uncertain inputs}
-The data for lift ($C_l$) and drag ($C_d$) polars are available at four locations along the blade radius at 11.87 m, 17.82 m, 28.97 m, and 35.53 m.
+The data for lift ($C_l$) and drag ($C_d$) polars are available at four locations along the blade radius: at 11.87 m, 17.82 m, 28.97 m, and 35.53 m.
 The lift and drag variables at these sections are numbered Cl1 - Cl4 and Cd1 - Cd4 respectively.
@@
 We use local perturbation for chord control points Ch3 - Ch7 and twist control points Tw2 - Tw7 control points, see Fig.\ \ref{perturbed_samples} (a) - (b), respectively.
-Each control point are perturbed using a uniform random number with bounds given by $\pm5\%$ of the magnitude of a control point (see equation \eqref{eqn:perturb_c}).
-For the lift and drag curves, all control points are perturbed using the same uniform random number resulting in samples of the lift and drag curves as shown in Fig.\ \ref{perturbed_samples} (c) - (d), respectively.
+The twist and chord are perturbed following equation \eqref{eqn:perturb_c}, whereas the lift and drag are perturbed according to equation \eqref{eqn:perturb_cl}.
+Examples of the resulting lift and drag curves are shown in Fig.\ \ref{perturbed_samples} (c) - (d), respectively.
 For simplicity, we only consider uniform distributions for perturbing the control points, however, we can also use other probability distributions for this purpose.
@@
 \section{Conclusions}\label{sec:conclusions}
-We have shown how Sobol indices computed using sparse adaptive polynomial expansion can be used for high-dimensional global sensitivity analysis of both geometric and model uncertainties.
-The identified sensitive model parameters will be utilized in future work to develop calibrated BEM models with built-in uncertainty estimates as part of the \textit{WindTrue} project, and is connected to IEA Task 29.
+We have shown how Sobol indices computed using sparse adaptive polynomial expansion can be used for high-dimensional global sensitivity analysis of both geometric and model uncertainties in BEM models.
+The proposed framework is generic and can deal with many types of (parametric uncertainties) The sensitive model parameters identified here will be utilized in future work to develop calibrated BEM models with built-in uncertainty estimates, as part of the \textit{WindTrue} project, and used to better understand the differences between BEM models and measurement data as studied in IEA Task 29.

### 2020-02-20 Papers/Torque_Paper/content.tex
minor updates
- reducing text
- updating references

@@
 
-As a consequence of the relatively strong model assumptions at the basis of BEM theory, the results from BEM codes can be subject to significant inaccuracies or uncertainties.
+As a consequence of the relatively strong model assumptions at the basis of BEM theory, the results from BEM codes can be subject to significant uncertainties.
 For example, the effect of sheared inflow \cite{Madsen2012} is not naturally accounted for in the theory and needs to be incorporated via correction terms.
-Other major model uncertainties in BEM models are for example the time constant in dynamic stall models, the wake correction factor, the tip loss model parameter, and the lift- and drag-polars used to compute local aerodynamic forces.
-Especially for increasing turbine sizes, these model parameters are foreseen to be not sufficiently accurate \cite{Sayed2019}.
-In other words, the uncertainty associated with the output of currently employed BEM models is rather large.
+Other major model uncertainties in BEM models are the time constant in dynamic stall models, the wake correction factor, the tip loss model parameter, and the lift- and drag-polars used to compute local aerodynamic forces.
+Especially for increasing turbine sizes, these model parameters are likely not sufficiently accurate \cite{Sayed2019}.
+%In other words, the uncertainty associated with the output of currently employed BEM models is rather large. 
 % to arrive at robust BEM models that include an uncertainty estimate associated
 
-Recently, several papers have addressed the uncertainty in BEM model output by performing forward uncertainty propagation or sensitivity studies, e.g.\ \cite{Echeverria2017,Matthaus2017,Murcia2018,Robertson2018,Bos2019a}.
+Recently, several papers have addressed the uncertainty in BEM model output by performing uncertainty propagation and sensitivity studies, e.g.\ \cite{Echeverria2017,Matthaus2017,Murcia2018,Robertson2018,Bos2019a}.
 In these studies, the focus in mainly on uncertainties in the external conditions (wind parameters) and/or uncertainty in the turbine specification (geometric parameters).
@@
 
-For the case of a rigid turbine, with a uniform inflow field, the main model uncertainties that arise in the BEM formulation are the following \cite{Hansen1993}:
+For the case of a rigid turbine, with a uniform inflow field, some of the main uncertainties that arise in the BEM formulation are \cite{Hansen1993}:
 \begin{itemize}
@@
 \item Three-dimensional flow effects also cause stall delay at the root; this is typically modeled via \textit{a dynamic stall model} and associated time constant;
-\item Yawed inflow is not naturally included in BEM and typically incorporated via \textit{a wake correction}.
+\item Yawed inflow is not naturally included and typically incorporated via \textit{a wake correction}.
 %\item Dynamic wake - Oye time lag constant
@@
 \end{itemize}
-Next to these model uncertainties we will consider the chord and twist distribution along the blade to be uncertain (geometric uncertainties).
+Next to these model uncertainties we will consider geometric uncertainties: the chord and twist distribution along the blade.
 
@@
 \end{equation}
-where $c_i$ is the weight of the $i$-th control point and $B_{i,p}(r)$ is the value of the B-spline corresponding to the $i$-th control point at $x$, with the subscript $p$ denoting its polynomial degree.
-In our case, $S(r)$ represent for example the chord as a function of the radial location along the blade.
+where $c_i$ is the weight of the $i$-th control point and $B_{i,p}(r)$ is the value of the B-spline corresponding to the $i$-th control point at radial location $r$ along the blade, with $p$ denoting the polynomial degree.
+In our case, $S(r)$ represents for example the chord length.
 A sequence of non-decreasing parameters $X = \{r_1,r_2,...,r_m\}$, called the \textit{knot vector}, determines the domain in which any control point is active.
 A control point $c_i$ is said to be active at a location when the corresponding B-spline $B_{i,p}(r)$ is non-zero at $r$.
-B-splines are recursive in polynomial degree, thus we can derive quadratic B-splines using linear B-splines, cubic from quadratic B-splines and so on.
+B-splines are recursive in polynomial degree: quadratic B-splines follow from linear B-splines, cubic from quadratic B-splines etc.
 Starting with a B-spline of degree 0 defined as
@@
 \end{equation}
-we can derive higher-order B-splines using the recurrence relation \cite{deBoor}:
+higher-order B-splines follow using the recurrence relation \cite{deBoor}:
 \begin{equation}\label{NURBS_recurrence}
@@
 \end{equation}
-Note that with increasing $p$ the interval in which a B-spline is non-zero also become larger.
-For a curve of polynomial degree $p$, there are $p+1$ active control points at any location $r$.
-For example, for a linear curve ($p=1$) only two adjacent control points are active at any location.
-Therefore, only a few nearby control points are active when computing the NURBS curve \eqref{NURB_curve}.
-These B-splines can be constructed efficiently using the de Boor's algorithm \cite{deBoor}.
-To compute high-order B-splines in \eqref{NURBS_recurrence}, we need to pad the knot vector $X$ by repeating the first and last elements $p$ times each.
+%Note that with increasing $p$ the interval in which a B-spline is non-zero also become larger. 
+%For a curve of polynomial degree $p$, there are $p+1$ active control points at any location $r$. For example, for a linear curve ($p=1$) only two adjacent control points are active at any location. 
+Only a few nearby control points are active when computing the NURBS curve \eqref{NURB_curve}; we use the de Boor's algorithm \cite{deBoor} for efficient construction. 
+%To compute high-order B-splines in \eqref{NURBS_recurrence}, we need to pad the knot vector $X$ by repeating the first and last elements $p$ times each. 
 In Fig.
@@
 
-In order to parameterize a given curve, e.g.
-the blade reference chord, we choose a degree $p$ and a number of knots $X =\{r_j\}_{j=1}^n$.
-The knots can be uniformly spaced or chosen heuristically (for advanced approaches for knot selection see\cite{LI2005791,razdan1999knot}).
+In order to parameterize a given curve, e.g.\ the blade reference chord, we choose a degree $p$ and a number of knots $X =\{r_j\}_{j=1}^n$.
+The knots can be uniformly spaced or chosen heuristically (for advanced knot selection approaches for knot selection see \cite{Li2005}).
 Next we sample $S_{ref}(r)$ at $X$ and also compute $B_{i,p}(r_j), \text{ for }i, j = 1,2, ..., n$.
@@
 \subsubsection{Sensitivity analysis results}
-In Fig.~\ref{sensitivity_analysis}, we show the total order Sobol indices as a measure of the sensitivity of the different geometric and model parameters on the power output computed using LARS with $M=19$ uncertain parameter chosen from Fig.~\ref{perturbed_samples}.
+In Fig.~\ref{fig:sensitivity_analysis_power} and \ref{fig:sensitivity_analysis_force}, we show the total order Sobol indices as a measure of the sensitivity of the different geometric and model parameters on the power output and the axial force, respectively, computed using LARS with $M=19$ uncertain parameters chosen from Fig.~\ref{perturbed_samples}.
 We found that $N=1024$ model evaluations were sufficient to achieve converged indices.
-In this case, we observe the important result that the uncertainty in model parameters is at least as important as the uncertainty in geometric parameters.
-Furthermore, within the geometric parameters, the chord distribution shows a significantly higher sensitivity compared to the twist variables.
+In this case, we observe that the uncertainty in model parameters is at least as important as the uncertainty in geometric parameters.
+Furthermore, within the geometric parameters, the chord distribution shows a significantly higher sensitivity than the twist variables.
+Lastly, we observe that the axial force is basically only sensitive to the lift coefficient at the outward blade sections, whereas the power is sensitive to both lift and drag coefficients, also at the inward section of the blade.
 
@@
 \centering
-\includegraphics[trim={1cm 0cm 0 0cm},clip, scale=0.4]{SA_Power_chord_twist_Cl_Cd.eps}
-\caption{}
-\label{sensitivity_analysis}
+\includegraphics[trim={2cm 0cm 0 0cm},clip, scale=0.4]{SA_Power_chord_twist_Cl_Cd.eps}
+\caption{Sensitivity of power output towards geometric and model uncertainties.}
+\label{fig:sensitivity_analysis_power}
+\end{figure}
+
+\begin{figure}[h!]
+\centering
+\includegraphics[trim={1cm 0cm 0 0cm},clip, scale=0.5]{SA_axial_force_chord_twist_Cl_Cd.eps}
+\caption{Sensitivity of axial force towards geometric and model uncertainties.}
+\label{fig:sensitivity_analysis_force}
 \end{figure}

### 2020-02-20 Papers/Torque_Paper/content.tex
minor updates
shortened description of PCE, several cleanups

@@
 We will employ the uncertainty quantification toolbox \texttt{UQLab} \cite{uqlab}, which computes the Sobol indices based on a sparse polynomial chaos expansion.
-
-The main novelty of the work lies in the study of the effect of model uncertainties rather than external conditions or geometric uncertainties.
-As such, the current sensitivity analysis is part of the so-called WindTrue project, in which the long-term goal is the development of calibrated BEM models, that possess a quantified level of uncertainty.
+The main novelty of the work lies in creating a generic framework for studying parametric uncertainteis, including both model uncertainties and geometric uncertainties.
+As such, the current sensitivity analysis is part of the so-called WindTrue project, in which the long-term goal is the development of calibrated BEM models.
 Anticipating on this goal, we are using the NM80 wind turbine model from the Tjaereborg wind farm, for which extensive measurement data is available through the DanAero project \cite{Troldborg2013} and IEA Task 29.
-The measurement data will enable the calibration step, and our sensitivity analysis will make clear which of the model parameters should be used in this calibration step.
+The current sensitivity analysis will make clear which of the model parameters are to be used in the calibration phase.
 
@@
 State-of-the-art wind-turbine models simulate the aeroelastic behavior of wind turbines by combining the concept of momentum conservation of the flow (BEM theory) with the equations of motion for the structure, possibly extended with the hydrodynamics of the sea and control algorithms \cite{Vorpahl2013}.
-In this work, we only concentrate on the first aspect, namely the prediction of flow and blade forces as given by the BEM method.
+In this work, we concentrate on the first aspect, namely the prediction of flow and blade forces as given by the BEM method.
 The particular BEM code that we use is ECN's \texttt{Aero-Module} \cite{Boorsma2012}.
-
-For the case of a rigid turbine, with a uniform inflow field, some of the main uncertainties that arise in the BEM formulation are \cite{Hansen1993}:
+For the case of a rigid turbine, with a uniform inflow field, the main uncertainties that arise in the BEM formulation are \cite{Hansen1993}:
 \begin{itemize}
@@
 In order to reduce this number, we parametrize the variation along the blade by using Non-Uniform Rational Basis Splines (NURBS) \cite{rogers2000}.
-In this way only a limited number of control points is needed to approximate a large variety of curves, so that the resulting number of uncertain parameters is relatively small.
+In this way a limited number of control points can approximate a large variety of curves, so that the resulting number of uncertain parameters is relatively small.
 %To perturb the geometrical parameters of the turbine blade, such as the reference chord or twist.
@@
 \centering
-\includegraphics[trim={0cm 0cm 0 0cm},clip, scale=1.3]{nurbs_approx.eps}
+\includegraphics[trim={0cm 0cm 1.5cm 0cm},clip, scale=1.25]{nurbs_approx.eps}
 \caption{Parametrization of chord using 9 control points and second-order B-splines.}
@@
 \begin{align}\label{sobol_terms}
-f_0 &= \int_{\mathcal{D}_{\mathbf{z}}}f(\mathbf{z})d\mathbf{z},\\
-f_i(z_i) &= \int_{{\mathcal{D}_{\mathbf{z}}}^{M-1}}f(\mathbf{z})d\mathbf{z}_{\sim \{i\}} - f_0,\\
-f_{ij}(z_i,z_j) &= \int_{{\mathcal{D}_{\mathbf{z}}}^{M-2}}f(\mathbf{z})d\mathbf{z}_{\sim \{i,j\}} - f_i(z_i) - f_j(z_j) - f_0,
+f_0 = \int_{\mathcal{D}_{\mathbf{z}}}f(\mathbf{z})d\mathbf{z}, \qquad  f_i(z_i)= \int_{{\mathcal{D}_{\mathbf{z}}}^{M-1}}f(\mathbf{z})d\mathbf{z}_{\sim \{i\}} - f_0,\\
+f_{ij}(z_i,z_j) = \int_{{\mathcal{D}_{\mathbf{z}}}^{M-2}}f(\mathbf{z})d\mathbf{z}_{\sim \{i,j\}} - f_i(z_i) - f_j(z_j) - f_0,
 \end{align}
@@
 Note that \eqref{sobol_decomp} is only valid for independent input parameters.
-We only consider a second-order expansion of \eqref{sobol_decomp}, as for many applications this is sufficient because higher-order terms have negligible effect on the output response.
+We consider a second-order expansion of \eqref{sobol_decomp}, which is generally sufficient for the purpose of this article.
 
 As an example, $Y$ can represent the power output of a wind turbine, with random inputs windspeed $(z_1)$, wind standard deviation $(z_2)$ and rotational speed $(z_3)$.
-$f_0$ then represents the mean power output considering all the three random inputs, $f_1(z_1)$ represent the independent contribution of windspeed on the power and $f_{1,2}(z_1,z_2)$ quantifies the interactions of wind speed and wind standard deviation on power.
+$f_0$ then represents the mean power output, $f_1(z_1)$ represent the independent contribution of windspeed on the power and $f_{1,2}(z_1,z_2)$ quantifies the interactions of wind speed and wind standard deviation on power.
 
-The definition of the variance is:
+Sobol indices are based on a variance decomposition \cite{SOBOL2001271}. The definition of the variance is:
 \begin{equation}\label{tot_var}
-D = \text{Var}[f(\mathbf{z})] = \int_{\mathcal{D}_{\mathbf{z}}} f^2(\mathbf{z})d\mathbf{z} - f_0^2.
+D = \text{Var}[f(\mathbf{z})] = \int_{\mathcal{D}_{\mathbf{z}}} f^2(\mathbf{z})d\mathbf{z} - f_0^2,
 \end{equation}
-As the expansion in \eqref{sobol_decomp} is not unique, some orthogonality conditions must be imposed  \cite{Rabitz1999,SOBOL2001271} which allows to express the total variance as
+which is rewritten as
+%As the expansion in \eqref{sobol_decomp} is not unique, some orthogonality conditions must be imposed  \cite{Rabitz1999,} which allows to express the total variance as
 \begin{equation}
@@
 \end{equation}
-where the first- and second-order partial variances are defined via the Sobol decomposition as
+where the first- and second-order partial variances are defined via decomposition \eqref{sobol_decomp} as
 \begin{align}
@@
 %\end{equation}
-The total effect of $z_i$ on the output $Y$ can be quantified using the total sensitivity index
-\begin{equation}
-S_{T_i} = S_i + \sum_{j=1}^M S_{ij}.
-\end{equation}
+The total effect of $z_i$ on the output $Y$ can be quantified using the total sensitivity index $S_{T_i} = S_i + \sum_{j=1}^M S_{ij}$.
 The total sensitivity indices can be interpreted as an importance measure for the parameter $z_i$, therefore a large $S_{T_i}$ implies that $z_i$ has a strong influence on $Y$.
@@
 On the other hand, Polynomial Chaos Expansions (PCE) based approaches provide a more efficient alternative for sensitivity analysis, and will be used here.
-As shown in \cite{SUDRET2008964}, one can compute the Sobol indices analytically by post-processing the PCE coefficients.
+As shown in \cite{SUDRET2008964}, one can compute the Sobol indices analytically by post-processing the PCE coefficients, whose computation will be detailed next.
 
 \subsection{PCE-based Sobol indices computation}
-The PCE approximation $f^{K}(\mathbf{z})$ of the computational model $Y$ can be defined as a weighted sum of multivariate polynomials in $\mathbf{z}$ \cite{RSmith}
+A PCE approximation $f^{K}(\mathbf{z})$ to the computational model $Y=f(\mathbf{z})$ can be defined as a weighted sum of multivariate polynomials in $\mathbf{z}$ \cite{RSmith}:
 \begin{equation}\label{eq:PCE}
@@
 \end{equation}
-where $\mathbf{k}\in \mathbb{N}_0^M$ is a $M$-dimensional multi-index with magnitude $|\mathbf{k}| = k_1+k_2 + ...
-+ k_M$ and $\Psi_{\mathbf{k}}(\mathbf{z})$ a multivariate polynomial basis, which is computed using a tensor product of univariate polynomials $\psi_{k_i}^{(i)}(z_i)$
+where $\mathbf{k}\in \mathbb{N}_0^M$ is multi-index with magnitude $|\mathbf{k}| = k_1+k_2 + ...
++ k_M$ and $\Psi_{\mathbf{k}}(\mathbf{z})$ a multivariate polynomial basis, computed using a tensor product of univariate polynomials $\psi_{k_i}^{(i)}(z_i)$
 \begin{equation}
@@
 where $k_i$ is the order of the univariate polynomial.
-The choice of the univariate orthogonal polynomial $\psi_{k_i}$ depends on the type of random variable, $z_i$.
-For example, for uniform random variables the Legendre family of polynomials is used \cite{Xiu2002}.
+The choice of univariate orthogonal polynomial $\psi_{k_i}$ depends on the type of the random variable $z_i$.
+For example, for uniform random variables the Legendre family of polynomials is used.
 %The multivariate polynomials $\Psi_{\mathbf{k}}(\mathbf{z})$ is orthogonal with respect to the joint probability distribution of random vector $\mathbf{z}$. 
-The total number of terms $K$ in \eqref{eq:PCE} can be based for example on a standard truncation scheme in which all polynomials in $M$ random variable with order less than or equal to $p$ are used.
-In this work, a more advanced truncation schemes known as the hyperbolic truncation scheme is used \cite{BlatmanThesis}.
+The total number of terms $K$ in \eqref{eq:PCE} depends on the truncation scheme; in a standard truncation scheme  all polynomials in $M$ random variable with order less than or equal to $p$ are used.
+In this work, a more advanced truncation scheme known as the hyperbolic truncation scheme is used \cite{BlatmanThesis}.
 %\textcolor{red}{In this work, we use the hyperbolic truncation scheme with a $qnorm = 0.75$.}
@@
 For a small number of dimensions this is an efficient approach.
-However, in the current work we will encounter more than 10 dimensions and quadrature methods are typically too expensive as they suffer strongly from the curse of dimensionality.
-We therefore focus on methods in which the sampling points can be chosen more freely: Ordinary Least Squares (OLS) and Least-Angle Regression (LARS).
+However, in the current work we will encounter more than 10 dimensions and quadrature methods become  too expensive as they suffer strongly from the curse of dimensionality.
+We therefore focus on methods in which the sampling points can be chosen more freely and adaptively: Ordinary Least Squares (OLS) and Least-Angle Regression (LARS).
 %In the current work where we have a limited computational budget, methods that minimizes the approximation error for a given number of samples are more practical.
@@
 \subsubsection{Ordinary Least Squares}
-Instead of using a projection-based method one can also use least squares based approaches to compute coefficients \cite{Berveiller2006}.
-The PCE expansion \eqref{eq:PCE} can be expressed as a vector-matrix product
+%Instead of using a projection-based method one can also use least squares based approaches to compute the coefficients \cite{Berveiller2006}. 
+The PCE expansion \eqref{eq:PCE} can be expressed in terms of the following vector-matrix product
 \begin{equation}\label{eq:LSM}
@@
 \begin{equation}\label{eq:PCE_OLS}
-\mathbf{\overline{w}} = \argmin_{\mathbf{w}} \mathbb{E}\left[\left(\mathbf{w}^T \boldsymbol{\Psi}(\mathbf{z}) - f(\mathbf{z})\right)^2\right]
+\mathbf{\overline{w}} = \argmin_{\mathbf{w}} \mathbb{E}\left[\left(\mathbf{w}^T \boldsymbol{\Psi}(\mathbf{z}) - f(\mathbf{z})\right)^2\right].
  \end{equation}
@@
 where $\mathbf{X}\in \mathbb{R}^{N\times K}$ is the design matrix where $\mathbf{X}_{i,j} = \Psi_{j}(\mathbf{z}_i)$, $i = 1,2, ..., N$ and $j = 0,1,2,..., K$.
-Although the OLS method allows to compute the coefficients using a limited number of model evaluations, one of the main drawbacks of this approach is that it may result in a PCE model that consists of high-order interaction terms that may lead to overfitting.
+Although the OLS method allows to compute the coefficients using a limited number of model evaluations, it may result in a PCE model that consists of high-order interaction terms that may lead to overfitting.
 For many engineering problems, low-order interactions between the input variables are most important.
-Therefore, the least squares form in \eqref{eq:PCE_OLS} can be regularized such that the minimization results is a low-order sparse PCE model, as discussed next.
+Therefore, the least squares form in \eqref{eq:PCE_OLS} can be regularized such that the minimization results in a low-order sparse PCE model, as discussed next.
 
 \subsubsection{Least Angle Regression}
-The LARS algorithm \cite{efron2004} is one of the methods that can be applied to compute sparse PCE models \cite{BlatmanThesis, BLATMAN2011}.
+The LARS algorithm \cite{efron2004} is one of the methods that can be applied to compute sparse PCE models \cite{BlatmanThesis}.
 The key idea is to find the best set of polynomials from a given candidate set that minimizes the regularized form of \eqref{eq:PCE_OLS}:
@@
 The number of iterations is given by the minimum of $\{K,N\}$.
-After every iteration a posteriori error, for instance the Leave-One-Out (LOO) cross-validation error, is computed.
+After every iteration an a-posteriori error, for instance the Leave-One-Out (LOO) cross-validation error, is computed.
 The active set with the smallest LOO is then chosen as the best sparse PCE model.
@@
 
-\subsection{Sobol index computation}
+%\subsection{Sobol index computation}
 
@@
 
-\subsection{NM80}
-\subsubsection{Operating conditions}
+%\subsection{NM80}
+\subsection{Operating conditions}
 As a reference wind turbine, we consider the 2MW NM80 onshore turbine from the DANAERO project \cite{Troldborg2013} with a blade radius of 38.8 m and a hub height of 57.19 m with the mean wind speed at hub height set as 6.1m/s.
@@
 
-\subsubsection{Choice of uncertain inputs}
+\subsection{Choice of uncertain inputs}
 The data for lift ($C_l$) and drag ($C_d$) polars are available at four locations along the blade radius: at 11.87 m, 17.82 m, 28.97 m, and 35.53 m.
@@
 
-\subsubsection{Verification of sensitivity analysis method}
+\subsection{Verification of sensitivity analysis method}
 We perform a convergence study for both OLS and LARS in order to determine which method converges fastest with increasing number of uncertain parameters and number of model runs.
@@
 
-\subsubsection{Sensitivity analysis results}
+\subsection{Sensitivity analysis results}
 In Fig.~\ref{fig:sensitivity_analysis_power} and \ref{fig:sensitivity_analysis_force}, we show the total order Sobol indices as a measure of the sensitivity of the different geometric and model parameters on the power output and the axial force, respectively, computed using LARS with $M=19$ uncertain parameters chosen from Fig.~\ref{perturbed_samples}.
@@
 \section{Conclusions}\label{sec:conclusions}
-We have shown how Sobol indices computed using sparse adaptive polynomial expansion can be used for high-dimensional global sensitivity analysis of both geometric and model uncertainties in BEM models.
+We have shown how Sobol indices computed using sparse adaptive polynomial chaos expansion can be used for high-dimensional global sensitivity analysis of both geometric and model uncertainties in BEM models.
 The proposed framework is generic and can deal with many types of (parametric uncertainties) The sensitive model parameters identified here will be utilized in future work to develop calibrated BEM models with built-in uncertainty estimates, as part of the \textit{WindTrue} project, and used to better understand the differences between BEM models and measurement data as studied in IEA Task 29.

### 2020-02-26 Papers/Torque_Paper/article_IOP.tex
updates section 4-5
minor updates to paper
* edited conclusion
* changed x to r
* description of RVs
minor updates
- reducing text
- updating references
full paper draft, sent to ECN for check

@@
 %\usepackage[toc]{appendix}
+\usepackage{placeins}
 \usepackage{hyperref}
@@
 \newcommand{\Lomd}{L^2(\Omega,\mathcal{D})}
+
+\DeclareMathOperator*{\argmax}{arg\,max}
+\DeclareMathOperator*{\argmin}{arg\,min}
 
@@
 
-\author{P.~Kumar$^{1}$, B.~Sanderse$^{1}$, K.~Boorsma$^{2}$, M.~Caboni$^{2}$} \ead{b.sanderse@cwi.nl}
+\author{P.~Kumar$^{1}$, B.~Sanderse$^{1}$, K.~Boorsma$ ^{2}$, M.~Caboni$^{2}$} \ead{b.sanderse@cwi.nl}
 \address{$^{1}$ Centrum Wiskunde \& Informatica (CWI), Amsterdam, The Netherlands}
@@
 \begin{abstract}
-A global sensitivity analysis of model parameters associated with the Blade Element Momentum (BEM) models is performed.
+A framework is presented for performing global sensitivity analysis of model parameters associated with the Blade Element Momentum (BEM) models.
 The Sobol indices based on a sparse polynomial  expansion are used as a measure of global sensitivities.
@@
 
+\vspace{-1cm}
 %\begin{keyword}
@@
 %\newpage
+\section*{References}
 \bibliography{\myreferences}

### 2020-02-26 Papers/Torque_Paper/content.tex
full paper draft, sent to ECN for check

@@
 
-Recently, several papers have addressed the uncertainty in BEM model output by performing uncertainty propagation and sensitivity studies, e.g.\ \cite{Echeverria2017,Matthaus2017,Murcia2018,Robertson2018,Bos2019a}.
+Recently, several papers have addressed the uncertainty in BEM model output by performing uncertainty propagation and sensitivity studies, e.g.\ \cite{Echeverria2017,Matthaus2017,Murcia2018,Robertson2018,Bos2019c}.
 In these studies, the focus in mainly on uncertainties in the external conditions (wind parameters) and/or uncertainty in the turbine specification (geometric parameters).
 Apart from understanding how uncertainty in the output is related to the different uncertain inputs, such sensitivity studies are very useful to reduce the number of parameters as needed for example in design optimization \cite{Echeverria2017}.
-However, the uncertainty in BEM model output as caused by uncertainty in the model formulation itself, e.g.\ through the values chosen for model parameters, has been given little attention (exceptions being the effect of aerodynamic properties studied in \cite{Bortolotti2019,Matthaus2017})\todo{this is also what we do}.
+However, the uncertainty in BEM model output as caused by uncertainty in the model formulation itself, e.g.\ through the values chosen for model parameters, has been given little attention.
+% (exceptions being the effect of aerodynamic properties studied in \cite{Bortolotti2019,Matthaus2017})\todo{this is also what we do}.
 
@@
 The main novelty of the work lies in creating a generic framework for studying parametric uncertainteis, including both model uncertainties and geometric uncertainties.
-As such, the current sensitivity analysis is part of the so-called WindTrue project, in which the long-term goal is the development of calibrated BEM models.
+This framework for sensitivity analysis is part of the so-called \textit{WindTrue} project, in which the long-term goal is the development of calibrated BEM models.
 Anticipating on this goal, we are using the NM80 wind turbine model from the Tjaereborg wind farm, for which extensive measurement data is available through the DanAero project \cite{Troldborg2013} and IEA Task 29.
-The current sensitivity analysis will make clear which of the model parameters are to be used in the calibration phase.
+The sensitivity analysis will make clear which of the model parameters are to be used in the calibration phase.
 
@@
 \end{itemize}
+In this paper we will focus on the first uncertainty -- uncertainty in lift and drag coefficients -- but we stress that the framework and methodology are directly applicable to other BEM model uncertainties as well.
+In a parallel paper, see \cite{Caboni2020}, the effect of other uncertainties such as the dynamic stall model will be investigated.
 Next to these model uncertainties we will consider geometric uncertainties: the chord and twist distribution along the blade.
@@
 Also note that when using a uniform distribution with bounds given by the relative magnitude of control points, we tend to obtain small perturbations for control points with values close to zero.
-\begin{figure}[h!]
+\begin{figure}[ht!]
 \centering
@@
 
+%\FloatBarrier
+
 \subsection{Sensitivity analysis results}
@@
 
+\FloatBarrier
+
 \section{Conclusions}\label{sec:conclusions}
-We have shown how Sobol indices computed using sparse adaptive polynomial chaos expansion can be used for high-dimensional global sensitivity analysis of both geometric and model uncertainties in BEM models.
-The proposed framework is generic and can deal with many types of (parametric uncertainties) The sensitive model parameters identified here will be utilized in future work to develop calibrated BEM models with built-in uncertainty estimates, as part of the \textit{WindTrue} project, and used to better understand the differences between BEM models and measurement data as studied in IEA Task 29.
+In this paper we have developed a framework for global sensitivity analysis using Sobol indices computed using sparse adaptive polynomial chaos expansions.
+The framework can be used for high-dimensional global sensitivity analysis of both geometric and model uncertainties in BEM models.
+As such, the proposed framework is generic and can deal with many types of (parametric) uncertainties, of which a few were investigated in this paper: twist, chord, lift coefficient and drag coefficient.
+By using NURBS parameterizations of chord, twist, lift and drag as a function of the blade radius, the number of uncertainties could be limited to 19.
+The results highlight amongst others the importance of the lift coefficient, especially for the axial force prediction.
+The identified sensitivities will be utilized in future work to develop calibrated BEM models with built-in uncertainty estimates, as part of the \textit{WindTrue} project, and used to better understand the differences between BEM models and measurement data as studied in IEA Task 29.

### 2020-03-03 Papers/Torque_Paper/article_IOP.tex
final Torque Paper as submitted

@@
 
-\title{A global sensitivity analysis of model uncertainty in aeroelastic wind turbine models}
+\title{Global sensitivity analysis of model uncertainty in aeroelastic wind turbine models}
 
@@
 A framework is presented for performing global sensitivity analysis of model parameters associated with the Blade Element Momentum (BEM) models.
-The Sobol indices based on a sparse polynomial  expansion are used as a measure of global sensitivities.
-The sensitivity analysis workflow is developed using the uncertainty quantification toolbox \texttt{UQLab} that is integrated with the ECN's \texttt{Aero-Module} aeroelastic code.
-Sensitivity studies are performed on the NM80 wind turbine model from the DANAERO project.
+Sobol indices based on adaptive sparse polynomial expansions are used as a measure of global sensitivities.
+The sensitivity analysis workflow is developed using the uncertainty quantification toolbox \texttt{UQLab} that is integrated with TNO's \texttt{Aero-Module} aeroelastic code.
+Uncertainties in chord, twist, and lift- and drag-coefficients have been parametrized through the use of NURBS curves.
+Sensitivity studies are performed on the NM80 wind turbine model from the DanAero project, for a case with 19 uncertainties in both model and geometry.
+The combination of parametrization and sparse adaptive polynomial chaos yields a new efficient framework for global sensitivity analysis of aeroelastic wind turbine models, paving the way to effective model calibration.
 \end{abstract}
@@
 
-%\section*{Acknowledgements}
+\section*{Acknowledgements}
+The research was sponsored by the Topsector Energy Subsidy from the Dutch Ministry of Economic Affairs and Climate.
 

### 2020-03-03 Papers/Torque_Paper/content.tex
final Torque Paper as submitted

@@
 Aeroelastic models such as the Blade Element Momentum (BEM) models \cite{HandBook} play a critical role in the design, development, and optimization of modern wind turbines.
-A large number of BEM models have been developed to predict turbine responses such as the structural loads and power output \cite{Vorpahl2013}.
+A large number of BEM models have been developed to predict turbine responses such as structural loads and power output \cite{Vorpahl2013}.
 
@@
 We will employ the uncertainty quantification toolbox \texttt{UQLab} \cite{uqlab}, which computes the Sobol indices based on a sparse polynomial chaos expansion.
-The main novelty of the work lies in creating a generic framework for studying parametric uncertainteis, including both model uncertainties and geometric uncertainties.
+The main novelty of the work lies in creating a generic framework that enables an advanced (global) sensitivity analysis of a relatively high number of uncertain parameters (e.g.\ 10-20), including both model and geometric uncertainties.
+The two key ingredients for this are efficient parameterization and the adaptive sparse polynomial expansion.
 This framework for sensitivity analysis is part of the so-called \textit{WindTrue} project, in which the long-term goal is the development of calibrated BEM models.
-Anticipating on this goal, we are using the NM80 wind turbine model from the Tjaereborg wind farm, for which extensive measurement data is available through the DanAero project \cite{Troldborg2013} and IEA Task 29.
+Anticipating on this goal, we are using the NM80 wind turbine model from the Tjaereborg wind farm, for which extensive measurement data is available through the DanAero project \cite{Troldborg2013} and IEA Task 29  Phase IV (Phase IV is a follow-up of Task 29 Phase III \cite{Boorsma2018} that is built around the DanAero experiment).
 The sensitivity analysis will make clear which of the model parameters are to be used in the calibration phase.
@@
 In this work, we concentrate on the first aspect, namely the prediction of flow and blade forces as given by the BEM method.
-The particular BEM code that we use is ECN's \texttt{Aero-Module} \cite{Boorsma2012}.
+The particular BEM code that we use is TNO's \texttt{Aero-Module} \cite{Boorsma2012}.
 For the case of a rigid turbine, with a uniform inflow field, the main uncertainties that arise in the BEM formulation are \cite{Hansen1993}:
@@
 %\item Sheared inflow
-\item The assumption of locally 2D flow along the blade is also invalid at the tip; this is typically corrected via \textit{a tip loss model};
-\item Three-dimensional flow effects also cause stall delay at the root; this is typically modeled via \textit{a dynamic stall model} and associated time constant;
-\item Yawed inflow is not naturally included and typically incorporated via \textit{a wake correction}.
+\item The assumption of an azimuthally uniform induction distribution; this is typically corrected via \textit{a tip loss model};
+\item Three-dimensional flow effects also cause stall delay at the root; this is typically modelled via \textit{a 3D correction or rotational augmentation model};
+\item Yawed inflow is not naturally included and incorporated via \textit{a skewed wake correction}.
 %\item Dynamic wake - Oye time lag constant
@@
 In this paper we will focus on the first uncertainty -- uncertainty in lift and drag coefficients -- but we stress that the framework and methodology are directly applicable to other BEM model uncertainties as well.
-In a parallel paper, see \cite{Caboni2020}, the effect of other uncertainties such as the dynamic stall model will be investigated.
-Next to these model uncertainties we will consider geometric uncertainties: the chord and twist distribution along the blade.
+In a parallel paper, see \cite{Caboni2020}, the effect of other uncertainties such as the unsteady airfoil aerodynamics model will be investigated.
+Next to these model uncertainties we will consider geometric uncertainties: the chord and twist distribution along the blade, to show the generic applicability of the framework.
 
@@
 \end{equation}
-and the perturbation is randomly sampled as $\Delta_{i} \sim \mathcal{U}[-0.05,0.05]$.
-This equation is applicable to both the twist- and chord distribution, where $\tilde{c}_{i}$ can denote either the value of the twist control points, $\text{Tw}_i$, or of the chord control points, $\text{Ch}_{i}$.
+and the perturbation is uniformly sampled as $\Delta_{i} \sim \mathcal{U}[-0.05,0.05]$.
+Other distributions can be specified in the framework; in future work we will further investigate how to choose these, for example based on design standards.
+This equation is applicable to both twist- and chord distributions: $\tilde{c}_{i}$ can denote either the value of the twist control points, $\text{Tw}_i$, or of the chord control points, $\text{Ch}_{i}$.
 For the lift- and drag polars, we use the following parametrization:
@@
 \end{equation}
-where $c_{i}(\alpha)$ can denote either the lift coefficient $\text{Cl}_{i}$ or drag coefficient $\text{Cd}_{i}$, and the subscript $\text{ref}$ denotes the reference curve.
+where $c_{i}(\alpha)$ can denote either the lift coefficient $\text{Cl}_{i}$ or drag coefficient $\text{Cd}_{i}$.%, and the subscript $\text{ref}$ denotes the reference curve.
 % $c_{\text{ref},i}(\alpha)$ can denote either the reference lift coefficient $C_{l,\text{ref},i}(\alpha)$ or drag coefficient $C_{d,\text{ref},i}(\alpha)$ at section $i$.
@@
 We employ the variance-based Sobol decomposition to perform global sensitivity analysis.
-This approach allows for quantification of the relative importance of the input parameters on a scale of $[0,1]$ known as \emph{Sobol indices}.
+This approach allows for quantification of the relative importance of the input parameters known as \emph{Sobol indices}.
 In a Sobol analysis, we express the total variance of the output in term of contributions from individual parameters and their combinations.
@@
 where $\mathbf{k}\in \mathbb{N}_0^M$ is multi-index with magnitude $|\mathbf{k}| = k_1+k_2 + ...
-+ k_M$ and $\Psi_{\mathbf{k}}(\mathbf{z})$ a multivariate polynomial basis, computed using a tensor product of univariate polynomials $\psi_{k_i}^{(i)}(z_i)$
-\begin{equation}
-\Psi_{\mathbf{k}}(\mathbf{z}) := \prod_{i=1}^M\psi_{k_i}^{(i)}(z_i),
-\end{equation}
++ k_M$ and $\Psi_{\mathbf{k}}(\mathbf{z})$ a multivariate polynomial basis, computed using a tensor product of univariate polynomials $\psi_{k_i}^{(i)}(z_i)$,
+%\begin{equation}
+%\Psi_{\mathbf{k}}(\mathbf{z}) := \prod_{i=1}^M\psi_{k_i}^{(i)}(z_i),
+%\end{equation}
 where $k_i$ is the order of the univariate polynomial.
@@
 The coefficient vector $\mathbf{\overline{w}}$ can be computed using the Ordinary Least Squares (OLS) method.
-Given $N$ model evaluations $\mathbf{Y} = \{f(\mathbf{z}_1), f(\mathbf{z}_2), ..., f(\mathbf{z}_N)\}$, we can obtain the coefficients as:
-\begin{equation}\label{eq:OLS}
-\overline{\mathbf{w}} = (\mathbf{X}^T\mathbf{X})^{-1}\mathbf{X}^T\mathbf{Y},
-\end{equation}
-where $\mathbf{X}\in \mathbb{R}^{N\times K}$ is the design matrix where $\mathbf{X}_{i,j} = \Psi_{j}(\mathbf{z}_i)$, $i = 1,2, ..., N$ and $j = 0,1,2,..., K$.
+Given $N$ model evaluations $\mathbf{Y} = \{f(\mathbf{z}_1), f(\mathbf{z}_2), ..., f(\mathbf{z}_N)\}$, we can obtain the coefficients as $\overline{\mathbf{w}} = (\mathbf{X}^T\mathbf{X})^{-1}\mathbf{X}^T\mathbf{Y}$, where $\mathbf{X}\in \mathbb{R}^{N\times K}$ is the design matrix where $\mathbf{X}_{i,j} = \Psi_{j}(\mathbf{z}_i)$, $i = 1,2, ..., N$ and $j = 0,1,2,..., K$.
 Although the OLS method allows to compute the coefficients using a limited number of model evaluations, it may result in a PCE model that consists of high-order interaction terms that may lead to overfitting.
@@
 This polynomial enters the active set.
-Next, we take the largest step in the direction of $\Psi_1(\mathbf{w})$ until some other polynomial $\Psi_2(\mathbf{w})$ from the candidate set, has as much correlation with the current residual.
+Next, we take the largest step in the direction of $\Psi_1(\mathbf{w})$ until another polynomial $\Psi_2(\mathbf{w})$ from the candidate set, has as much correlation with the current residual.
 The second polynomial is now also added to the active set.
@@
 The active set with the smallest LOO is then chosen as the best sparse PCE model.
-
-One of the main advantage of the LARS algorithm is that it works well in cases when $K>>N$, i.e. the number of polynomials in the candidate set is much larger than the number of model evaluations.
+One of the main advantage of the LARS algorithm is that it works well in cases when $K>>N$, i.e.
+the number of polynomials in the candidate set is much larger than the number of model evaluations.
 
@@
 As a reference wind turbine, we consider the 2MW NM80 onshore turbine from the DANAERO project \cite{Troldborg2013} with a blade radius of 38.8 m and a hub height of 57.19 m with the mean wind speed at hub height set as 6.1m/s.
-Furthermore, the RPM is set at 12.3$/\text{min.}$, the blade pitch angle at 0.15 $\deg$ and the turbine yaw angle is zero.
+Furthermore, the rotational speed is set at 12.3 rpm, the blade pitch angle at 0.15 $\deg$ and the turbine yaw angle is zero.
 
@@
  \end{subfigure}
-    \caption{Convergence of OLS and LARS with increasing number of uncertain inputs $M$ and model evaluations $N$.}
+    \caption{Convergence of OLS and LARS with increasing number of uncertain inputs $M$ (chord control points) and model evaluations $N$.}
 \label{convergence_dim}
@@
 Lastly, we observe that the axial force is basically only sensitive to the lift coefficient at the outward blade sections, whereas the power is sensitive to both lift and drag coefficients, also at the inward section of the blade.
+We note that these findings are testcase-specific and might not generalize directly to other turbines or other operational settings.
+Furthermore, the reported sensitivities depend on the variance that is chosen for the uncertain parameters; in future work we will relate these to values known from design standards.
 
@@
 
-\FloatBarrier
+%\FloatBarrier
 
 \section{Conclusions}\label{sec:conclusions}
-In this paper we have developed a framework for global sensitivity analysis using Sobol indices computed using sparse adaptive polynomial chaos expansions.
+In this paper we have developed a framework for global sensitivity analysis using Sobol indices computed using adaptive sparse polynomial chaos expansions.
 The framework can be used for high-dimensional global sensitivity analysis of both geometric and model uncertainties in BEM models.
 As such, the proposed framework is generic and can deal with many types of (parametric) uncertainties, of which a few were investigated in this paper: twist, chord, lift coefficient and drag coefficient.
-By using NURBS parameterizations of chord, twist, lift and drag as a function of the blade radius, the number of uncertainties could be limited to 19.
+By using NURBS parameterizations of chord, twist, lift and drag as a function of the blade radius, the number of uncertainties was limited but remains considerable (19 uncertain variables).
+Our approach (adaptive sparse polynomial chaos expansion) is able to efficiently quantify this high-dimensional problem.
 The results highlight amongst others the importance of the lift coefficient, especially for the axial force prediction.

### 2020-05-05 Papers/Calibration/main.tex
Latex document

%% %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% Template for a conference paper, prepared for the
%% Food and Resource Economics Department - IFAS
%% UNIVERSITY OF FLORIDA
%% %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% Version 1.0 // November 2019
%% %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% Ariel Soto-Caro
%%  - asotocaro@ufl.edu
%%  - arielsotocaro@gmail.com
%% %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\documentclass[11pt]{article}
\usepackage{UF_FRED_paper_style}

\usepackage{lipsum}  %% Package to create dummy text (comment or erase before start)

%% ===============================================
%% Setting the line spacing (3 options: only pick one)
% \doublespacing
% \singlespacing
\onehalfspacing
%% ===============================================

\setlength{\droptitle}{-5em} %% Don't touch

% %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% SET THE TITLE
% %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

% TITLE:
\title{Bayesian calibration of AeroModule code using DanAero data
% \thanks{Selected Paper prepared for presentation at the 201X Agricultural \& Applied Economics Association Annual Meeting}
}

% AUTHORS:
\author{Vinit Dighe\\% Name author
    \href{mailto:vinit@cwi.nl}{\texttt{vinit@cwi.nl}} %% Email author 1 
\and Benjamin Sanderse\\% Name author
    \href{mailto:b.sanderse@cwi.nl}{\texttt{b.sanderse@cwi.nl}} %% Email author 2
% \and Third Author\\% Name author
%     \href{mailto:thirdauthor@ufl.edu}{\texttt{thirdauthor@ufl.edu}}%% Email author 3
%\and Forth Author\\% Name author
%    \href{mailto:forthuthor@ufl.edu}{\texttt{forthuthor@ufl.edu}}%% Email author 4
    }

% DATE:
\date{\today}

% %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{document}
% %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% ABSTRACT
% %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
{\setstretch{.8}
\maketitle
% %%%%%%%%%%%%%%%%%%
\begin{abstract}
% CONTENT OF ABS HERE--------------------------------------

\noindent A Bayesian framework using UQLab has been applied to calibrate the 3D correction factor used in the  axial force model implemented in the AeroModule code.   %% Dummy text for abstract. Erase before use.

% END CONTENT ABS------------------------------------------
\noindent
% \textit{\textbf{Keywords: }%
% key1; key2; key3; key4.} \\ %% <-- Keywords HERE!
% \noindent
% \textit{\textbf{JEL Classification: }%
% Q12; C22; D81.} %% <-- JEL code HERE!

\end{abstract}
}

% %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% BODY OF THE DOCUMENT
% %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

% --------------------
\section{Introduction}

The wind energy industry is experiencing remarkable growths annually. Despite the great progress made, further cost reductions in wind turbine technology are necessary for wind energy to reach its full potential in terms of the large‐scale supply of electricity. Improving the reliability of aerodynamic models embedded in the design software currently used in industry is indispensable to guarantee reductions in the cost of wind energy.

Due to its relatively high computational efficiency compared to free‐wake vortex methods and CFD, the Blade‐Element‐Momentum (BEM) theory still forms the basis for many aerodynamic models. Yet various experimental campaigns have demonstrated that BEM‐based design codes are not always sufficiently reliable for predicting the aerodynamic load distributions on the wind turbine blades \citep{UAE_exp}. In many situations, the reliability of such codes was found to be unacceptable, in particular when comparing the predicted results by different aerodynamics/aeroelastic codes from various universities/institutions with the measurement data obtained from the UAE experimental campaigns. In some cases, deviations of the BEM predictions from the measurements exceeded 200\%, even though the simplest operating conditions of a wind turbine were being considered (i.e. uniform wind-speed and constant rotor speed, blade pitch and yaw angle).

Wind tunnel tests on model turbines are indispensable to have a better understanding of the underlying physics and to improve engineering models for design codes. The controlled environment offered by a wind tunnel provides a set of measurements that is free from the uncertainties caused by the different atmospheric effects that are always present in open field tests of turbines. To improve the predictions of BEM‐based design codes, more reliable airfoil data models and inflow correction models are required. However, using the experimental data to improve these models is not an easy task. Two major problems are encountered: Firstly, measurement data is usually rare and limited. Secondly, there is a difficulty in determining accurately the angle of attack for the operating turbine.

In this investigation, we describe an approach for combining
observations from DanAero field experiments with predictions obtained using the TNO’s AeroModule aeroelastic (BEM) code  to carry out statistical
inference. Of particular interest here is determining uncertainty in the resulting predictions. This typically involves calibration of parameters in the computer model in response to the sensitivity analysis study carried by \citet{torque2020}.

In recent times, there has been increasing efforts in using Bayesian approach for the calibration of computer models. This is because of its ability to quantify uncertainties in input parameters while at the
same time reducing discrepancies between simulation
output and physical measurements. In the Bayesian
calibration of computer models, Markov Chain
Monte Carlo (MCMC) methods are a common way
for sampling from the posterior distributions of the
calibration parameters. Its widespread use can be
attributed to its ease of use in a wide variety of
problems.

\section{DanAero experiment}

\begin{figure}
\centering
\includegraphics[width=0.65\textwidth]{exp.png}
\caption{Experimental blade forces denoting the mean normal forces along the blade radius $R$}
\label{pt1}
\end{figure}

A measurement campaign was carried out as part of the DanAero MW experiment \citep{DanAero}. A 2.3MW NM80 turbine located at the Tjæreborg Enge site and a nearby met mast were both instrumented with various sensors. ScaniValve system was used for the surface pressure measurements on the turbine blade.

Besides the data obtained directly from the Tjæreborg Enge site,  a number of wind tunnel tests on different airfoils of NM80 turbine were also carried out in different wind tunnels \citep{tunnel}. The objective of the wind tunnel tests was to measure the lift, drag, moment, normal force coefficient and tangential force coefficient at four blade sections (13m, 19m, 30m, and 37m) of the LM38.8 blade (38.8m) which were instrumented with pressure taps.

The Tjæreborg experiment  database contains 35 Hz measurements and 10 minute statistics (mean, minimum, maximum and standard deviation) of all the data acquired by the DaqWin system as well as down sampled surface pressure measurements from five selected pressure taps at each of the four blade sections. All the data from the  experiments are available in data files with formats as explained in \citep{DanAero,tunnel}

\section{BEM model - AeroModule}

The aeroelastic BEM code - AeroModule  is used to model the aerodynamic and aeroelastic behaviour of wind turbine blades by combining the concept of momentum conservation of the flow (aerodynamic analysis) with the equations
of motion (structural analysis). In this work, we focus on the first aspect, namely the prediction of aerodynamic blade forces as given by the BEM method. For the purpose of calibration, normal aerodynamic forces (mean values) measured under non-sheared and non-yawed inflow conditions  has been selected from the measurements as described in \citep{DanAero}. The normal aerodynamic force on the blade surface is  calculated based on the following equation:

\begin{equation}
    F_n = 0.5\rho W^{2}c({{}{C}_l}_{3D} sin\beta + {{}{C}_d}_{3D} cos\beta),
\label{eq1}
\end{equation}

\noindent where $\rho$ is the fluid density, $W$ is the relative flow angle, $c$ is the chord length and $\beta$ is the relative flow angle. The 3D correction factor \citep{snel} has been applied to account for the effects of rotation on the lift and drag coefficients ($C_l$ and $C_d$). Then, the lift coefficient in equation \ref{eq1}  takes the following form:

\begin{equation}
    {{}{C}_l}_{3D} = 3.1(\frac{c}{r})^2(\frac{\Omega r}{W})^2({{}{C}_l}_{pot} - {{}{C}_l}_{2D})
\end{equation}

\noindent where $c$ is the local chord, $r$ is the local radius, $\Omega$ is the rotational speed of the rotor, ${{}{C}_l}_{pot}$ is the potential flow lift coefficient and ${{}{C}_l}_{2D}$ is the 2D lift coefficient.

\section{Sensitivity analysis}

Before calibrating the BEM model, sensitivity analysis was performed to identify the parameters that have the most influence over the model’s output ($F_n$). The objective is to reduce the number of calibration parameters and choose only important parameters in the calibration process. This not only reduces computation cost but also
helps mitigate over-fitting. \citet{torque2020} observed that the model's output is basically only sensitive to the lift coefficient $C_l$ at the outward blade sections.

\begin{figure}
\centering
\includegraphics[width=0.70\textwidth]{ Sobol.pdf}
\label{pt1}
\caption{Sensitivity of axial force towards geometric and model uncertainties.}
\end{figure}

\section{Lift coefficient parametrization}

\citet{snel} derived a so-called `3D correction’ method  that gives an increase of the aerodynamic lift coefficient for the effects of rotation. The 3D factor 3.1 was originally used to fit with the DanAero measurements. Following the comparison of the UAE rotor experiment with the BEM code predictions, \citet{UAE_exp}  concluded that this factor could
as well be reduced to 3. Because it is
known that the term $(\frac{c}{r})^2$ is an empirical fit which gives a stronger dependency on $(\frac{c}{r})$ than
what is found from theory, it was decided to calibrate the the  3D factor; viz. a constant value for the range of $\alpha$ considered. Assuming a constant value of k, the $C_l$ curve at point 4 is  parametrized:

\begin{equation}
    {{}{C}_l}_{3D} = k(\alpha)(3.1 \pm \Delta)
    \label{eq3}
\end{equation}

\noindent where $k$ = $(\frac{c}{r})^2(\frac{\Omega r}{W})^2({{}{C}_l}_{pot} - {{}{C}_l}_{2D})$ as a function of angle of attack $\alpha$.

\begin{figure}
\centering
\includegraphics[width=0.65\textwidth]{ snel.png}
\label{pt1}
\caption{Realization of lift polar at point 4 using uniform distribution of $\Delta$}
\end{figure}

\section{Bayesian calibration}

Bayesian methods are used to estimate model parameter $\theta$  by combining two sources of information: prior information about $\theta$ and observations $D$ on output variables $x$. The prior information of $\theta$ is based on expert knowledge, literature review or by measuring parameters directly in the field or laboratory. In our case, $D$  as a function of $x$ are field measurements from the DanAero experiment. Bayes theorem makes it possible to combine the two
sources of information in order to calibrate $\theta$. The first step is to assign a probability distribution to $\theta$, representing our prior uncertainty about the value. In our case, we specified lower and upper bounds of the $\theta$'s uncertainty, defining the prior parameter distributions as uniform. The aim of Bayesian calibration is to reduce this uncertainty by using the $D$, thereby producing the posterior distribution for $\theta$. This is achieved by multiplying the prior with the likelihood function, which is the
probability of the $D$ given the $\theta$. The likelihood function is determined by the probability distribution of errors in
observations. To generate a representative sample of $\theta$ from the posterior distribution, we used  Markov Chain Monte Carlo (MCMC) methods.

\begin{center}
\begin{tabular}{ |c|c| } 
 \hline
 Symbol & Description  \\ 
 \hline
$D$ & Axial force at corresponding radial locations   \\ 
 $x$ & Angle of attack  \\ 
 $\theta$ & 3D factor\\
 \hline
\end{tabular}
\end{center}

\subsection{Metropolis-Hastings algorithm}

In the Metropolis–Hastings (MH) algorithm, a chain is initialized at a certain seed point $\theta^{(0)}$  belonging to an admissible domain. At iteration $t$ from the current point $x^{(t)}$, the algorithm randomly draws a new `candidate' parameter:

    \begin{equation*}
        \theta^{(*)} = \theta^{(t)} + \delta
    \end{equation*}
    \noindent where $\delta$ is a random vector generated from an uniform distribution on the interval (3.0,  3.2).

    Subsequently, the candidate is accepted with probability, if:

    \begin{equation*}
        \alpha = min \bigg\{1, \frac{p(\theta^{(*)}|D)}{p(\theta^{(t)}|D)}\bigg\},
    \end{equation*}

    % \begin{equation*}
    %     a = \frac{p(\theta^{*}|D)}{p(\theta_{t}|D)} = \frac{p(D|\theta^{*}) p(\theta^{*})}{p(D|\theta_{i-1})p(\theta_{i-1})}
    % \end{equation*}

    \noindent and rejected otherwise. It is important to note that the model evidence term cancels out from the acceptance probability. This avoids the calculation of the often intractable
integral. The candidate point $\theta^{(*)}$ is always accepted if its posterior value is no lower than the posterior value of $\theta^{(t-1)}$. Once the chain has attained the
N iterations, the chain must have converged to the target distribution
which is the posterior $\theta$ distribution.

\subsection{Adaptive Metropolis algorithm}

The basic idea of Adaptive Metropolis (AM) algorithm is to update the proposal distribution $p(\theta^{(*)}|D)$ by using the knowledge  acquired about the target distribution. Otherwise the definition of the AM algorithm is identical to the MH algorithm.

Suppose, therefore, that at time t we have sampled the states $x_{0}$, $x_{1}$, . . . , $x_{t}$, where $x_{0}$ is the initial state. Then, the posterior probability of the new 'candidate' parameter $\theta^{*}$ is accepted:

\begin{equation*}
        \alpha = min \bigg\{1, \frac{p(\theta^{(*)}|D)}{p(\theta^{(t)}|D)}\bigg\}, otherwise 
    \end{equation*}

\noindent where the proposal distribution $p(\theta^{(*)}|D)$ employed in AM algorithm is a Gaussian distribution with mean at the current point $\theta^{(t)}$ and covariance $C^{(t)}$ = $C^{(t)}$($x_{0}$, $x_{1}$, . . . , $x_{t}$). The AM algorithm is particularly useful over MH algorithm when a large number of samples from the same distribution is necessary (Bayesian calibration), and in CPU intensive applications, for example, optimization problems.

\subsection{Affine invariant ensemble algorithm}

MH and AM algorithm perform poorly when the target (i.e., posterior) distribution shows
strong correlation between the parameters. The performance of these algorithms can typically only be improved by considerable amount of tuning. The affine invariant ensemble algorithm (AIES)  reduces this problem. AIES algorithm has the property of being invariant to affine transformations to the target distribution. This means that if there exists an affine transformation of the difficult-to-sample
(by standard MCMC methods) target distribution to an easier-to-sample target distribution,
AIES samples both distributions equally easily without explicitly requiring this affine transformation.

The affine invariance property is achieved by generating proposals according to a so-called stretch move. This refers to proposing the new 'candidate' parameter by:

\begin{equation*}
    \theta^{*} = \theta^{(t)} + Z \hspace{1cm} where \hspace{1cm} Z \sim p(z) = \begin{cases}
    \frac{1}{\sqrt{z}(2\sqrt{a}-\frac{2}{\sqrt{a}})},& \text{if } z \epsilon [\frac{1}{a},a],\\
    0,              & \text{otherwise}.
\end{cases}
\end{equation*}

\noindent This requires sampling from the distribution $p(z)$ defined by the tuning parameter $a$, which is often set to $a$ = 2. The
candidate $\theta^{(*)}$ is then accepted with probability:

\begin{equation*}
        \alpha = min \bigg\{1, z^{M-1} \frac{p(\theta^{(*)}|D)}{p(\theta^{(t)}|D)}\bigg\},
    \end{equation*}

\noindent where $M$ is a scaled identity matrix. A practical advantage of the AIES algorithm is that it only has a single scalar tuning parameter $a$, and on the other hand, due to
its sequential nature, the algorithm cannot be parallelized which makes it comparably slower in comparison to the MH and AM algorithms.

\section{Results and discussion}
\begin{figure}
\centering
\includegraphics[width=0.7\textwidth]{prior.png}
\caption{Prior distribution of the parameters used in the axial force model.}
\label{prior}
\end{figure}

The information from the AeroModule run was utilised
to construct the prior distributions shown in Figure \ref{prior}. We compare the effectiveness of three MCMC algorithms: MH, AM and AIES. The comparison was carried out using the following metrics:

\begin{itemize}
    \item Trace plots: Trace plots are plots of the Markov chains versus the sample steps and can be useful for assessing convergence. As chains are typically initialized at random points,  the distribution of points can give valuable insights about convergence. 
    \item Gelman-Rubin diagnostics (\^{R}): A quantitative approach to assess convergence was introduced by  \citet{gelman}. \^{R} is the ratio of
between independent chain variance to  the combination of all chain variance
and is based on the concept that if multiple
chains have converged, there should be little variability between independent and the combination of all chains. For convergence, \^{R} should be approximately 1 $\pm$ 0.1. 
\end{itemize}

For comparison, the same number of iterations were run for all three algorithms. The first 50\% samples are discarded as burn-in to reduce the influence of the starting values.

\begin{figure}[htp]

\centering
\includegraphics[width=.3\textwidth]{ trace_mh.png}\hfill
\includegraphics[width=.3\textwidth]{ trace_am.png}\hfill
\includegraphics[width=.3\textwidth]{ trace.png}

\caption{Trace plots for the same number of iterations obtained with MH (left), AM (center) and AIES (right) algorithms.}
\label{trace}

\end{figure}

\begin{figure}
\centering
\includegraphics[width=0.45\textwidth]{ post1.png}
\includegraphics[width=0.45\textwidth]{ y1.png}
\caption{Posterior distribution of $\delta$ (left) along with posterior distribution of $D$ (right).}
\label{post}
\end{figure}

Figure \ref{trace}  provides a visual comparison of
the trace plots (including burn-in)
with samples generated by the three different MCMC algorithms. MH and AM algorithms demonstrate
bad mixing for the calibration parameters $\theta$, indicating that the algorithm does not sufficiently
explore the parameter space. After 1000 iterations
the calibration parameters for MH and AM algorithms have \^{R} = 2.276 \& 5.96 respectively. AIES algorithm performs
better, with the calibration parameters $\theta$ achieving
adequate convergence (\^{R} = 1.04) after 1000 iterations.
Trace plot (Figure \ref{trace} (right)) also shows good mixing. As with the AIES algorithm, the posterior distribution for $\Delta$ is plotted in Figure \ref{post}. Then, as in equation \ref{eq3}, the calibrated  $\theta$ = 3.1 - 0.14 = 2.96.

\section{Conclusions}
Bayesian calibration framework using UQLab was tested using the DanAero data (axial force meaurement) to calibrate the 3D factor used in the AeroModule code. The governing forward model (axial force model) was obtained. This is followed by using the existing sensitivity analysis study to reduce the number of calibration parameters. The effectiveness of three MCMC algorithms (MH, AM and AIES) were
evaluated. Following the Bayesian calibration, the calibrated forward model suggests the 3D factor value equal to 2.96.
\medskip

\bibliography{references.bib}

\end{document}

### 2020-06-02 Papers/Torque_Paper/revision/content.tex
added Torque revision; updated folder structure

--- 2020-03-03 Papers/Torque_Paper/content.tex
@@
 Although local approaches are widely employed due to their ease of implementation and low computational cost, they are unable to quantify the global behaviour of nonlinearly parametrized models such as aeroelastic models.
-Global sensitivity approaches, on the other hand, consider the entire range of input values to compute output sensitivities.
-Therefore, a global sensitivity analysis is more suitable for the aeroelastic models considered in this work.
+Global sensitivity approaches, on the other hand, consider the entire range of input values to compute output sensitivities, and will be employed in this work.
+\hl{As this technique is still relatively new in the wind engineering community, we will describe it in some detail in the following sections.}
 %In engineering, sensitivity analysis can be employed for several reasons: to determine the stability and robustness of a computational model for input parameters, for simplification of stochastic models by fixing the insensitive parameters, and to guide data acquisition campaigns and experimental design to refine the data on sensitive parameters.
@@
 \subsection{Sobol analysis}
-We employ the variance-based Sobol decomposition to perform global sensitivity analysis.
-This approach allows for quantification of the relative importance of the input parameters known as \emph{Sobol indices}.
-In a Sobol analysis, we express the total variance of the output in term of contributions from individual parameters and their combinations.
+We employ the variance-based Sobol decomposition to perform global sensitivity analysis and quantify the relative importance of input parameters via \emph{Sobol indices}.
+The total variance of the output is expressed in term of contributions from individual parameters and their combinations.
 Consider a generic BEM model $f(\mathbf{z})$, depending on uncertain parameters $\mathbf{z}$ and returning a quantity of interest $Y$ (e.g.\ power, loads, moments):
@@
 where ${\mathbf{z}} = [z_1, z_2, ..., z_M]\in \mathcal{D}_{\mathbf{z}}\in \mathbb{R}^M$ is a vector of random variables (parameters).
-For ease of notation, we assume these input parameters to be uniformly distributed i.e.
-$z_i \sim \mathcal{U}(0,1)$ and the support of the input set is $\mathcal{D}_{\mathbf{z}} = [0,1]^M$ where $M$ is the total number of input parameters.
-The Sobol decomposition is based on the following hierarchical representation \cite{RSmith}:
+For ease of notation, we assume these input parameters to be uniformly distributed, i.e.\ $z_i \sim \mathcal{U}(0,1)$, and $\mathcal{D}_{\mathbf{z}} = [0,1]^M$ where $M$ is the total number of input parameters.
+The Sobol decomposition is based on the following representation \cite{RSmith}:
 \begin{equation}\label{sobol_decomp}
@@
 \end{equation}
-Here the zeroth-order function $f_0$ is the mean response of $f$, the first-order univariate functions $f_i(z_i)$ quantify independent contributions due to each random input, and second-order functions $f_{i,j}(z_i,z_j)$ represent the effect of interaction between $z_i$ and $z_j$ on the response $f$.
-Formally, the zeroth-, first- and second-order terms are defined as:
+The zeroth-order function $f_0$ is the mean response of $f$, the first-order univariate functions $f_i(z_i)$ quantify independent contributions due to each random input, and second-order functions $f_{i,j}(z_i,z_j)$ represent the effect of interaction between $z_i$ and $z_j$ on the response $f$:
 \begin{align}\label{sobol_terms}
@@
 \end{align}
-where ${\mathcal{D}_{\mathbf{z}}}^{M-1} = [0,1]^{M-1}$ and the notation $\mathbf{z}_{\sim \{i,j\}}$ denotes the vector having all the components of $\mathbf{z}$ except in the set $\{i,j\}$.
+where the notation $\mathbf{z}_{\sim \{i,j\}}$ denotes the vector having all the components of $\mathbf{z}$ except in the set $\{i,j\}$.
 Higher-order terms are interpreted in a similar manner.
@@
 
-Sobol indices are based on a variance decomposition \cite{SOBOL2001271}. The definition of the variance is:
+Sobol indices are based on the decompositon of the variance \cite{SOBOL2001271}, defined as:
 \begin{equation}\label{tot_var}
@@
 The total effect of $z_i$ on the output $Y$ can be quantified using the total sensitivity index $S_{T_i} = S_i + \sum_{j=1}^M S_{ij}$.
-The total sensitivity indices can be interpreted as an importance measure for the parameter $z_i$, therefore a large $S_{T_i}$ implies that $z_i$ has a strong influence on $Y$.
+The total sensitivity indices can be interpreted as an importance measure for the parameter $z_i$: a large $S_{T_i}$ implies that $z_i$ has a strong influence on $Y$.
 The computation of $S_{T_i}$ requires the approximation of partial variances defined in \eqref{partial_variance2}.
-Using the Monte Carlo method for computing these variances can be prohibitive for computationally expensive models $Y$.
-On the other hand, Polynomial Chaos Expansions (PCE) based approaches provide a more efficient alternative for sensitivity analysis, and will be used here.
-As shown in \cite{SUDRET2008964}, one can compute the Sobol indices analytically by post-processing the PCE coefficients, whose computation will be detailed next.
+Using the Monte Carlo method for computing these variances is computationally very expensive, and we therefore use instead a Polynomial Chaos Expansions (PCE) based approach.
+As shown in \cite{SUDRET2008964}, one can compute the Sobol indices analytically by post-processing the PCE coefficients.
 
@@
 The choice of univariate orthogonal polynomial $\psi_{k_i}$ depends on the type of the random variable $z_i$.
-For example, for uniform random variables the Legendre family of polynomials is used.
+For the uniform random variables appearing in this work Legendre polynomials are employed.
 %The multivariate polynomials $\Psi_{\mathbf{k}}(\mathbf{z})$ is orthogonal with respect to the joint probability distribution of random vector $\mathbf{z}$. 
-The total number of terms $K$ in \eqref{eq:PCE} depends on the truncation scheme; in a standard truncation scheme  all polynomials in $M$ random variable with order less than or equal to $p$ are used.
-In this work, a more advanced truncation scheme known as the hyperbolic truncation scheme is used \cite{BlatmanThesis}.
-%\textcolor{red}{In this work, we use the hyperbolic truncation scheme with a $qnorm = 0.75$.}
-
+The total number of terms $K$ in \eqref{eq:PCE} depends on the truncation scheme; in this work, a hyperbolic truncation scheme is used \cite{BlatmanThesis}.
 The PCE coefficients $w_{\mathbf{k}}$ corresponding to $\Psi_{\mathbf{k}}(\mathbf{z})$ can be computed using a pseudo-spectral approach by using quadrature methods.
@@
 Given $N$ model evaluations $\mathbf{Y} = \{f(\mathbf{z}_1), f(\mathbf{z}_2), ..., f(\mathbf{z}_N)\}$, we can obtain the coefficients as $\overline{\mathbf{w}} = (\mathbf{X}^T\mathbf{X})^{-1}\mathbf{X}^T\mathbf{Y}$, where $\mathbf{X}\in \mathbb{R}^{N\times K}$ is the design matrix where $\mathbf{X}_{i,j} = \Psi_{j}(\mathbf{z}_i)$, $i = 1,2, ..., N$ and $j = 0,1,2,..., K$.
-Although the OLS method allows to compute the coefficients using a limited number of model evaluations, it may result in a PCE model that consists of high-order interaction terms that may lead to overfitting.
+Although the OLS method allows to compute the coefficients using a limited number of model evaluations, it may result in a PCE model that consists of high-order interaction terms leading to overfitting.
 For many engineering problems, low-order interactions between the input variables are most important.
@@
 \end{equation}
-The LARS algorithm in the context of PCE can be summarized as follows.
+The LARS algorithm in the context of PCE starts with all PCE coefficients set to zero, and then iteratively selects polynomials based on the correlation with the current residual.
+After every iteration an a-posteriori error, for instance the Leave-One-Out (LOO) cross-validation error, is computed.
+The active set with the smallest LOO is then chosen as the best sparse PCE model.
+One of the main advantages of the LARS algorithm is that it works well in cases when the number of polynomials in the candidate set is much larger than the number of model evaluations.
+
+%The LARS algorithm in the context of PCE can be summarized as follows.
 We start with all PCE coefficients for the candidate polynomials set to zero, and find the polynomial, say $\Psi_1(\mathbf{w})$ most correlated with the model evaluations.
@@
 The reference value of these polars is obtained from the wind-tunnel experiment with 3D corrections.
-Random samples of chord, twist, lift- and drag-polars are obtained by perturbing the control points with a uniformly distributed random variable.
+Random samples of chord, twist, lift- and drag-polars are obtained by perturbing the control points with a uniformly distributed random variable \hl{based on Latin Hypercube Sampling}.
 
@@
 The twist and chord are perturbed following equation \eqref{eqn:perturb_c}, whereas the lift and drag are perturbed according to equation \eqref{eqn:perturb_cl}.
-Examples of the resulting lift and drag curves are shown in Fig.\ \ref{perturbed_samples} (c) - (d), respectively.
+\hl{Note that the multiplication of $(1+\Delta_i)$ with the unperturbed value (equations \eqref{eqn:perturb_c}, \eqref{eqn:perturb_cl}) has some physical intuition (especially for the lift and drag curves), but causes relatively large input uncertainty for the chord at the inboard sections.
+Similarly, we tend to obtain small perturbations for control points with values close to zero.} Examples of the resulting lift and drag curves are shown in Fig.\ \ref{perturbed_samples} (c) - (d), respectively.
 For simplicity, we only consider uniform distributions for perturbing the control points, however, we can also use other probability distributions for this purpose.
-Also note that when using a uniform distribution with bounds given by the relative magnitude of control points, we tend to obtain small perturbations for control points with values close to zero.
 \begin{figure}[ht!]
@@
 In Fig.\ \ref{convergence_dim}, we show for both OLS and LARS the convergence behaviour of the total Sobol indices of the power output with increasing number of model evaluations $N$ (runs of the AeroModule), and with increasing number of uncertain inputs $M$ (4, 6 and 8 control points).
-We see that both methods scale well with increasing dimensions, however, LARS exhibits faster convergence than OLS in high dimensions due to its adaptive nature, and will be the method of choice in the subsequent tests.
-For both OLS and LARS, we start with a small $N$ and double it with every iteration until converged values of Sobol indices are obtained.
+We see that both methods scale well with increasing dimensions, however, LARS exhibits faster convergence than OLS in high dimensions due to its adaptive nature.
+\hl{This is consistent with results from literature \cite{uqlab}, and therefore LARS will be the method of choice in the subsequent tests.} For both OLS and LARS, we start with a small $N$ and double it with every iteration until converged values of Sobol indices are obtained.
 %}  \todo{what is a reasonable choice of $N$, depending on $M$?}
@@
 We note that these findings are testcase-specific and might not generalize directly to other turbines or other operational settings.
-Furthermore, the reported sensitivities depend on the variance that is chosen for the uncertain parameters; in future work we will relate these to values known from design standards.
+Furthermore, the reported sensitivities depend on the variance that is chosen for the uncertain parameters.
+For example, \hl{the large sensitivity of the power with respect to uncertainty in Ch2 as shown in figure \ref{fig:sensitivity_analysis_power} is mainly due to the fact that the effective input uncertainty is larger here than at the outboard stations because of the multiplication with the local chord value (see equation \eqref{eqn:perturb_c}).
+Similarly, the low sensitivity with respect to twist as apparent in figures \ref{fig:sensitivity_analysis_power}-\ref{fig:sensitivity_analysis_force} is due to the fact that the effective perturbation in the twist is rather small: around 0.5 degree at the mid-span location (see equation \eqref{eqn:perturb_c}).} In future work we will relate these perturbations to values known from design standards to arrive at more realistic predictions.
 

### 2020-06-02 Calibration/BS_tests/linear_regression.m

@@ -13,8 +13,8 @@ rng default
 % select which methods to run
 runOLS = 1; % Ordinary Least Squares
 runUQLab = 1; % UQLab
-runIQR = 1; % Implicit quadrature rule
-runSampling = 1; % Brute-force sampling of prior 
+runIQR = 0; % Implicit quadrature rule
+runSampling = 0; % Brute-force sampling of prior 
 
 %% input settings
 p = 2; % dimension of parameters to be calibrated

### 2020-06-02 Papers/Torque_Paper/revision/content.tex
rebuttal and small changes

@@
 
-Recently, several papers have addressed the uncertainty in BEM model output by performing uncertainty propagation and sensitivity studies, e.g.\ \cite{Echeverria2017,Matthaus2017,Murcia2018,Robertson2018,Bos2019c}.
+Recently, several papers have addressed the uncertainty in BEM model output by performing uncertainty propagation and sensitivity studies, e.g.\ \cite{Petrone2011b,Echeverria2017,Matthaus2017,Murcia2018,Robertson2018,Bos2019c}.
 In these studies, the focus in mainly on uncertainties in the external conditions (wind parameters) and/or uncertainty in the turbine specification (geometric parameters).
@@
 The twist and chord are perturbed following equation \eqref{eqn:perturb_c}, whereas the lift and drag are perturbed according to equation \eqref{eqn:perturb_cl}.
-\hl{Note that the multiplication of $(1+\Delta_i)$ with the unperturbed value (equations \eqref{eqn:perturb_c}, \eqref{eqn:perturb_cl}) has some physical intuition (especially for the lift and drag curves), but causes relatively large input uncertainty for the chord at the inboard sections.
+\hl{Note that the multiplication of $(1+\Delta_i)$ with the unperturbed value (equations \eqref{eqn:perturb_c}, \eqref{eqn:perturb_cl}) has some physical intuition (especially for the lift and drag curves), but causes relatively large input uncertainty for the chord at the inboard sections (see dashed lines in figure \ref{perturbed_samples}).
 Similarly, we tend to obtain small perturbations for control points with values close to zero.} Examples of the resulting lift and drag curves are shown in Fig.\ \ref{perturbed_samples} (c) - (d), respectively.
@@
 Furthermore, the reported sensitivities depend on the variance that is chosen for the uncertain parameters.
-For example, \hl{the large sensitivity of the power with respect to uncertainty in Ch2 as shown in figure \ref{fig:sensitivity_analysis_power} is mainly due to the fact that the effective input uncertainty is larger here than at the outboard stations because of the multiplication with the local chord value (see equation \eqref{eqn:perturb_c}).
-Similarly, the low sensitivity with respect to twist as apparent in figures \ref{fig:sensitivity_analysis_power}-\ref{fig:sensitivity_analysis_force} is due to the fact that the effective perturbation in the twist is rather small: around 0.5 degree at the mid-span location (see equation \eqref{eqn:perturb_c}).} In future work we will relate these perturbations to values known from design standards to arrive at more realistic predictions.
+For example, \hl{the large sensitivity of the power with respect to uncertainty in Ch2 as shown in figure \ref{fig:sensitivity_analysis_power} is mainly due to the fact that the effective input uncertainty is larger here than at the outboard stations because of the multiplication with the local chord value (see equation \eqref{eqn:perturb_c} and figure \ref{perturbed_samples}).
+Similarly, the low sensitivity with respect to twist as apparent in figures \ref{fig:sensitivity_analysis_power}-\ref{fig:sensitivity_analysis_force} is due to the fact that the effective perturbation in the twist is rather small: around 0.5 degree at the mid-span location.} In future work we will relate these perturbations to values known from design standards to arrive at more realistic predictions.
 
@@
 \end{figure}
-
 \begin{figure}[h!]
@@
 The framework can be used for high-dimensional global sensitivity analysis of both geometric and model uncertainties in BEM models.
-As such, the proposed framework is generic and can deal with many types of (parametric) uncertainties, of which a few were investigated in this paper: twist, chord, lift coefficient and drag coefficient.
+The proposed framework is generic and can deal with many types of (parametric) uncertainties, of which a few were investigated in this paper: twist, chord, lift coefficient and drag coefficient.
 By using NURBS parameterizations of chord, twist, lift and drag as a function of the blade radius, the number of uncertainties was limited but remains considerable (19 uncertain variables).
